### 2012-11-02 paper/paper.tex
start writing a paper.

\documentclass[9pt]{sigplanconf}
\usepackage{url}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{framed}
\usepackage{color}
\usepackage{url}
\usepackage{alltt}
\usepackage{tabularx}
\usepackage{subfigure}
\usepackage{fancyhdr}
\usepackage{paralist}

\usepackage{pgf}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}

%% \include{math}

%% \newenvironment{code}{\begin{framed}\begin{alltt}\scriptsize}{\end{alltt}
%%   \end{framed}}
\newenvironment{code}{\begin{alltt}\footnotesize}{\end{alltt}}
\newenvironment{cols}{\begin{tabular}{m{3.6cm}|m{3.6cm}}{Haskell} &
    {\scriptsize Copilot}\\\hline}{\end{tabular}}

\usepackage{ifthen}
\newboolean{submission}  %set to true for the submission version
%\setboolean{submission}{false}
\setboolean{submission}{true}
\ifthenelse
{\boolean{submission}}
{ \newcommand{\todo}[1]{ } } %hide todo
{ \newcommand{\todo}[1]{ {\color{blue}$<<$#1$>>$}
 }}
%\usepackage{fancyhdr}

\begin{document}

%% \conferenceinfo{ICFP'12,} {September 9--15, 2012, Copenhagen, Denmark.}
%% \CopyrightYear{2012}
%% \copyrightdata{978-1-4503-1054-3/12/09}

%\titlebanner{banner above paper title}        % These are ignored unless
%\preprintfooter{}   % 'preprint' option specified.

\title{SmartCheck}
%\subtitle{Subtitle Text, if any}

\authorinfo{Lee Pike}
           {Galois, Inc.}
           {leepike@galois.com}
\maketitle

\begin{abstract}
foobar
\end{abstract}

%% \category{D.2.4}{Software/Program Verification}{Reliability}

%% \terms
%% Languages, Verification

%% \keywords
%% embedded domain-specific language, compiler, verification

\section{Introduction}\label{sec:intro}

\paragraph{Outline.}

%--------------------------------------------------------------------------------

\begin{code}
fib :: [Word32]
fib = [0,1] ++ zipWith (+) fib (drop 1 fib)
\end{code}
%

%--------------------------------------------------------------------------------
\section{Related Work}
\label{sec:related}

\cite{}

%--------------------------------------------------------------------------------
\section{Conclusions}
\label{sec:conclusions}

%% \balancecolumns

\bibliographystyle{abbrvnat}
\bibliography{paper}

\end{document}

### 2012-11-10 paper/paper.tex
paper updates.

@@
     {\scriptsize Copilot}\\\hline}{\end{tabular}}
+
+\newcommand{\ttp}[1]{\texttt{#1}}
 
@@
 
+The QuickCheck testing framework was a revolutionary step-forward in
+type-directed random testing~\cite{qc}.  Nowadays, QuickCheck is a standard
+stage of a program development cycle in Haskell.
+
+Because QuickCheck generates random values for testing, counterexamples it finds
+may not be minimal.  In their original QuickCheck paper~\cite{qc}, the authors
+report the following user experience by Andy Gill:
+%
+\begin{quote}
+Sometimes the counter examples found are very large and it is difficult to go
+back to the property and understand why it is a counter example.
+\end{quote}
+%
+\noindent
+QuickCheck defines a type class \ttp{Arbitrary} that defines a function
+\ttp{arbitrary} for generating random values of a given type.  Gill added another
+method to the type class:
+%
+\begin{code}
+smaller :: a -> [a] 
+\end{code}
+%
+\noindent
+The purpose of \ttp{smaller} is to generate strictly smaller values (according to
+some measure) from a given counterexample to attempt to find a smaller
+counterexample.  Today, \ttp{smaller} is called \ttp{shrink}.
+
+One of the dirty little secrets of QuickCheck is that even for simple data types
+and simple properties, counterexamples can be much larger than the minimal
+counterexamples.  Intuitively, there are two reasons: (1) because values are
+generated randomly, a smaller value that produces a counterexample is skipped;
+and (2) for a composite type, larger values are generated in tandem.  For
+example, consider the following data type and property:
+%
+\begin{code}
+data A = A Int deriving Show
+
+instance Arbitrary A where
+  arbitrary = liftM A arbitrary
+
+prop :: [A] -> Property
+prop x = property $ length x < 5
+\end{code}
+%
+\noindent
+where \ttp{property} is a QuickCheck library function that lifts a \ttp{Bool} into
+the \ttp{Property} type.  (We wrap \ttp{Int} in a constructor to prevent QuickCheck
+from using a default implementation of \ttp{shrink} for \ttp{Int}.)  A typical
+counterexample returned by QuickCheck is
+%
+\begin{code}
+> [A (-2),A (-9),A (-26),A (-28),A (-22)]  
+\end{code}
+%
+\noindent
+Whereas the smallest counterexample is
+%
+\begin{code}
+> [A 0,A 0,A 0,A 0,A 0]
+\end{code}
+%
+\noindent
+The problem is that while QuickCheck is generating larger lists, it is also
+generating larger integers.
+
+SmallCheck is another testing framework that is guaranteed to return the a
+smallest counterexample, if one exists~\cite{sc}.  SmallCheck does this by
+enumerating all possible inputs, ordered from smallest to largest, up to some
+user-defined bound.  While SmallCheck is effective for testing many programs and
+properties (in accordance with the small model principle~\cite{jackson}),
+counterexamples deep in the state space may be infeasible to uncover.  For
+example, consider again our data type \ttp{A} defined above, used in the
+following property:
+%
+\begin{code}
+data B = B A A A A deriving Show
+
+prop :: B -> Bool
+prop (B a b c d) = and $ map f [a, b, c, d]
+  where f (A x) = x < 100
+\end{code}
+%
+\noindent
+With SmallCheck, a property is checked up to a depth set by the user, measuring
+how large values are.  For example, for type \ttp{Int}, the depth \ttp{d} is
+defined as the integers in the list \ttp{(-d)..d]}.  For a product type,
+  SmallCheck generates the cross product of its fields.  In the example above,
+  SmallCheck (or Lazy SmallCheck) will find a counterexample at depth 101,
+  requiring approximately $100^4$ test cases, which is infeasible to compute.\footnote{This
+  is an upper-bound, since Lazy SmallCheck eliminates some redundant test-cases,
+  but computing a counterexample is still infeasible.}
+
+quantified away
+
+and straightforward
+implementations of shrink are ineffective.
+
+writing good shrink functions is tedious, difficult, or
+both.
+
+In principle, for a given counterexample, \ttp{shrink} can be implemented to
+find a local minimum within the neighborhood of the counterexample.
+
+One of the
+dirty little secrets of QuickCheck is that even for simple data types and simple
+properties, counterexamples can be large, and straightforward implementations of
+\ttp{shrink} can miss finding minimal counterexamples, by a lot.
+
+%% For example, consider a trivial arithmetic expression language, containing
+%% constants, addition, and division:
+%% %
+%% \begin{code}
+%% data Exp = C Int
+%%          | Add Exp Exp
+%%          | Div Exp Exp
+%% \end{code}
+%% %
+%% \noindent
+%% And consider the following evaluation function that returns \ttp{Nothing} if the
+%% expression contains division by zero and
+
+%% \todo{Example}
+
+\todo{then explain why---need to understand property}
+
+\todo{SmallCheck is guaranteed to find minimal counterexamples.}
+
+\todo{insight: shrink based on arbitrary}
+
+\url{http://stackoverflow.com/questions/8788542/how-do-i-get-good-small-shrinks-out-of-quickcheck}
+
 \paragraph{Outline.}
@@
 
+\todo{\cite{syb} paper is motivated by generic shrink}
+\todo{shrink paper rehger pointed to ``Automatic isolation of compiler errors'' from 1994}
+
 \cite{}
@@
 \label{sec:conclusions}
+
+\section*{Acknowledgements}
+\todo{Thank Rehger for comments/pointing to paper.}
 

### 2012-11-12 examples/Paper/Paper.hs
paper benchmarks.

@@ -1,0 +1,202 @@
+{-# LANGUAGE DeriveDataTypeable #-}
+{-# LANGUAGE DeriveGeneric #-}
+
+-- | Try to generate a very large counterexample.
+
+module Paper where
+
+import Test.SmartCheck
+import Test.SmartCheck.Reduce
+import Test.QuickCheck
+import Test.LazySmallCheck hiding (Property, test, (==>))
+import qualified Test.LazySmallCheck as S
+
+import GHC.Generics hiding (P, C)
+import Data.Typeable
+import Control.Monad
+
+import Data.Int
+import Data.List
+import Data.Maybe
+import Data.Time
+import System.Environment
+
+-----------------------------------------------------------------
+
+-- Let's try to generate a product type of long lists when all we need is a
+-- single element to have a long list.
+
+-- Container so that we don't have base types.
+data A = A Int16
+  deriving (Read, Show, Typeable, Generic)
+
+instance Serial Int16 where
+  series d = drawnFrom [(-d')..d']
+    where d' = fromIntegral d
+
+instance SubTypes A
+
+instance Arbitrary A where
+  arbitrary = liftM A arbitrary
+  shrink (A i) = map A (shrink i)
+
+instance Serial A where
+  series = cons1 A 
+data B = B [A] [A] [A] [A]
+  deriving (Read, Show, Typeable, Generic)
+
+instance SubTypes B
+
+-- qc/shrink takes over 1m seconds
+instance Arbitrary B where
+  arbitrary = liftM4 B arbitrary arbitrary arbitrary arbitrary
+  shrink (B a b c d) = 
+    if defShrink 
+      then [ B w x y z | 
+             w <- tk a
+           , x <- tk b
+           , y <- tk c
+           , z <- tk d ]
+      else []
+    where tk x = take 10 (shrink x)
+
+instance Serial B where
+  series = cons4 B 
+
+size :: B -> Int
+size (B a b c d) = length a + length b + length c + length d
+
+add :: [A] -> Int16
+add = sum . map (\(A i) -> i)
+
+pre :: B -> Bool
+pre (B a b c d) = and $ map pre' [a, b, c, d]
+  where
+  pre' x = add x < 16
+
+post :: B -> Bool
+post (B a b c d) = add a + add b + add c + add d < 64
+
+prop :: B -> Property
+prop p = pre p ==> post p
+
[…]

### 2012-11-12 examples/Div0.hs
Example tweaks.

@@ -14,48 +14,50 @@ import Data.Typeable
 
 -----------------------------------------------------------------
 
-data M = C Int
-       | A M M
-       | D M M
+data Exp = C Int
+         | Add Exp Exp
+         | Div Exp Exp
   deriving (Read, Show, Typeable, Generic)
 
-instance SubTypes M 
+instance SubTypes Exp 
 
-eval :: M -> Maybe Int
+eval :: Exp -> Maybe Int
 eval (C i) = Just i
-eval (A a b) = do
-  i <- eval a 
-  j <- eval b
-  return $ i + j
-eval (D a b) = 
-  if eval b == Just 0 then Nothing 
-    else do i <- eval a 
-            j <- eval b
-            return $ i `div` j
+eval (Add e0 e1) =
+  liftM2 (+) (eval e0) (eval e1)
+eval (Div e0 e1) = 
+  let e = eval e1 in 
+  if e == Just 0 then Nothing 
+    else liftM2 div (eval e0) e
 
-instance Arbitrary M where
+instance Arbitrary Exp where
   arbitrary = sized mkM
     where
     mkM 0 = liftM C arbitrary
-    mkM n = oneof [ liftM2 A mkM' mkM' 
-                  , liftM2 D mkM' mkM' ]
+    mkM n = oneof [ liftM2 Add mkM' mkM' 
+                  , liftM2 Div mkM' mkM' ]
       where mkM' = mkM =<< choose (0,n-1)
 
-  -- shrink (C _)   = []
-  -- shrink (A a b) = [a, b]
-  -- shrink (D a b) = [a, b]
+  shrink (C i)       = map C (shrink i)
+  shrink (Add e0 e1) = [e0, e1]
+  shrink (Div e0 e1) = [e0, e1]
 
 -- property: so long as 0 isn't in the divisor, we won't try to divide by 0.
 -- It's false: something might evaluate to 0 still.
-div_prop :: M -> Property
-div_prop m = divSubTerms m ==> eval m /= Nothing
+div_prop :: Exp -> Property
+--div_prop e = divSubTerms e ==> eval e /= Nothing
+div_prop e = property $ case x of
+                          Nothing -> True
+                          Just True -> True
+                          _       -> False
+  where x = fmap (< 1) (eval e)
 
   -- precondition: no dividand in a subterm can be 0.
-divSubTerms :: M -> Bool
-divSubTerms (C _)       = True
-divSubTerms (D _ (C 0)) = False
[…]

### 2012-11-12 examples/Large.hs

@@ -7,13 +7,13 @@ module Large where
 
 import Test.SmartCheck
 import Test.QuickCheck
-import Test.LazySmallCheck hiding (Property, test)
-import qualified Test.LazySmallCheck as S
+import Test.LazySmallCheck hiding (Property, test, (==>))
 
 import GHC.Generics hiding (P, C)
 import Data.Typeable
 import Control.Monad
-import Data.List
+
+import Data.Int
 
 -----------------------------------------------------------------
 
@@ -21,47 +21,93 @@ import Data.List
 -- single element to have a long list.
 
 -- Container so that we don't have base types.
-data C = C Int
+data A = A Int16
   deriving (Read, Show, Typeable, Generic)
 
-instance SubTypes C
+instance Serial Int16 where
+  series d = drawnFrom [(-d')..d']
+    where d' = fromIntegral d
 
-instance Arbitrary C where
-  arbitrary = liftM C arbitrary
-  shrink (C i) = map C (shrink i)
+instance SubTypes A
 
-instance Serial C where
-  series = cons1 C
+instance Arbitrary A where
+  arbitrary = liftM A arbitrary
+  shrink (A i) = map A (shrink i)
 
-data P = P [C] [C] [C] [C]
+instance Serial A where
+  series = cons1 A
+
+data B = B [A] [A] [A] [A]
   deriving (Read, Show, Typeable, Generic)
 
-instance SubTypes P
+instance SubTypes B
 
-instance Arbitrary P where
-  arbitrary = liftM4 P arbitrary arbitrary arbitrary arbitrary
---  shrink (P a b c d) = [ P w x y z | w <- shrink a, x <- shrink b, y <- shrink c, z <- shrink d ]
+-- qc/shrink takes over 1m seconds
+instance Arbitrary B where
+  arbitrary = liftM4 B arbitrary arbitrary arbitrary arbitrary
+--  shrink (B a b c d) = [ B w x y z | w <- shrink a, x <- shrink b, y <- shrink c, z <- shrink d ]
 
-instance Serial P where
-  series = cons4 P 
+instance Serial B where
+  series = cons4 B 
 
-sumC :: [C] -> Int
-sumC = foldl' (\acc (C c) -> acc + c) 0 
+add :: [A] -> Int16
+add = sum . map (\(A i) -> i)
 
-test :: P -> Bool
-test (P a b _ _) = sumC a - sumC b < 4096
+pre :: B -> Bool
[…]

### 2012-11-12 paper/paper.tex
paper updates.

@@
 type-directed random testing~\cite{qc}.  Nowadays, QuickCheck is a standard
-stage of a program development cycle in Haskell.
-
-Because QuickCheck generates random values for testing, counterexamples it finds
-may not be minimal.  In their original QuickCheck paper~\cite{qc}, the authors
-report the following user experience by Andy Gill:
+stage of a program development cycle in Haskell.  Because QuickCheck generates
+random values for testing, counterexamples it finds may not be substantially
+larger than a minimal counterexample.  In their original QuickCheck
+paper~\cite{qc}, the authors report the following user experience by Andy Gill:
 %
@@
 
-One of the dirty little secrets of QuickCheck is that even for simple data types
-and simple properties, counterexamples can be much larger than the minimal
-counterexamples.  Intuitively, there are two reasons: (1) because values are
-generated randomly, a smaller value that produces a counterexample is skipped;
-and (2) for a composite type, larger values are generated in tandem.  For
-example, consider the following data type and property:
+This trouble with randomly generating large counterexamples is compounded in
+composite data types, for which increasingly larger sub-values are generated in
+tandem.  For example, consider the following data type and property:
 %
@@
 prop :: [A] -> Property
-prop x = property $ length x < 5
+prop x = property (length x < 5)
 \end{code}
@@
 properties (in accordance with the small model principle~\cite{jackson}),
-counterexamples deep in the state space may be infeasible to uncover.  For
+counterexamples to even relatively simple properties may be infeasible to
+uncover for product types.  For
 example, consider again our data type \ttp{A} defined above, used in the
@@
   SmallCheck (or Lazy SmallCheck) will find a counterexample at depth 101,
-  requiring approximately $100^4$ test cases, which is infeasible to compute.\footnote{This
-  is an upper-bound, since Lazy SmallCheck eliminates some redundant test-cases,
-  but computing a counterexample is still infeasible.}
+  requiring on the order of $200^4$ test cases to be executed; executing $10e^6$
+  tests requires just under four minutes on a Pentium i7 at
+  2.7GHz.\footnote{This is an upper-bound, since Lazy SmallCheck eliminates some
+    redundant test-cases.}
+
+Furthermore, even if a minimal counterexample is found, regardless of the testing
+framework used, it tells the user nothing about related counterexamples.  That
+is, it would be nice for a counterexample to be generalized so that the user is
+presented with a formula characterizing a set of counterexamples.
+
+For
+example,
 

### 2012-11-12 src/Test/SmartCheck.hs
export runQC for benchmarking.

@@ -5,6 +5,7 @@
 
 module Test.SmartCheck 
   ( smartCheck
+  , runQC
   , module Test.SmartCheck.Types
   ) where
 

### 2012-11-13 paper/paper.tex
paper updates.

--- 2012-11-02 paper/paper.tex
@@
 \usepackage{pgf}
-\usepackage{tikz}
-\usetikzlibrary{arrows,automata}
+%% \usepackage{tikz}
+%% \usetikzlibrary{arrows,automata}
+
+%% \usepackage{amsmath}
+%% \usepackage{amsthm}
+
+%% \newtheoremstyle{lst}
+%%   {\topsep}   % ABOVESPACE
+%%   {\topsep}   % BELOWSPACE
+%%   {\itshape}  % BODYFONT
+%%   {0pt}       % INDENT (empty value is the same as 0pt)
+%%   {} % HEADFONT
+%%   {}         % HEADPUNCT
+%%   {5pt plus 1pt minus 1pt} % HEADSPACE
+%%   {(\thmnumber{#2})}          % CUSTOM-HEAD-SPEC
+%% \theoremstyle{lst}
+%% \newtheorem{codenum}{lst}
+
+\usepackage{float}
+\floatstyle{boxed} 
+\restylefloat{figure}
 
@@
 
-%% \newenvironment{code}{\begin{framed}\begin{alltt}\scriptsize}{\end{alltt}
-%%   \end{framed}}
-\newenvironment{code}{\begin{alltt}\footnotesize}{\end{alltt}}
-\newenvironment{cols}{\begin{tabular}{m{3.6cm}|m{3.6cm}}{Haskell} &
-    {\scriptsize Copilot}\\\hline}{\end{tabular}}
+\newenvironment{code}{\begin{alltt}\scriptsize}{\end{alltt}}
+%% \newenvironment{cols}{\begin{tabular}{m{3.6cm}|m{3.6cm}}{Haskell} &
+%%     {\scriptsize Copilot}\\\hline}{\end{tabular}}
+
+\newcommand{\ttp}[1]{\texttt{#1}}
 
@@
 
+The QuickCheck testing framework was a revolutionary step-forward in
+type-directed random testing~\cite{qc}.  Nowadays, QuickCheck is a standard
+aspect of the Haskell development cycle.
+
+Because QuickCheck generates
+random values for testing, counterexamples it finds may not be substantially
+larger than a minimal counterexample.  In their original QuickCheck
+paper~\cite{qc}, the authors report the following user experience by Andy Gill:
+%
+\begin{quote}
+Sometimes the counter examples found are very large and it is difficult to go
+back to the property and understand why it is a counter example.
+\end{quote}
+%
+\noindent
+QuickCheck defines a type class \ttp{Arbitrary} that defines a function
+\ttp{arbitrary} for generating random values of a given type.  Gill added another
+method to the type class:
+%
+\begin{code}
+smaller :: a -> [a] 
+\end{code}
+%
+\noindent
+The purpose of \ttp{smaller} is to generate strictly smaller values (according to
+some measure) from a given counterexample to attempt to find a smaller
+counterexample.  Today, \ttp{smaller} is called \ttp{shrink}.
+
+\begin{figure}[ht]
+\begin{code}
+data A = A Int16 deriving Show
+
+data B = B [A] [A] [A] [A] deriving Show
+
+-- QuickCheck instances
+instance Arbitrary A where
+  arbitrary    = liftM A arbitrary
+  shrink (A i) = map A (shrink i)
+
+instance Arbitrary B where
+  arbitrary = liftM4 B arbitrary arbitrary arbitrary arbitrary
+  shrink (B a b c d) = 
+    [ B w x y z | w <- shrink a, x <- shrink b
+                , y <- shrink c, z <- shrink d ]
+
+-- SmallCheck instances
+instance Serial Int16 where
+  series d = drawnFrom [(-d')..d'] where d' = fromIntegral d
+
+instance Serial A where series = cons1 A 
+instance Serial B where series = cons4 B
+
+-- Predicates for defining properties
+pre :: B -> Bool
+pre (B a b c d) = and \$ map pre' [a, b, c, d]
+  where pre' x = add x < 16
+
+post :: B -> Bool
+post (B a b c d) = add a + add b + add c + add d < 64
+  where add :: [A] -> Int16
+        add = sum . map (\(\backslash\)(A i) -> i)
+
+-- Property
+prop :: B -> Property
+prop p = pre p ==> post p
+\end{code}
+  \caption{QuickCheck and SmallCheck for a product type input.}
+  \label{fig:initial}
+\end{figure}
+
+While \ttp{shrink} is user-defined and can have any type-correct implementation,
+it is typically defined using structural recursion for composite data types.
+For example, consider the \ttp{shrink} instances given to data types \ttp{A} and
+\ttp{B} in Figure~\ref{fig:initial}.  Indeed, the principal motivation for one
+of Haskell's first generics package was for providing generic definitions for
+\ttp{shrink}~\cite{syb}.
+
+However, this straightforward approach to defining \ttp{shrink} instances may
+either miss minimal counterexamples, be too inefficient to be practical, or
+both.\footnote{E.g., see the user question posed on StackOverflow:
+  \url{http://stackoverflow.com/questions/8788542/how-do-i-get-good-small-shrinks-out-of-quickcheck}.}
+For example, consider again Figure~\ref{fig:initial}.  The data type \ttp{B} is
+a product type over lists of 16-bit signed integers, wrapped by a constructor.
+Consider the property \ttp{prop} in the figure, stating that if the sum of each
+list in a field of \ttp{B} is less than 16, then the sum of the four fields of
+\ttp{B} is less than 64.  \ttp{prop} seems reasonable at first glance, until
+one realizes that due to underflow, the property can be violated.  For example,
+consider the value
+%
+\begin{code}
+B [A (-32769)] [] [] []
+\end{code}
+%
+\noindent
+Without shrinking, QuickCheck finds large examples to the property; on average,
+a value contains 70 \ttp{Int16} values!
+
+Thus, it pays to define \ttp{shrink}.  Unfortunately, even for a simple property
+and program like this one, the definition for \ttp{shrink} given in
+Figure~\ref{fig:initial} produces an intractable number of potential
+counterexamples to test; with the definitions and property provided, which also
+rely on QuickCheck's default instances for the list type, the list of potential
+counterexamples generated by \ttp{shrink} can contain more than $10^{10}$
+elements.  This is an enormous state space to search for a smaller example!
+
+It is not always straightforward to determine how to make a more efficient
+implementation for \ttp{shrink}.  Shrinking presents a programmer's dilemma:
+shrinking is used to better understand a counterexample to understand why a
+property failed; but to implement an efficient shrink function in general
+requires knowing how to generate counterexamples to the property.
+
+Typically, a user tries a brute-force approach to reduce the number of potential
+counterexamples, for example by truncating lists using the \ttp{take n} function
+that returns the first $n$ elements of a list.  The trade-off is quicker
+shrinking with a lower probability of finding a smaller counterexample.  For
+example, redefining shrink for type \ttp{B} as follows
+%
+\begin{code}
+shrink (B a b c d) = [ B w x y z | w <- tk a, x <- tk b
+                                 , y <- tk c, z <- tk d ]
+  where tk x = take 10 (shrink x)
+\end{code}
+%
+\noindent
+controls the exponential blowup of the state-space by taking the first ten
+elements of each list, so the cross product is limited to 1000 elements.  The
+downside is that potentially smaller counterexamples may be omitted.
+
+What if we omit the need for shrinking counterexamples altogether?  SmallCheck
+is another testing framework for Haskell.  SmallCheck is guaranteed to return
+the a smallest counterexample, if one exists~\cite{sc}.  SmallCheck does this by
+enumerating all possible inputs, ordered from smallest to largest, up to some
+user-defined bound.  While SmallCheck is effective for testing many programs and
+properties (in accordance with the \emph{small scope hypothesis}~\cite{jackson}),
+counterexamples to even relatively simple properties may be infeasible to
+discover due to state-space explosion.
+
+With SmallCheck, a property is checked up to a depth set by the user, measuring
+how large values are.  For example, for type \ttp{Int}, the depth $d$ is defined
+as the integers in the list enumerating from $-d$ to $d$ (i.e.,
+\ttp{[(-d)..d]}).  For a product type, SmallCheck must check values produced by
+taking the cross-product of the type's fields at a given depth.  In
+Figure~\ref{fig:initial}, we have defined instances for generating Lazy
+SmallCheck (a more efficient version of SmallCheck) tests by defining instances
+of the \ttp{Serial} class.  For the property \ttp{prop} defined above (suitably
+redefined to use SmallCheck's types), SmallCheck must check to depth 16 to find
+the first counterexample.  Unfortunately, after a couple hours of testing, Lazy
+SmallCheck is still checking values at depth six, and the number of tests scales
+exponentially with respect to the depth.\footnote{All tests in this paper are
+  performed on a four-core Pentium~i7 running at 2.7GHz with 8GB RAM on
+  Fedora~16.  This test, and others unless noted, are performed using
+  interpreted Haskell under GHC 7.4.2.}
+
+Finally, even if we could overcome the problems described above with QuickCheck
+and SmallCheck, we are left with two others.  First, when a
+counterexample is discovered, one might wish to search for different
+counterexamples.  There currently is no automated way to add a predicate
+characterizing a counterexample as preconditions to a property to strengthen the
+property.  Second, and more generally, a counterexample is usually
+representative of other counterexamples, but that generalization is left to the
+user to discover for herself.  Really, what would be useful would be not just a
+minimal counterexample, but a quantified formula characterizing a set of
+counterexamples.  Such a formula has the added benefit of replacing a
+potentially large sub-term of a composite data type value with a quantifier.
+
+\begin{figure}[ht]
+\scriptsize
+  \begin{center}
+    \begin{tabular}{|r||c|c|c|c|}
+\hline 
+ & QC (none) & QC (10) & QC (20) & SmartCheck \\      
+\hline \hline 
+Max.  & 0.15s & 21.51s & 125.37s & 1.96s\\
+\hline
+Mean  & 0.07s & 1.21s & 3.80s & 0.30s\\
+\hline
+Median & 0.07s & 0.48s & 0.52s & 0.24s\\
+\hline
+Mean size & 70 & 34 & 32 & 5\\
+\hline
+    \end{tabular}
+  \end{center}
+  \caption{Results for 100 tests of the program in Figure~\ref{fig:initial}
+    using interpreted Haskell.}
+  \label{fig:results}
+\end{figure}
+
+\paragraph{SmartCheck}
+Motivated by these limitations of QuickCheck and SmallCheck, we have developed
+\emph{SmartCheck}.  SmartCheck extends QuickCheck to efficiently and generically
+find small counterexamples.  To motivate the benefit of SmartCheck, consider the
+table in Figure~\ref{fig:results}.  The table characterizes testing the program
+shown in Figure~\ref{fig:initial} that we have described so far.  Four runs are
+presented.  In the first three, we use QuickCheck to generate counterexamples.
+The first column contains the results of using QuickCheck without shrinking to
+provide a baseline.  The next two columns show the result of using QuickCheck
+with definitions of \ttp{shrink} that limit the lists of fields of \ttp{B} to 10
+(the ``QC (10)'' tests) and 20 (the ``QC (20)'' tests) elements, respectively,
+using a definition like the truncated implementation of \ttp{shrink} for the
+type \ttp{B} discussed earlier.  In each row, we show the maximum, mean, and
+median number of seconds required to by QuickCheck to generate a counterexample
+over 100 runs.  In the last row, we show the mean size of the final value
+returned, measured by the taking the cumulative length of the lists of the
+fields of \ttp{B}.  Notice that while typically counterexamples are returned in
+a fraction of a second, there are outliers taking 21 seconds and 125 seconds for
+the two tests.  Both tests roughly cut the size of the counterexample in half;
+increasing the number of potential counterexamples can result in a significant
+performance penalty for this example without much benefit.
+
+In the last column are the results under the same condition for SmartCheck.  Not
+only does it execute faster than QuickCheck with shrinking, it has significantly
+smaller outliers.  The most striking difference, however, is the size of the
+counterexample returned, reducing on average the size of a counterexample found
+by QuickCheck without shrinking by a factor of 14.
+
+The purpose of this paper is to explain first how to efficiently and generically
+generate small counterexamples for algebraic data types and second how to
+generalize the counterexamples..
+
+\paragraph{Contributions}
+This paper, and the library it describes, makes the following contributions:
+
+\begin{itemize}
+
+\item Using generic programming techniques, we describe efficient counterexample
+  reduction strategies for algebraic data types.
+
+\item We describe an approach to counterexample generalization, to present the
+  user with a formula describing a set of counterexamples to a property.
+
+\item We present an approach to generically strengthen a property with a
+  precondition that characterizes an already-discovered counterexample for the
+  purpose of finding new, dissimilar counterexamples.
+
+\item More generally, this paper is the first to explore the idea of efficient,
+  generic counterexample reduction and generalization for functional program testing.
+\end{itemize}
+
 \paragraph{Outline.}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\newpage
+
+\todo{insight: shrink based on arbitrary}
+
+\section{Shrinking Data}
+
+\section{Generalizing Data}
+
+\section{Implementation}
+
+\section{Experiments}
+
 %--------------------------------------------------------------------------------
-
-\begin{code}
-fib :: [Word32]
-fib = [0,1] ++ zipWith (+) fib (drop 1 fib)
-\end{code}
-%
 
@@
 \label{sec:related}
+
+\todo{\cite{syb} paper is motivated by generic shrink}
+\todo{shrink paper rehger pointed to ``Automatic isolation of compiler errors'' from 1994}
 
@@
 
+\section*{Acknowledgements}
+\todo{Thank Rehger for comments/pointing to paper.}
+
 %% \balancecolumns

### 2012-11-13 examples/Paper/Paper.hs

@@ -42,6 +42,7 @@ instance Arbitrary A where
 
 instance Serial A where
   series = cons1 A 
+
 data B = B [A] [A] [A] [A]
   deriving (Read, Show, Typeable, Generic)
 
@@ -71,14 +72,13 @@ add = sum . map (\(A i) -> i)
 
 pre :: B -> Bool
 pre (B a b c d) = and $ map pre' [a, b, c, d]
-  where
-  pre' x = add x < 16
-
+  where pre' x = add x < 16
+  
 post :: B -> Bool
 post (B a b c d) = add a + add b + add c + add d < 64
 
-prop :: B -> Property
-prop p = pre p ==> post p
+prop_qc :: B -> Property
+prop_qc p = pre p ==> post p
 
 prop_sc :: B -> Bool
 prop_sc p = pre p S.==> post p
@@ -130,20 +130,20 @@ test' run = do
 
 runSC :: IO (Maybe B)
 runSC = do
-  res <- runQC stdArgs prop
+  res <- runQC stdArgs prop_qc
   case res of 
     Nothing -> return Nothing
-    Just r  -> do smartRun scStdArgs r prop >>= return . Just
+    Just r  -> do smartRun scStdArgs r prop_qc >>= return . Just
                   
 
 --------------------------------------------------------------------------------
 
 defShrink :: Bool
-defShrink = True
+defShrink = False
 
 main :: IO ()
 main = do 
-  test $ runQC stdArgs prop
+  test $ runQC stdArgs prop_qc
 --  test runSC 
 
   -- _ <- test' $ smallCheck 7 prop_sc >> return (Just $ B [] [] [] [])
@@ -153,15 +153,17 @@ main = do
 {-
 -- RESULTS
 
+-- maybe a 2x noise in results
+
 ----------------------------------------
 QC, no shrinking:
 ----------------------------------------
 
 Num  : 100
-Max  : 0.646203s
-Avg  : 0.31339717s
-Med  : 0.302749s
-Size : 72.45
+Max  : 0.154356s
+Avg  : 0.0733841s
+Med  : 0.07041s
+Size : 70.25
 
 ----------------------------------------
 QC, shrinking:
@@ -180,23 +182,14 @@ Max  : 125.366509s
 Avg  : 3.79998622s
 Med  : 0.516224s
 Size : 32.34
+
 ----------------------------------------
 SC:
 ----------------------------------------
 Num  : 100
[…]

### 2012-12-03 paper/paper.tex
paper updates.

@@
 \usepackage{pgf}
+
+\usepackage{listings}
+\lstset{ %
+  language=Haskell,                 % the language of the code
+  basicstyle=\scriptsize,           % the size of the fonts that are used
+  %% for the code
+  %%firstnumber=0, % index to start numbering at
+  %%numbers=left,                      % where to put the line-numbers
+  %% numberstyle=\tiny\color{gray},  % the style that is used for the
+  %% line-numbers
+  stepnumber=5,                   % the step between two line-numbers. If it's
+                                  % 1, each line 
+  %%                              % will be numbered
+  numbersep=5pt,                  % how far the line-numbers are from the code
+  xleftmargin=0ex,
+  %% backgroundcolor=\color{white},      % choose the background color. You must
+  %% add \usepackage{color}
+  %% showspaces=false,               % show spaces adding particular underscores
+  %% showstringspaces=false,         % underline spaces within strings
+  %% showtabs=false,                 % show tabs within strings adding
+  %% particular underscores
+  %% frame=single,                   % adds a frame around the code
+  %% rulecolor=\color{black},        % if not set, the frame-color may be
+  %% changed on line-breaks within not-black text (e.g. commens (green here))
+  %% tabsize=2,                      % sets default tabsize to 2 spaces
+  %% captionpos=b,                   % sets the caption-position to bottom
+  %% breaklines=true,                % sets automatic line breaking
+  %% breakatwhitespace=false,        % sets if automatic breaks should only
+  %% happen at whitespace
+  %% title=\lstname,                   % show the filename of files included
+  %% with \lstinputlisting;
+  %%                                 % also try caption instead of title
+  %% keywordstyle=\color{blue},          % keyword style
+  %% commentstyle=\color{dkgreen},       % comment style
+  %% stringstyle=\color{mauve},          % string literal style
+  escapeinside={@*}{*@},            % if you want to add a comment within your
+                                    % code
+  %% morekeywords={*,...}           % if you want to add more keywords to
+  %% the set
+}
+
 %% \usepackage{tikz}
@@
 
+\newcommand{\sub}[1]{\(\sb{#1}\)}
+
 \begin{document}
@@
 type-directed random testing~\cite{qc}.  Nowadays, QuickCheck is a standard
-aspect of the Haskell development cycle.
-
-Because QuickCheck generates
-random values for testing, counterexamples it finds may not be substantially
-larger than a minimal counterexample.  In their original QuickCheck
-paper~\cite{qc}, the authors report the following user experience by Andy Gill:
+aspect of the Haskell development cycle.  Because QuickCheck generates random
+values for testing, counterexamples it finds may not be substantially larger
+than a minimal counterexample.  In their original QuickCheck paper~\cite{qc},
+the authors report the following user experience by Andy Gill:
 %
 \begin{quote}
-Sometimes the counter examples found are very large and it is difficult to go
-back to the property and understand why it is a counter example.
+Sometimes the counterexamples found are very large and it is difficult to go
+back to the property and understand why it is a counterexample.
 \end{quote}
@@
 \noindent
-QuickCheck defines a type class \ttp{Arbitrary} that defines a function
+QuickCheck defines a type class \ttp{Arbitrary} that presents a method
 \ttp{arbitrary} for generating random values of a given type.  Gill added another
@@
 \end{code}
-  \caption{QuickCheck and SmallCheck for a product type input.}
+  \caption{QuickCheck and SmallCheck for a product-type input.}
   \label{fig:initial}
@@
 For example, consider the \ttp{shrink} instances given to data types \ttp{A} and
-\ttp{B} in Figure~\ref{fig:initial}.  Indeed, the principal motivation for one
+\ttp{B} in Figure~\ref{fig:initial}.  Indeed, because of the regular structure
+of \ttp{shrink} implementations, the principal motivation for one
 of Haskell's first generics package was for providing generic definitions for
@@
 either miss minimal counterexamples, be too inefficient to be practical, or
-both.\footnote{E.g., see the user question posed on StackOverflow:
+both.\footnote{E.g., for a recent real-world case, see the user question posed on StackOverflow:
   \url{http://stackoverflow.com/questions/8788542/how-do-i-get-good-small-shrinks-out-of-quickcheck}.}
@@
 Without shrinking, QuickCheck finds large examples to the property; on average,
-a value contains 70 \ttp{Int16} values!
+QuickCheck returns a value contains 70 \ttp{Int16} values!
 
@@
 counterexamples generated by \ttp{shrink} can contain more than $10^{10}$
-elements.  This is an enormous state space to search for a smaller example!
+elements.  This is an enormous state space to search for a smaller example!  In
+particular, if a counterexample is not found does not finish in a reasonable
+amount of time.
 
@@
 Typically, a user tries a brute-force approach to reduce the number of potential
-counterexamples, for example by truncating lists using the \ttp{take n} function
+counterexamples, for example, by truncating lists using the \ttp{take n} function
 that returns the first $n$ elements of a list.  The trade-off is quicker
@@
 %
-\begin{code}
+\lstset{ caption={Truncated shrinking function for data type \ttp{B}.}
+       , label=lst:newshrink
+       ,captionpos=b }
+\begin{lstlisting}
+len = 10
+
 shrink (B a b c d) = [ B w x y z | w <- tk a, x <- tk b
                                  , y <- tk c, z <- tk d ]
-  where tk x = take 10 (shrink x)
-\end{code}
+  where tk x = take len (shrink x)
+\end{lstlisting}
+%% \begin{code}
+%% \end{code}
 %
@@
 
-What if we omit the need for shrinking counterexamples altogether?  SmallCheck
-is another testing framework for Haskell.  SmallCheck is guaranteed to return
-the a smallest counterexample, if one exists~\cite{sc}.  SmallCheck does this by
-enumerating all possible inputs, ordered from smallest to largest, up to some
-user-defined bound.  While SmallCheck is effective for testing many programs and
-properties (in accordance with the \emph{small scope hypothesis}~\cite{jackson}),
+The programmer might ask herself: ``What if we omit the need for shrinking
+counterexamples altogether?''  SmallCheck is another testing framework for
+Haskell that does just this.  SmallCheck is guaranteed to return the a smallest
+counterexample, if one exists~\cite{sc}.  SmallCheck does this by enumerating
+all possible inputs, ordered from smallest to largest, up to some user-defined
+bound.  While SmallCheck is effective for testing many programs and properties
+(in accordance with the \emph{small scope hypothesis}~\cite{jackson}),
 counterexamples to even relatively simple properties may be infeasible to
@@
 Finally, even if we could overcome the problems described above with QuickCheck
-and SmallCheck, we are left with two others.  First, when a
-counterexample is discovered, one might wish to search for different
-counterexamples.  There currently is no automated way to add a predicate
-characterizing a counterexample as preconditions to a property to strengthen the
-property.  Second, and more generally, a counterexample is usually
-representative of other counterexamples, but that generalization is left to the
-user to discover for herself.  Really, what would be useful would be not just a
+and SmallCheck, we are left with two others.  
+\begin{itemize}
+  \item First, when a counterexample is discovered to some property, say with
+    type
+%
+\begin{code}
+prop :: B -> Bool      
+\end{code}
+%
+\noindent
+    one might wish to search for different counterexamples to \ttp{prop}.
+    Ideally, the testing framework would automatically add a predicate
+    characterizing (and perhaps generalizing) a counterexample 
+%
+    \begin{code}
+cex :: B -> Bool      
+    \end{code}
+%
+\noindent as a precondition to \ttp{prop} to define a new property
+\begin{code}
+prop' b = not (cex b) ==> prop b
+\end{code}
+%
+\noindent
+where \ttp{==>} is implication to generate new counterexamples to the same property.
+
+  \item Second, and more generally, a counterexample is usually representative
+    of other counterexamples, but that generalization is left to the user to
+    discover for herself.  Really, what would be useful would be not just a
     minimal counterexample, but a quantified formula characterizing a set of
@@
     potentially large sub-term of a composite data type value with a quantifier.
+\end{itemize}
 
@@
 \emph{SmartCheck}.  SmartCheck extends QuickCheck to efficiently and generically
-find small counterexamples.  To motivate the benefit of SmartCheck, consider the
-table in Figure~\ref{fig:results}.  The table characterizes testing the program
-shown in Figure~\ref{fig:initial} that we have described so far.  Four runs are
+find small counterexamples as well as generalize them as described in the
+preceding paragraph.  To motivate the benefit of SmartCheck, consider the table
+in Figure~\ref{fig:results}.  The table characterizes testing the program shown
+in Figure~\ref{fig:initial} that we have described so far.  Four trials are
 presented.  In the first three, we use QuickCheck to generate counterexamples.
@@
 with definitions of \ttp{shrink} that limit the lists of fields of \ttp{B} to 10
-(the ``QC (10)'' tests) and 20 (the ``QC (20)'' tests) elements, respectively,
-using a definition like the truncated implementation of \ttp{shrink} for the
-type \ttp{B} discussed earlier.  In each row, we show the maximum, mean, and
-median number of seconds required to by QuickCheck to generate a counterexample
+(Listing~\ref{lst:newshrink}) and 20 elements (Listing~\ref{lst:newshrink}, with
+\ttp{len = 20}), respectively.  (Recall that with the original definition for
+shrink from Figure~\ref{fig:initial}, some counterexamples cause QuickCheck to
+not return after hours of computation.)  In each row, we show the maximum, mean,
+and median number of seconds required by QuickCheck to generate a counterexample
 over 100 runs.  In the last row, we show the mean size of the final value
@@
 
-In the last column are the results under the same condition for SmartCheck.  Not
-only does it execute faster than QuickCheck with shrinking, it has significantly
-smaller outliers.  The most striking difference, however, is the size of the
-counterexample returned, reducing on average the size of a counterexample found
-by QuickCheck without shrinking by a factor of 14.
+In the last column are the results for SmartCheck.  Not only does it execute
+faster than QuickCheck with shrinking, it has significantly smaller outliers.
+The most striking difference, however, is the size of the counterexample
+returned, reducing on average the size of a counterexample found by QuickCheck
+without shrinking by a factor of 14.
 
@@
 generate small counterexamples for algebraic data types and second how to
-generalize the counterexamples..
+generalize the counterexamples.
 
@@
 
-\item Using generic programming techniques, we describe efficient counterexample
-  reduction strategies for algebraic data types.
+\item an efficient counterexample reduction strategy for algebraic
+  data types;
 
-\item We describe an approach to counterexample generalization, to present the
-  user with a formula describing a set of counterexamples to a property.
+\item an approach to counterexample generalization, to present the
+  user with a formula describing a set of counterexamples to a property;
 
-\item We present an approach to generically strengthen a property with a
-  precondition that characterizes an already-discovered counterexample for the
-  purpose of finding new, dissimilar counterexamples.
+\item an approach to generically strengthen a property with a precondition that
+  characterizes an already-discovered counterexample for the purpose of finding
+  new, dissimilar counterexamples;
 
-\item More generally, this paper is the first to explore the idea of efficient,
-  generic counterexample reduction and generalization for functional program testing.
+\item an implementation based on generic programming that is not only more
+  efficient that hand-written \ttp{shrink} instances in many cases but that
+  generates the instances automatically for the user.
+
+%% \item More generally, this paper is the first to explore the idea of
+%%   efficient, generic counterexample reduction and generalization for
+%%   functional program testing.
 \end{itemize}
+
+%% \noindent Moreover, the inability to quickly uncover understandable
+%% counterexamples is a real-world problem with QuickCheck (or other functional
+%% languages).  It works great when your tests pass,
 
@@
 
-\newpage
+\section{Shrinking Data}
+In this section, we describe how to efficiently shrink data values.  Before we
+begin, recall the basic approach implemented by the \ttp{shrink} method:
+generate a list of values, each of which is smaller than the current
+counterexample.  Each of the new values generated has nothing in common with the
+original counterexample, since each one is freshly generated by recursive
+size-constrained calls to \ttp{arbitrary}.
 
-\todo{insight: shrink based on arbitrary}
+Consider the problem of finding a small counterexample as a graph search
+problem.  The set of vertexes are all possible well-typed values (the same as
+the counterexample).  There is an edge from vertex $v$ to $v'$ if $v'$ is
+smaller (by some measure).  The goal is to fine a vertex from the set of
+smallest counterexamples in the graph.
 
-\section{Shrinking Data}
+The approach taken by \ttp{shrink} is the most naive graph search algorithm:
+it simply generates a random set of potential vertexes to check in the graph.
+However, it does not exploit that the original counterexample \ttp{cex} is
+already in a neighborhood of likely counterexamples.  So rather than choose
+random vertexes, it may be more efficient to test other values that are closely
+related to \ttp{cex}.  By restricting the search to \ttp{cex}'s neighborhood, a
+smaller counterexample might be missed, and only a local minimum is returned.
+However, with a large state space, finding a global minimum by random sampling
+vertexes in the graph is possible but too inefficient, as we saw from the
+example in the introduction.
+
+\paragraph{Smart reduction algorithm}
+The algorithm we present for efficiently searching for new counterexamples can
+be thought of as a recursive breadth-first traversal over the structure of the
+current counterexample value, where the value is represented as a tree.  At each
+node, during the traversal, we generate arbitrary \emph{structurally} smaller
+sub-values and build a new value from that, leaving the remainder of the tree
+unchanged.  Intuitively, by a structurally smaller value, we mean one with fewer
+constructors.  We continue until we reach a fixed-point.
+
+To present the algorithm, consider the \ttp{SubTypes} type class below for which
+QuickCheck's \ttp{Arbitrary} class is a super-class.\footnote{This presentation
+  deviates from the implementation; for example, we elide details regarding
+  error handling.}
+%
+\begin{code}
+class Arbitrary a => SubTypes a where
+  size    :: a -> Int
+  index   :: a -> Int -> Maybe SubVal
+  replace :: a -> Int -> SubVal -> a
+\end{code}
+%
+\noindent
+\ttp{SubTypes} has three methods.  The \ttp{size} method returns the size of a
+value, \ttp{index} returns a sub-value at an breadth-first index in a value, and
+\ttp{replace} replaces a sub-value at a particular focus.  If the index is
+out-of-bounds (too large or negative), \ttp{index} return \ttp{Nothing}.  The
+method \ttp{replace} returns the original value for an out-of-bounds index or an
+incorrectly-typed replacement value.  The \ttp{SubVal} type is an existential
+data type, used as a generic container for sub-values from a counterexample.
+%
+\begin{code}
+data SubVal = forall a. (Arbitrary a, SubTypes a) => SubVal a
+\end{code}
+%
+\noindent
+For example, consider a binary tree type
+%
+\begin{code}
+data Tree = L | B Tree Tree
+\end{code}
+% 
+\noindent
+and the value \ttp{tree}, labeled with the indexes of its sub-values
+%
+\begin{code}
+tree = B\sub{0} (B\sub{1} L\sub{3} 
+             (B\sub{4} L\sub{6} L\sub{8}))
+          (B\sub{2} L\sub{5} L\sub{7})
+\end{code}
+%
+\noindent
+Here are example applications of \ttp{SubTypes}; in the following, we show
+the indexes with respect to the \ttp{tree} value:
+%
+\begin{code}
+size tree = 9
+
+index tree 4 = (Just . SubVal) (B\sub{4} L\sub{6} L\sub{8})
+
+index tree 12 = Nothing
+
+replace tree 2 (SubVal L) = 
+  B\sub{0} (B\sub{1} L\sub{3} 
+        (B\sub{4} L\sub{6} L\sub{8}))
+     L\sub{2}
+\end{code}
+%
+\noindent
+We will sometimes refer to unwrapped value returned by \ttp{index a i} as the
+\emph{$i$th sub-value of $a$}, so for example,
+
+\begin{figure}[ht]
+  \lstset{caption={}
+%         ,xleftmargin=5ex,firstnumber=0,numbers=left
+         }
+  \begin{lstlisting}
+reduce :: SubTypes a => a -> Property a -> IO a
+reduce cex prop = reduce' 1
+  where
+  reduce' idx =
+    case index cex idx of 
+      Nothing -> return cex
+      Just v  -> 
+        case test cex idx v prop of
+          Nothing   -> reduce' (idx + 1)
+          Just cex' -> reduce cex' prop
+
+test :: SubTypes a 
+     => a -> Int -> SubVal -> Property a -> Maybe a
+test cex idx v prop = go maxTries
+  where
+  go 0 = Nothing
+  go n = let v'   = newVal v in
+         let cex' = replace cex idx v' in
+         if prop cex' then go (n - 1) else Just cex'
+
+  newVal (SubVal a) = sizedArbitrary a (size a)
+  \end{lstlisting}
+  \caption{Counterexample reduction algorithm\label{fig:reduction}}
+\end{figure}
+
+Figure~\ref{fig:reduction} shows the reduction algorithm.  The function
+\ttp{reduce} takes a counterexample \ttp{cex} and the property \ttp{prop}, which
+for our purposes, we can suppose has the type \ttp{a -> Bool}.  The reduction
+begins at the first sub-value of \ttp{cex}.  Why not begin with \ttp{cex}
+itself?  The idea is that
+
+\todo{optimization: check value directly}
+
+%% For example, consider the following types that give us a binary tree
+%% %
+%% \begin{code}
+%% data T = N Int
+%%        | B T T
+%% \end{code}
+%% %
+%% \noindent
+%% and suppose that the property is that the sum of the leaves is less than 10.
+%% Starting from the following counterexample
+%% %
+%% \lstset{caption={A counterexample value.}
+%%        ,label=lst:val}
+%% \begin{lstlisting}
+%% B (B (N 2) (N 3)) 
+%%   (B (N 4) (B (N 5) (N 6)))
+%% \end{lstlisting}
+%% %
+%% \noindent
+%% Consider the sub-value \ttp{B (N 4) (B (N 5) (N 6))}.  Some example values of
+%% structurally smaller well-typed values include the following:
+%% %
+%% \begin{code}
+%% N 2
+%% B (N 8) (N 7)
+%% B (N 0) (N 1)
+%% \end{code}
+%% %
+%% \noindent
+%% Replacing the sub-value, we get the following new values, each of which is
+%% smaller than the value in Listing~\ref{lst:val}:
+%% %
+%% \begin{code}
+%% B (B (N 2) (N 3)) 
+%%   (N 2)
+
+%% B (B (N 2) (N 3)) 
+%%   (B (N 8) (N 7))
+
+%% B (B (N 2) (N 3)) 
+%%   (B (N 0) (N 1))
+%% \end{code}
+
+In this work, we limit ourselves to algebraic data types.
+
+\todo{GADTs?}
 

### 2012-12-03 paper/Paper.hs

@@ -1,0 +1,47 @@
+{-# LANGUAGE ExistentialQuantification, ScopedTypeVariables #-}
+
+module Paper where
+
+class Arbitrary a where
+  -- arbitrary :: a
+
+type Property a = a -> Bool
+
+-- sized :: [a] -> Int -> [a]
+-- sized = undefined
+
+data SubVal = forall a. (Arbitrary a, SubTypes a) => SubVal a
+
+class Arbitrary a => SubTypes a where
+  size    :: a -> Int
+  index   :: a -> Int -> Maybe SubVal
+  replace :: a -> Int -> SubVal -> a
+
+reduce :: SubTypes a => a -> Property a -> IO a
+reduce cex prop = reduce' 1
+  where
+  reduce' idx = 
+    case index cex idx of 
+      Nothing -> return cex
+      Just v  -> 
+        case test cex idx v prop of
+          Nothing   -> reduce' (idx + 1)
+          Just cex' -> reduce cex' prop
+
+test :: SubTypes a 
+     => a -> Int -> SubVal -> Property a -> Maybe a
+test cex idx v prop = go maxTries
+  where
+  go 0 = Nothing
+  go n = let v'   = newVal v in
+         let cex' = replace cex idx v' in
+         if prop cex' then go (n - 1) else Just cex'
+
+  newVal (SubVal a) = sizedArbitrary a (size a)
+
+sizedArbitrary :: forall a. SubTypes a => a -> Int -> SubVal
+sizedArbitrary _ _ = SubVal (undefined :: a)
+--sizedArbitrary a n = SubVal (sized arbitrary n)
+
+maxTries :: Int
+maxTries = 100    

### 2012-12-04 paper/paper.tex
paper updates.

@@
-\documentclass[9pt]{sigplanconf}
+% can be 9pt
+\documentclass[10pt]{sigplanconf}
 \usepackage{url}
@@
   language=Haskell,                 % the language of the code
-  basicstyle=\scriptsize,           % the size of the fonts that are used
+  %%basicstyle= \scriptsize,           % the size of the fonts that are used
   %% for the code
@@
   %% line-numbers
-  stepnumber=5,                   % the step between two line-numbers. If it's
+  %% stepnumber=5,                   % the step between two line-numbers. If it's
                                   % 1, each line 
   %%                              % will be numbered
-  numbersep=5pt,                  % how far the line-numbers are from the code
-  xleftmargin=0ex,
+  %% numbersep=5pt,                  % how far the line-numbers are from the code
+  %% xleftmargin=0ex,
   %% backgroundcolor=\color{white},      % choose the background color. You must
@@
 
-\newenvironment{code}{\begin{alltt}\scriptsize}{\end{alltt}}
+\newenvironment{code}{\begin{alltt}}{\end{alltt}}
 %% \newenvironment{cols}{\begin{tabular}{m{3.6cm}|m{3.6cm}}{Haskell} &
-%%     {\scriptsize Copilot}\\\hline}{\end{tabular}}
 
@@
 instance Arbitrary B where
-  arbitrary = liftM4 B arbitrary arbitrary arbitrary arbitrary
+  arbitrary = liftM4 B arbitrary arbitrary 
+                       arbitrary arbitrary
   shrink (B a b c d) = 
-    [ B w x y z | w <- shrink a, x <- shrink b
-                , y <- shrink c, z <- shrink d ]
+    [ B w x y z | w <- shrink a
+                , x <- shrink b
+                , y <- shrink c
+                , z <- shrink d ]
 
@@
 instance Serial Int16 where
-  series d = drawnFrom [(-d')..d'] where d' = fromIntegral d
+  series d = drawnFrom [(-d')..d'] 
+    where d' = fromIntegral d
 
@@
 post :: B -> Bool
-post (B a b c d) = add a + add b + add c + add d < 64
+post (B a b c d) = 
+  add a + add b + add c + add d < 64
   where add :: [A] -> Int16
@@
 %
-\lstset{ caption={Truncated shrinking function for data type \ttp{B}.}
-       , label=lst:newshrink
-       ,captionpos=b }
-\begin{lstlisting}
+
+%% \begin{minipage}{\columnwidth}
+%% \lstset{ caption={Truncated shrinking function for data type \ttp{B}.}
+%%        , label=lst:newshrink
+%%        , captionpos=b }
+%% \begin{lstlisting}
+
+\begin{figure}[ht]
+\begin{code}
 len = 10
 
-shrink (B a b c d) = [ B w x y z | w <- tk a, x <- tk b
+shrink (B a b c d) = 
+  [ B w x y z | w <- tk a, x <- tk b
               , y <- tk c, z <- tk d ]
   where tk x = take len (shrink x)
-\end{lstlisting}
-%% \begin{code}
-%% \end{code}
+\end{code}
+  \caption{Truncated shrinking function for data type \ttp{B}.}
+  \label{lst:newshrink}
+\end{figure}
+
+%% \end{lstlisting}
+%% \end{minipage}
 %
@@
 \noindent
-    one might wish to search for different counterexamples to \ttp{prop}.
-    Ideally, the testing framework would automatically add a predicate
-    characterizing (and perhaps generalizing) a counterexample 
+one might wish to search for different counterexamples to \ttp{prop}.  Ideally,
+the testing framework would automatically add a predicate characterizing (and
+perhaps generalizing) a counterexample
 %
@@
 %
-\noindent as a precondition to \ttp{prop} to define a new property
+\noindent
+as a precondition to \ttp{prop} to define a new property
 \begin{code}
@@
 \begin{figure}[ht]
-\scriptsize
+\footnotesize
   \begin{center}
@@
 with definitions of \ttp{shrink} that limit the lists of fields of \ttp{B} to 10
-(Listing~\ref{lst:newshrink}) and 20 elements (Listing~\ref{lst:newshrink}, with
+(Figure~\ref{lst:newshrink}) and 20 elements (Figure~\ref{lst:newshrink}, with
 \ttp{len = 20}), respectively.  (Recall that with the original definition for
@@
 
-\paragraph{Smart reduction algorithm}
+\subsection{Reduction Algorithm Overview}
 The algorithm we present for efficiently searching for new counterexamples can
@@
 \begin{code}
-data SubVal = forall a. (Arbitrary a, SubTypes a) => SubVal a
+data SubVal = 
+  forall a. (Arbitrary a, SubTypes a) => SubVal a
 \end{code}
@@
         (B\sub{4} L\sub{6} L\sub{8}))
-     L\sub{2}
+     L
 \end{code}
@@
 We will sometimes refer to unwrapped value returned by \ttp{index a i} as the
-\emph{$i$th sub-value of $a$}, so for example,
+\emph{$i$th sub-value of $a$}, so for example, \ttp{B\sub{4} L\sub{6} L\sub{8}}
+is the 4th sub-value of \ttp{tree}.
 
@@
   \begin{lstlisting}
-reduce :: SubTypes a => a -> Property a -> IO a
+reduce :: SubTypes a 
+       => a -> (a -> Property) -> IO a
 reduce cex prop = reduce' 1
@@
       Nothing -> return cex
-      Just v  -> 
-        case test cex idx v prop of
-          Nothing   -> reduce' (idx + 1)
+      Just v  -> do
+        vs <- newVals v
+        case test cex idx vs prop of
+          Nothing   -> reduce' (idx+1)
           Just cex' -> reduce cex' prop
 
+  newVals (SubVal a) = 
+    sequence (replicate maxTries s)
+    where s =     sizedArbitrary a 
+              >>= return . SubVal
+
 test :: SubTypes a 
-     => a -> Int -> SubVal -> Property a -> Maybe a
-test cex idx v prop = go maxTries
+     => a -> Int -> [SubVal] 
+     -> (a -> Property) -> Maybe a
+test cex idx vs prop = go vs
   where
-  go 0 = Nothing
-  go n = let v'   = newVal v in
-         let cex' = replace cex idx v' in
-         if prop cex' then go (n - 1) else Just cex'
-
-  newVal (SubVal a) = sizedArbitrary a (size a)
+  go []      = Nothing
+  go (v:vs') = 
+    let cex' = replace cex idx v in
+    if pass prop cex' then go vs' 
+      else Just cex'
   \end{lstlisting}
@@
 Figure~\ref{fig:reduction} shows the reduction algorithm.  The function
-\ttp{reduce} takes a counterexample \ttp{cex} and the property \ttp{prop}, which
-for our purposes, we can suppose has the type \ttp{a -> Bool}.  The reduction
-begins at the first sub-value of \ttp{cex}.  Why not begin with \ttp{cex}
-itself?  The idea is that
+\ttp{reduce} takes a counterexample \ttp{cex} and the property \ttp{prop}.  At
+the beginning of execution, we assume that a counterexample has been obtained
+somehow; our implementation relies on QuickCheck as a ``front-end'' to generate
+a counterexample that the algorithm automatically shrinks.  The reduction begins
+at the first sub-value of \ttp{cex}, \ttp{v}.  When the index \ttp{idx} becomes
+out-of-bounds and returns \ttp{Nothing}, the algorithm terminates.  Otherwise, a
+list of new random values are generated.
+%
+\begin{code}
+sizedArbitrary :: SubTypes a => a -> IO a  
+\end{code}
+%
+generates a new value \ttp{v'} having the same type as \ttp{v} and strictly
+smaller (with respect to the \ttp{size} method) than \ttp{v}.
+\ttp{sizedArbitrary} returns a value inside the \ttp{IO} monad since it depends
+on a random seed for generation (we could pass in the generator seed, but for
+simplicity, we assume one is generated locally).  The constant \ttp{maxTries} is
+the maximum number of tries to discover a new counterexample by replacing
+\ttp{v} in \ttp{cex} and testing it.  The result of \ttp{pass prop cex'} for
+%
+\begin{code}
+pass :: (a -> Property) -> a -> Bool
+\end{code}
+%
+holds if \ttp{cex'} satisfies the property \ttp{prop}.  If no failure is found,
+we move to the next sub-value of \ttp{cex} and continue.  However, if a new
+smaller counterexample \ttp{cex'} is found, we start a new breadth-first
+traversal of \ttp{cex'}, attempting to shrink it further.
 
-\todo{optimization: check value directly}
+\subsection{Reduction Algorithm Details and Optimizations}
+The reduction algorithm description above omits some minor details and
+optimizations we describe here.
+
+\paragraph{Sum-type hypothesis}
+First, recall that the algorithm begins at the \emph{first} sub-value of the
+counterexample rather than the 0th sub-value.  The 0th sub-value of \ttp{cex} is
+\ttp{cex} itself.  No invariant of the algorithm would be violated by beginning
+with the 0th sub-value; recall that it would generate strictly smaller values
+than \ttp{cex}.  However, we begin with a strict sub-value because of a
+hypothesis we refer to as the \emph{sum-type hypothesis}.  The hypothesis is
+that in the search space of possible values, if a counterexample of the form
+\ttp{C}~$\vec{x}$, where \ttp{C} is some constructor and $\vec{x}$ are its
+arguments, is discovered, then it is more likely to find another ``nearby''
+counterexample of the form \ttp{C}~$\vec{y}$, where $\vec{y}$ are different
+arguments to \ttp{C}.  For example, consider a property about unintended
+variable capture over a language's parse tree represented by a sum type with
+constructors for module imports, function definitions, and global-variable
+assignments, respectively.  A function definition counterexample can only be
+reduced to smaller function definition counterexamples, the only construct in
+which variable capture is possible.
+
+Another way to motivate the hypothesis is that searching the space of possible
+counterexamples by generating \emph{any} well-typed value that is strictly
+smaller than the original counterexample is something we have seen before: it is
+precisely the naive approach to implementing \ttp{shrink}!
+
+\paragraph{Conditional properties}
+A QuickCheck property may be a conditional of the form
+%
+\begin{code}
+prop a = pre a ==> prop' a
+\end{code}
+%
+\noindent
+such that \ttp{prop} is true if the precondition \ttp{pre a} is false or
+\ttp{prop' a} holds.  In practice, conditionals are used to restrict the set of
+values tested; values that fail the precondition are discarded.  Thus, the call
+to \ttp{pass} in the reduction algorithm only considers a potential
+counterexample \ttp{cex} to have failed if it passes the property's
+precondition and fails the consequent.
+
+\paragraph{Counterexample search}
+We make two optimizations in searching for counterexamples.  First, just like
+QuickCheck's \ttp{arbitrary} method does, our \ttp{sizedArbitrary} function
+successfully larger counterexamples when generating new values with which to
+replace a sub-value.
+
+Furthermore, we can instrument a low-cost shortcut check in the algorithm.
+Sometimes, a counterexample fails a property due to a sub-value nested deep
+inside the counterexample.  For example, consider a simple calculator language
+containing constant \ttp{Int}s, addition, and division:
+%
+\begin{code}
+data Exp = C Int
+         | Add Exp Exp
+         | Div Exp Exp  
+\end{code}
+%
+\noindent
+and consider an evaluator for the language that returns \ttp{Nothing} if
+division-by-zero is encountered and \ttp{Just} the value otherwise:
+%
+\begin{code}
+eval :: Exp -> Maybe Int
+eval (C i) = Just i
+eval (Add e0 e1) =
+  liftM2 (+) (eval e0) (eval e1)
+eval (Div e0 e1) = 
+  let e = eval e1 in 
+  if e == Just 0 then Nothing 
+    else liftM2 div (eval e0) e
+\end{code}
+%
+\noindent
+Suppose we restrict the set of programs with the following predicate:
+%
+\begin{code}
+divSubTerms :: Exp -> Bool
+divSubTerms (C _)         = True
+divSubTerms (Div _ (C 0)) = False
+divSubTerms (Add e0 e1)   =    divSubTerms e0 
+                            && divSubTerms e1
+divSubTerms (Div e0 e1)   =    divSubTerms e0 
+                            && divSubTerms e1  
+\end{code}
+%
+and then test the property
+%
+\begin{code}
+div_prop e = divSubTerms e ==> eval e /= Nothing  
+\end{code}
+%
+\noindent
+The property \ttp{div\_prop} turns out to be false; for example, 
+%
+\begin{code}
+Div (C 1) (Div (C 0) (C 1))
+\end{code}
+%
+\noindent
+is a counterexample.  In the predicate case
+%
+\begin{code}
+divSubTerms (Div _ (C 0)) = False  
+\end{code}
+%
+\noindent
+we forgot to recur on the divisor.  However, a typical counterexample returned
+by QuickCheck, without \ttp{shrink} implemented, is as follows:
+%
+\begin{quote}
+\ttp{Div (Div (C (-22)) (Div (C 2) (C 13))) (Div (Div (Div (C (-11)) (C 14))
+  (Add (Div (C (-4)) (C (-3))) (Div (C (-25)) (C (-25))))) (Add (Add (C 22) (Div
+  (C (-30)) (C (-14)))) (C 13)))}
+\end{quote}
+%
+\noindent
+To butcher an expression, finding the cause of failure is like searching for a
+divide-by-zero in the haystack!  But it turns out that the culprit is a small
+sub-value:
+%
+\begin{code}
+Div (C (-22)) (Div (C 2) (C 13))  
+\end{code}
+%
+
+Generalizing this scenario, during the reduction algorithm's breadth-first
+search through a counterexample \ttp{cex}'s sub-values, we may happen upon a
+sub-value \ttp{cex'} that has the same type as \ttp{cex} and fails the property
+(while passing any preconditions).  In this case, we can return \ttp{cex'}
+directly, and rerun the reduction algorithm on \ttp{cex'}.
+
+\todo{violates sum-type hypothesis but cheap---not testing generated values,
+  just direct subvalues.  Call this the sub-value hypothesis: direct sub-values
+  of a counterexample may also be a counterexample.}
+
+\paragraph{Bounding counterexample exploration}
+\todo{limit depth}
+\todo{limit size of generated structures}
+
+\paragraph{Base types}
 
@@
 
-\section{Generalizing Data}
+\section{Counterexample Generalization}
 
 \section{Implementation}
+
+\todo{generics}
 
@@
 
-\todo{\cite{syb} paper is motivated by generic shrink}
 \todo{shrink paper rehger pointed to ``Automatic isolation of compiler errors'' from 1994}

### 2012-12-04 paper/Paper.hs

@@ -5,7 +5,9 @@ module Paper where
 class Arbitrary a where
   -- arbitrary :: a
 
-type Property a = a -> Bool
+--type Property a = a -> Bool
+
+data Property
 
 -- sized :: [a] -> Int -> [a]
 -- sized = undefined
@@ -17,31 +19,62 @@ class Arbitrary a => SubTypes a where
   index   :: a -> Int -> Maybe SubVal
   replace :: a -> Int -> SubVal -> a
 
-reduce :: SubTypes a => a -> Property a -> IO a
+reduce :: SubTypes a => a -> (a -> Property) -> IO a
 reduce cex prop = reduce' 1
   where
   reduce' idx = 
     case index cex idx of 
       Nothing -> return cex
-      Just v  -> 
-        case test cex idx v prop of
+      Just v  -> do
+        vs <- newVals v
+        case test cex idx vs prop of
           Nothing   -> reduce' (idx + 1)
           Just cex' -> reduce cex' prop
 
+  newVals (SubVal a) = sequence (replicate maxTries s)
+    where s = sizedArbitrary a >>= return . SubVal
+
 test :: SubTypes a 
-     => a -> Int -> SubVal -> Property a -> Maybe a
-test cex idx v prop = go maxTries
+     => a -> Int -> [SubVal] -> (a -> Property) -> Maybe a
+test cex idx vs prop = go vs
   where
-  go 0 = Nothing
-  go n = let v'   = newVal v in
-         let cex' = replace cex idx v' in
-         if prop cex' then go (n - 1) else Just cex'
+  go []     = Nothing
+  go (v:vs') = let cex' = replace cex idx v in
+               if pass prop cex' then go vs' else Just cex'
 
-  newVal (SubVal a) = sizedArbitrary a (size a)
-
-sizedArbitrary :: forall a. SubTypes a => a -> Int -> SubVal
-sizedArbitrary _ _ = SubVal (undefined :: a)
+sizedArbitrary :: forall a. SubTypes a => a -> IO a
+sizedArbitrary _ = return (undefined :: a)
 --sizedArbitrary a n = SubVal (sized arbitrary n)
 
 maxTries :: Int
-maxTries = 100    
+maxTries = 100
+
+pass :: (a -> Property) -> a -> Bool
+pass _ _ = True
+
+
+--------------------------------------------------------------------------------
+
+data Tree = L | B Tree Tree
+
[…]

### 2012-12-04 examples/Div0.hs
source wibbles.

@@ -27,7 +27,7 @@ eval (Add e0 e1) =
   liftM2 (+) (eval e0) (eval e1)
 eval (Div e0 e1) = 
   let e = eval e1 in 
-  if e == Just 0 then Nothing 
+  if e == Just 0 then trace ("dom " ++ show e1) Nothing 
     else liftM2 div (eval e0) e
 
 instance Arbitrary Exp where
@@ -38,19 +38,19 @@ instance Arbitrary Exp where
                   , liftM2 Div mkM' mkM' ]
       where mkM' = mkM =<< choose (0,n-1)
 
-  shrink (C i)       = map C (shrink i)
-  shrink (Add e0 e1) = [e0, e1]
-  shrink (Div e0 e1) = [e0, e1]
+  -- shrink (C i)       = map C (shrink i)
+  -- shrink (Add e0 e1) = [e0, e1]
+  -- shrink (Div e0 e1) = [e0, e1]
 
 -- property: so long as 0 isn't in the divisor, we won't try to divide by 0.
 -- It's false: something might evaluate to 0 still.
 div_prop :: Exp -> Property
---div_prop e = divSubTerms e ==> eval e /= Nothing
-div_prop e = property $ case x of
-                          Nothing -> True
-                          Just True -> True
-                          _       -> False
-  where x = fmap (< 1) (eval e)
+div_prop e = divSubTerms e ==> eval e /= Nothing
+-- div_prop e = property $ case x of
+--                           Nothing -> True
+--                           Just True -> True
+--                           _       -> False
+--   where x = fmap (< 1) (eval e)
 
   -- precondition: no dividand in a subterm can be 0.
 divSubTerms :: Exp -> Bool

### 2012-12-04 src/Test/SmartCheck/Reduce.hs

@@ -15,7 +15,7 @@ import Data.Typeable
 import Data.Tree
 import Data.Maybe
 
----------------------------------------------------------------------------------
+--------------------------------------------------------------------------------
 
 -- Smarter than shrinks.  Does substitution.  m is a value that failed QC that's
 -- been shrunk.  We substitute successive children with strictly smaller (and
@@ -30,7 +30,7 @@ smartRun args res prop = do
   print new
   return new
 
----------------------------------------------------------------------------------
+--------------------------------------------------------------------------------
 
 -- | Breadth-first traversal of d, trying to shrink it with *strictly* smaller
 -- children.  We replace d whenever a successful shrink is found and try again.
@@ -92,7 +92,7 @@ smartShrink args d prop =
       extractAndTest :: a -> IO (Maybe a)
       extractAndTest y = do 
         res <- resultify notProp y
-        return $ resultToMaybe res
+        return $ traceShow resultToMaybe res
 
 resultToMaybe :: Result a -> Maybe a
 resultToMaybe res =
@@ -101,7 +101,7 @@ resultToMaybe res =
     FailedProp    -> Nothing
     Result n      -> Just n
 
----------------------------------------------------------------------------------
+--------------------------------------------------------------------------------
 
 -- | Get the maximum depth of d's subforest at idx.  Intuitively, it's the
 -- maximum number of constructors you have *below* the constructor at idx.  So
@@ -123,4 +123,4 @@ subValSize d idx = maybe 0 id (fmap depth forestIdx)
   forestIdx :: Maybe [Tree Bool]
   forestIdx = fmap subForest $ getIdxForest (mkSubstForest d True) idx
 
----------------------------------------------------------------------------------
+--------------------------------------------------------------------------------

### 2012-12-05 paper/paper.tex
paper mods.

@@
 
+newVals :: SubVal -> IO [SubVal]
 newVals (SubVal a) = 
@@
   \end{lstlisting}
-  \caption{Counterexample reduction algorithm\label{fig:reduction}}
+  \caption{Counterexample reduction algorithm.\label{fig:reduction}}
 \end{figure}
@@
 %
+\noindent
 generates a new value \ttp{v'} having the same type as \ttp{v} and strictly
-smaller (with respect to the \ttp{size} method) than \ttp{v}.
+smaller (with respect to the \ttp{size} method) than \ttp{v}.  Just like
+QuickCheck's \ttp{arbitrary} method, \ttp{sizedArbitrary} generates successfully
+larger counterexamples when generating new values with which to replace a
+sub-value.
+
 \ttp{sizedArbitrary} returns a value inside the \ttp{IO} monad since it depends
@@
 %
+\noindent
 holds if \ttp{cex'} satisfies the property \ttp{prop}.  If no failure is found,
@@
 
-\paragraph{Sum-type hypothesis}
+\paragraph{Sum-type counterexample hypothesis}
 First, recall that the algorithm begins at the \emph{first} sub-value of the
@@
 than \ttp{cex}.  However, we begin with a strict sub-value because of a
-hypothesis we refer to as the \emph{sum-type hypothesis}.  The hypothesis is
-that in the search space of possible values, if a counterexample of the form
-\ttp{C}~$\vec{x}$, where \ttp{C} is some constructor and $\vec{x}$ are its
-arguments, is discovered, then it is more likely to find another ``nearby''
-counterexample of the form \ttp{C}~$\vec{y}$, where $\vec{y}$ are different
-arguments to \ttp{C}.  For example, consider a property about unintended
-variable capture over a language's parse tree represented by a sum type with
-constructors for module imports, function definitions, and global-variable
-assignments, respectively.  A function definition counterexample can only be
-reduced to smaller function definition counterexamples, the only construct in
-which variable capture is possible.
+hypothesis we refer to as the \emph{sum-type counterexample hypothesis}.  The
+hypothesis is that in the search space of possible values, if a counterexample
+of the form \ttp{C}~$\vec{x}$, where \ttp{C} is some constructor and $\vec{x}$
+are its arguments, is discovered, then it is more likely to find another
+``nearby'' counterexample of the form \ttp{C}~$\vec{y}$, where $\vec{y}$ are
+different arguments to \ttp{C}.  For example, consider a property about
+unintended variable capture over a language's parse tree represented by a sum
+type with constructors for module imports, function definitions, and
+global-variable assignments, respectively.  A function definition counterexample
+can only be reduced to smaller function definition counterexamples, the only
+construct in which variable capture is possible.
 
@@
 
-\paragraph{Counterexample search}
-We make two optimizations in searching for counterexamples.  First, just like
-QuickCheck's \ttp{arbitrary} method does, our \ttp{sizedArbitrary} function
-successfully larger counterexamples when generating new values with which to
-replace a sub-value.
+\paragraph{Sub-value counterexample hypothesis}
+Sometimes, a counterexample fails a property due to a sub-value nested deep
+inside the counterexample.  The rest of the value is irrelevant.  We call this
+the \emph{sub-value counterexample hypothesis.}  Thus, one way to efficiently
+search the space of potential counterexamples is to test a counterexample's
+(well-typed) sub-values.
 
-Furthermore, we can instrument a low-cost shortcut check in the algorithm.
-Sometimes, a counterexample fails a property due to a sub-value nested deep
-inside the counterexample.  For example, consider a simple calculator language
-containing constant \ttp{Int}s, addition, and division:
+For example, consider a simple calculator language containing constant
+\ttp{Int}s, addition, and division:
 %
@@
 \noindent
-and consider an evaluator for the language that returns \ttp{Nothing} if
+and consider an interpreter for the language that returns \ttp{Nothing} if
 division-by-zero is encountered and \ttp{Just} the value otherwise:
@@
 
+\begin{figure}
+\begin{lstlisting}
+reduce0 :: forall a . SubTypes a 
+        => a -> (a -> Property) -> IO a
+reduce0 cex prop = reduce' 1
+  where
+  reduce' idx = 
+    case index cex idx of 
+      Nothing -> return cex
+      Just v  -> 
+        case testHole v of
+          Just v' -> reduce0 v' prop
+          Nothing -> do
+            vs <- newVals v
+            case test cex idx vs prop of
+              Nothing   -> 
+                reduce' (idx+1)
+              Just cex' -> 
+                reduce0 cex' prop
+
+  testHole (SubVal a) = 
+    case cast a :: Maybe a of
+      Nothing -> Nothing
+      Just a' -> 
+        if pass prop a' then Nothing
+          else Just a'
+\end{lstlisting}
+  \caption{Reduction algorithm with the sub-value counterexample optimization.}
+  \label{fig:reduce0}
+\end{figure}
+
 Generalizing this scenario, during the reduction algorithm's breadth-first
@@
 (while passing any preconditions).  In this case, we can return \ttp{cex'}
-directly, and rerun the reduction algorithm on \ttp{cex'}.
-
-\todo{violates sum-type hypothesis but cheap---not testing generated values,
-  just direct subvalues.  Call this the sub-value hypothesis: direct sub-values
-  of a counterexample may also be a counterexample.}
+directly, and rerun the reduction algorithm on \ttp{cex'}.  In
+Figure~\ref{fig:reduce0}, we show an updated reduction algorithm, \ttp{reduce0}
+that implements this optimization.  The function \ttp{testHole} tests the
+current sub-value and if it fails the property, the function tests the sub-value
+directly.
 

### 2012-12-05 paper/Paper.hs

@@ -2,6 +2,9 @@
 
 module Paper where
 
+cast :: SubTypes a => a -> Maybe b
+cast _ = undefined
+
 class Arbitrary a where
   -- arbitrary :: a
 
@@ -19,28 +22,35 @@ class Arbitrary a => SubTypes a where
   index   :: a -> Int -> Maybe SubVal
   replace :: a -> Int -> SubVal -> a
 
-reduce :: SubTypes a => a -> (a -> Property) -> IO a
+reduce :: SubTypes a 
+       => a -> (a -> Property) -> IO a
 reduce cex prop = reduce' 1
   where
-  reduce' idx = 
+  reduce' idx =
     case index cex idx of 
       Nothing -> return cex
       Just v  -> do
         vs <- newVals v
         case test cex idx vs prop of
-          Nothing   -> reduce' (idx + 1)
+          Nothing   -> reduce' (idx+1)
           Just cex' -> reduce cex' prop
 
-  newVals (SubVal a) = sequence (replicate maxTries s)
-    where s = sizedArbitrary a >>= return . SubVal
+newVals :: SubVal -> IO [SubVal]
+newVals (SubVal a) = 
+  sequence (replicate maxTries s)
+  where s =     sizedArbitrary a 
+            >>= return . SubVal
 
 test :: SubTypes a 
-     => a -> Int -> [SubVal] -> (a -> Property) -> Maybe a
+     => a -> Int -> [SubVal] 
+     -> (a -> Property) -> Maybe a
 test cex idx vs prop = go vs
   where
-  go []     = Nothing
-  go (v:vs') = let cex' = replace cex idx v in
-               if pass prop cex' then go vs' else Just cex'
+  go []      = Nothing
+  go (v:vs') = 
+    let cex' = replace cex idx v in
+    if pass prop cex' then go vs' 
+      else Just cex'
 
 sizedArbitrary :: forall a. SubTypes a => a -> IO a
 sizedArbitrary _ = return (undefined :: a)
@@ -52,7 +62,31 @@ maxTries = 100
 pass :: (a -> Property) -> a -> Bool
 pass _ _ = True
 
+reduce0 :: forall a . SubTypes a 
+        => a -> (a -> Property) -> IO a
+reduce0 cex prop = reduce' 1
+  where
+  reduce' idx = 
+    case index cex idx of 
+      Nothing -> return cex
+      Just v  -> 
+        case testHole v of
+          Just v' -> reduce0 v' prop
+          Nothing -> do
+            vs <- newVals v
[…]

### 2012-12-05 examples/Div0.hs
code wibbles.

@@ -27,7 +27,7 @@ eval (Add e0 e1) =
   liftM2 (+) (eval e0) (eval e1)
 eval (Div e0 e1) = 
   let e = eval e1 in 
-  if e == Just 0 then trace ("dom " ++ show e1) Nothing 
+  if e == Just 0 then Nothing 
     else liftM2 div (eval e0) e
 
 instance Arbitrary Exp where

### 2012-12-05 src/Test/SmartCheck/Reduce.hs

@@ -92,7 +92,7 @@ smartShrink args d prop =
       extractAndTest :: a -> IO (Maybe a)
       extractAndTest y = do 
         res <- resultify notProp y
-        return $ traceShow resultToMaybe res
+        return $ resultToMaybe res
 
 resultToMaybe :: Result a -> Maybe a
 resultToMaybe res =

### 2012-12-10 paper/paper.tex
done with reduction algorithm description.

@@
 
-\title{SmartCheck}
+\title{SmartCheck: Automatic and Efficient Counterexample Reduction and Generalization}
 %\subtitle{Subtitle Text, if any}
@@
 \begin{abstract}
-foobar
+QuickCheck is a powerful library for automatic test-case generation.  Because
+QuickCheck performs random testing, it can discover very large counterexamples,
+even when enumerating test-cases is infeasible, as is done with SmallCheck.
+QuickCheck provides an interface for the user to write a \emph{shrink} function
+to attempt to reduce the size of a counterexample discovered by QuickCheck.
+Implementations of \emph{shrink} often require significant boilerplate and are
+not ``smart'', just generating structurally smaller values.  We introduce
+\emph{SmartCheck}, an automatic efficient counterexample reduction algorithm.
+SmartCheck reduces algebraic data using generic search heuristics to efficiently
+find smaller counterexamples.  In addition to shrinking, SmartCheck generalizes
+counterexamples by returning a formula that abstracts over individual
+counterexamples.  SmartCheck has been implemented for Haskell and tested on a
+variety of Haskell programs.
 \end{abstract}
@@
 
-\begin{figure}[ht]
+\begin{table}[ht]
 \footnotesize
@@
     using interpreted Haskell.}
-  \label{fig:results}
-\end{figure}
+  \label{table:results}
+\end{table}
 
@@
 preceding paragraph.  To motivate the benefit of SmartCheck, consider the table
-in Figure~\ref{fig:results}.  The table characterizes testing the program shown
+in Table~\ref{fig:results}.  The table characterizes testing the program shown
 in Figure~\ref{fig:initial} that we have described so far.  Four trials are
@@
 
-\subsection{Reduction Algorithm Overview}
+\subsection{Reduction Algorithm Overview}\label{sec:reduct}
 The algorithm we present for efficiently searching for new counterexamples can
@@
 
-\paragraph{Bounding counterexample exploration}
-\todo{limit depth}
-\todo{limit size of generated structures}
-
 \paragraph{Base types}
 
-%% For example, consider the following types that give us a binary tree
-%% %
-%% \begin{code}
-%% data T = N Int
-%%        | B T T
-%% \end{code}
-%% %
-%% \noindent
-%% and suppose that the property is that the sum of the leaves is less than 10.
-%% Starting from the following counterexample
-%% %
-%% \lstset{caption={A counterexample value.}
-%%        ,label=lst:val}
-%% \begin{lstlisting}
-%% B (B (N 2) (N 3)) 
-%%   (B (N 4) (B (N 5) (N 6)))
-%% \end{lstlisting}
-%% %
-%% \noindent
-%% Consider the sub-value \ttp{B (N 4) (B (N 5) (N 6))}.  Some example values of
-%% structurally smaller well-typed values include the following:
-%% %
-%% \begin{code}
-%% N 2
-%% B (N 8) (N 7)
-%% B (N 0) (N 1)
-%% \end{code}
-%% %
-%% \noindent
-%% Replacing the sub-value, we get the following new values, each of which is
-%% smaller than the value in Listing~\ref{lst:val}:
-%% %
-%% \begin{code}
-%% B (B (N 2) (N 3)) 
-%%   (N 2)
+The number of constructors in an algebraic data value primarily determines how
+large and understandable it is.  In our experience, the size of fields that come
+from primitive types, like \ttp{Int} or \ttp{Char}, have less of a role in
+understanding a counterexample.
 
-%% B (B (N 2) (N 3)) 
-%%   (B (N 8) (N 7))
+Consequently, we distinguish a class of types referred to as \emph{base types}.
+If the reduction algorithm encounters a base type, it is ignored.  By default,
+this includes numeric types (e.g., \ttp{Int}, \ttp{Word}, \ttp{Float}, etc.,
+\ttp{Bool}, and \ttp{Char}.  We also treat lists specially: lists of base types
+are considered to be base types themselves.  This prevents the algorithm from
+analyzing, for example, strings.  A list of a non-base-types is shrunk by the
+algorithm.
 
-%% B (B (N 2) (N 3)) 
-%%   (B (N 0) (N 1))
-%% \end{code}
+By providing custom instances, the user can declare any substructure in a data
+type to be a base type.  Doing so can help make SmartCheck more efficient if the
+user knows that some portion of the structure cannot be shrunk or is irrelevant
+to the property.
 
-In this work, we limit ourselves to algebraic data types.
+\paragraph{Bounding counterexample exploration}
 
-\todo{GADTs?}
+SmartCheck's implementation contains flags to allow the user to customize its
+behavior.  Three flags that are relevant to the reduction algorithm are the
+following:
+%
+\begin{code}
+scMaxFailure :: Int
+scMaxSize    :: Int
+scMaxDepth   :: Maybe Int  
+\end{code}
+%
+\noindent
+The ``\ttp{sc}'' prefix stands for \emph{S}mart\emph{C}heck and is used to
+prevent name clashes with QuickCheck's flags.  \ttp{scMaxFailure} controls the
+number of values generated by the reduction algorithm for each sub-value
+analyzed.  \ttp{scMaxSize} controls the maximum size of values generated to
+replace sub-values by the reduction algorithm.  Thus, new sub-values must be
+strictly smaller than the minimum of \ttp{scMaxSize} and the size of the
+sub-value being replaced.  (The size of the value is with respect to the
+\ttp{size} method in the \ttp{SubTypes} class.)  Finally, \ttp{scMaxDepth}
+determines the maximum depth in the counterexample the reduction algorithm
+should analyze.  A value of \ttp{Nothing} means that the counterexample should
+be exhaustively reduced, as the algorithm is presented in
+Figure~\ref{fig:reduction}.  For example, the depth of a list is determined by
+its length and the depth of the binary tree defined in Section~\ref{sec:reduct}
+is determined by the function \ttp{depth}:
+%
+\begin{code}
+depth L         = 0
+depth (B t0 t1) = 1 + max (depth t0) (depth t1)
+\end{code}
+%
+\noindent
+Of the flags, \ttp{scMaxDepth} is the most important for controlling
+efficiency, particularly for large product types with significant ``fan out''.
+The number of sub-values of a value product type can grow exponentially with
+respect to the depth.  Furthermore, note that as the reduction algorithm
+descends further, there is less chance to reduce the size of the value overall,
+since smaller and smaller sub-values are replaced.
 
-\section{Counterexample Generalization}
+\section{Counterexample Generalization} \label{sec:generalization}
+Suppose we are testing a value.
 
@@
 
+\todo{GADTs? don't work}
+
 \section{Experiments}
 
-%--------------------------------------------------------------------------------
+\todo{xmonad}
 
-%--------------------------------------------------------------------------------
+%-------------------------------------------------------------------------------
+
+%-------------------------------------------------------------------------------
 \section{Related Work}
@@
 
-%--------------------------------------------------------------------------------
+%-------------------------------------------------------------------------------
 \section{Conclusions}

### 2012-12-24 paper/paper.tex
paper: updates.

@@
 \section{Introduction}\label{sec:intro}
-
 The QuickCheck testing framework was a revolutionary step-forward in
@@
 either miss minimal counterexamples, be too inefficient to be practical, or
-both.\footnote{E.g., for a recent real-world case, see the user question posed on StackOverflow:
+both.\footnote{E.g., for a recent real-world case, see the user question posed
+  on StackOverflow:
   \url{http://stackoverflow.com/questions/8788542/how-do-i-get-good-small-shrinks-out-of-quickcheck}.}
@@
 list in a field of \ttp{B} is less than 16, then the sum of the four fields of
-\ttp{B} is less than 64.  \ttp{prop} seems reasonable at first glance, until
-one realizes that due to underflow, the property can be violated.  For example,
-consider the value
+\ttp{B} is less than 64.  The property \ttp{prop} seems reasonable at first
+glance, until one realizes that due to underflow, the property can be violated.
+For example, consider the value
 %
@@
 
+\todo{check paper against this section again, especially generalization sec.}
 \paragraph{Contributions}
@@
 \begin{code}
+type Idx = Int
+
 class Arbitrary a => SubTypes a where
-  size    :: a -> Int
-  index   :: a -> Int -> Maybe SubVal
-  replace :: a -> Int -> SubVal -> a
+  size    :: a -> Idx
+  index   :: a -> Idx -> Maybe SubVal
+  replace :: a -> Idx -> SubVal -> a
 \end{code}
@@
 data SubVal =
-  forall a. (Arbitrary a, SubTypes a) => SubVal a
+  forall a. SubTypes a => SubVal a
 \end{code}
@@
          }
-  \begin{lstlisting}
+  \begin{code}
 reduce :: SubTypes a
@@
 newVals (SubVal a) =
-  sequence (replicate maxTries s)
-  where s =     sizedArbitrary a 
-            >>= return . SubVal
+  sequence (replicate max s)
+  where
+  s = sizedArbitrary a >>= return . SubVal
 
 test :: SubTypes a
-     => a -> Int -> [SubVal] 
+     => a -> Idx -> [SubVal]
      -> (a -> Property) -> Maybe a
@@
       else Just cex'
-  \end{lstlisting}
+  \end{code}
   \caption{Counterexample reduction algorithm.\label{fig:reduction}}
@@
 For example, consider a simple calculator language containing constant
-\ttp{Int}s, addition, and division:
+\ttp{Ints}, addition, and division:
 %
@@
 \noindent
-we forgot to recur on the divisor.  However, a typical counterexample returned
-by QuickCheck, without \ttp{shrink} implemented, is as follows:
-%
-\begin{quote}
+we forgot to recur on the divisor.
+
+\begin{figure}[ht!]
 \ttp{Div (Div (C (-22)) (Div (C 2) (C 13))) (Div (Div (Div (C (-11)) (C 14))
@@
   (C (-30)) (C (-14)))) (C 13)))}
-\end{quote}
-%
-\noindent
-To butcher an expression, finding the cause of failure is like searching for a
-divide-by-zero in the haystack!  But it turns out that the culprit is a small
-sub-value:
+  \caption{Typical counterexample from QuickCheck with no user-defined
+    \ttp{shrink} function.}
+  \label{fig:div-cex}
+\end{figure}
+
+However, a typical counterexample returned by QuickCheck, without \ttp{shrink}
+implemented, is shown in Figure~\ref{fig:div-cex}.  To butcher an expression,
+finding the cause of failure is like searching for a divide-by-zero in the
+haystack!  But it turns out that the culprit is a small sub-value:
 %
@@
 \begin{figure}
-\begin{lstlisting}
+\begin{code}
 reduce0 :: forall a . SubTypes a
@@
       Nothing -> return cex
-      Just v  -> 
-        case testHole v of
+      Just v  -> case testHole v of
                    Just v' -> reduce0 v' prop
-          Nothing -> do
+                   Nothing -> test' v idx
+
+  test' v idx = do
     vs <- newVals v
     case test cex idx vs prop of
-              Nothing   -> 
-                reduce' (idx+1)
-              Just cex' -> 
-                reduce0 cex' prop
+      Nothing   -> reduce' (idx+1)
+      Just cex' -> reduce0 cex' prop
 
@@
           else Just a'
-\end{lstlisting}
+\end{code}
   \caption{Reduction algorithm with the sub-value counterexample optimization.}
@@
 depth L         = 0
-depth (B t0 t1) = 1 + max (depth t0) (depth t1)
+depth (B t0 t1) =
+  1 + max (depth t0) (depth t1)
 \end{code}
@@
 \section{Counterexample Generalization} \label{sec:generalization}
-Suppose we are testing a value.
+
+Small counterexamples make debugging easier, but they are just half the battle.
+To go from a specific counterexample to the required fix in a program, the
+programmer must have a flash of insight in which she generalizes the
+counterexample to a set of counterexamples for which the program and property
+fails.  For example, in the introduction, we realized that the specific
+counterexample
+%
+\begin{code}
+B [A (-32769)] [] [] []
+\end{code}
+%
+\noindent
+is representative of the set of counterexamples in which underflow causes the
+sum of the fields to be greater or equal to 64.
+
+The generalization step is an important under-appreciated step in the debugging
+process.  For counterexamples that are large, deeply-nested values, the
+generalization step can be difficult.  One way we might help the programmer is
+to provide her not with a single counterexample but with a quantified formula
+that characterizes a set of counterexamples.  Doing so has two benefits.  First,
+it already generalizes a counterexample, perhaps helping the programmer's
+insight in the generalization step.  Second, it reduces the noise in favor of
+the signal by abstracting away portions of large counterexample that are
+irrelevant to why it violates the property.
+
+The characterization of counterexamples that most helps the programmer seems to
+should strike a middle ground.  At the one extreme, a single counterexample is
+can be thought of as a closed formula containing only literals that
+under-approximates the set of counterexamples.  At the other extreme, the
+property itself being tested is a formula that over-approximates the failing
+inputs.  We describe two kinds of formula that fall between these two extremes.
+
+\subsection{Universal Sub-Value Generalization}
+An example best explains the first kind of quantified formulas we generate.
+Compare the following formula to the counterexample in Figure~\ref{fig:div-cex}:
+%
+\begin{code}
+\ttp{forall x, y . Div (Div x (Div (C 2) (C 13))) y}
+\end{code}
+%
+\noindent
+Here, the formula claims that for all well-typed values \ttp{x} and \ttp{y}, the
+property fails.  Not only does the generalization assist the programmer's
+insight, but it reduces the sheer size of the counterexample, and it can do so significantly.
+
+\begin{figure}
+  \begin{code}
+extrapolate :: SubTypes a
+       => a -> (a -> Property) -> IO [Idx]
+extrapolate cex prop = extrapolate' 1 []
+  where
+  extrapolate' idx idxs =
+    case index cex idx of
+      Nothing -> return idxs
+      Just v  -> do
+        vs <- extrapVals v
+        if allFail cex idx vs prop
+          then extrapolate' (idx+1) (idx:idxs)
+          else extrapolate' (idx+1) idxs
+
+extrapVals :: SubVal -> IO [SubVal]
+extrapVals (SubVal a) =
+  sequence (replicate maxExtrap s)
+  where
+  s = unsizedArbitrary a >>= return . SubVal
+
+allFail :: SubTypes a
+     => a -> Idx -> [SubVal]
+     -> (a -> Property) -> Bool
+allFail cex idx vs prop =
+  (not (null res)) && and res
+  where
+  res  = catMaybes (map (fail prop) vals)
+  vals = map (replace cex idx) vs
+  \end{code}
+  \caption{Universal sub-value generation algorithm.}
+  \label{fig:universal}
+\end{figure}
+
+We call the kind of formula just shown \emph{universal sub-value generalization}
+and is implemented in SmartCheck.  The algorithm is shown in
+Figure~\ref{fig:universal}.  The algorithm is similar to the reduction algorithm
+in Figure~\ref{fig:reduction} (and in the implementation, the algorithms are
+generalized and combined).  The function \ttp{extrapolate} takes a
+counterexample \ttp{cex} and a property.  It returns a list of indexes to be
+universally quantified.  This algorithm also performs a breadth-first search
+over the sub-values of the counterexample.  For each sub-value, the algorithm
+generates new sub-values, replaces them in \ttp{cex}, and tests them.  If every
+new counterexample non-trivially fails (i.e., does not just fail the
+precondition) the property \ttp{prop}, then we extrapolate, claiming that for
+\emph{any} new value replacing that sub-value in \ttp{cex}, the property will
+fail.
+
+The function \ttp{extrapVals} generates the new sub-values, much like
+\ttp{newVal} in the reduction algorithm.  Here, \ttp{maxExtrap} well-typed
+values are generated, and \ttp{unsizedArbitrary} is the analogue of
+\ttp{sizedArbitrary}, but does not limit the size of values generated.
+
+In the definition of \ttp{allFail},
+%
+\begin{code}
+fail :: (a -> Property) -> a -> Maybe Bool
+\end{code}
+%
+\noindent
+is the roughly the dual of \ttp{pass} in the reduction algorithm: it returns
+\ttp{Just True} if a value satisfies the property and passes the property's
+precondition \ttp{pre}, \ttp{Just False} if the value fails the property, and
+\ttp{Nothing} if the value fails \ttp{pre}.
+
+Thus,
+
+\subsection{Existential Sub-Value Generalization}
+
+\subsection{Automated Invariant Strengthening}
+
+\begin{figure}[ht!]
+  \begin{center}
+    \includegraphics[scale=0.5]{Figs/cex-gen}
+  \end{center}
+  \caption{Counterexample generalization.}
+  \label{fig:cex-gen}
+\end{figure}
+
+\subsection{Generalization Details}
+
+\todo{flags, optimizations}
 

### 2012-12-24 paper/Paper.hs

@@ -2,59 +2,64 @@
 
 module Paper where
 
+import Prelude hiding (fail)
+import Data.Maybe (catMaybes)
+
 cast :: SubTypes a => a -> Maybe b
 cast _ = undefined
 
 class Arbitrary a where
-  -- arbitrary :: a
+--  arbitrary :: IO a
 
---type Property a = a -> Bool
+sizedArbitrary :: forall a. SubTypes a => a -> IO a
+sizedArbitrary _ = return (undefined :: a)
+
+unsizedArbitrary :: forall a. SubTypes a => a -> IO a
+unsizedArbitrary _ = return (undefined :: a)
 
 data Property
 
 -- sized :: [a] -> Int -> [a]
 -- sized = undefined
 
-data SubVal = forall a. (Arbitrary a, SubTypes a) => SubVal a
+data SubVal = forall a. SubTypes a => SubVal a
+
+type Idx = Int
 
 class Arbitrary a => SubTypes a where
-  size    :: a -> Int
-  index   :: a -> Int -> Maybe SubVal
-  replace :: a -> Int -> SubVal -> a
+  size    :: a -> Idx
+  index   :: a -> Idx -> Maybe SubVal
+  replace :: a -> Idx -> SubVal -> a
 
-reduce :: SubTypes a 
+reduce :: SubTypes a
        => a -> (a -> Property) -> IO a
 reduce cex prop = reduce' 1
   where
   reduce' idx =
-    case index cex idx of 
+    case index cex idx of
       Nothing -> return cex
       Just v  -> do
-        vs <- newVals v
+        vs <-newVals v
         case test cex idx vs prop of
           Nothing   -> reduce' (idx+1)
           Just cex' -> reduce cex' prop
 
 newVals :: SubVal -> IO [SubVal]
-newVals (SubVal a) = 
+newVals (SubVal a) =
   sequence (replicate maxTries s)
-  where s =     sizedArbitrary a 
-            >>= return . SubVal
+  where
+  s = sizedArbitrary a >>= return . SubVal
 
-test :: SubTypes a 
-     => a -> Int -> [SubVal] 
+test :: SubTypes a
+     => a -> Idx -> [SubVal]
      -> (a -> Property) -> Maybe a
 test cex idx vs prop = go vs
   where
   go []      = Nothing
-  go (v:vs') = 
+  go (v:vs') =
     let cex' = replace cex idx v in
-    if pass prop cex' then go vs' 
+    if pass prop cex' then go vs'
       else Just cex'
-
[…]

### 2012-12-27 paper/paper.tex
paper: updates on extrapolation.

@@
 
+\todo{note constants are user-defined arguments}
+
 \section{Introduction}\label{sec:intro}
@@
 For example, consider the \ttp{shrink} instances given to data types \ttp{A} and
-\ttp{B} in Figure~\ref{fig:initial}.  Indeed, because of the regular structure
-of \ttp{shrink} implementations, the principal motivation for one
-of Haskell's first generics package was for providing generic definitions for
+\ttp{B} in Figure~\ref{fig:initial}.\footnote{All examples and algorithms in
+  this paper are in Haskell~\cite{haskell98}}.  Indeed, because of the regular
+structure of \ttp{shrink} implementations, the principal motivation for one of
+Haskell's first generics package was for providing generic definitions for
 \ttp{shrink}~\cite{syb}.
@@
 type Idx = Int
+type Size = Int
 
 class Arbitrary a => SubTypes a where
-  size    :: a -> Idx
+  size    :: a -> Size
   index   :: a -> Idx -> Maybe SubVal
@@
       Just v  -> do
-        vs <- newVals v
+        vs <- newVals v (getSize v) scMaxFailure
         case test cex idx vs prop of
@@
 
-newVals :: SubVal -> IO [SubVal]
-newVals (SubVal a) =
-  sequence (replicate max s)
+getSize :: SubVal -> Size
+getSize (SubVal a) = size a
+
+newVals :: SubVal -> Size -> Int -> IO [SubVal]
+newVals (SubVal a) sz tries =
+  sequence (replicate tries s)
   where
-  s = sizedArbitrary a >>= return . SubVal
+  s  =     sizedArbitrary sz a
+       >>= return . SubVal
 
@@
 on a random seed for generation (we could pass in the generator seed, but for
-simplicity, we assume one is generated locally).  The constant \ttp{maxTries} is
+simplicity, we assume one is generated locally).  The constant \ttp{scMaxFailure} is
 the maximum number of tries to discover a new counterexample by replacing
@@
   test' v idx = do
-    vs <- newVals v
+    vs <- newVals v scMaxFailure
     case test cex idx vs prop of
@@
 property fails.  Not only does the generalization assist the programmer's
-insight, but it reduces the sheer size of the counterexample, and it can do so significantly.
+insight, but it reduces the sheer size of the counterexample, and it can do so
+significantly.  We call the kind of formula just shown \emph{universal sub-value
+  generalization} and is implemented in SmartCheck.
+
+An extrapolation algorithm performs universal sub-value generalization.  The
+basic idea is as follows: for a counterexample \ttp{cex} and a property
+\ttp{prop}, a breadth-first search over the sub-values of the \ttp{cex} is
+performed.  For each sub-value, the algorithm generates new sub-values, replaces
+them in \ttp{cex} to create a list of new potential counterexamples.  If enough
+new counterexamples non-trivially fails \ttp{prop} (i.e., does not just fail the
+precondition), and non satisfy \ttp{prop}, then we extrapolate, claiming that
+for \emph{any} new value replacing that sub-value in \ttp{cex}, the property
+will fail.
 
@@
       Just v  -> do
-        vs <- extrapVals v
+        vs <- newVals v scMaxSize scMaxSuccess
         if allFail cex idx vs prop
@@
           else extrapolate' (idx+1) idxs
-
-extrapVals :: SubVal -> IO [SubVal]
-extrapVals (SubVal a) =
-  sequence (replicate maxExtrap s)
-  where
-  s = unsizedArbitrary a >>= return . SubVal
 
@@
 allFail cex idx vs prop =
-  (not (null res)) && and res
+  length res >= scMinExtrap && and res
   where
-  res  = catMaybes (map (fail prop) vals)
-  vals = map (replace cex idx) vs
+  res  = catMaybes (map go vs)
+  go   = fail prop . replace cex idx
   \end{code}
@@
 
-We call the kind of formula just shown \emph{universal sub-value generalization}
-and is implemented in SmartCheck.  The algorithm is shown in
-Figure~\ref{fig:universal}.  The algorithm is similar to the reduction algorithm
-in Figure~\ref{fig:reduction} (and in the implementation, the algorithms are
-generalized and combined).  The function \ttp{extrapolate} takes a
-counterexample \ttp{cex} and a property.  It returns a list of indexes to be
-universally quantified.  This algorithm also performs a breadth-first search
-over the sub-values of the counterexample.  For each sub-value, the algorithm
-generates new sub-values, replaces them in \ttp{cex}, and tests them.  If every
-new counterexample non-trivially fails (i.e., does not just fail the
-precondition) the property \ttp{prop}, then we extrapolate, claiming that for
-\emph{any} new value replacing that sub-value in \ttp{cex}, the property will
-fail.
-
-The function \ttp{extrapVals} generates the new sub-values, much like
-\ttp{newVal} in the reduction algorithm.  Here, \ttp{maxExtrap} well-typed
-values are generated, and \ttp{unsizedArbitrary} is the analogue of
-\ttp{sizedArbitrary}, but does not limit the size of values generated.
-
-In the definition of \ttp{allFail},
+The extrapolation algorithm is shown in Figure~\ref{fig:universal}.  The
+algorithm is similar to the reduction algorithm in Figure~\ref{fig:reduction},
+and in the implementation, the algorithms are generalized and combined.  The
+function \ttp{extrapolate} returns a list of indexes to be generalized in the
+original counterexample.  New sub-values are generated by \ttp{newVals}, shown
+in Figure~\ref{fig:reduction}.  Here, we bound the size of values to generate by
+the constant \ttp{scMaxSize}, which is independent of the size of the particular
+sub-value.
+
+The function \ttp{allFail} takes a counterexample \ttp{cex}, an index into
+\ttp{cex}, a list of new sub-values, and a property.  It returns true if enough
+new values pass the property's precondition and fail the property.  The function
 %
@@
 \noindent
-is the roughly the dual of \ttp{pass} in the reduction algorithm: it returns
-\ttp{Just True} if a value satisfies the property and passes the property's
-precondition \ttp{pre}, \ttp{Just False} if the value fails the property, and
-\ttp{Nothing} if the value fails \ttp{pre}.
-
-Thus,
+is the roughly the dual of \ttp{pass} in the reduction algorithm: \ttp{fail prop
+  cex} returns \ttp{Just True} if \ttp{cex} passes \ttp{prop}'s precondition but
+fails the property; \ttp{Just False} if \ttp{cex} non-trivially satisfies
+\ttp{prop}, and \ttp{Nothing} if \ttp{cex} fails \ttp{prop}'s precondition.  The
+constant \ttp{scMinExtrap} is the minimum number of \ttp{Just False} results
+required from \ttp{fail} to extrapolate from failed tests to a universal claim.
 
@@
 
+\todo{check all option names}
+
 \subsection{Generalization Details}

### 2012-12-27 paper/Paper.hs

@@ -4,6 +4,7 @@ module Paper where
 
 import Prelude hiding (fail)
 import Data.Maybe (catMaybes)
+import qualified Test.QuickCheck as Q
 
 cast :: SubTypes a => a -> Maybe b
 cast _ = undefined
@@ -11,11 +12,48 @@ cast _ = undefined
 class Arbitrary a where
 --  arbitrary :: IO a
 
-sizedArbitrary :: forall a. SubTypes a => a -> IO a
-sizedArbitrary _ = return (undefined :: a)
+sizedArbitrary :: forall a. SubTypes a => Size -> a -> IO a
+sizedArbitrary sz _ = return ((undefined sz) :: a)
 
-unsizedArbitrary :: forall a. SubTypes a => a -> IO a
-unsizedArbitrary _ = return (undefined :: a)
+--------------------------------------------------------------------------------
+
+data Format = PrintTree | PrintString
+  deriving (Eq, Read, Show)
+
+-- SmartCheck arguments
+format       :: Format    -- ^ How to show extrapolated formula
+format       = undefined
+qcArgs       :: Q.Args    -- ^ QuickCheck arguments
+qcArgs       = undefined
+qc           :: Bool      -- ^ Should we run QuickCheck?  (If not, you are
+                          -- expected to pass in data to analyze.)
+qc           = undefined
+extrap       :: Bool      -- ^ Should we extrapolate?
+extrap       = undefined
+constrGen    :: Bool      -- ^ Should we try to generalize constructors?
+constrGen    = undefined
+scMaxSuccess :: Int       -- ^ How hard (number of rounds) to look for failures
+                          -- during the extrapolation and constructor
+                          -- generalization stages.
+scMaxSuccess = undefined
+scMaxFailure :: Int       -- ^ How hard (number of rounds) to look for failure
+                          -- in the reduction stage.
+scMaxFailure = undefined
+scMaxSize    :: Int       -- ^ Maximum size of data to generate, in terms of the
+                          -- size parameter of QuickCheck's Arbitrary instance
+                          -- for your data.
+scMaxSize    = undefined
+scMaxDepth   :: Maybe Int -- ^ How many levels into the structure of the failed
+                          -- value should we descend when reducing or
[…]

### 2012-12-27 src/Test/SmartCheck/Extrapolate.hs
SC: add minimum passed test argument.

@@ -53,10 +53,11 @@ extrapolate args d origProp ds = do
                  (scMaxSize args) origProp
 
   -- Control-flow.
+
+  -- None of the tries satisfy prop (but something passed the precondition).
+  -- Prevent recurring down this tree, since we can generalize.
   next _ (i, FailedProp) forest' idx idxs
-    -- None of the tries satisfy prop (but something passed the precondition).
-    -- Prevent recurring down this tree, since we can generalize.
-    | scMinExtrap args <= i =
+    | scMinExtrap args < i =
         nextIter (forestReplaceChildren forest' idx False) idx (idx : idxs)
   next _ _ forest' idx idxs = nextIter forest' idx idxs
 

### 2012-12-27 src/Test/SmartCheck/Types.hs

@@ -99,7 +99,7 @@ scStdArgs = ScArgs { format       = PrintTree
                    , scMaxFailure = 100
                    , scMaxSize    = 10
                    , scMaxDepth   = Nothing
-                   , scMinExtrap  = 0
+                   , scMinExtrap  = 10
                    }
 
 --------------------------------------------------------------------------------

### 2012-12-28 paper/paper.tex
paper: extrapolation section.

@@
 
-\todo{note constants are user-defined arguments}
+\todo{note constants are user-defined arguments, other simplifications}
 
@@
 aspect of the Haskell development cycle.  Because QuickCheck generates random
-values for testing, counterexamples it finds may not be substantially larger
+values for testing, counterexamples it finds may be substantially larger
 than a minimal counterexample.  In their original QuickCheck paper~\cite{qc},
@@
 \ttp{B} in Figure~\ref{fig:initial}.\footnote{All examples and algorithms in
-  this paper are in Haskell~\cite{haskell98}}.  Indeed, because of the regular
+  this paper are presented in Haskell~\cite{haskell98}.}  Indeed, because of the regular
 structure of \ttp{shrink} implementations, the principal motivation for one of
@@
 either miss minimal counterexamples, be too inefficient to be practical, or
-both.\footnote{E.g., for a recent real-world case, see the user question posed
-  on StackOverflow:
+both.\footnote{For a recent real-world case, see the user question posed on
+  StackOverflow:
   \url{http://stackoverflow.com/questions/8788542/how-do-i-get-good-small-shrinks-out-of-quickcheck}.}
@@
 Consider the property \ttp{prop} in the figure, stating that if the sum of each
-list in a field of \ttp{B} is less than 16, then the sum of the four fields of
-\ttp{B} is less than 64.  The property \ttp{prop} seems reasonable at first
-glance, until one realizes that due to underflow, the property can be violated.
-For example, consider the value
+list in a field of \ttp{B} is less than 16, then the sum of the four fields is
+less than 64.  The property \ttp{prop} seems reasonable at first glance, until
+one realizes that due to underflow, the property can be violated.  For example,
+consider the value
 %
@@
 counterexamples generated by \ttp{shrink} can contain more than $10^{10}$
-elements.  This is an enormous state space to search for a smaller example!  In
-particular, if a counterexample is not found does not finish in a reasonable
-amount of time.
+elements.  This is an enormous state space to search for a smaller example and
+can cause the shrinking phase to take orders-of-magnitude more time than the
+counterexample discovery phase.
 
@@
 shrinking is used to better understand a counterexample to understand why a
-property failed; but to implement an efficient shrink function in general
+property failed, but to implement an efficient shrink function in general
 requires knowing how to generate counterexamples to the property.
@@
 Figure~\ref{fig:initial}, we have defined instances for generating Lazy
-SmallCheck (a more efficient version of SmallCheck) tests by defining instances
-of the \ttp{Serial} class.  For the property \ttp{prop} defined above (suitably
+SmallCheck (a variant of SmallCheck) tests by defining instances
+for the \ttp{Serial} class.  For the property \ttp{prop} defined above (suitably
 redefined to use SmallCheck's types), SmallCheck must check to depth 16 to find
@@
 preceding paragraph.  To motivate the benefit of SmartCheck, consider the table
-in Table~\ref{fig:results}.  The table characterizes testing the program shown
-in Figure~\ref{fig:initial} that we have described so far.  Four trials are
-presented.  In the first three, we use QuickCheck to generate counterexamples.
-The first column contains the results of using QuickCheck without shrinking to
-provide a baseline.  The next two columns show the result of using QuickCheck
-with definitions of \ttp{shrink} that limit the lists of fields of \ttp{B} to 10
-(Figure~\ref{lst:newshrink}) and 20 elements (Figure~\ref{lst:newshrink}, with
-\ttp{len = 20}), respectively.  (Recall that with the original definition for
-shrink from Figure~\ref{fig:initial}, some counterexamples cause QuickCheck to
-not return after hours of computation.)  In each row, we show the maximum, mean,
-and median number of seconds required by QuickCheck to generate a counterexample
-over 100 runs.  In the last row, we show the mean size of the final value
-returned, measured by the taking the cumulative length of the lists of the
-fields of \ttp{B}.  Notice that while typically counterexamples are returned in
-a fraction of a second, there are outliers taking 21 seconds and 125 seconds for
-the two tests.  Both tests roughly cut the size of the counterexample in half;
-increasing the number of potential counterexamples can result in a significant
-performance penalty for this example without much benefit.
+in Table~\ref{table:results}, giving the results for using QuickCheck and
+SmartCheck on the program shown in Figure~\ref{fig:initial} (we omit SmallCheck,
+since as we noted, it is infeasible for an example like this).  In the first
+three columns, the results of using QuickCheck to generate counterexamples are
+shown.  The first column contains the results of using QuickCheck without
+shrinking to provide a baseline.  The next two columns show the result of using
+QuickCheck with definitions of \ttp{shrink} that limit the lists of fields of
+\ttp{B} to 10 (Figure~\ref{lst:newshrink}) and 20 elements
+(Figure~\ref{lst:newshrink}, with \ttp{len = 20}), respectively.  (Recall that
+with the original definition for shrink from Figure~\ref{fig:initial}, some
+counterexamples cause QuickCheck to not return after hours of computation.)  In
+each row, we show the maximum, mean, and median number of seconds required by
+QuickCheck to generate a counterexample over 100 runs.  In the last row, we show
+the mean size of the final value returned, measured by the taking the cumulative
+length of the lists of the fields of \ttp{B}.  Notice that while typically
+counterexamples are returned in a fraction of a second, there are outliers
+taking 21 seconds and 125 seconds for the two tests.  Both tests roughly cut the
+size of the counterexample in half; increasing the number of potential
+counterexamples can result in a significant performance penalty for this example
+without much benefit.
 
@@
 
-test :: SubTypes a
-     => a -> Idx -> [SubVal]
+test :: SubTypes a => a -> Idx -> [SubVal]
      -> (a -> Property) -> Maybe a
@@
 
-\paragraph{Sum-type counterexample hypothesis}
+\subsubsection{Sum-type counterexample hypothesis}
 First, recall that the algorithm begins at the \emph{first} sub-value of the
@@
 
-\paragraph{Conditional properties}
+\subsubsection{Conditional properties}
 A QuickCheck property may be a conditional of the form
@@
 
-\paragraph{Sub-value counterexample hypothesis}
+\subsubsection{Sub-value counterexample hypothesis}\label{sec:subval}
 Sometimes, a counterexample fails a property due to a sub-value nested deep
@@
 \begin{code}
-div_prop e = divSubTerms e ==> eval e /= Nothing
-\end{code}
-%
-\noindent
-The property \ttp{div\_prop} turns out to be false; for example,
+prop_div e = divSubTerms e ==> eval e /= Nothing
+\end{code}
+%
+\noindent
+The property \ttp{prop\_div} turns out to be false; for example,
 %
@@
 
-\paragraph{Base types}
+\subsubsection{Base types}
 
@@
 
-\paragraph{Bounding counterexample exploration}
+\subsubsection{Bounding counterexample exploration}
 
@@
   where
-  extrapolate' idx idxs =
+  extrapolate' idx idxs
+    | or (map (subTree cex idx) idxs) =
+      extrapolate' (idx+1) idxs
+  extrapolate' idx idxs
+    | otherwise                       =
       case index cex idx of
@@
 function \ttp{extrapolate} returns a list of indexes to be generalized in the
-original counterexample.  New sub-values are generated by \ttp{newVals}, shown
-in Figure~\ref{fig:reduction}.  Here, we bound the size of values to generate by
+original counterexample.  In the recursive function \ttp{extrapolate'}, there is
+a function guard in which a call to \ttp{subTree}
+%
+\begin{code}
+subTree :: SubTypes a => a -> Idx -> Idx -> Bool
+\end{code}
+%
+\ttp{subTree cex idx0 idx1} is true if in \ttp{cex}, the value at index
+\ttp{idx0} is a child of index \ttp{idx1}, in a tree representation of the
+data.  This prevents the algorithm from trying to generalize sub-values that are
+abstracted away already since their parents have been generalized.
+
+New sub-values are generated by \ttp{newVals}, shown in
+Figure~\ref{fig:reduction}.  Here, we bound the size of values to generate by
 the constant \ttp{scMaxSize}, which is independent of the size of the particular
@@
 
+\todo{Add discussion here}
+
 \subsection{Existential Sub-Value Generalization}
+
+Sum types denote choice in a data type.  Sometimes, a property over a sum type
+fails because there is a bug for some of the variants but not others.  For
+example, recall the calculator language from Section~\ref{sec:subval}.  The
+property \ttp{prop\_div} fails only for values that contain a variant tagged
+with the \ttp{Div} constructor.  For example, consider the following counterexample:
+%
+\begin{code}
+Div (Add (C 20) (C 30))
+    (Add (C 89) (C -89))
+\end{code}
+%
+\noindent
+The counterexample fails because the divsor evaluates to zero.  Using universal
+sub-value generalization described above, the counterexample is generalized as
+follows:
+%
+\begin{code}
+forall x . Div x (Add (C 89) (C -89))
+\end{code}
+%
+\noindent
+So we know there is something particular about the \ttp{Add (C 89) (C -89)},
+because there are some values that we can replace this sub-value with in the
+original counterexample causing it to satisfy the property.  But we might wonder
+if there is something special about variants tagged by the \ttp{Add}
+constructor, or might we finding failing sub-values with the other variants
+tagged by the constructors \ttp{C} and \ttp{Div}?  If so, then we know there is
+not some bug being triggered by \ttp{Add} variants specifically, so it is more
+likely that the bug involves how the second field of the \ttp{Div} variant is
+evaluated.
+
+We therefore introduce another kind of generalization we call \emph{existential
+  sub-value generalization}.  In this generalization, if there a counterexample
+containing every possible variant as a sub-value, then we abstract it.  For
+example, the following all characterize counterexamples:
+%
+\begin{code}
+forall x . Div x (Add (C 89) (C -89))
+forall x . Div x (C 0)
+forall x . Div x (Div (C 0) (C 1))
+\end{code}
+%
+\noindent
+Because there are only three variants in the type, we have found counterexamples
+built from each of them in the second field of the outermost \ttp{Div}
+constructor.  We therefore return a formula like the following:
+%
+\begin{code}
+forall values x .
+  forall constructors C\(\sb{0}\) .
+    there exist arguments \(\stackrel{\rightarrow}{y}\) .
+      such that Div x (C\(\sb{0} \stackrel{\rightarrow}{y}\))
+\end{code}
+%
 

### 2012-12-28 paper/Paper.hs

@@ -93,8 +93,7 @@ newVals (SubVal a) sz tries =
   s  =     sizedArbitrary sz a
        >>= return . SubVal
 
-test :: SubTypes a
-     => a -> Idx -> [SubVal]
+test :: SubTypes a => a -> Idx -> [SubVal]
      -> (a -> Property) -> Maybe a
 test cex idx vs prop = go vs
   where
@@ -140,18 +139,25 @@ reduce0 cex prop = reduce' 1
 
 --------------------------------------------------------------------------------
 
+subTree :: SubTypes a => a -> Idx -> Idx -> Bool
+subTree _ _ _ = undefined
+
 extrapolate :: SubTypes a
        => a -> (a -> Property) -> IO [Idx]
 extrapolate cex prop = extrapolate' 1 []
   where
-  extrapolate' idx idxs =
-    case index cex idx of
-      Nothing -> return idxs
-      Just v  -> do
-        vs <- newVals v scMaxSize scMaxSuccess
-        if allFail cex idx vs prop
-          then extrapolate' (idx+1) (idx:idxs)
-          else extrapolate' (idx+1) idxs
+  extrapolate' idx idxs
+    | or (map (subTree cex idx) idxs) =
+      extrapolate' (idx+1) idxs
+  extrapolate' idx idxs
+    | otherwise                       =
+      case index cex idx of
+        Nothing -> return idxs
+        Just v  -> do
+          vs <- newVals v scMaxSize scMaxSuccess
+          if allFail cex idx vs prop
+            then extrapolate' (idx+1) (idx:idxs)
+            else extrapolate' (idx+1) idxs
 
 allFail :: SubTypes a
      => a -> Idx -> [SubVal]

### 2012-12-28 examples/Div0.hs
Div0 example: wibble.

@@ -19,22 +19,22 @@ data Exp = C Int
          | Div Exp Exp
   deriving (Read, Show, Typeable, Generic)
 
-instance SubTypes Exp 
+instance SubTypes Exp
 
 eval :: Exp -> Maybe Int
 eval (C i) = Just i
 eval (Add e0 e1) =
   liftM2 (+) (eval e0) (eval e1)
-eval (Div e0 e1) = 
-  let e = eval e1 in 
-  if e == Just 0 then Nothing 
+eval (Div e0 e1) =
+  let e = eval e1 in
+  if e == Just 0 then Nothing
     else liftM2 div (eval e0) e
 
 instance Arbitrary Exp where
   arbitrary = sized mkM
     where
     mkM 0 = liftM C arbitrary
-    mkM n = oneof [ liftM2 Add mkM' mkM' 
+    mkM n = oneof [ liftM2 Add mkM' mkM'
                   , liftM2 Div mkM' mkM' ]
       where mkM' = mkM =<< choose (0,n-1)
 
@@ -44,9 +44,9 @@ instance Arbitrary Exp where
 
 -- property: so long as 0 isn't in the divisor, we won't try to divide by 0.
 -- It's false: something might evaluate to 0 still.
-div_prop :: Exp -> Property
-div_prop e = divSubTerms e ==> eval e /= Nothing
--- div_prop e = property $ case x of
+prop_div :: Exp -> Property
+prop_div e = divSubTerms e ==> eval e /= Nothing
+-- prop_div e = property $ case x of
 --                           Nothing -> True
 --                           Just True -> True
 --                           _       -> False
@@ -62,15 +62,15 @@ divSubTerms (Div e0 e1)   = divSubTerms e0 && divSubTerms e1
 -- div0 (A _ _) = property False
 -- div0 _       = property True
 
--- test_prop m = case eval m of
+-- prop_test m = case eval m of
 --                 Nothing -> True
 --                 Just i -> i < 5
-              
+
 
 divTest :: IO ()
-divTest = smartCheck args div_prop
-  where 
-  args = scStdArgs { qcArgs  = stdArgs 
+divTest = smartCheck args prop_div
+  where
+  args = scStdArgs { qcArgs  = stdArgs
                                 -- { maxSuccess = 1000
                                 -- , maxSize    = 20  }
                    , format  = PrintString

### 2012-12-28 src/Test/SmartCheck.hs
SC: fix bug in reading in counterexamples.

@@ -1,9 +1,9 @@
 {-# LANGUAGE FlexibleContexts #-}
-{-# LANGUAGE ScopedTypeVariables #-} 
+{-# LANGUAGE ScopedTypeVariables #-}
 
 -- | Interface module.
 
-module Test.SmartCheck 
+module Test.SmartCheck
   ( smartCheck
   , runQC
   , module Test.SmartCheck.Types
@@ -39,9 +39,9 @@ smartCheck args propT = do
     -- Run standard QuickCheck or read in value.
     res <- if qc args then runQC (qcArgs args) prop'
              else do smartPrtLn "Input value to SmartCheck:"
-                     getLine >>= return . read
-    case res of 
-      Nothing -> smartPrtLn "No value to smart-shrink; done." 
+                     fmap Just (readLn :: IO a)
+    case res of
+      Nothing -> smartPrtLn "No value to smart-shrink; done."
       Just r  -> go r
 
     where
@@ -58,17 +58,17 @@ smartCheck args propT = do
 
       -- If we asked to extrapolate constructors, do so.
       cs  <- if constrGen args
-               then     constrsGen args d prop (concat . maybeToList $ fmap fst vs) 
+               then     constrsGen args d prop (concat . maybeToList $ fmap fst vs)
                     >>= return . Just
                else return Nothing
 
       -- If either kind of extrapolation pass yielded fruit, prettyprint it.
-      if nonEmpty d vs cs 
+      if nonEmpty d vs cs
         then output d (repls vs cs)
         else smartPrtLn "Could not extrapolate a new value."
 
       -- Ask the user if she wants to try again.
-      putStrLn $ "Attempt to find a new counterexample?\n" 
+      putStrLn $ "Attempt to find a new counterexample?\n"
                    ++ "  ('Enter' to continue;"
                    ++ " any character then 'Enter' to quit.)"
       s <- getLine
@@ -92,7 +92,7 @@ smartCheck args propT = do
                    Nothing        -> False
         csIdxs = case cs of
                    Just idxs -> not $ null idxs
-                   Nothing   -> False 
+                   Nothing   -> False
 
 ---------------------------------------------------------------------------------
 
@@ -104,7 +104,7 @@ runQC args prop = do
   case res of
     -- XXX C'mon, QuickCheck, let me grab the result in a sane way rather than
     -- parsing a string!
[…]

### 2012-12-29 paper/paper.tex
paper: updates.

@@
 
+data SubVal =
+  forall a. SubTypes a => SubVal a
+
 class Arbitrary a => SubTypes a where
@@
   replace :: a -> Idx -> SubVal -> a
-\end{code}
-%
-\noindent
-\ttp{SubTypes} has three methods.  The \ttp{size} method returns the size of a
-value, \ttp{index} returns a sub-value at an breadth-first index in a value, and
-\ttp{replace} replaces a sub-value at a particular focus.  If the index is
-out-of-bounds (too large or negative), \ttp{index} return \ttp{Nothing}.  The
-method \ttp{replace} returns the original value for an out-of-bounds index or an
-incorrectly-typed replacement value.  The \ttp{SubVal} type is an existential
-data type, used as a generic container for sub-values from a counterexample.
-%
-\begin{code}
-data SubVal =
-  forall a. SubTypes a => SubVal a
-\end{code}
-%
-\noindent
-For example, consider a binary tree type
+  constr  :: a -> String
+  constrs :: a -> [String]
+\end{code}
+%
+\noindent
+\ttp{SubTypes} has five methods.  The \ttp{size} method returns the size of a
+value, \ttp{index} returns a sub-value at an breadth-first index in a value,
+\ttp{replace} replaces a sub-value at a particular focus, \ttp{constr} returns
+the constructor tagging the value, and \ttp{constrs} returns a list of unique
+identifiers for the constructors associated with each variant of the type.
+
+For example, consider a binary tree type:
 %
@@
 \noindent
-Here are example applications of \ttp{SubTypes}; in the following, we show
-the indexes with respect to the \ttp{tree} value:
+Here are example applications of \ttp{SubTypes} methods; in the following, we
+show the indexes with respect to the \ttp{tree} value:
 %
@@
 index tree 4 = (Just . SubVal) (B\sub{4} L\sub{6} L\sub{8})
-
 index tree 12 = Nothing
@@
      L
-\end{code}
-%
-\noindent
-We will sometimes refer to unwrapped value returned by \ttp{index a i} as the
-\emph{$i$th sub-value of $a$}, so for example, \ttp{B\sub{4} L\sub{6} L\sub{8}}
-is the 4th sub-value of \ttp{tree}.
+
+constr  tree = ["B"]
+constrs tree = ["B", "L"]
+constrs L    = ["B", "L"]
+\end{code}
+%
+\noindent
+We will sometimes refer to the unwrapped value returned by \ttp{index a i} as
+the \emph{$i$th sub-value of $a$}, so for example, \ttp{B\sub{4} L\sub{6}
+  L\sub{8}} is the 4th sub-value of \ttp{tree}.  An invariant of \ttp{index} is
+that for any value \ttp{a}, consider the least $i \geq 0$ such that \ttp{index a
+  i == Nothing}.  For all $0 \leq j < i$, \ttp{index a j /= Nothing}.  We use
+this invariant as a termination case in recursive algorithms over the sub-values
+of a value.
+
+The method \ttp{replace} returns the original value for an out-of-bounds index
+or an incorrectly-typed replacement value.  The \ttp{SubVal} type is an
+existential data type, used as a generic container for sub-values from a
+counterexample.
+
+Finally, while \ttp{constr} returns a string representation of the constructor
+associated with the specific variant passed to it, \ttp{constrs} ignores the
+value and returns the same list of constructors for any value from the type.
+(In our implementation, \ttp{constrs} depends on GHC Generics~\cite{generics},
+which we describe in Section~\ref{sec:implementation}.  For simplicity, we omit
+here Generics-specific super-class constraints on the \ttp{SubTypes} class.)
 
@@
   \begin{code}
+getSize :: SubVal -> Size
+getSize (SubVal a) = size a
+
+newVals :: SubVal -> Size -> Int -> IO [SubVal]
+newVals (SubVal a) sz tries =
+  replicateM tries s
+  where
+  s  = liftM SubVal (sizedArbitrary sz a)
+
 reduce :: SubTypes a
@@
       Just v  -> do
-        vs <- newVals v (getSize v) scMaxFailure
+        vs <- newVals v (getSize v) scMaxReduce
         case test cex idx vs prop of
@@
           Just cex' -> reduce cex' prop
-
-getSize :: SubVal -> Size
-getSize (SubVal a) = size a
-
-newVals :: SubVal -> Size -> Int -> IO [SubVal]
-newVals (SubVal a) sz tries =
-  sequence (replicate tries s)
-  where
-  s  =     sizedArbitrary sz a
-       >>= return . SubVal
 
@@
 
+\todo{explain getSize}
+
 \ttp{sizedArbitrary} returns a value inside the \ttp{IO} monad since it depends
 on a random seed for generation (we could pass in the generator seed, but for
-simplicity, we assume one is generated locally).  The constant \ttp{scMaxFailure} is
+simplicity, we assume one is generated locally).  The constant \ttp{scMaxReduce} is
 the maximum number of tries to discover a new counterexample by replacing
@@
   test' v idx = do
-    vs <- newVals v scMaxFailure
+    vs <- newVals v (getSize v) scMaxReduce
     case test cex idx vs prop of
@@
 \begin{code}
-scMaxFailure :: Int
+scMaxReduce :: Int
 scMaxSize   :: Int
@@
 \noindent
-The ``\ttp{sc}'' prefix stands for \emph{S}mart\emph{C}heck and is used to
-prevent name clashes with QuickCheck's flags.  \ttp{scMaxFailure} controls the
-number of values generated by the reduction algorithm for each sub-value
-analyzed.  \ttp{scMaxSize} controls the maximum size of values generated to
-replace sub-values by the reduction algorithm.  Thus, new sub-values must be
-strictly smaller than the minimum of \ttp{scMaxSize} and the size of the
-sub-value being replaced.  (The size of the value is with respect to the
-\ttp{size} method in the \ttp{SubTypes} class.)  Finally, \ttp{scMaxDepth}
-determines the maximum depth in the counterexample the reduction algorithm
-should analyze.  A value of \ttp{Nothing} means that the counterexample should
-be exhaustively reduced, as the algorithm is presented in
+The \ttp{scMaxReduce} flag controls the number of values generated by the
+reduction algorithm for each sub-value analyzed.  \ttp{scMaxSize} controls the
+maximum size of values generated to replace sub-values by the reduction
+algorithm.  Thus, new sub-values must be strictly smaller than the minimum of
+\ttp{scMaxSize} and the size of the sub-value being replaced.  (The size of the
+value is with respect to the \ttp{size} method in the \ttp{SubTypes} class.)
+Finally, \ttp{scMaxDepth} determines the maximum depth in the counterexample the
+reduction algorithm should analyze.  A value of \ttp{Nothing} means that the
+counterexample should be exhaustively reduced, as the algorithm is presented in
 Figure~\ref{fig:reduction}.  For example, the depth of a list is determined by
@@
   \begin{code}
+subTrees :: SubTypes a => a -> Idx -> [Idx] -> Bool
+subTrees cex idx = any (subTree cex idx)
+
 extrapolate :: SubTypes a
@@
   extrapolate' idx idxs
-    | or (map (subTree cex idx) idxs) =
-      extrapolate' (idx+1) idxs
-  extrapolate' idx idxs
-    | otherwise                       =
+    | subTrees cex idx idxs
+    = extrapolate' (idx+1) idxs
+  extrapolate' idx idxs =
     case index cex idx of
@@
       Just v  -> do
-          vs <- newVals v scMaxSize scMaxSuccess
-          if allFail cex idx vs prop
-            then extrapolate' (idx+1) (idx:idxs)
-            else extrapolate' (idx+1) idxs
+        vs <- newVals v scMaxSize scMaxExtrap
+        extrapolate' (idx+1)
+          (if allFail cex idx vs prop
+             then idx:idxs else idxs)
 
@@
   where
-  res  = catMaybes (map go vs)
+  res  = mapMaybe go vs
   go   = fail prop . replace cex idx
@@
 
+\begin{figure}
+  \begin{code}
+subConstr :: SubVal -> String
+subConstr (SubVal a) = constr a
+
+subConstrs :: SubVal -> [String]
+subConstrs (SubVal a) = constrs a
+
+sumTest :: SubTypes a
+          => a -> (a -> Property) -> [Idx] -> IO [Idx]
+sumTest cex prop exIdxs = sumTest' 1 []
+  where
+  sumTest' idx idxs
+    | subTrees cex idx (exIdxs ++ idxs)
+    = sumTest' (idx+1) idxs
+  sumTest' idx idxs =
+    case index cex idx of
+      Nothing -> return idxs
+      Just v  -> do
+        vs <- newVals v scMaxSize scConstrMax
+        sumTest' (idx+1)
+          (if constrFail cex idx vs prop
+                (subConstr v) (subConstrs v)
+             then idx:idxs else idxs)
+
+constrFail :: SubTypes a => a -> Idx -> [SubVal]
+  -> (a -> Property) -> String -> [String] -> Bool
+constrFail cex idx vs prop con allCons =
+  constrFail' [con] vs
+  where
+  constrFail' cons vs'
+    | length cons == length allCons
+    = True
+    | null vs'
+    = False
+    | go v == Just True
+    = constrFail' cs (tail vs')
+    | otherwise
+    = constrFail' cons (tail vs')
+
+    where
+    v  = head vs'
+    c' = subConstr v
+    cs = if c' `elem` allCons
+           then cons else c':cons
+
+  go = fail prop . replace cex idx
+  \end{code}
+  \caption{Universal sub-value generation algorithm.}
+  \label{fig:constrs}
+\end{figure}
+
+The algorithm
+
 \subsection{Automated Invariant Strengthening}
@@
 
-\section{Implementation}
+\section{Implementation}\label{sec:implementation}
 
@@
 \todo{GADTs? don't work}
+\todo{add lazy testing---proof value is unused}
 
@@
 \todo{xmonad}
+\todo{use SC to test itself?}
 
@@
 
+\todo{future work: higher-kinded data, functions}
+
 \section*{Acknowledgements}

### 2012-12-29 paper/Paper.hs

@@ -1,21 +1,18 @@
 {-# LANGUAGE ExistentialQuantification, ScopedTypeVariables #-}
 
+-- | Model of the SmartCheck algorithms explained in the paper.  We leave a
+-- number of functions undefined here.
+
 module Paper where
 
 import Prelude hiding (fail)
-import Data.Maybe (catMaybes)
+import Data.Maybe (mapMaybe)
+import Control.Monad (liftM, replicateM)
 import qualified Test.QuickCheck as Q
 
-cast :: SubTypes a => a -> Maybe b
-cast _ = undefined
-
-class Arbitrary a where
---  arbitrary :: IO a
-
-sizedArbitrary :: forall a. SubTypes a => Size -> a -> IO a
-sizedArbitrary sz _ = return ((undefined sz) :: a)
-
 --------------------------------------------------------------------------------
+
+-- Arguments
 
 data Format = PrintTree | PrintString
   deriving (Eq, Read, Show)
@@ -28,17 +25,6 @@ qcArgs       = undefined
 qc           :: Bool      -- ^ Should we run QuickCheck?  (If not, you are
                           -- expected to pass in data to analyze.)
 qc           = undefined
-extrap       :: Bool      -- ^ Should we extrapolate?
-extrap       = undefined
-constrGen    :: Bool      -- ^ Should we try to generalize constructors?
-constrGen    = undefined
-scMaxSuccess :: Int       -- ^ How hard (number of rounds) to look for failures
-                          -- during the extrapolation and constructor
-                          -- generalization stages.
-scMaxSuccess = undefined
-scMaxFailure :: Int       -- ^ How hard (number of rounds) to look for failure
-                          -- in the reduction stage.
-scMaxFailure = undefined
 scMaxSize    :: Int       -- ^ Maximum size of data to generate, in terms of the
                           -- size parameter of QuickCheck's Arbitrary instance
                           -- for your data.
@@ -48,27 +34,73 @@ scMaxDepth   :: Maybe Int -- ^ How many levels into the structure of the failed
                           -- generalizing?  Nothing means we go down to base
                           -- types.
 scMaxDepth   = undefined
-
+scMaxReduce  :: Int       -- ^ How hard (number of rounds) to look for failure
[…]

### 2012-12-30 examples/Paper/Paper.hs
Paper.hs: wibbles.

@@ -27,7 +27,12 @@ import System.Environment
 -- single element to have a long list.
 
 -- Container so that we don't have base types.
+
+-----------------------------------------------------------------
 data A = A Int16
+  deriving (Read, Show, Typeable, Generic)
+
+data B = B [A] [A] [A] [A]
   deriving (Read, Show, Typeable, Generic)
 
 instance Serial Int16 where
@@ -41,19 +46,18 @@ instance Arbitrary A where
   shrink (A i) = map A (shrink i)
 
 instance Serial A where
-  series = cons1 A 
+  series = cons1 A
 
-data B = B [A] [A] [A] [A]
-  deriving (Read, Show, Typeable, Generic)
 
 instance SubTypes B
 
 -- qc/shrink takes over 1m seconds
 instance Arbitrary B where
-  arbitrary = liftM4 B arbitrary arbitrary arbitrary arbitrary
-  shrink (B a b c d) = 
-    if defShrink 
-      then [ B w x y z | 
+  arbitrary = liftM4 B arbitrary arbitrary
+                       arbitrary arbitrary
+  shrink (B a b c d) =
+    if defShrink
+      then [ B w x y z |
              w <- tk a
            , x <- tk b
            , y <- tk c
@@ -62,18 +66,14 @@ instance Arbitrary B where
     where tk x = take 10 (shrink x)
 
 instance Serial B where
-  series = cons4 B 
-
-size :: B -> Int
-size (B a b c d) = length a + length b + length c + length d
+  series = cons4 B
 
 add :: [A] -> Int16
 add = sum . map (\(A i) -> i)
 
 pre :: B -> Bool
-pre (B a b c d) = and $ map pre' [a, b, c, d]
-  where pre' x = add x < 16
-  
+pre (B a b c d) = all (\x -> add x < 16) [a, b, c, d]
+
 post :: B -> Bool
 post (B a b c d) = add a + add b + add c + add d < 64
 
@@ -86,7 +86,7 @@ prop_sc p = pre p S.==> post p
 test :: IO (Maybe B) -> IO ()
 test run = do
   [arg] <- getArgs
-  let rnds = (read arg) :: Int
+  let rnds = read arg :: Int
   res <- replicateM rnds (test' run)
 
   putStrLn ""
@@ -108,20 +108,20 @@ test run = do
 type Res = Maybe (NominalDiffTime, Int)
 
 test' :: IO (Maybe B) -> IO Res
-test' run = do 
+test' run = do
 
[…]

### 2012-12-30 paper/paper.tex
paper: outline, refactoring.

@@
 
+\todo{define variant, tag}
 \todo{note constants are user-defined arguments, other simplifications}
@@
 The QuickCheck testing framework was a revolutionary step-forward in
-type-directed random testing~\cite{qc}.  Nowadays, QuickCheck is a standard
-aspect of the Haskell development cycle.  Because QuickCheck generates random
-values for testing, counterexamples it finds may be substantially larger
-than a minimal counterexample.  In their original QuickCheck paper~\cite{qc},
-the authors report the following user experience by Andy Gill:
+type-directed random testing~\cite{qc}.  Originally designed for Haskell,
+QuickCheck has been ported to other languages and is a now a widely-used testing
+tool.  Because QuickCheck generates random values for testing, counterexamples
+it finds may be substantially larger than a minimal counterexample.  In their
+original QuickCheck paper~\cite{qc}, the authors report the following user
+experience by Andy Gill:
 %
@@
 -- Predicates for defining properties
+add :: [A] -> Int16
+add = sum . map (\(\backslash\)(A i) -> i)
+
 pre :: B -> Bool
-pre (B a b c d) = and \$ map pre' [a, b, c, d]
-  where pre' x = add x < 16
+pre (B a b c d) =
+  all (\(\backslash\) -> add x < 16) [a, b, c, d]
 
@@
   add a + add b + add c + add d < 64
-  where add :: [A] -> Int16
-        add = sum . map (\(\backslash\)(A i) -> i)
 
@@
 Figure~\ref{fig:initial} produces an intractable number of potential
-counterexamples to test; with the definitions and property provided, which also
-rely on QuickCheck's default instances for the list type, the list of potential
-counterexamples generated by \ttp{shrink} can contain more than $10^{10}$
-elements.  This is an enormous state space to search for a smaller example and
-can cause the shrinking phase to take orders-of-magnitude more time than the
-counterexample discovery phase.
+counterexamples to test; with the definitions and property provided (together
+with QuickCheck's default \ttp{Arbitrary} instance for lists), the list of
+potential counterexamples generated by \ttp{shrink} can contain more than
+$10^{10}$ elements.  This is an enormous state space to search for a smaller
+example and can cause the shrinking phase to take orders-of-magnitude more time
+than the counterexample discovery phase.
 
@@
 Typically, a user tries a brute-force approach to reduce the number of potential
-counterexamples, for example, by truncating lists using the \ttp{take n} function
+counterexamples, for example, by truncating lists using the (\ttp{take n}) function
 that returns the first $n$ elements of a list.  The trade-off is quicker
@@
 SmartCheck on the program shown in Figure~\ref{fig:initial} (we omit SmallCheck,
-since as we noted, it is infeasible for an example like this).  In the first
-three columns, the results of using QuickCheck to generate counterexamples are
-shown.  The first column contains the results of using QuickCheck without
+since as we noted, it is not even feasible for an example like this).  In the
+first three columns, the results of using QuickCheck to generate counterexamples
+are shown.  The first column contains the results of using QuickCheck without
 shrinking to provide a baseline.  The next two columns show the result of using
@@
 \paragraph{Outline.}
+The paper precedes as follows.  We define a type class and some definitions in
+Section~\ref{sec:preliminaries} used in the remainder of the paper.  In
+Section~\ref{sec:shrinking}, we describe SmartCheck's reduction algorithm for
+efficiently shrinking large counterexamples.  In
+Section~\ref{sec:generalization}, we turn our attention to the problem of
+counterexample generalization.  Here, we present two algorithms for generalizing
+counterexamples and constructing formulas for describing them.  We also discuss
+a method for property strengthening to discover new counterexamples.  We
+describe implementation-specific details in Section~\ref{sec:implementation},
+and the results of some experiments in using our tool are presented in
+Section~\ref{sec:experiments}.  Related work is given in
+Section~\ref{sec:related}.  Finally, concluding remarks and future work are
+presented in Section~\ref{sec:conclusions}.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Shrinking Data}
+\section{Preliminaries}\label{sec:preliminaries}
+
+To present the algorithms in the following sections, we provide some definitions
+here.  Our presentation simplifies the implementation
+(Section~\ref{sec:implementation}) somewhat to improve the presentation.
+Consider the following types and type class.
+%
+\begin{code}
+type Idx    = Int
+type Size   = Int
+data SubVal = forall a. SubTypes a => SubVal a
+
+class Arbitrary a => SubTypes a where
+  size    :: a -> Size
+  index   :: a -> Idx -> Maybe SubVal
+  replace :: a -> Idx -> SubVal -> a
+  constr  :: a -> String
+  constrs :: a -> [String]
+\end{code}
+%
+\noindent
+The \ttp{SubTypes} type class requires QuickCheck's \ttp{Arbitrary} as a
+super-class.  \ttp{SubTypes} has five methods.  The \ttp{size} method returns
+the size of a value, \ttp{index} returns a sub-value at an breadth-first index
+in a value, \ttp{replace} replaces a sub-value at a particular focus,
+\ttp{constr} returns the constructor tagging the value, and \ttp{constrs}
+returns a list of unique identifiers for the constructors associated with each
+variant of the type.
+
+For example, consider a binary tree type:
+%
+\begin{code}
+data Tree = L | B Tree Tree
+\end{code}
+%
+\noindent
+and the value \ttp{tree}, labeled with the indexes of its sub-values
+%
+\begin{code}
+tree = B\sub{0} (B\sub{1} L\sub{3}
+             (B\sub{4} L\sub{6} L\sub{8}))
+          (B\sub{2} L\sub{5} L\sub{7})
+\end{code}
+%
+\noindent
+Here are example applications of \ttp{SubTypes} methods; in the following, we
+show the indexes with respect to the \ttp{tree} value:
+%
+\begin{code}
+size tree = 9
+
+index tree 4  = (Just . SubVal) (B\sub{4} L\sub{6} L\sub{8})
+index tree 12 = Nothing
+
+replace tree 2 (SubVal L) =
+  B\sub{0} (B\sub{1} L\sub{3}
+        (B\sub{4} L\sub{6} L\sub{8}))
+     L
+
+constr  tree = ["B"]
+constrs tree = ["B", "L"]
+constrs L    = ["B", "L"]
+\end{code}
+%
+\noindent
+We will sometimes refer to the unwrapped value returned by \ttp{index a i} as
+the \emph{$i$th sub-value of $a$}, so for example, \ttp{B\sub{4} L\sub{6}
+  L\sub{8}} is the 4th sub-value of \ttp{tree}.  An invariant of \ttp{index} is
+that for any value \ttp{a}, consider the least $i \geq 0$ such that $$\ttp{index
+  a i} = \ttp{Nothing}$$ For all $0 \leq j < i$, $$\ttp{index a j} \neq
+\ttp{Nothing}$$ We use this invariant as a termination case in recursive
+algorithms over the sub-values of a value.
+
+The method \ttp{replace} returns the original value for an out-of-bounds index
+or an incorrectly-typed replacement value.  The \ttp{SubVal} type is an
+existential data type, used as a generic container for sub-values from a
+counterexample.
+
+Finally, while \ttp{constr} returns a string representation of the constructor
+associated with the specific variant passed to it, \ttp{constrs} ignores the
+value and returns the same list of constructors for any value from the type.
+(In our implementation, \ttp{constrs} depends on GHC Generics~\cite{generics},
+which we describe in Section~\ref{sec:implementation}.  For simplicity, we omit
+here Generics-specific super-class constraints on the \ttp{SubTypes} class.)
+
+% ------------------------------------------------------------------------------
+
+\section{Shrinking Data}\label{sec:shrinking}
 In this section, we describe how to efficiently shrink data values.  Before we
-begin, recall the basic approach implemented by the \ttp{shrink} method:
-generate a list of values, each of which is smaller than the current
-counterexample.  Each of the new values generated has nothing in common with the
-original counterexample, since each one is freshly generated by recursive
-size-constrained calls to \ttp{arbitrary}.
+begin, recall the basic idea behind the \ttp{shrink} method of the
+\ttp{Arbitrary} class: generate a list of values, each of which is smaller than
+the current counterexample.  Each of the new values generated typically bears no
+relationship to the original counterexample other than being smaller.
 
 Consider the problem of finding a small counterexample as a graph search
-problem.  The set of vertexes are all possible well-typed values (the same as
-the counterexample).  There is an edge from vertex $v$ to $v'$ if $v'$ is
-smaller (by some measure).  The goal is to fine a vertex from the set of
-smallest counterexamples in the graph.
-
-The approach taken by \ttp{shrink} is the most naive graph search algorithm:
-it simply generates a random set of potential vertexes to check in the graph.
+problem.  The set of vertexes are all possible well-typed values.  There is an
+edge from vertex $v$ to $v'$ if $v'$ is smaller than $v$ by some measure.  The
+goal is to traverse the graph, starting from a counterexample, to find a vertex
+from the set of smallest counterexamples in the graph.
+
+The approach taken by \ttp{shrink} is the most naive graph search algorithm: it
+simply generates a random set of potential vertexes to check in the graph.
 However, it does not exploit that the original counterexample \ttp{cex} is
-already in a neighborhood of likely counterexamples.  So rather than choose
-random vertexes, it may be more efficient to test other values that are closely
-related to \ttp{cex}.  By restricting the search to \ttp{cex}'s neighborhood, a
-smaller counterexample might be missed, and only a local minimum is returned.
-However, with a large state space, finding a global minimum by random sampling
-vertexes in the graph is possible but too inefficient, as we saw from the
-example in the introduction.
+already in a neighborhood of other likely counterexamples.  So rather than
+choose random vertexes, it may be more efficient to test other values that are
+closely related to \ttp{cex}.  By restricting the search to \ttp{cex}'s
+neighborhood, a smaller counterexample might be missed, and only a local minimum
+is returned.  However, with a large state space, finding a global minimum by
+random sampling vertexes in the graph is possible but too inefficient, as we saw
+from the example in the introduction.
 
@@
 constructors.  We continue until we reach a fixed-point.
-
-To present the algorithm, consider the \ttp{SubTypes} type class below for which
-QuickCheck's \ttp{Arbitrary} class is a super-class.\footnote{This presentation
-  deviates from the implementation; for example, we elide details regarding
-  error handling.}
-%
-\begin{code}
-type Idx = Int
-type Size = Int
-
-data SubVal =
-  forall a. SubTypes a => SubVal a
-
-class Arbitrary a => SubTypes a where
-  size    :: a -> Size
-  index   :: a -> Idx -> Maybe SubVal
-  replace :: a -> Idx -> SubVal -> a
-  constr  :: a -> String
-  constrs :: a -> [String]
-\end{code}
-%
-\noindent
-\ttp{SubTypes} has five methods.  The \ttp{size} method returns the size of a
-value, \ttp{index} returns a sub-value at an breadth-first index in a value,
-\ttp{replace} replaces a sub-value at a particular focus, \ttp{constr} returns
-the constructor tagging the value, and \ttp{constrs} returns a list of unique
-identifiers for the constructors associated with each variant of the type.
-
-For example, consider a binary tree type:
-%
-\begin{code}
-data Tree = L | B Tree Tree
-\end{code}
-%
-\noindent
-and the value \ttp{tree}, labeled with the indexes of its sub-values
-%
-\begin{code}
-tree = B\sub{0} (B\sub{1} L\sub{3}
-             (B\sub{4} L\sub{6} L\sub{8}))
-          (B\sub{2} L\sub{5} L\sub{7})
-\end{code}
-%
-\noindent
-Here are example applications of \ttp{SubTypes} methods; in the following, we
-show the indexes with respect to the \ttp{tree} value:
-%
-\begin{code}
-size tree = 9
-
-index tree 4 = (Just . SubVal) (B\sub{4} L\sub{6} L\sub{8})
-index tree 12 = Nothing
-
-replace tree 2 (SubVal L) =
-  B\sub{0} (B\sub{1} L\sub{3}
-        (B\sub{4} L\sub{6} L\sub{8}))
-     L
-
-constr  tree = ["B"]
-constrs tree = ["B", "L"]
-constrs L    = ["B", "L"]
-\end{code}
-%
-\noindent
-We will sometimes refer to the unwrapped value returned by \ttp{index a i} as
-the \emph{$i$th sub-value of $a$}, so for example, \ttp{B\sub{4} L\sub{6}
-  L\sub{8}} is the 4th sub-value of \ttp{tree}.  An invariant of \ttp{index} is
-that for any value \ttp{a}, consider the least $i \geq 0$ such that \ttp{index a
-  i == Nothing}.  For all $0 \leq j < i$, \ttp{index a j /= Nothing}.  We use
-this invariant as a termination case in recursive algorithms over the sub-values
-of a value.
-
-The method \ttp{replace} returns the original value for an out-of-bounds index
-or an incorrectly-typed replacement value.  The \ttp{SubVal} type is an
-existential data type, used as a generic container for sub-values from a
-counterexample.
-
-Finally, while \ttp{constr} returns a string representation of the constructor
-associated with the specific variant passed to it, \ttp{constrs} ignores the
-value and returns the same list of constructors for any value from the type.
-(In our implementation, \ttp{constrs} depends on GHC Generics~\cite{generics},
-which we describe in Section~\ref{sec:implementation}.  For simplicity, we omit
-here Generics-specific super-class constraints on the \ttp{SubTypes} class.)
 
@@
 
-\section{Counterexample Generalization} \label{sec:generalization}
+\section{Counterexample Generalization}\label{sec:generalization}
 
@@
 
-\subsection{Universal Sub-Value Generalization}
+\subsection{Universal Sub-Value Generalization}\label{sec:universal}
 An example best explains the first kind of quantified formulas we generate.
@@
     = extrapolate' (idx+1) idxs
-  extrapolate' idx idxs =
-    case index cex idx of
+    | otherwise
+    = case index cex idx of
         Nothing -> return idxs
@@
     = sumTest' (idx+1) idxs
-  sumTest' idx idxs =
-    case index cex idx of
+    | otherwise
+    = case index cex idx of
         Nothing -> return idxs
@@
     | go v == Just True
-    = constrFail' cs (tail vs')
+    = constrFail' (c:cons) (tail vs')
     | otherwise
@@
     v  = head vs'
-    c' = subConstr v
-    cs = if c' `elem` allCons
-           then cons else c':cons
-
-  go = fail prop . replace cex idx
+    c  = subConstr v
+    go = fail prop' . replace cex idx
+    prop' a = c `notElem` cons ==> prop a
   \end{code}
@@
 
-The algorithm
+The function \ttp{sumTest} takes a counterexample, a property, and a list of
+indexes that have already been generalized---perhaps by the extrapolation
+algorithm in Figure~\ref{fig:universal}.  The list of course may be empty if no
+sub-values have been previously extrapolated.  Like with the other algorithms,
+this algorithm also performs a breadth-first search over a counterexample.  In a
+call to \ttp{subTrees}, discussed in Section~\ref{sec:universal}, the guard
+prevents constructor generalization if the current index is a sub-value of a
+previously generalized value.  Otherwise, a list of well-typed new values are
+generated by a call to \ttp{newVals}, as shown in Figure~\ref{fig:reduction}.
+In the arguments to \ttp{newVals}, we bound the size of values generated with
+\ttp{scMaxSize} as before, and bound the number of values generated with the
+constant \ttp{scConstrMax}.  Because values are randomly generated, for ``wide''
+sum-types (i.e., with a large number of constructors), \ttp{scConstrMax} should
+be large enough to ensure with high probability that each variant is generated.
+
+The function \ttp{constrFail} returns true if we replace the sub-value at index
+\ttp{idx} in counterexample \ttp{cex} with every possible variant given the type
+and construct a counterexample to the property.  There are three guards to the
+recursive function \ttp{constrFail'}: the first guard holds if the list of
+constructors tagging variants in which a counterexample is found is equal in
+size to the list of all possible constructors for the type.  The second guard
+tests whether the set of test values is null; if so (and if the first guard
+fails), then we have exhausted test values before finding all possible failing
+variants.  Third, for a specific sub-value \ttp{v}, we test whether it fails the
+property.  If so, we add its constructor to the list of constructors.  Note in
+the definition of \ttp{prop'}, we add an additional precondition that the
+current constructor is not an element of constructors already seen.  Thus,
+\ttp{go v} returns \ttp{Just True} if \ttp{replace cex idx v} passes this
+precondition (and any other preconditions of \ttp{prop}), but fails the
+property.
 
@@
 
-\section{Experiments}
+\section{Experiments}\label{sec:experiments}
 
@@
 %-------------------------------------------------------------------------------
-\section{Related Work}
-\label{sec:related}
+\section{Related Work}\label{sec:related}
 
@@
 
+\todo{avoid complexities
+  \url{http://stackoverflow.com/questions/14006005/idiomatic-way-to-shrink-a-record-in-quickcheck}}
 \cite{}
@@
 %-------------------------------------------------------------------------------
-\section{Conclusions}
-\label{sec:conclusions}
+\section{Conclusions}\label{sec:conclusions}
 

### 2012-12-30 paper/Paper.hs

@@ -8,7 +8,7 @@ module Paper where
 import Prelude hiding (fail)
 import Data.Maybe (mapMaybe)
 import Control.Monad (liftM, replicateM)
-import qualified Test.QuickCheck as Q
+import Test.QuickCheck
 
 --------------------------------------------------------------------------------
 
@@ -20,7 +20,7 @@ data Format = PrintTree | PrintString
 -- SmartCheck arguments
 format       :: Format    -- ^ How to show extrapolated formula
 format       = undefined
-qcArgs       :: Q.Args    -- ^ QuickCheck arguments
+qcArgs       :: Args    -- ^ QuickCheck arguments
 qcArgs       = undefined
 qc           :: Bool      -- ^ Should we run QuickCheck?  (If not, you are
                           -- expected to pass in data to analyze.)
@@ -57,14 +57,14 @@ scConstrMax  = undefined
 
 -- Types
 
-data Property
+-- data Property
 
 data SubVal = forall a. SubTypes a => SubVal a
 
 type Size = Int
 type Idx = Int
 
-class Q.Arbitrary a => SubTypes a where
+class Arbitrary a => SubTypes a where
   size    :: a -> Size
   index   :: a -> Idx -> Maybe SubVal
   replace :: a -> Idx -> SubVal -> a
@@ -163,14 +163,14 @@ extrapolate cex prop = extrapolate' 1 []
   extrapolate' idx idxs
     | subTrees cex idx idxs
     = extrapolate' (idx+1) idxs
-  extrapolate' idx idxs =
-    case index cex idx of
-      Nothing -> return idxs
-      Just v  -> do
-        vs <- newVals v scMaxSize scMaxExtrap
-        extrapolate' (idx+1)
-          (if allFail cex idx vs prop
-             then idx:idxs else idxs)
+    | otherwise
+    = case index cex idx of
+        Nothing -> return idxs
+        Just v  -> do
+          vs <- newVals v scMaxSize scMaxExtrap
+          extrapolate' (idx+1)
+            (if allFail cex idx vs prop
+               then idx:idxs else idxs)
 
 allFail :: SubTypes a
         => a -> Idx -> [SubVal]
@@ -196,15 +196,15 @@ sumTest cex prop exIdxs = sumTest' 1 []
   sumTest' idx idxs
     | subTrees cex idx (exIdxs ++ idxs)
     = sumTest' (idx+1) idxs
-  sumTest' idx idxs =
-    case index cex idx of
-      Nothing -> return idxs
-      Just v  -> do
[…]

### 2012-12-30 paper/paper.tex
paper: more on precondition generalization.

@@
 
+\todo{make clear that preconditions shouldn't be considered logical preconditions}
+
 To present the algorithms in the following sections, we provide some definitions
@@
 
-\todo{Add discussion here}
+\todo{Add discussion here, note may be overapprox, since testing}
+
+\subsection{Automated Precondition Strengthening}
+
+The extrapolation algorithm generalizes a counterexample, but in the
+``neighborhood'' of he original counterexample.  In particular, all
+generalizations are from the same variant as the original counterexample.  To
+help the programmer in the generalization step, we would also like a way to test
+the property again, ensuring we get counterexamples (if they exist) outside of
+the neighborhood of the original one.
+
+\begin{figure}[ht!]
+  \begin{center}
+    \includegraphics[scale=0.5]{Figs/cex-gen}
+   \end{center}
+  \caption{Counterexample generalization.}
+  \label{fig:cex-gen}
+\end{figure}
+
+Figure~\ref{fig:cex-gen} illustrates the situation in which we have some values
+and a property in the case that some values fail the property of the form
+(\ttp{pre ==> prop}).  Our goal is to find additional counterexamples in the
+shaded region, in which \ttp{pre} is satisfied, \ttp{prop} is not satisfied, and
+the counterexample is not already characterized by a universal sub-value
+generalization formula.
+
+\todo{talk about matches shape algorithm, real-eval-print}
 
@@
 
-\subsection{Automated Invariant Strengthening}
-
-\begin{figure}[ht!]
-  \begin{center}
-    \includegraphics[scale=0.5]{Figs/cex-gen}
-   \end{center}
-  \caption{Counterexample generalization.}
-  \label{fig:cex-gen}
-\end{figure}
-
-\todo{check all option names}
-
 \subsection{Generalization Details}

### 2013-01-01 paper/paper.tex
paper: updates.

@@
   constrs  :: a -> [String]
+  baseType :: a -> Bool
 \end{code}
@@
 The \ttp{SubTypes} type class requires QuickCheck's \ttp{Arbitrary} as a
-super-class.  \ttp{SubTypes} has five methods.  The \ttp{size} method returns
-the size of a value, \ttp{index} returns a sub-value at an breadth-first index
-in a value, \ttp{replace} replaces a sub-value at a particular focus,
-\ttp{constr} returns the constructor tagging the value, and \ttp{constrs}
-returns a list of unique identifiers for the constructors associated with each
-variant of the type.
-
+super-class.  \ttp{SubTypes} has the following methods:
+\begin{itemize}
+\item \ttp{size} returns the size of a value,
+\item \ttp{index} returns a sub-value at an breadth-first index in a value,
+\item \ttp{replace} replaces a sub-value at a particular focus,
+\item \ttp{constr} returns the constructor tagging the value,
+\item \ttp{constrs} returns a list of unique identifiers for the constructors
+  associated with each variant of the type,
+\item \ttp{baseType} returns whether the type of the value is an ``interesting
+  type'', which generally means it has structure (e.g., a sum or product type).
+  We discuss the purpose of base types in Section~\ref{sec:base}.
+\end{itemize}
+
+\noindent
 For example, consider a binary tree type:
@@
 constrs L    = ["B", "L"]
+
+baseType (3 :: Int) = True
+baseType L          = False
 \end{code}
@@
 
-Finally, while \ttp{constr} returns a string representation of the constructor
+While \ttp{constr} returns a string representation of the constructor
 associated with the specific variant passed to it, \ttp{constrs} ignores the
@@
 here Generics-specific super-class constraints on the \ttp{SubTypes} class.)
+
+Finally, \ttp{baseType} is true for our tree type but false for a structureless
+type, such as \ttp{Int}.
 
@@
 
-\subsubsection{Base types}
+\subsubsection{Base types}\label{sec:base}
 
@@
 
-\todo{Add discussion here, note may be overapprox, since testing}
+The extrapolation algorithm is unsound insofar as it extrapolations from a set
+of counterexamples to a universal claim, similar to QuickSpec~\cite{qs}.  By
+tuning the parameters, the risk of an unsound generalization is reduced.
 
 \subsection{Automated Precondition Strengthening}
-
 The extrapolation algorithm generalizes a counterexample, but in the
@@
 the counterexample is not already characterized by a universal sub-value
-generalization formula.
+generalization formula.  Iterating, we increasingly cover the shaded region
+with new formulas until no more uncovered counterexamples exist or the testing
+tool fails to find new counterexamples.
+
+\begin{figure}
+
+  \caption{Shape matching algorithm.}
+  \label{fig:matches}
+\end{figure}
 
@@
 
+Unlike universal sub-value generalization, existential sub-value generalization
+is sound.  The existential claim is only that for each variant, there exists at
+least one counterexample, so inductive reasoning is not required.
+
 \subsection{Generalization Details}
@@
 \todo{use SC to test itself?}
+\todo{experiments? scale vs. broad data, linear data?}
 
@@
 
+\todo{Use Bayes' rule to do extrapolation?}
+
 \todo{future work: higher-kinded data, functions}

### 2013-01-01 paper/Paper.hs

@@ -65,11 +65,12 @@ type Size = Int
 type Idx = Int
 
 class Arbitrary a => SubTypes a where
-  size    :: a -> Size
-  index   :: a -> Idx -> Maybe SubVal
-  replace :: a -> Idx -> SubVal -> a
-  constr  :: a -> String
-  constrs :: a -> [String]
+  size     :: a -> Size
+  index    :: a -> Idx -> Maybe SubVal
+  replace  :: a -> Idx -> SubVal -> a
+  constr   :: a -> String
+  constrs  :: a -> [String]
+  baseType :: a -> Bool
 
 --------------------------------------------------------------------------------
 
@@ -228,3 +229,25 @@ constrFail cex idx vs prop con allCons =
     prop' a = c `notElem` cons ==> prop a
 
 --------------------------------------------------------------------------------
+
+-- matchesShapes :: SubTypes a => ScArgs -> a -> [a] -> [Idx] -> Bool
+-- matchesShapes args d ds idxs = foldl' f False ds
+--   where
+--   f True _   = True
+--   f False d' = matchesShape args d d' idxs
+
+matchesShape :: SubTypes a => a -> a -> [Idx] -> Bool
+matchesShape args a b idxs = test (subT a, subT b) && repIdxs
+  where
+  repIdxs = case foldl' f (Just b) idxs of
+              Nothing -> False
+              Just b' -> all test $ zip (nextLevel a) (nextLevel b')
+
+  f mb idx = do
+    b' <- mb
+    v  <- getAtIdx a idx (scMaxDepth args)
+    replace b' idx v
+
+  nextLevel x = map rootLabel (subTypes x)
+
+  test (SubT x, SubT y)  = baseType x || toConstr x == toConstr y

### 2013-01-06 paper/paper.tex
paper: implementation and usage.

@@
 
-\subsection{Automated Precondition Strengthening}
+\subsection{Automated Precondition Strengthening}\label{sec:precondition}
 The extrapolation algorithm generalizes a counterexample, but in the
@@
 
-\section{Implementation}\label{sec:implementation}
-
-\todo{generics}
-
-\todo{GADTs? don't work}
-\todo{add lazy testing---proof value is unused}
+\section{Implementation and Usage}\label{sec:implementation}
+
+The implementation of SmartCheck is written in Haskell and is designed to test
+Haskell programs.  The source code is licensed BSD3 and is freely
+available.\footnote{\url{https://github.com/leepike/SmartCheck.git}}
+
+SmartCheck generically operates over arbitrary algebraic data and so uses a
+generics library to encode generic traversals.  Specifically, SmartCheck uses
+``GHC Generics'', a recently-developed generics library for
+Haskell~\cite{generics}.  The generics library allows algebraic data types to
+automatically derive the type class \ttp{SubTypes} presented in
+Section~\ref{sec:preliminaries}.  One limitation with GHC Generics is that it
+does not (currently) support generalized algebraic data types~\cite{gadts}.
+
+The data type to be tested must derive the \ttp{Read}, \ttp{Show},
+\ttp{Typeable}, and \ttp{Generic} type classes.  Deriving \ttp{Typeable} and
+\ttp{Generic} require using the GHC language extensions \ttp{DeriveDataTypeable}
+and \ttp{DeriveGeneric}, respectively.  \ttp{Typeable} is used for dynamic
+typing in defining the \ttp{replace} method of \ttp{SubTypes} since it is
+unknown at compile-time whether the value to be replaced and replacing value
+have the same types.
+
+Then, the user simply declares, for a data type \ttp{D},
+%
+\begin{code}
+instance Subtypes D
+\end{code}
+%
+
+SmartCheck does not implement a counterexample discovery algorithm.  While any
+method can be used to discover an initial counterexample, SmartCheck by default
+uses QuickCheck to discover an initial counterexample.  If QuickCheck is used,
+instances for \ttp{Arbitrary} must be provided.
+
+A property may be a function taking any number of arguments, but SmartCheck only
+operates on a single input during testing, which is assumed to be the first
+argument to the function.  Our experience is that often a program operates over
+one large structure together with smaller inputs.  While multiple inputs could
+be analyzed simultaneously by SmartCheck, it is not clear how to clearly
+represent multiple large counterexamples to the user---the implementation
+currently prints to standard out.  (In future work, a web-based interface could help.)
+
+A read-eval-print loop is presented to the user, allowing her to iterate shrink
+and generalize counterexamples, and then generate new counterexamples after
+strengthening the property's precondition as described in
+Figure~\ref{sec:precondition}.
+
+SmartCheck is executed using
+%
+\begin{code}
+> smartCheck args prop
+\end{code}
+%
+\noindent
+where \ttp{args} are the arguments passed in, and \ttp{prop} is the property
+being tested.  The arguments is a record type that includes the values described
+above to control SmartCheck's behavior.
 
@@
 
+\todo{stuff from SC repo}
+
 %-------------------------------------------------------------------------------
-\section{Conclusions}\label{sec:conclusions}
-
+\section{Conclusions and Future Work}\label{sec:conclusions}
+
+\todo{how to display counterexamples}
 \todo{Use Bayes' rule to do extrapolation?}
@@
 \todo{future work: higher-kinded data, functions}
+
+\todo{add lazy testing---proof value is unused}
 

### 2013-01-14 paper/paper.tex
paper: proofreading changes.

@@
 \begin{abstract}
-QuickCheck is a powerful library for automatic test-case generation.  Because
-QuickCheck performs random testing, it can discover very large counterexamples,
-even when enumerating test-cases is infeasible, as is done with SmallCheck.
-QuickCheck provides an interface for the user to write a \emph{shrink} function
-to attempt to reduce the size of a counterexample discovered by QuickCheck.
-Implementations of \emph{shrink} often require significant boilerplate and are
-not ``smart'', just generating structurally smaller values.  We introduce
+QuickCheck, developed by Klassen and Hughes, is a powerful library for automatic
+test-case generation.  Because QuickCheck performs random testing, it can
+discover very large counterexamples.  QuickCheck provides an interface for the
+user to write \emph{shrink} functions to attempt to reduce the size of
+counterexamples discovered by QuickCheck.  Implementations of \emph{shrink}
+often require significant boilerplate and usually just generate smaller tests
+than the discovered counterexample.  Therefore, shrinking can be inefficient and
+poor at discovering new, smaller counterexamples.  In this paper, we introduce
 \emph{SmartCheck}, an automatic efficient counterexample reduction algorithm.
@@
 counterexamples by returning a formula that abstracts over individual
-counterexamples.  SmartCheck has been implemented for Haskell and tested on a
-variety of Haskell programs.
+counterexamples and make it easier for the programmer to uncover multiple
+counterexamples at once.  SmartCheck has been implemented for Haskell and is
+freely available.
 \end{abstract}
@@
 \todo{note constants are user-defined arguments, other simplifications}
+\todo{thread args through}
+\todo{termination for each algorithm}
+\todo{fromMaybe over pattern matching?}
 
@@
 
--- SmallCheck instances
-instance Serial Int16 where
-  series d = drawnFrom [(-d')..d']
-    where d' = fromIntegral d
-
-instance Serial A where series = cons1 A
-instance Serial B where series = cons4 B
-
 -- Predicates for defining properties
@@
 \ttp{B} in Figure~\ref{fig:initial}.\footnote{All examples and algorithms in
-  this paper are presented in Haskell~\cite{haskell98}.}  Indeed, because of the regular
-structure of \ttp{shrink} implementations, the principal motivation for one of
-Haskell's first generics package was for providing generic definitions for
+  this paper are presented in Haskell~\cite{haskell98}.}  For \ttp{A}, we map
+the constructor \ttp{A} over a list of new \ttp{Int16}s returned by
+\ttp{shrink}; for \ttp{B}, we use Haskell's list comprehension notation to
+return a list of new values.  Indeed, because of the regular structure of
+\ttp{shrink} implementations, the principal motivation for one of Haskell's
+first generics package was for providing generic definitions for
 \ttp{shrink}~\cite{syb}.
@@
 list in a field of \ttp{B} is less than 16, then the sum of the four fields is
-less than 64.  The property \ttp{prop} seems reasonable at first glance, until
+less than 64.  The property seems reasonable at first glance, until
 one realizes that due to underflow, the property can be violated.  For example,
@@
 potential counterexamples generated by \ttp{shrink} can contain more than
-$10^{10}$ elements.  This is an enormous state space to search for a smaller
+$10^{10}$ elements \todo{true?}.  This is an enormous state space to search for a smaller
 example and can cause the shrinking phase to take orders-of-magnitude more time
@@
 
-Typically, a user tries a brute-force approach to reduce the number of potential
-counterexamples, for example, by truncating lists using the (\ttp{take n}) function
-that returns the first $n$ elements of a list.  The trade-off is quicker
-shrinking with a lower probability of finding a smaller counterexample.  For
-example, redefining shrink for type \ttp{B} as follows
+Typically, a programmer tries a brute-force approach to reduce the number of
+potential counterexamples; for example, she may try to truncate lists using the
+(\ttp{take n}) function that returns the first $n$ elements of a list.  The
+trade-off is quicker shrinking with a lower probability of finding a smaller
+counterexample.  For example, redefining shrink for type \ttp{B} as follows
 %
@@
 \noindent
-controls the exponential blowup of the state-space by taking the first ten
-elements of each list, so the cross product is limited to 1000 elements.  The
+controls the exponential blowup of the input-space by taking the first ten
+elements of each list, so the cross product is limited to 10,000 elements.  The
 downside is that potentially smaller counterexamples may be omitted.
@@
 counterexamples altogether?''  SmallCheck is another testing framework for
-Haskell that does just this.  SmallCheck is guaranteed to return the a smallest
+Haskell that does just this.  SmallCheck is guaranteed to return a smallest
 counterexample, if one exists~\cite{sc}.  SmallCheck does this by enumerating
@@
 counterexamples to even relatively simple properties may be infeasible to
-discover due to state-space explosion.
-
-With SmallCheck, a property is checked up to a depth set by the user, measuring
-how large values are.  For example, for type \ttp{Int}, the depth $d$ is defined
-as the integers in the list enumerating from $-d$ to $d$ (i.e.,
+discover due to input-space explosion.
+
+\begin{figure}
+  \begin{code}
+instance Serial Int16 where
+  series d = drawnFrom [(-d')..d']
+    where d' = fromIntegral d
+
+instance Serial A where series = cons1 A
+instance Serial B where series = cons4 B
+  \end{code}
+  \caption{SmallCheck instances for the data types in
+    Figure~\ref{fig:initial}.}
+  \label{fig:smallshrink}
+\end{figure}
+
+With SmallCheck, a property is checked up to a depth set by the user that
+measures how large values are.  For example, for type \ttp{Int}, the depth $d$
+is defined as the integers in the list enumerating from $-d$ to $d$ (i.e.,
 \ttp{[(-d)..d]}).  For a product type, SmallCheck must check values produced by
-taking the cross-product of the type's fields at a given depth.  In
-Figure~\ref{fig:initial}, we have defined instances for generating Lazy
-SmallCheck (a variant of SmallCheck) tests by defining instances
-for the \ttp{Serial} class.  For the property \ttp{prop} defined above (suitably
-redefined to use SmallCheck's types), SmallCheck must check to depth 16 to find
-the first counterexample.  Unfortunately, after a couple hours of testing, Lazy
-SmallCheck is still checking values at depth six, and the number of tests scales
+taking the cross-product of the type's fields at a given depth.
+Figure~\ref{fig:smallshrink} contains functions for generating Lazy SmallCheck
+(a variant of SmallCheck) tests by defining instances for the \ttp{Serial}
+class.  For the property \ttp{prop} defined above (suitably redefined to use
+SmallCheck's types), SmallCheck must check to depth 16 to find the first
+counterexample.  Unfortunately, after a couple hours of testing, Lazy SmallCheck
+is still checking values at depth six, and the number of tests scales
 exponentially with respect to the depth.\footnote{All tests in this paper are
@@
   Fedora~16.  This test, and others unless noted, are performed using
-  interpreted Haskell under GHC 7.4.2.}
+  interpreted Haskell under GHC 7.4.2.\todo{redo tests?}}
 
@@
 \begin{itemize}
-  \item First, when a counterexample is discovered to some property, say with
+  \item First, after discovering a counterexample to some property, say with
     type
@@
 \noindent
-where \ttp{==>} is implication to generate new counterexamples to the same property.
+where \ttp{==>} is QuickCheck's implication operator that discards tests that
+violate the precondition.
 
@@
 \ttp{B} to 10 (Figure~\ref{lst:newshrink}) and 20 elements
-(Figure~\ref{lst:newshrink}, with \ttp{len = 20}), respectively.  (Recall that
-with the original definition for shrink from Figure~\ref{fig:initial}, some
-counterexamples cause QuickCheck to not return after hours of computation.)  In
-each row, we show the maximum, mean, and median number of seconds required by
-QuickCheck to generate a counterexample over 100 runs.  In the last row, we show
-the mean size of the final value returned, measured by the taking the cumulative
-length of the lists of the fields of \ttp{B}.  Notice that while typically
-counterexamples are returned in a fraction of a second, there are outliers
-taking 21 seconds and 125 seconds for the two tests.  Both tests roughly cut the
-size of the counterexample in half; increasing the number of potential
-counterexamples can result in a significant performance penalty for this example
-without much benefit.
+(Figure~\ref{lst:newshrink}, with \ttp{len = 20}), respectively.  (With the
+original definition for shrink from Figure~\ref{fig:initial}, some
+counterexamples cause QuickCheck to not return during the shrinking phase after
+hours of computation, so we omit these.)
+
+  In each row, we show the maximum, mean, and median
+number of seconds required by QuickCheck to generate a counterexample over 100
+runs.  In the last row, we show the mean size of the final value returned,
+measured by the taking the cumulative length of the lists of the fields of
+\ttp{B}.  Notice that while typically counterexamples are returned in a fraction
+of a second, there are outliers taking 21 seconds and 125 seconds for the two
+tests.  Both tests roughly cut the size of the counterexample in half;
+increasing the number of potential counterexamples can result in a significant
+performance penalty for this example without much benefit.
 
@@
 \todo{check paper against this section again, especially generalization sec.}
+
 \paragraph{Contributions}
@@
 \item an efficient counterexample reduction strategy for algebraic
-  data types;
-
-\item an approach to counterexample generalization, to present the
-  user with a formula describing a set of counterexamples to a property;
+  data types (Section~\ref{sec:shrinking});
+
+\item an approach to counterexample generalization, to present the user with a
+  formula describing a set of counterexamples to a property
+  (Section~\ref{sec:generalization});
 
@@
   characterizes an already-discovered counterexample for the purpose of finding
-  new, dissimilar counterexamples;
+  new, dissimilar counterexamples(Section~\ref{sec:generalization});
 
@@
   efficient that hand-written \ttp{shrink} instances in many cases but that
-  generates the instances automatically for the user.
+  generates the instances automatically for the user
+  (Section~\ref{sec:implementation}).
 
@@
 
+\noindent
+In Section~\ref{experiments}, we discuss some of our experiences with using
+SmartCheck, including checking properties from Xmonad.
+
 %% \noindent Moreover, the inability to quickly uncover understandable
@@
 
-\paragraph{Outline.}
-The paper precedes as follows.  We define a type class and some definitions in
-Section~\ref{sec:preliminaries} used in the remainder of the paper.  In
-Section~\ref{sec:shrinking}, we describe SmartCheck's reduction algorithm for
-efficiently shrinking large counterexamples.  In
-Section~\ref{sec:generalization}, we turn our attention to the problem of
-counterexample generalization.  Here, we present two algorithms for generalizing
-counterexamples and constructing formulas for describing them.  We also discuss
-a method for property strengthening to discover new counterexamples.  We
-describe implementation-specific details in Section~\ref{sec:implementation},
-and the results of some experiments in using our tool are presented in
-Section~\ref{sec:experiments}.  Related work is given in
-Section~\ref{sec:related}.  Finally, concluding remarks and future work are
-presented in Section~\ref{sec:conclusions}.
+%% \paragraph{Outline.}
+%% The paper precedes as follows.  We define a type class and some definitions in
+%% Section~\ref{sec:preliminaries} used in the remainder of the paper.  In
+%% Section~\ref{sec:shrinking}, we describe SmartCheck's reduction algorithm for
+%% efficiently shrinking large counterexamples.  In
+%% Section~\ref{sec:generalization}, we turn our attention to the problem of
+%% counterexample generalization.  Here, we present two algorithms for generalizing
+%% counterexamples and constructing formulas for describing them.  We also discuss
+%% a method for property strengthening to discover new counterexamples.  We
+%% describe implementation-specific details in Section~\ref{sec:implementation},
+%% and the results of some experiments in using our tool are presented in
+%% Section~\ref{sec:experiments}.  Related work is given in
+%% Section~\ref{sec:related}.  Finally, concluding remarks and future work are
+%% presented in Section~\ref{sec:conclusions}.
 
@@
 
-\todo{make clear that preconditions shouldn't be considered logical preconditions}
+\todo{make clear that preconditions shouldn't be considered logical
+  preconditions}
+
+We focus on \emph{algebraic data types}~\cite{}.  Algebraic data types are
+represented as ``sums of products''.  We sometimes refer to the elements of a
+sum type as a \emph{variant} that is \emph{tagged} by a constructor.  For
+example, the type
+%
+\begin{code}
+Maybe a = Just a | Nothing
+\end{code}
+%
+\noindent
+contains two variants, tagged by the constructors \ttp{Just} and \ttp{Nothing},
+respectively.
 
@@
 \item \ttp{size} returns the size of a value,
-\item \ttp{index} returns a sub-value at an breadth-first index in a value,
-\item \ttp{replace} replaces a sub-value at a particular focus,
-\item \ttp{constr} returns the constructor tagging the value,
-\item \ttp{constrs} returns a list of unique identifiers for the constructors
-  associated with each variant of the type,
+\item \ttp{index} returns a \emph{sub-value} at an breadth-first index in a
+  value,
+\item \ttp{replace} replaces a sub-value at a particular focus (returning the
+  original value if the index is out-of-bounds),
+\item \ttp{constr} returns a string representation of the constructor tagging
+  the value,
+\item \ttp{constrs} returns the list of all possible constructors from the
+  value's type,
 \item \ttp{baseType} returns whether the type of the value is an ``interesting
@@
 \noindent
-and the value \ttp{tree}, labeled with the indexes of its sub-values
+and the value \ttp{tree}, labeled with indexes in a breadth-first order:
 %
@@
 \noindent
-We will sometimes refer to the unwrapped value returned by \ttp{index a i} as
-the \emph{$i$th sub-value of $a$}, so for example, \ttp{B\sub{4} L\sub{6}
-  L\sub{8}} is the 4th sub-value of \ttp{tree}.  An invariant of \ttp{index} is
-that for any value \ttp{a}, consider the least $i \geq 0$ such that $$\ttp{index
-  a i} = \ttp{Nothing}$$ For all $0 \leq j < i$, $$\ttp{index a j} \neq
-\ttp{Nothing}$$ We use this invariant as a termination case in recursive
-algorithms over the sub-values of a value.
-
-The method \ttp{replace} returns the original value for an out-of-bounds index
-or an incorrectly-typed replacement value.  The \ttp{SubVal} type is an
-existential data type, used as a generic container for sub-values from a
-counterexample.
-
-While \ttp{constr} returns a string representation of the constructor
-associated with the specific variant passed to it, \ttp{constrs} ignores the
-value and returns the same list of constructors for any value from the type.
-(In our implementation, \ttp{constrs} depends on GHC Generics~\cite{generics},
-which we describe in Section~\ref{sec:implementation}.  For simplicity, we omit
-here Generics-specific super-class constraints on the \ttp{SubTypes} class.)
-
-Finally, \ttp{baseType} is true for our tree type but false for a structureless
-type, such as \ttp{Int}.
+The \ttp{SubVal} type is an existential data type, used as a generic container
+for sub-values from a counterexample.  We will sometimes refer to the unwrapped
+value returned by \ttp{index a i} as the \emph{$i$th sub-value of $a$}, so for
+example, (\ttp{B\sub{4} L\sub{6} L\sub{8}}) is the 4th sub-value of \ttp{tree}.
+An invariant of \ttp{index} is that for any value \ttp{a}, and for the smallest
+$i \geq 0$ such that
+%
+\begin{code}
+index a i == Nothing
+\end{code}
+%
+\noindent
+then for all $0 \leq j < i$,
+%
+\begin{code}
+index a j /= Nothing
+\end{code}
+\noindent
+We use this invariant as a termination case in recursive algorithms over the
+sub-values of a value.
+
+In our implementation, \ttp{constr} and \ttp{constrs} depends on GHC
+Generics~\cite{generics}, which we describe in Section~\ref{sec:implementation}.
+For simplicity, we omit here Generics-specific super-class constraints on the
+\ttp{SubTypes} class.
+
+%% Finally, \ttp{baseType} is true for our tree type but false for a
+%% structureless type, such as \ttp{Int}.
 
@@
 neighborhood, a smaller counterexample might be missed, and only a local minimum
-is returned.  However, with a large state space, finding a global minimum by
+is returned.  However, with a large set of values, finding a global minimum by
 random sampling vertexes in the graph is possible but too inefficient, as we saw
@@
 sub-values and build a new value from that, leaving the remainder of the tree
-unchanged.  Intuitively, by a structurally smaller value, we mean one with fewer
-constructors.  We continue until we reach a fixed-point.
+unchanged.  Intuitively, by a \emph{structurally smaller value}, we mean one
+with fewer constructors.  We continue until we reach a fixed-point.
 
@@
 a counterexample that the algorithm automatically shrinks.  The reduction begins
-at the first sub-value of \ttp{cex}, \ttp{v}.  When the index \ttp{idx} becomes
-out-of-bounds and returns \ttp{Nothing}, the algorithm terminates.  Otherwise, a
-list of new random values are generated.
+at the first proper sub-value of \ttp{cex}; call it \ttp{v}.  When the index
+\ttp{idx} becomes out-of-bounds and returns \ttp{Nothing}, the algorithm
+terminates.  Otherwise, a list of new random values are generated.
 %
@@
 \noindent
-generates a new value \ttp{v'} having the same type as \ttp{v} and strictly
-smaller (with respect to the \ttp{size} method) than \ttp{v}.  Just like
-QuickCheck's \ttp{arbitrary} method, \ttp{sizedArbitrary} generates successfully
-larger counterexamples when generating new values with which to replace a
-sub-value.
+generates a new value \ttp{v'} having the same type as \ttp{v} and that is
+strictly smaller (with respect to the \ttp{size} method) than \ttp{v}.  Just
+like QuickCheck's \ttp{arbitrary} method, \ttp{sizedArbitrary} generates
+successfully larger counterexamples when generating new values with which to
+replace a sub-value.
 
@@
 on a random seed for generation (we could pass in the generator seed, but for
-simplicity, we assume one is generated locally).  The constant \ttp{scMaxReduce} is
-the maximum number of tries to discover a new counterexample by replacing
-\ttp{v} in \ttp{cex} and testing it.  The result of \ttp{pass prop cex'} for
+simplicity, let us assume one is generated locally).  The constant
+\ttp{scMaxReduce} is the maximum number of tries to discover a new
+counterexample by replacing \ttp{v} in \ttp{cex} and testing it.  The result of
+\ttp{pass prop cex'} for
 %
@@
 \noindent
-holds if \ttp{cex'} satisfies the property \ttp{prop}.  If no failure is found,
-we move to the next sub-value of \ttp{cex} and continue.  However, if a new
-smaller counterexample \ttp{cex'} is found, we start a new breadth-first
-traversal of \ttp{cex'}, attempting to shrink it further.
-
-\subsection{Reduction Algorithm Details and Optimizations}
-The reduction algorithm description above omits some minor details and
-optimizations we describe here.
-
-\subsubsection{Sum-type counterexample hypothesis}
-First, recall that the algorithm begins at the \emph{first} sub-value of the
-counterexample rather than the 0th sub-value.  The 0th sub-value of \ttp{cex} is
-\ttp{cex} itself.  No invariant of the algorithm would be violated by beginning
-with the 0th sub-value; recall that it would generate strictly smaller values
-than \ttp{cex}.  However, we begin with a strict sub-value because of a
-hypothesis we refer to as the \emph{sum-type counterexample hypothesis}.  The
-hypothesis is that in the search space of possible values, if a counterexample
-of the form \ttp{C}~$\vec{x}$, where \ttp{C} is some constructor and $\vec{x}$
-are its arguments, is discovered, then it is more likely to find another
-``nearby'' counterexample of the form \ttp{C}~$\vec{y}$, where $\vec{y}$ are
-different arguments to \ttp{C}.  For example, consider a property about
-unintended variable capture over a language's parse tree represented by a sum
-type with constructors for module imports, function definitions, and
-global-variable assignments, respectively.  A function definition counterexample
-can only be reduced to smaller function definition counterexamples, the only
-construct in which variable capture is possible.
-
-Another way to motivate the hypothesis is that searching the space of possible
-counterexamples by generating \emph{any} well-typed value that is strictly
-smaller than the original counterexample is something we have seen before: it is
-precisely the naive approach to implementing \ttp{shrink}!
-
-\subsubsection{Conditional properties}
-A QuickCheck property may be a conditional of the form
-%
-\begin{code}
-prop a = pre a ==> prop' a
-\end{code}
-%
-\noindent
-such that \ttp{prop} is true if the precondition \ttp{pre a} is false or
-\ttp{prop' a} holds.  In practice, conditionals are used to restrict the set of
-values tested; values that fail the precondition are discarded.  Thus, the call
-to \ttp{pass} in the reduction algorithm only considers a potential
-counterexample \ttp{cex} to have failed if it passes the property's
-precondition and fails the consequent.
-
-\subsubsection{Sub-value counterexample hypothesis}\label{sec:subval}
+holds if \ttp{cex'} satisfies the property \ttp{prop}.  The property may be a
+conditional, in which case the value must pass the precondition as well as the
+consequent for \ttp{pass} to return \ttp{True}.  If no failure is found, we move
+to the next sub-value of \ttp{cex} and continue.  However, if a new smaller
+counterexample \ttp{cex'} is found, we start a new breadth-first traversal of
+\ttp{cex'}, attempting to shrink it further.  The algorithm is guaranteed to
+terminate: intuitively, the measure for the function is that either the index
+increases or the size of the counterexample being evaluated decreases.
+
+\subsection{Reduction Algorithm Details}
+Having describe the reduction algorithm, there are two important details about
+its design we describe below.
+
+\subsubsection{Variant Counterexample Hypothesis}
+A motivation for the design of the reduction algorithm is something we cal the
+\emph{variant counterexample hypothesis}.  Recall that the algorithm begins at
+the \emph{first} sub-value of the counterexample rather than the 0th sub-value.
+The 0th sub-value of \ttp{cex} is \ttp{cex} itself.  No invariant of the
+algorithm would be violated by beginning with the 0th sub-value, and in
+particular, the algorithm would still terminate.  However, we begin with a
+strict sub-value because of the hypothesis.  The hypothesis is that in the
+search space of possible values from a given type \ttp{T}, starting with a
+particular counterexample \ttp{cex}, we are more likely to find other
+counterexamples from the same variant as \ttp{cex}.  For example, consider a
+property about unintended variable capture over a language's parse tree
+represented by a sum type with constructors for module imports, function
+definitions, and global-variable assignments, respectively.  A function
+definition counterexample can only be reduced to smaller function definition
+counterexamples, the only construct in which variable capture is possible.
+
+%% Another way to motivate the hypothesis is that searching the space of
+%% possible counterexamples by generating \emph{any} well-typed value that is
+%% strictly smaller than the original counterexample is something we have seen
+%% before: it is precisely the naive approach to implementing \ttp{shrink}!
+
+%% \subsubsection{Conditional Properties}
+%% A QuickCheck property may be a conditional of the form
+%% %
+%% \begin{code}
+%% prop a = pre a ==> prop' a
+%% \end{code}
+%% %
+%% \noindent
+%% such that \ttp{prop} is true if the precondition (\ttp{pre a}) is false or
+%% (\ttp{prop' a}) holds.  In practice, conditionals are used to restrict the set
+%% of values tested; values that fail the precondition are discarded.  Thus, the
+%% call to \ttp{pass} in the reduction algorithm only considers a potential
+%% counterexample \ttp{cex} to have failed if it passes the property's precondition
+%% and fails the consequent.
+
+\subsubsection{Base Types}\label{sec:base}
+
+The number of constructors in an algebraic data value primarily determines how
+large and understandable it is.  In our experience, the size of fields that come
+from primitive types, like \ttp{Int} or \ttp{Char}, have less of a role in
+understanding a counterexample.  However, a testing algorithm might spend
+significant time testing new values in which primitive types are reduces.
+
+Consequently, we distinguish a class of types referred to as \emph{base types}.
+If the reduction algorithm encounters a base type, it is ignored.  By default,
+this includes numeric types (e.g., \ttp{Int}, \ttp{Word}, \ttp{Float}, etc.,
+\ttp{Bool}, and \ttp{Char}.  We also treat recursive container types like lists
+specially: lists of base types are considered to be base types themselves.  This
+prevents the algorithm from analyzing, for example, strings.  A list of a
+non-base-types is shrunk by the algorithm.
+
+By providing custom instances, the user can declare any substructure in a data
+type to be a base type.  Doing so effectively treats values from that type as
+``black boxes'' for SmartCheck.  Doing so can help make SmartCheck more
+efficient if the user knows that some portion of the structure cannot be shrunk
+or is irrelevant to the property.
+
+\subsection{Reduction Algorithm Optimizations}
+The reduction algorithm description above omits some details and optimizations
+we describe here.
+
+\subsubsection{Sub-value Counterexample Hypothesis}\label{sec:subval}
 Sometimes, a counterexample fails a property due to a sub-value nested deep
@@
 
-For example, consider a simple calculator language containing constant
-\ttp{Ints}, addition, and division:
+For example, consider a simple calculator language containing constants,
+addition, and division:
 %
@@
 \noindent
-and consider an interpreter for the language that returns \ttp{Nothing} if
-division-by-zero is encountered and \ttp{Just} the value otherwise:
-%
-\begin{code}
-eval :: Exp -> Maybe Int
-eval (C i) = Just i
-eval (Add e0 e1) =
-  liftM2 (+) (eval e0) (eval e1)
-eval (Div e0 e1) =
-  let e = eval e1 in
-  if e == Just 0 then Nothing
-    else liftM2 div (eval e0) e
-\end{code}
-%
-\noindent
-Suppose we restrict the set of programs with the following predicate:
-%
-\begin{code}
-divSubTerms :: Exp -> Bool
-divSubTerms (C _)         = True
-divSubTerms (Div _ (C 0)) = False
-divSubTerms (Add e0 e1)   =    divSubTerms e0
-                            && divSubTerms e1
-divSubTerms (Div e0 e1)   =    divSubTerms e0
-                            && divSubTerms e1
-\end{code}
-%
-and then test the property
-%
-\begin{code}
-prop_div e = divSubTerms e ==> eval e /= Nothing
-\end{code}
-%
-\noindent
-The property \ttp{prop\_div} turns out to be false; for example,
-%
-\begin{code}
-Div (C 1) (Div (C 0) (C 1))
-\end{code}
-%
-\noindent
-is a counterexample.  In the predicate case
-%
-\begin{code}
-divSubTerms (Div _ (C 0)) = False
-\end{code}
-%
-\noindent
-we forgot to recur on the divisor.
-
-\begin{figure}[ht!]
-\ttp{Div (Div (C (-22)) (Div (C 2) (C 13))) (Div (Div (Div (C (-11)) (C 14))
-  (Add (Div (C (-4)) (C (-3))) (Div (C (-25)) (C (-25))))) (Add (Add (C 22) (Div
-  (C (-30)) (C (-14)))) (C 13)))}
-  \caption{Typical counterexample from QuickCheck with no user-defined
-    \ttp{shrink} function.}
-  \label{fig:div-cex}
-\end{figure}
-
-However, a typical counterexample returned by QuickCheck, without \ttp{shrink}
-implemented, is shown in Figure~\ref{fig:div-cex}.  To butcher an expression,
-finding the cause of failure is like searching for a divide-by-zero in the
-haystack!  But it turns out that the culprit is a small sub-value:
-%
-\begin{code}
-Div (C (-22)) (Div (C 2) (C 13))
-\end{code}
-%
+Consider a program that takes an \ttp{Exp} as an argument that throws an error
+if an evaluation of an \ttp{Exp} value contains a division-by-zero.  Testing for
+the occurrence of the error, we might have a counterexample like the following:
+%
+\medskip%
+\begin{sloppypar}
+\noindent%
+\ttp{Add (Div (C 5) (C (-12))) (Add (Add (C 2) (C 4)) (Add (C 7) (Div (C 3) (Add
+(C (-5)) (C 5)))))}
+\end{sloppypar}
+\medskip%
+In the counterexample, the culprit is a deeply-buried sub-value,
+%
+\begin{code}
+Div (C 3) (Add (C (-5)) (C 5))
+\end{code}
+%
+Thus, when attempting to shrink an \ttp{Exp} value, it pays to test whether a
+sub-value itself fails the property.
+
+%% and consider an interpreter for the language that returns \ttp{Nothing} if
+%% division-by-zero is encountered and (\ttp{Just x}), where \ttp{x} is the result
+%% of the computation, otherwise:
+%% %
+%% \begin{code}
+%% eval :: Exp -> Maybe Int
+%% eval (C i) = Just i
+%% eval (Add e0 e1) =
+%%   liftM2 (+) (eval e0) (eval e1)
+%% eval (Div e0 e1) =
+%%   let e = eval e1 in
+%%   if e == Just 0 then Nothing
+%%     else liftM2 div (eval e0) e
+%% \end{code}
+%% %
+%% \noindent
+%% Suppose we restrict the set of programs with the following predicate:
+%% %
+%% \begin{code}
+%% divSubTerms :: Exp -> Bool
+%% divSubTerms (C _)         = True
+%% divSubTerms (Div _ (C 0)) = False
+%% divSubTerms (Add e0 e1)   =    divSubTerms e0
+%%                             && divSubTerms e1
+%% divSubTerms (Div e0 e1)   =    divSubTerms e0
+%%                             && divSubTerms e1
+%% \end{code}
+%% %
+%% and then test the property
+%% %
+%% \begin{code}
+%% prop_div e = divSubTerms e ==> eval e /= Nothing
+%% \end{code}
+%% %
+%% \noindent
+%% The property \ttp{prop\_div} turns out to be false; for example,
+%% %
+%% \begin{code}
+%% Div (C 1) (Div (C 0) (C 1))
+%% \end{code}
+%% %
+%% \noindent
+%% is a counterexample.  In the predicate case
+%% %
+%% \begin{code}
+%% divSubTerms (Div _ (C 0)) = False
+%% \end{code}
+%% %
+%% \noindent
+%% we forgot to recur on the divisor.
+
+%% %% \begin{figure}[ht!]
+%% %%   \caption{Typical counterexample from QuickCheck with no user-defined
+%% %%     \ttp{shrink} function.}
+%% %%   \label{fig:div-cex}
+%% %% \end{figure}
+
+%% A typical counterexample returned by QuickCheck, with a \ttp{shrink} function
+%% implemented, is as follows:%
+%% %
+%% \medskip
+
+%% \begin{sloppypar}
+%% %\makebox[\textwidth]{
+%% \noindent%
+%% \ttp{Div (Add (Div (Div (C (-12)) (C 22)) (Add (C 10) (C (-5)))) (Div (Add (C
+%%   30) (C 12)) (C (-18)))) (Div (Div (C (-20)) (C 32)) (Div (C 10) (C (-8))))}
+%% \end{sloppypar}
+%% \medskip%
+
+%% \noindent
+%% It turns out that the culprit is the sub-value
+%% %
+%% \begin{code}
+%% Div (Div (C (-20)) (C 32)) (Div (C 10) (C (-8)))
+%% \end{code}
+%% %
+%% \noindent
+%% Thus, while analyzing the counterexample above
 
@@
 
-Generalizing this scenario, during the reduction algorithm's breadth-first
+Generalizing the scenario, during the reduction algorithm's breadth-first
 search through a counterexample \ttp{cex}'s sub-values, we may happen upon a
@@
 
-\subsubsection{Base types}\label{sec:base}
-
-The number of constructors in an algebraic data value primarily determines how
-large and understandable it is.  In our experience, the size of fields that come
-from primitive types, like \ttp{Int} or \ttp{Char}, have less of a role in
-understanding a counterexample.
-
-Consequently, we distinguish a class of types referred to as \emph{base types}.
-If the reduction algorithm encounters a base type, it is ignored.  By default,
-this includes numeric types (e.g., \ttp{Int}, \ttp{Word}, \ttp{Float}, etc.,
-\ttp{Bool}, and \ttp{Char}.  We also treat lists specially: lists of base types
-are considered to be base types themselves.  This prevents the algorithm from
-analyzing, for example, strings.  A list of a non-base-types is shrunk by the
-algorithm.
-
-By providing custom instances, the user can declare any substructure in a data
-type to be a base type.  Doing so can help make SmartCheck more efficient if the
-user knows that some portion of the structure cannot be shrunk or is irrelevant
-to the property.
-
-\subsubsection{Bounding counterexample exploration}
+\subsubsection{Bounding Counterexample Exploration}
 
@@
 
+\todo{default instances for all common types, including those provided by QC}
+
 \section{Experiments}\label{sec:experiments}

### 2013-01-15 paper/paper.tex
Paper updates.

@@
 \noindent%
-\ttp{Add (Div (C 5) (C (-12))) (Add (Add (C 2) (C 4)) (Add (C 7) (Div (C 3) (Add
-(C (-5)) (C 5)))))}
+\ttp{Add (Div (C 5) (C (-12))) (Add (Add (C 2) (C 4)) (Add (C 7) (Div (Add (C 7)
+  (C 3)) (Add (C (-5)) (C 5)))))}
 \end{sloppypar}
 \medskip%
-In the counterexample, the culprit is a deeply-buried sub-value,
-%
-\begin{code}
-Div (C 3) (Add (C (-5)) (C 5))
+In the counterexample, the culprit is a buried sub-value,
+%
+\begin{code}
+Div (Add (C 7) (C 3)) (Add (C (-5)) (C 5))
 \end{code}
@@
 Figure~\ref{fig:reduction}.  For example, the depth of a list is determined by
-its length and the depth of the binary tree defined in Section~\ref{sec:reduct}
+its length, and the depth of the binary tree defined in Section~\ref{sec:reduct}
 is determined by the function \ttp{depth}:
@@
 to provide her not with a single counterexample but with a quantified formula
-that characterizes a set of counterexamples.  Doing so has two benefits.  First,
-it already generalizes a counterexample, perhaps helping the programmer's
-insight in the generalization step.  Second, it reduces the noise in favor of
-the signal by abstracting away portions of large counterexample that are
-irrelevant to why it violates the property.
-
-The characterization of counterexamples that most helps the programmer seems to
-should strike a middle ground.  At the one extreme, a single counterexample is
-can be thought of as a closed formula containing only literals that
-under-approximates the set of counterexamples.  At the other extreme, the
-property itself being tested is a formula that over-approximates the failing
+that characterizes a set of counterexamples.  In particular, a formula reduces
+the noise in favor of the signal by abstracting away portions of large
+counterexample that are irrelevant to why it violates the property.
+
+The characterization of counterexamples that most helps the programmer should
+strike a middle ground.  A single counterexample is too specific.  On the other
+hand, the property itself is a formula that over-approximates the failing
 inputs.  We describe two kinds of formula that fall between these two extremes.
@@
 An example best explains the first kind of quantified formulas we generate.
-Compare the following formula to the counterexample in Figure~\ref{fig:div-cex}:
-%
-\begin{code}
-\ttp{forall x, y . Div (Div x (Div (C 2) (C 13))) y}
-\end{code}
-%
-\noindent
-Here, the formula claims that for all well-typed values \ttp{x} and \ttp{y}, the
-property fails.  Not only does the generalization assist the programmer's
-insight, but it reduces the sheer size of the counterexample, and it can do so
-significantly.  We call the kind of formula just shown \emph{universal sub-value
-  generalization} and is implemented in SmartCheck.
+Recall the calculator language from Section~\ref{sec:subval}.  Recall the
+culprit buried within a counterexample, causing it to result in a
+division-by-zero error:
+%
+\begin{code}
+Div (Add (C 7) (C 3)) (Add (C (-5)) (C 5))
+\end{code}
+%
+\noindent
+Notice that any value substituted in the dividend results in an error.  We can
+quantify the error then as follows:
+%
+\begin{code}
+forall x . Div x (Add (C (-5)) (C 5))
+\end{code}
+%
+\noindent
+Not only does the generalization assist the programmer's insight, but it reduces
+the sheer size of the counterexample, and it can do so significantly.  We call
+the kind of formula just shown \emph{universal sub-value generalization} and is
+implemented in SmartCheck.
 
@@
 them in \ttp{cex} to create a list of new potential counterexamples.  If enough
-new counterexamples non-trivially fails \ttp{prop} (i.e., does not just fail the
-precondition), and non satisfy \ttp{prop}, then we extrapolate, claiming that
-for \emph{any} new value replacing that sub-value in \ttp{cex}, the property
-will fail.
+new counterexamples non-trivially fail \ttp{prop} (i.e., fail the property's
+consequent but satisfy its antecendent), then we extrapolate, claiming that for
+\emph{any} new value replacing that sub-value in \ttp{cex}, the property will
+fail.
 
@@
 original counterexample.  In the recursive function \ttp{extrapolate'}, there is
-a function guard in which a call to \ttp{subTree}
+a function guard in which a call
+%
+\begin{code}
+subTree cex idx0 idx1
+\end{code}
+%
+\noindent
+where \ttp{subTree} has the type
 %
@@
 %
-\ttp{subTree cex idx0 idx1} is true if in \ttp{cex}, the value at index
-\ttp{idx0} is a child of index \ttp{idx1}, in a tree representation of the
-data.  This prevents the algorithm from trying to generalize sub-values that are
-abstracted away already since their parents have been generalized.
+\noindent
+is true if in \ttp{cex}, the value at index \ttp{idx0}, is a child of index
+\ttp{idx1}, in a tree representation of the data.  This prevents the algorithm
+from trying to generalize sub-values that are abstracted away already since
+their parents have been generalized.
 
@@
 \noindent
-is the roughly the dual of \ttp{pass} in the reduction algorithm: \ttp{fail prop
-  cex} returns \ttp{Just True} if \ttp{cex} passes \ttp{prop}'s precondition but
-fails the property; \ttp{Just False} if \ttp{cex} non-trivially satisfies
-\ttp{prop}, and \ttp{Nothing} if \ttp{cex} fails \ttp{prop}'s precondition.  The
-constant \ttp{scMinExtrap} is the minimum number of \ttp{Just False} results
-required from \ttp{fail} to extrapolate from failed tests to a universal claim.
+is the roughly the dual of \ttp{pass} in the reduction algorithm: (\ttp{fail
+  prop cex}) returns (\ttp{Just True}) if \ttp{cex} passes \ttp{prop}'s
+precondition but fails the property; (\ttp{Just False}) if \ttp{cex}
+non-trivially satisfies \ttp{prop}, and \ttp{Nothing} if \ttp{cex} fails
+\ttp{prop}'s precondition.  The constant \ttp{scMinExtrap} is the minimum number
+of \ttp{Just False} results required from \ttp{fail} to extrapolate from failed
+tests to a universal claim.
 
@@
 property \ttp{prop\_div} fails only for values that contain a variant tagged
-with the \ttp{Div} constructor.  For example, consider the following counterexample:
-%
-\begin{code}
-Div (Add (C 20) (C 30))
-    (Add (C 89) (C -89))
-\end{code}
-%
-\noindent
-The counterexample fails because the divsor evaluates to zero.  Using universal
-sub-value generalization described above, the counterexample is generalized as
-follows:
-%
-\begin{code}
-forall x . Div x (Add (C 89) (C -89))
-\end{code}
-%
-\noindent
-So we know there is something particular about the \ttp{Add (C 89) (C -89)},
-because there are some values that we can replace this sub-value with in the
-original counterexample causing it to satisfy the property.  But we might wonder
-if there is something special about variants tagged by the \ttp{Add}
-constructor, or might we finding failing sub-values with the other variants
-tagged by the constructors \ttp{C} and \ttp{Div}?  If so, then we know there is
-not some bug being triggered by \ttp{Add} variants specifically, so it is more
-likely that the bug involves how the second field of the \ttp{Div} variant is
-evaluated.
+with the \ttp{Div} constructor.  Recall again the generalized counterexample
+from Section~\ref{sec:generalization}:
+%
+\begin{code}
+forall x . Div x (Add (C (-5)) (C 5))
+\end{code}
+%
+\noindent
+Because the divisor does not generalize, we know there is something special
+about it that causes failure.  But we might wonder if there is something special
+about variants tagged by the \ttp{Add} constructor, or might we finding failing
+sub-values with the other variants.
 
@@
 \begin{code}
-forall x . Div x (Add (C 89) (C -89))
+forall x . Div x (Add (C (-5)) (C 5))
 forall x . Div x (C 0)
@@
 Because there are only three variants in the type, we have found counterexamples
-built from each of them in the second field of the outermost \ttp{Div}
-constructor.  We therefore return a formula like the following:
+built from each of them in the divisor.  We therefore return a formula like the
+following:
 %

### 2013-01-27 examples/Div0.hs
Div0: helper functions.

@@ -38,9 +38,9 @@ instance Arbitrary Exp where
                   , liftM2 Div mkM' mkM' ]
       where mkM' = mkM =<< choose (0,n-1)
 
-  -- shrink (C i)       = map C (shrink i)
-  -- shrink (Add e0 e1) = [e0, e1]
-  -- shrink (Div e0 e1) = [e0, e1]
+  shrink (C i)       = map C (shrink i)
+  shrink (Add e0 e1) = [e0, e1]
+  shrink (Div e0 e1) = [e0, e1]
 
 -- property: so long as 0 isn't in the divisor, we won't try to divide by 0.
 -- It's false: something might evaluate to 0 still.
@@ -66,7 +66,6 @@ divSubTerms (Div e0 e1)   = divSubTerms e0 && divSubTerms e1
 --                 Nothing -> True
 --                 Just i -> i < 5
 
-
 divTest :: IO ()
 divTest = smartCheck args prop_div
   where
@@ -74,6 +73,23 @@ divTest = smartCheck args prop_div
                                 -- { maxSuccess = 1000
                                 -- , maxSize    = 20  }
                    , format  = PrintString
+                   , extrap  = True
                    }
 
----------------------------------------------------------------------------------
+-- Get the minimal offending sub-value.
+findVal :: Exp -> (Exp,Exp)
+findVal (Div e0 e1)
+  | eval e1 == Just 0     = (e0,e1)
+  | eval e1 == Nothing    = findVal e1
+  | otherwise             = findVal e0
+findVal a@(Add e0 e1)
+  | eval e0 == Nothing    = findVal e0
+  | eval e1 == Nothing    = findVal e1
+  | eval a == Just 0      = (a,a)
+findVal _                 = error "not possible"
+
+divSubValue :: Exp
+divSubValue =
+  Add (Div (C 5) (C (-12))) (Add (Add (C 2) (C 4)) (Add (C 7) (Div (C 3) (Add (C (-5)) (C 5)))))
+
+--------------------------------------------------------------------------------

### 2013-01-27 examples/LambdaCalc.hs
examples: wibbles, cleanups.

@@ -6,7 +6,6 @@
 module LambdaCalc where
 
 import Data.List
-import Data.Tree
 import Data.Typeable
 
 import Control.Monad
@@ -25,7 +24,7 @@ data Expr
         deriving (Eq, Read, Show, Typeable, Generic)
 
 freeVars :: Expr -> [Sym]
-freeVars (Var s) = [s]
+freeVars (Var v) = [v]
 freeVars (App f a) = freeVars f `union` freeVars a
 freeVars (Lam i e) = freeVars e \\ [i]
 
@@ -44,37 +43,41 @@ subst v x b = sub b
                 Lam i (sub e)
         fvx = freeVars x
         cloneSym e i = loop i
-           where loop i' = if i' `elem` vars then loop (i ++ "'") else i'
-                 vars = fvx ++ freeVars e
+           where loop i' = if i' `elem` vs then loop (i ++ "'") else i'
+                 vs = fvx ++ freeVars e
 
 substVar :: Sym -> Sym -> Expr -> Expr
-substVar s s' e = subst s (Var s') e
+substVar v v' e = subst v (Var v') e
 
 alphaEq :: Expr -> Expr -> Bool
 alphaEq (Var v)   (Var v')    = v == v'
 alphaEq (App f a) (App f' a') = alphaEq f f' && alphaEq a a'
-alphaEq (Lam s e) (Lam s' e') = alphaEq e (substVar s' s e')
+alphaEq (Lam v e) (Lam v' e') = alphaEq e (substVar v' v e')
 alphaEq _ _ = False
 
 nf :: Expr -> Expr
 nf ee = spine ee []
   where spine (App f a) as = spine f (a:as)
-        spine (Lam s e) [] = Lam s (nf e)
-        spine (Lam s e) (a:as) = spine (subst s a e) as
+        spine (Lam v e) [] = Lam v (nf e)
+        spine (Lam v e) (a:as) = spine (subst v a e) as
         spine f as = app f as
         app f as = foldl App f (map nf as)
 
 betaEq :: Expr -> Expr -> Bool
 betaEq e1 e2 = alphaEq (nf e1) (nf e2)
 
+z,s,m,n :: Expr
 [z,s,m,n] = map (Var . (:[])) "zsmn"
+app2 :: Expr -> Expr -> Expr -> Expr
 app2 f x y = App (App f x) y
+zero, one, two, three, plus :: Expr
 zero  = Lam "s" $ Lam "z" z
[…]

### 2013-01-27 examples/MutualRecData.hs

@@ -8,15 +8,14 @@ import Test.SmartCheck
 import Test.QuickCheck hiding (Result)
 
 import Data.Tree
-import Control.Monad.State 
+import Control.Monad.State
 import Data.Typeable
 
 import GHC.Generics
 
----------------------------------------------------------------------------------
+--------------------------------------------------------------------------------
 
-data M = M N N Int
-       | P
+data M = M N N Int | P
   deriving (Typeable, Show, Eq, Read, Generic)
 
 instance SubTypes M
@@ -26,41 +25,36 @@ data N = N M Int String
 
 instance SubTypes N
 
----------------------------------------------------------------------------------
+--------------------------------------------------------------------------------
 
 instance Arbitrary M where
-  arbitrary = 
-    sized $ \n -> if n == 0 then return P 
+  arbitrary =
+    sized $ \n -> if n == 0 then return P
                     else oneof [ return P
-                               , liftM3 M (resize (n-1) arbitrary) 
-                                          (resize (n-1) arbitrary) 
-                                          arbitrary 
+                               , liftM3 M (resize (n-1) arbitrary)
+                                          (resize (n-1) arbitrary)
+                                          arbitrary
                                ]
 
 instance Arbitrary N where
   arbitrary = liftM3 N arbitrary arbitrary arbitrary
 
----------------------------------------------------------------------------------
+--------------------------------------------------------------------------------
 
 prop0 :: M -> Bool
 prop0 (M _ _ a) = a < 100
 prop0 _         = True
 
 mutRecTest :: IO ()
-mutRecTest = smartCheck args p
-  where 
-  p    = \a -> property (prop0 a)
+mutRecTest = smartCheck args (property . prop0)
+  where
   args = scStdArgs { qcArgs = stdArgs {maxSuccess = 1000} }
 
----------------------------------------------------------------------------------
-
-testProp :: M -> Bool
-testProp (M _ _ i) = i > 100
-testProp _         = True
+--------------------------------------------------------------------------------
 
 xx :: M
-xx = M (N (M (N P 1 "goo") (N P 7 "foo") 8) 3 "hi") (N P 4 "bye") 6    
+xx = M (N (M (N P 1 "goo") (N P 7 "foo") 8) 3 "hi") (N P 4 "bye") 6
 yy :: Forest Int
 yy = [Node 0 [Node 1 [], Node 2 []], Node 3 [Node 4 [], Node 5 [Node 6 []]]]
 
----------------------------------------------------------------------------------
+--------------------------------------------------------------------------------

### 2013-01-27 examples/Tests.hs

@@ -1,4 +1,4 @@
-module Tests where
+module Main where
 
 import Div0
 import MutualRecData

### 2013-01-27 paper/paper.tex
paper updates.

@@
 \newboolean{submission}  %set to true for the submission version
-%\setboolean{submission}{false}
-\setboolean{submission}{true}
+\setboolean{submission}{false}
+%\setboolean{submission}{true}
 \ifthenelse
@@
 \todo{fromMaybe over pattern matching?}
+\todo{show quantification in intro}
+\todo{benchmark feat}
 
@@
 list in a field of \ttp{B} is less than 16, then the sum of the four fields is
-less than 64.  The property seems reasonable at first glance, until
-one realizes that due to underflow, the property can be violated.  For example,
-consider the value
+less than 64.  The property seems reasonable at first glance, until one realizes
+that due to underflow, the property can be violated.  For example, consider the
+value
 %
@@
 \noindent
-In Section~\ref{experiments}, we discuss some of our experiences with using
+In Section~\ref{sec:experiments}, we discuss some of our experiences with using
 SmartCheck, including checking properties from Xmonad.
@@
 
+\todo{correspond to dimension 0 from\cite{agata}}
 The number of constructors in an algebraic data value primarily determines how
@@
 \begin{code}
-forall x . Div x (Add (C (-5)) (C 5))
-forall x . Div x (C 0)
-forall x . Div x (Div (C 0) (C 1))
+Div (C 7) (Add (C (-5)) (C 5))
+Div (Add (C 3) (C 2)) (C 0)
+Div (C (-2)) (Div (C 0) (C 1))
 \end{code}
@@
 
-\todo{avoid complexities
-  \url{http://stackoverflow.com/questions/14006005/idiomatic-way-to-shrink-a-record-in-quickcheck}}
-\cite{}
+Closely related to SmartCheck is the recent Haskell library Feat.  Feat
+enumerates algebraic data types in Haskell and indexing them
+quickly~\cite{feat}.  For example, Feat returns the $10^{1000}$ value in the
+enumeration of lists of Booleans in under a second.  Feat combines some of the
+advantages of SmallCheck and QuickCheck:
+
+Feat provides a functional representation of enumerations,
+allowing for fast indexing of large values in an enumeration.
+
+\todo{avoid complexities \url{http://stackoverflow.com/questions/14006005/idiomatic-way-to-shrink-a-record-in-quickcheck}%
+%
+The SYB paper:%
+  \url{http://research.microsoft.com/en-us/um/people/simonpj/Papers/hmap/index.htm}%
+  Has shrink example%
+%
+* QuickSpec:%
+  \url{http://www.cse.chalmers.se/~nicsma/quickspec.pdf}%
+%
+\url{http://blog.regehr.org/archives/697} --- check out C-reduce (Regehr) and Delta%
+  debugging (\url{http://www.st.cs.uni-saarland.de/dd/}).  C-reduce is in PLDI.%
+%
+* Also check out \url{http://www.whyprogramsfail.com/}%
+%
+* GenCheck \url{https://github.com/JacquesCarette/GenCheck}%
+%
+* Dart-check: \url{http://code.google.com/p/dart-immutable/} (and G+ post by Paul%
+  Brauner)%
+}
 

### 2013-01-27 src/Test/SmartCheck.hs
SC: bug fix in Boolean test on indexes.

@@ -65,9 +65,9 @@ smartCheck args propT = do
       let replIdxs = Replace valIdxs csIdxs
 
       -- If either kind of extrapolation pass yielded fruit, prettyprint it.
-      if (extrap args && null valIdxs) || (constrGen args && null csIdxs)
-        then smartPrtLn "Could not extrapolate a new value."
-        else output d replIdxs
+      if not $ null (valIdxs ++ csIdxs)
+        then output d replIdxs
+        else smartPrtLn "Could not extrapolate a new value."
 
       -- Ask the user if she wants to try again.
       putStrLn $ "Attempt to find a new counterexample?\n"

### 2013-01-27 src/Test/SmartCheck/SmartGen.hs
wibbles.

@@ -26,9 +26,9 @@ iterateArbIdx :: SubTypes a
               => a -> (Idx, Maybe Int) -> Int -> Int
               -> (a -> Q.Property) -> IO (Int, Result a)
 iterateArbIdx d (idx, max) tries sz prop =
-  case getAtIdx d idx max of
-    Nothing  -> errorMsg "iterateArb 0"
-    Just ext -> iterateArb d ext idx tries sz prop
+  maybe (errorMsg "iterateArb 0")
+        (\ext -> iterateArb d ext idx tries sz prop)
+        (getAtIdx d idx max)
 
 -- | Replace the hole in d indexed by idx with a bunch of random values, and
 -- test the new d against the property.  Returns the first new d (the full d but
@@ -76,14 +76,14 @@ iterateArb d ext idx tries max prop = do
     (size, g0) = randomR (0, currMax) g
     s = sample ext g size
     sample SubT { unSubT = v } = newVal v
-    rec res' = iterateArb' res' g0 (try + 1)
-                 -- XXX what ratio is right to increase size of values?  This
-                 -- gives us exponentail growth, but remember we're randomly
-                 -- chosing within the range of [0, max], so many values are
-                 -- significantly smaller than the max.  Plus we reset the size
-                 -- whenever we get a value that's too big.  Note the need for
-                 -- (+ 1), since we seed with 0.
-                 ((currMax + 1) * 2)
+    rec res' =
+      iterateArb' res' g0 (try + 1)
+        -- XXX what ratio is right to increase size of values?  This gives us
+        -- exponentail growth, but remember we're randomly chosing within the
+        -- range of [0, max], so many values are significantly smaller than the
+        -- max.  Plus we reset the size whenever we get a value that's too big.
+        -- Note the need for (+ 1), since we seed with 0.
+        ((currMax + 1) * 2)
 
 --------------------------------------------------------------------------------
 

### 2013-02-03 paper/paper.tex
Updates: paper, examples.

@@
 
-\newenvironment{code}{\begin{alltt}}{\end{alltt}}
+\newenvironment{code}{\begin{alltt}\small}{\end{alltt}}
 %% \newenvironment{cols}{\begin{tabular}{m{3.6cm}|m{3.6cm}}{Haskell} &
@@
 
-\todo{define variant, tag}
-\todo{note constants are user-defined arguments, other simplifications}
-\todo{thread args through}
+\todo{complexity of each algorithm}
 \todo{termination for each algorithm}
-\todo{fromMaybe over pattern matching?}
 \todo{show quantification in intro}
-\todo{benchmark feat}
+\todo{benchmark feat -- send to author}
 
@@
 \noindent
-The purpose of \ttp{smaller} is to generate strictly smaller values (according to
-some measure) from a given counterexample to attempt to find a smaller
-counterexample.  Today, \ttp{smaller} is called \ttp{shrink}.
+The purpose of \ttp{smaller} is to generate strictly smaller values (according
+to some measure) from a given counterexample.  These new values are then tested
+to attempt to find a smaller counterexample.  Today, \ttp{smaller} is called
+\ttp{shrink}.
 
@@
 \begin{code}
-data A = A Int16 deriving Show
-
-data B = B [A] [A] [A] [A] deriving Show
-
--- QuickCheck instances
-instance Arbitrary A where
-  arbitrary    = liftM A arbitrary
-  shrink (A i) = map A (shrink i)
-
-instance Arbitrary B where
-  arbitrary = liftM4 B arbitrary arbitrary
-                       arbitrary arbitrary
-  shrink (B a b c d) =
-    [ B w x y z | w <- shrink a
-                , x <- shrink b
-                , y <- shrink c
-                , z <- shrink d ]
-
--- Predicates for defining properties
-add :: [A] -> Int16
-add = sum . map (\(\backslash\)(A i) -> i)
-
-pre :: B -> Bool
-pre (B a b c d) =
-  all (\(\backslash\) -> add x < 16) [a, b, c, d]
-
-post :: B -> Bool
-post (B a b c d) =
-  add a + add b + add c + add d < 64
-
--- Property
-prop :: B -> Property
-prop p = pre p ==> post p
+data N a = N a
+type I   = [N Int16]
+data T   = T (N Word8) I I I I
+
+toList :: T -> [[Int16]]
+toList (T w i0 i1 i2 i3) =
+  [go w] : (map . map) go [i0, i1, i2, i3]
+  where go (N i) = fromIntegral i
+
+pre :: T -> Bool
+pre t = all ((>) 256 . sum) (toList t)
+
+post :: T -> Bool
+post t = (sum . concat) (toList t) < 5 * 256
+
+prop :: T -> Property
+prop t = pre t ==> post t
 \end{code}
@@
 it is typically defined using structural recursion for composite data types.
-For example, consider the \ttp{shrink} instances given to data types \ttp{A} and
-\ttp{B} in Figure~\ref{fig:initial}.\footnote{All examples and algorithms in
-  this paper are presented in Haskell~\cite{haskell98}.}  For \ttp{A}, we map
-the constructor \ttp{A} over a list of new \ttp{Int16}s returned by
-\ttp{shrink}; for \ttp{B}, we use Haskell's list comprehension notation to
-return a list of new values.  Indeed, because of the regular structure of
-\ttp{shrink} implementations, the principal motivation for one of Haskell's
-first generics package was for providing generic definitions for
-\ttp{shrink}~\cite{syb}.
-
-However, this straightforward approach to defining \ttp{shrink} instances may
-either miss minimal counterexamples, be too inefficient to be practical, or
-both.\footnote{For a recent real-world case, see the user question posed on
-  StackOverflow:
+Indeed, because of the regular structure of \ttp{shrink} implementations, the
+principal motivation for one of Haskell's first generics package was for
+providing generic definitions for \ttp{shrink}~\cite{syb}.  However, simple
+structural recursion may either miss smaller counterexamples, be too inefficient
+to be practical, or both.\footnote{For a recent real-world case, see the user
+  question posed on StackOverflow:
   \url{http://stackoverflow.com/questions/8788542/how-do-i-get-good-small-shrinks-out-of-quickcheck}.}
-For example, consider again Figure~\ref{fig:initial}.  The data type \ttp{B} is
-a product type over lists of 16-bit signed integers, wrapped by a constructor.
-Consider the property \ttp{prop} in the figure, stating that if the sum of each
-list in a field of \ttp{B} is less than 16, then the sum of the four fields is
-less than 64.  The property seems reasonable at first glance, until one realizes
-that due to underflow, the property can be violated.  For example, consider the
-value
-%
-\begin{code}
-B [A (-32769)] [] [] []
-\end{code}
-%
-\noindent
-Without shrinking, QuickCheck finds large examples to the property; on average,
-QuickCheck returns a value contains 70 \ttp{Int16} values!
-
-Thus, it pays to define \ttp{shrink}.  Unfortunately, even for a simple property
-and program like this one, the definition for \ttp{shrink} given in
-Figure~\ref{fig:initial} produces an intractable number of potential
-counterexamples to test; with the definitions and property provided (together
-with QuickCheck's default \ttp{Arbitrary} instance for lists), the list of
-potential counterexamples generated by \ttp{shrink} can contain more than
-$10^{10}$ elements \todo{true?}.  This is an enormous state space to search for a smaller
-example and can cause the shrinking phase to take orders-of-magnitude more time
-than the counterexample discovery phase.
+
+Consider the example in Figure~\ref{fig:initial}.\footnote{All examples and
+  algorithms in this paper are presented in Haskell.  We strive to use idiomatic
+  Haskell, and only rely on the language extensions of existential type
+  quantification and scoped type variables.~\cite{haskell98}.}  Data type
+\ttp{T} is a product type containing four lists of \ttp{Int16}s and a single
+\ttp{Word8}.  For our purposes, it is not important what \ttp{T} models, but
+for concreteness, suppose it models the values in four linked lists in an
+operating system, together with an 8-bit status register.
+
+Now suppose we are modeling some program that serializes values of type \ttp{T}.
+The input to the program satisfies the invariant \ttp{pre}, that the sum of
+values in each list of \ttp{Int16}s is less than or equal to 256 (and by
+definition, the \ttp{Word8} value is less than or equal to 256).  Assuming this,
+we want to show \ttp{post} holds, that the sum of all the values from \ttp{T} is
+less than $5 * 256$.  At first glance, the property seems reasonable, but we
+have forgotten about underflows.  For example, the value
+%
+\begin{code}
+T (N 10) [N (-20000)] [N (-20000)] [] []
+\end{code}
+%
+fails \ttp{prop} (\ttp{==>} is implication from the QuickCheck library).
+Unfortunately, a typical counterexample returned by QuickCheck is not quite so
+simple.  With standard settings, it returns a value containing just under 70
+\ttp{N} constructors, and sometimes values contain over 100 constructors.  Thus,
+it pays to define \ttp{shrink}!
+
+A typical definition of \ttp{shrink} for a data type like \ttp{T} is to take the
+cross-product of shrunk values over the arguments to the constructor \ttp{T}.
+This can be expressed using Haskell's list-comprehension notation:
+%
+\begin{code}
+shrink (T w i0 i1 i2 i3) =
+  [ T a b c d e | a <- shrink w,  b <- shrink i0
+                , c <- shrink i1, d <- shrink i2
+                , e <- shrink i3 ]
+\end{code}
+%
+\noindent
+The result of (\ttp{shrink t}) with this definition is null if any list
+contained in \ttp{t} is null.  Still, for typical counterexamples returned by
+QuickCheck that do not contain empty lists, together with QuickCheck's default
+\ttp{Arbitrary} instances, the length of potential counterexamples returned by
+\ttp{shrink} can easily exceed $10^{10}$, which is an intractable number of
+tests for QuickCheck to analyze.  The time complexity is compounded by the
+shrink stage being recursive, so when a new smaller counterexample is found, it
+is also shrunk.  In practice, QuickCheck appears to loop indefinitely.
 
@@
 trade-off is quicker shrinking with a lower probability of finding a smaller
-counterexample.  For example, redefining shrink for type \ttp{B} as follows
-%
-
-%% \begin{minipage}{\columnwidth}
-%% \lstset{ caption={Truncated shrinking function for data type \ttp{B}.}
-%%        , label=lst:newshrink
-%%        , captionpos=b }
-%% \begin{lstlisting}
-
-\begin{figure}[ht]
+counterexample.  For example, we might redefine shrink as follows:
+%
+%\begin{figure}[ht]
 \begin{code}
@@
 
-shrink (B a b c d) =
-  [ B w x y z | w <- tk a, x <- tk b
-              , y <- tk c, z <- tk d ]
+shrink (T w i0 i1 i2 i3) =
+  [ T a b c d e | a <- tk w
+                , b <- tk i0, c <- tk i1
+                , d <- tk i2, e <- tk i3 ]
   where tk x = take len (shrink x)
 \end{code}
-  \caption{Truncated shrinking function for data type \ttp{B}.}
-  \label{lst:newshrink}
-\end{figure}
+%%   \caption{Truncated shrinking function for data type \ttp{B}.}
+%%   \label{lst:newshrink}
+%% \end{figure}
 
@@
 \noindent
-controls the exponential blowup of the input-space by taking the first ten
-elements of each list, so the cross product is limited to 10,000 elements.  The
-downside is that potentially smaller counterexamples may be omitted.
+let us call this version ``truncated shrink''.  Truncating the controls the
+blowup of the input-space; the number of possible new values is $10^5$ in this
+case.  While truncation controls the blowup, the downside is that potentially
+smaller counterexamples may be omitted.
 
@@
 
-\begin{figure}
-  \begin{code}
-instance Serial Int16 where
-  series d = drawnFrom [(-d')..d']
-    where d' = fromIntegral d
-
-instance Serial A where series = cons1 A
-instance Serial B where series = cons4 B
-  \end{code}
-  \caption{SmallCheck instances for the data types in
-    Figure~\ref{fig:initial}.}
-  \label{fig:smallshrink}
-\end{figure}
+%% \begin{figure}
+%%   \begin{code}
+%% instance Serial Int16 where
+%%   series d = drawnFrom [(-d')..d']
+%%     where d' = fromIntegral d
+
+%% instance Serial A where series = cons1 A
+%% instance Serial B where series = cons4 B
+%%   \end{code}
+%%   \caption{SmallCheck instances for the data types in
+%%     Figure~\ref{fig:initial}.}
+%%   \label{fig:smallshrink}
+%% \end{figure}
 
 With SmallCheck, a property is checked up to a depth set by the user that
-measures how large values are.  For example, for type \ttp{Int}, the depth $d$
-is defined as the integers in the list enumerating from $-d$ to $d$ (i.e.,
-\ttp{[(-d)..d]}).  For a product type, SmallCheck must check values produced by
-taking the cross-product of the type's fields at a given depth.
-Figure~\ref{fig:smallshrink} contains functions for generating Lazy SmallCheck
-(a variant of SmallCheck) tests by defining instances for the \ttp{Serial}
-class.  For the property \ttp{prop} defined above (suitably redefined to use
-SmallCheck's types), SmallCheck must check to depth 16 to find the first
-counterexample.  Unfortunately, after a couple hours of testing, Lazy SmallCheck
-is still checking values at depth six, and the number of tests scales
-exponentially with respect to the depth.\footnote{All tests in this paper are
-  performed on a four-core Pentium~i7 running at 2.7GHz with 8GB RAM on
-  Fedora~16.  This test, and others unless noted, are performed using
-  interpreted Haskell under GHC 7.4.2.\todo{redo tests?}}
+measures how large values are (typically, the depth is a number of constructors
+in the value).  For a product type, SmallCheck must check values produced by
+taking the cross-product of the type's fields at a given depth.  SmallCheck is
+unable to enumerate all possible values of type \ttp{T} to discover a
+counterexample, due to combinatorial explosion.  (A related library named Feat
+combines some aspects of SmallCheck and QuickCheck~\cite{feat}; we discuss it
+more in the related work (Section~\ref{sec:related}).
+
+%% is unable to verify the property from Figure~\ref{fig:initial}, for much the
+%% same reason that \ttp{shrink} is inefficient
+
+%% For the property \ttp{prop} defined in Figure~\ref{fig:initial} (suitably
+%% redefined to use SmallCheck's types).
+
+%% to depth 16 to find the first counterexample.  Unfortunately, after a couple
+%% hours of testing, Lazy SmallCheck is still checking values at depth six, and
+%% the number of tests scales with respect to the depth.\footnote{All tests in
+%% this paper are performed on a four-core Pentium~i7 running at 2.7GHz with 8GB
+%% RAM on Fedora~16.  This test, and others unless noted, are performed using
+%% interpreted Haskell under GHC 7.4.2.\todo{redo tests?}}
+
+%% For example, for type \ttp{Int}, the depth $d$ is defined as the integers in
+%% the list enumerating from $-d$ to $d$ (i.e., \ttp{[(-d)..d]}).
+%% Figure~\ref{fig:smallshrink} contains functions for generating Lazy
+%% SmallCheck (a variant of SmallCheck) tests by defining instances for the
+%% \ttp{Serial} class.
 
@@
 \begin{itemize}
-  \item First, after discovering a counterexample to some property, say with
-    type
-%
-\begin{code}
-prop :: B -> Bool
-\end{code}
-%
-\noindent
-one might wish to search for different counterexamples to \ttp{prop}.  Ideally,
-the testing framework would automatically add a predicate characterizing (and
-perhaps generalizing) a counterexample
-%
-\begin{code}
-cex :: B -> Bool
-\end{code}
-%
-\noindent
-as a precondition to \ttp{prop} to define a new property
-\begin{code}
-prop' b = not (cex b) ==> prop b
-\end{code}
-%
-\noindent
-where \ttp{==>} is QuickCheck's implication operator that discards tests that
-violate the precondition.
-
-  \item Second, and more generally, a counterexample is usually representative
-    of other counterexamples, but that generalization is left to the user to
-    discover for herself.  Really, what would be useful would be not just a
-    minimal counterexample, but a quantified formula characterizing a set of
-    counterexamples.  Such a formula has the added benefit of replacing a
-    potentially large sub-term of a composite data type value with a quantifier.
+  \item A counterexample is usually representative of other counterexamples, but
+    that generalization is left to the user to discover for herself.  For
+    example, consider the counterexample presented earlier:
+%
+\begin{code}
+T (N 10) [N (-20000)] [N (-20000)] [] []
+\end{code}
+%
+\noindent
+For any \ttp{Word8}, the counterexample still holds.  Thus, we can be more
+informative and return a quantified formula rather than a single counterexample:
+%
+\begin{code}
+forall x . T x [N (-20000)] [N (-20000)] [] []
+\end{code}
+%
+\noindent
+Such a formula has the added benefit of replacing a potentially large sub-term
+of a composite data type value with a quantifier, reducing the sheer size of the term.
+
+  \item Second, during testing, it is sometimes helpful to cover the input space
+    and by discovering different counterexamples to \ttp{prop}.  In the example
+    in Figure~\ref{fig:initial}, one counterexample is fairly representative of
+    others, but this is not true in general (e.g., consider properties about
+    abstract syntax trees).  Ideally, the testing framework would automatically
+    add a predicate characterizing a counterexample as the precondition to
+    \ttp{prop} that strengthens the property.  That is, for a specific
+    counterexample \ttp{ex}, we have a predicate \ttp{predEx} that holds if its
+    argument is another counterexample that is similar to (i.e., some
+    generalization of) \ttp{ex}.  Then we test against a new property
+%
+\begin{code}
+prop' a = not (cex a) ==> prop a
+\end{code}
+%
+%% \noindent
+%% where \ttp{==>} is QuickCheck's implication operator that discards tests that
+%% violate the precondition.
 \end{itemize}
@@
 find small counterexamples as well as generalize them as described in the
-preceding paragraph.  To motivate the benefit of SmartCheck, consider the table
-in Table~\ref{table:results}, giving the results for using QuickCheck and
+preceding paragraph.  To motivate the benefit of SmartCheck, consider
+Table~\ref{table:results}, giving the results for using QuickCheck and
 SmartCheck on the program shown in Figure~\ref{fig:initial} (we omit SmallCheck,
@@
 shrinking to provide a baseline.  The next two columns show the result of using
-QuickCheck with definitions of \ttp{shrink} that limit the lists of fields of
-\ttp{B} to 10 (Figure~\ref{lst:newshrink}) and 20 elements
-(Figure~\ref{lst:newshrink}, with \ttp{len = 20}), respectively.  (With the
-original definition for shrink from Figure~\ref{fig:initial}, some
-counterexamples cause QuickCheck to not return during the shrinking phase after
-hours of computation, so we omit these.)
-
-  In each row, we show the maximum, mean, and median
-number of seconds required by QuickCheck to generate a counterexample over 100
-runs.  In the last row, we show the mean size of the final value returned,
-measured by the taking the cumulative length of the lists of the fields of
-\ttp{B}.  Notice that while typically counterexamples are returned in a fraction
-of a second, there are outliers taking 21 seconds and 125 seconds for the two
-tests.  Both tests roughly cut the size of the counterexample in half;
-increasing the number of potential counterexamples can result in a significant
-performance penalty for this example without much benefit.
+QuickCheck with definitions of truncated \ttp{shrink} that limit the number of
+smaller values to 10 and 20 elements, respectively.  (With the original
+definition for shrink from Figure~\ref{fig:initial}, some counterexamples cause
+QuickCheck to not return during the shrinking phase after hours of computation,
+so we omit these.)
+
+In each row, we show the maximum, mean, and median number of seconds required by
+QuickCheck to generate a counterexample over 100 runs.  In the last row, we show
+the mean size of the final value returned, measured by the taking the cumulative
+length of the lists of the fields of \ttp{B}.  Notice that while typically
+counterexamples are returned in a fraction of a second, there are outliers
+taking 21 seconds and 125 seconds for the two tests.  Both tests roughly cut the
+size of the counterexample in half; increasing the number of potential
+counterexamples can result in a significant performance penalty for this example
+without much benefit.
 
@@
   characterizes an already-discovered counterexample for the purpose of finding
-  new, dissimilar counterexamples(Section~\ref{sec:generalization});
+  new, dissimilar counterexamples (Section~\ref{sec:generalization});
 
@@
 \subsection{Reduction Algorithm Overview}\label{sec:reduct}
-The algorithm we present for efficiently searching for new counterexamples can
-be thought of as a recursive breadth-first traversal over the structure of the
-current counterexample value, where the value is represented as a tree.  At each
-node, during the traversal, we generate arbitrary \emph{structurally} smaller
-sub-values and build a new value from that, leaving the remainder of the tree
-unchanged.  Intuitively, by a \emph{structurally smaller value}, we mean one
-with fewer constructors.  We continue until we reach a fixed-point.
+The algorithm we present for efficiently searching for new counterexamples is an
+instance of greedy breadth-first search over a tree structures, representing a
+value.  At each node, during the traversal, we generate arbitrary
+\emph{structurally} smaller sub-values and build a new value from that, leaving
+the remainder of the tree unchanged.  Intuitively, by a \emph{structurally
+  smaller value}, we mean one with fewer constructors.  We continue until we
+reach a fixed-point.
 
@@
 terminate: intuitively, the measure for the function is that either the index
-increases or the size of the counterexample being evaluated decreases.
+increases or the size of the counterexample being evaluated decreases.  The
+complexity of the algorithm is $\Omega(n^2)$, where $n$ is the number of
+constructors in the counterexample.  The analysis assumes generating new
+sub-values and testing them is done in constant time.
 
@@
   \begin{center}
-    \includegraphics[scale=0.5]{Figs/cex-gen}
+    \includegraphics[scale=0.49]{Figs/cex-gen}
    \end{center}
@@
 \begin{figure}
-
+  \begin{code}
+matchesShapes :: SubTypes a => a -> [(a,([Idx],[Idx]))] -> Bool
+matchesShapes d = any (matchesShape d)
+
+matchesShape :: SubTypes a => a -> (a, ([Idx],[Idx])) -> Bool
+matchesShape a (b, (idxVals, idxConstrs))
+  | constr a /= constr b = False
+  | Just a' <- aRepl         = let x = subForest (subVals a') in
+                               let y = subForest (subVals b) in
+                               all foldEqConstrs (zip x y)
+  | otherwise                = False
+
+  where
+  foldEqConstrs :: (Tree SubVal, Tree SubVal) -> Bool
+  foldEqConstrs (Node (SubVal l0) sts0, Node (SubVal l1) sts1)
+    | constr l0 == constr l1 = next
+    | otherwise              = False
+    where next = all foldEqConstrs (zip sts0 sts1)
+
+  updateA idx d = fmap (replace d idx) (index b idx)
+  aRepl = foldl go (Just a) (idxVals ++ idxConstrs)
+    where go ma idx = maybe Nothing (updateA idx) ma
+  \end{code}
   \caption{Shape matching algorithm.}
@@
 
-\todo{shrink paper rehger pointed to ``Automatic isolation of compiler errors'' from 1994}
-
-Closely related to SmartCheck is the recent Haskell library Feat.  Feat
-enumerates algebraic data types in Haskell and indexing them
-quickly~\cite{feat}.  For example, Feat returns the $10^{1000}$ value in the
-enumeration of lists of Booleans in under a second.  Feat combines some of the
-advantages of SmallCheck and QuickCheck:
+\todo{shrink paper rehger pointed to ``Automatic isolation of compiler errors''
+  from 1994}
+\todo{\url{http://www.cs.utah.edu/~regehr/papers/pldi12-preprint.pdf}}
+\url{http://blog.regehr.org/archives/697} --- check out C-reduce (Regehr) and Delta%
+  debugging (\url{http://www.st.cs.uni-saarland.de/dd/}).  C-reduce is in PLDI.%
+
+Related to SmartCheck is the recent Haskell library Feat~\cite{feat}.  Feat
+provides automated enumerations of algebraic data types in Haskell, allowing for
+fast access to very large indexes.  For example, from the enumeration of
+(\ttp{[Bool]})
+%
+\begin{code}
+[[],[False],[True],[False,False],[False,True] ...
+\end{code}
+%
+\noindent
+Accessing the $10^{1000}$th element takes under $0.1$ seconds in interpreted
+Haskell.  Feat combines some advantages of SmallCheck and QuickCheck, since the
+user can choose to exhaustively test an enumeration up to some depth, like with
+SmallCheck, or she can create a uniform distribution of test cases up to some
+depth.
+
+combines some of the advantages of SmallCheck and QuickCheck:
 
@@
 %
-\url{http://blog.regehr.org/archives/697} --- check out C-reduce (Regehr) and Delta%
-  debugging (\url{http://www.st.cs.uni-saarland.de/dd/}).  C-reduce is in PLDI.%
 %
@@
 }
+%
+\url{http://www.haskell.org/haskellwiki/Research_papers/Testing_and_correctness}
 

### 2013-02-03 paper/Example/InitEx.hs

@@ -91,10 +91,10 @@ add :: [A] -> Int16
 add = sum . map (\(A i) -> i)
 
 pre :: B -> Bool
-pre (B a b c d) = all (\x -> add x < 16) [a, b, c, d]
+pre (B a b c d) = all (\x -> add x < 5000) [a, b, c, d]
 
 post :: B -> Bool
-post (B a b c d) = add a + add b + add c + add d < 64
+post (B a b c d) = add a + add b + add c + add d < 20000
 
 prop_qc :: B -> Property
 prop_qc p = pre p ==> post p
@@ -169,4 +169,4 @@ main = do
 #endif
 
 smtChk :: IO ()
-smtChk = smartCheck scStdArgs {scMaxExtrap = 20, format = PrintString } prop_qc
+smtChk = smartCheck scStdArgs {scMaxExtrap = 30, scMinExtrap = 25, format = PrintString } prop_qc

### 2013-02-03 paper/Example/InitEx2.hs

@@ -1,0 +1,197 @@
+{-# LANGUAGE CPP #-}
+{-# LANGUAGE TemplateHaskell #-}
+{-# LANGUAGE DeriveDataTypeable #-}
+{-# LANGUAGE FlexibleInstances #-}
+{-# LANGUAGE DeriveGeneric #-}
+
+-- | Try to generate a very large counterexample.
+
+module Main where
+
+import Test.SmartCheck
+import Test.SmartCheck.Reduce
+import Test.QuickCheck
+import Test.LazySmallCheck hiding (Property, test, (==>))
+import qualified Test.LazySmallCheck as S
+
+import GHC.Generics hiding (P, C)
+import Data.Typeable
+import Control.Monad
+
+import Data.Int
+import Data.Word
+import Data.List
+import Data.Maybe
+import Data.Time
+import System.Environment
+
+import Test.Feat
+
+-----------------------------------------------------------------
+
+-- Let's try to generate a product type of long lists when all we need is a
+-- single element to have a long list.
+
+-- Container so that we don't have base types.
+
+-----------------------------------------------------------------
+
+data N a = N a
+  deriving (Read, Show, Typeable, Generic)
+
+type I = [N Int16]
+
+data T = T (N Word8) I I I I
+  deriving (Read, Show, Typeable, Generic)
+
+-- SmallCheck --------------------------
+enum :: (Enum b, Integral a, Num b) => a -> [b]
+enum d = [(-d')..d']
+  where d' = fromIntegral d
+
+instance Serial Int16 where
+  series = drawnFrom . enum
+
+instance Serial Word8 where
+  series = drawnFrom . enum
+
+instance Serial a => Serial (N a) where
+  series = cons1 N
+
+instance Serial T where
+  series = cons5 T
+
+-- SmallCheck --------------------------
+
+-- SmartCheck --------------------------
+instance SubTypes I
+instance SubTypes a => SubTypes (N a)
+instance SubTypes T
+-- SmartCheck --------------------------
+
+instance Arbitrary a => Arbitrary (N a) where
+  arbitrary = liftM N arbitrary
+  shrink (N i) = map N (shrink i)
+
+-- qc/shrink takes over 1m seconds
+instance Arbitrary T where
+#ifdef feat
+  arbitrary = sized uniform
+#else
+  arbitrary = liftM5 T arbitrary arbitrary
[…]

### 2013-02-03 paper/Paper.hs

@@ -3,12 +3,13 @@
 -- | Model of the SmartCheck algorithms explained in the paper.  We leave a
 -- number of functions undefined here.
 
--- Limit: col 46
+-- Limit: col 50
 
 module Paper where
 
 import Prelude hiding (fail)
 import Data.Maybe (mapMaybe)
+import Data.Tree
 import Control.Monad (liftM, replicateM)
 import Test.QuickCheck
 
@@ -20,47 +21,35 @@ data Format = PrintTree | PrintString
   deriving (Eq, Read, Show)
 
 -- SmartCheck arguments
-format       :: Format    -- ^ How to show extrapolated formula
-format       = undefined
-qcArgs       :: Args    -- ^ QuickCheck arguments
-qcArgs       = undefined
-qc           :: Bool      -- ^ Should we run QuickCheck?  (If not, you are
-                          -- expected to pass in data to analyze.)
-qc           = undefined
-scMaxSize    :: Int       -- ^ Maximum size of data to generate, in terms of the
-                          -- size parameter of QuickCheck's Arbitrary instance
-                          -- for your data.
-scMaxSize    = undefined
-scMaxDepth   :: Maybe Int -- ^ How many levels into the structure of the failed
-                          -- value should we descend when reducing or
-                          -- generalizing?  Nothing means we go down to base
-                          -- types.
-scMaxDepth   = undefined
-scMaxReduce  :: Int       -- ^ How hard (number of rounds) to look for failure
-                          -- in the reduction stage.
-scMaxReduce  = undefined
-extrap       :: Bool      -- ^ Should we extrapolate?
-extrap       = undefined
-scMaxExtrap  :: Int       -- ^ How hard (number of rounds) to look for failures
-                          -- during the extrapolation and constructor
-                          -- generalization stages.
-scMaxExtrap  = undefined
-scMinExtrap  :: Int       -- ^ Minimum number of times a property's precondition
-                          -- must be passed to generalize it.
-scMinExtrap  = undefined
-constrGen    :: Bool      -- ^ Should we try to generalize constructors?
-constrGen    = undefined
-scConstrMax  :: Int       -- ^ How hard (number of rounds) to look for failing
[…]

### 2013-02-03 src/Test/SmartCheck/Args.hs

@@ -71,11 +71,11 @@ scStdArgs = ScArgs { format       = PrintTree
                    , scMaxReduce  = 100
                    ---------------------
                    , extrap       = True
-                   , scMaxExtrap  = 100
+                   , scMaxExtrap  = 20
                    , scMinExtrap  = 10
                    ---------------------
                    , constrGen    = True
-                   , scConstrMax  = 100
+                   , scConstrMax  = 20
                    }
 
 --------------------------------------------------------------------------------

### 2013-02-10 paper/paper.tex
paper updates.

@@
 user to write \emph{shrink} functions to attempt to reduce the size of
-counterexamples discovered by QuickCheck.  Implementations of \emph{shrink}
-often require significant boilerplate and usually just generate smaller tests
-than the discovered counterexample.  Therefore, shrinking can be inefficient and
-poor at discovering new, smaller counterexamples.  In this paper, we introduce
+counterexamples discovered by QuickCheck.  Implementations of shrink often
+require significant boilerplate and usually just generate smaller tests than the
+discovered counterexample.  Therefore, shrinking can be inefficient and poor at
+discovering new, smaller counterexamples.  In this paper, we introduce
 \emph{SmartCheck}, an automatic efficient counterexample reduction algorithm.
@@
 
+In industrial uses, shrinking is essential.  Hughes, the inventor of QuickCheck
+and the CEO of Quiviq, a company that commercializes the technology, has noted
+that ``without it [shrinking], randomly generated failing cases would often be
+so large as to be almost useless.''~\cite{qcjh}.  Hughes~\emph{et al.} give an
+extended example in which shrinking is essential in debugging telecom
+software~\cite{telecom}.
+
+Typically, effective shrinking on complex data requires a good understanding of
+how shrinking in QuickCheck works and the semantics of the property and program
+being evaluated.  Bad definitions can be so inefficient that they are unusable.
+Shrinking presents a programmer's dilemma: shrinking is used to better
+understand a counterexample to understand why a property failed, but to
+implement an efficient shrink function in general requires knowing how to
+generate counterexamples to the property.
+
 \begin{figure}[ht]
@@
 
-While \ttp{shrink} is user-defined and can have any type-correct implementation,
-it is typically defined using structural recursion for composite data types.
-Indeed, because of the regular structure of \ttp{shrink} implementations, the
-principal motivation for one of Haskell's first generics package was for
-providing generic definitions for \ttp{shrink}~\cite{syb}.  However, simple
-structural recursion may either miss smaller counterexamples, be too inefficient
-to be practical, or both.\footnote{For a recent real-world case, see the user
-  question posed on StackOverflow:
-  \url{http://stackoverflow.com/questions/8788542/how-do-i-get-good-small-shrinks-out-of-quickcheck}.}
-
-Consider the example in Figure~\ref{fig:initial}.\footnote{All examples and
-  algorithms in this paper are presented in Haskell.  We strive to use idiomatic
-  Haskell, and only rely on the language extensions of existential type
-  quantification and scoped type variables.~\cite{haskell98}.}  Data type
+To illustrate, consider the example in Figure~\ref{fig:initial}.\footnote{All
+  examples and algorithms in this paper are presented in Haskell.  We strive to
+  use idiomatic Haskell, and only rely on the language extensions of existential
+  type quantification and scoped type variables.~\cite{haskell98}.}  Data type
 \ttp{T} is a product type containing four lists of \ttp{Int16}s and a single
-\ttp{Word8}.  For our purposes, it is not important what \ttp{T} models, but
-for concreteness, suppose it models the values in four linked lists in an
-operating system, together with an 8-bit status register.
+\ttp{Word8}.  For our purposes, it is not important what \ttp{T} models, but for
+concreteness, suppose it models the values in four linked lists in an operating
+system, together with an 8-bit status register.
 
@@
 we want to show \ttp{post} holds, that the sum of all the values from \ttp{T} is
-less than $5 * 256$.  At first glance, the property seems reasonable, but we
-have forgotten about underflows.  For example, the value
+less than $5 * 256$, where five is the number of fields in \ttp{T}.  At first
+glance, the property seems reasonable.  But we have forgotten about underflow;
+for example, the value
 %
@@
 %
+\noindent
 fails \ttp{prop} (\ttp{==>} is implication from the QuickCheck library).
-Unfortunately, a typical counterexample returned by QuickCheck is not quite so
-simple.  With standard settings, it returns a value containing just under 70
-\ttp{N} constructors, and sometimes values contain over 100 constructors.  Thus,
-it pays to define \ttp{shrink}!
-
-A typical definition of \ttp{shrink} for a data type like \ttp{T} is to take the
-cross-product of shrunk values over the arguments to the constructor \ttp{T}.
-This can be expressed using Haskell's list-comprehension notation:
+
+Despite the simplicity of the example, a typical counterexample returned by
+QuickCheck is not quite so simple.  With standard settings, the median
+counterexample discovered contains around 70 \ttp{N} constructors, and
+counterexamples can contain over 100 constructors.  Thus, it pays to define
+\ttp{shrink}!
+
+We might first naively try to shrink counterexamples for a data type like \ttp{T}
+by taking the cross-product of shrunk values over the arguments to the
+constructor \ttp{T}.  This can be expressed using Haskell's list-comprehension
+notation:
 %
@@
 \noindent
-The result of (\ttp{shrink t}) with this definition is null if any list
-contained in \ttp{t} is null.  Still, for typical counterexamples returned by
-QuickCheck that do not contain empty lists, together with QuickCheck's default
-\ttp{Arbitrary} instances, the length of potential counterexamples returned by
-\ttp{shrink} can easily exceed $10^{10}$, which is an intractable number of
-tests for QuickCheck to analyze.  The time complexity is compounded by the
-shrink stage being recursive, so when a new smaller counterexample is found, it
-is also shrunk.  In practice, QuickCheck appears to loop indefinitely.
-
-It is not always straightforward to determine how to make a more efficient
-implementation for \ttp{shrink}.  Shrinking presents a programmer's dilemma:
-shrinking is used to better understand a counterexample to understand why a
-property failed, but to implement an efficient shrink function in general
-requires knowing how to generate counterexamples to the property.
-
-Typically, a programmer tries a brute-force approach to reduce the number of
-potential counterexamples; for example, she may try to truncate lists using the
+While the above definition appears reasonable, there are two problems with it.
+First, the result of (\ttp{shrink t}) is null if any list contained in \ttp{t}
+is null, then \ttp{t} will not be shrunk.  More troublesome is that with
+QuickCheck's default \ttp{Arbitrary} instances, the length of potential
+counterexamples returned by \ttp{shrink} can easily exceed $10^{10}$, which is
+an intractable number of tests for QuickCheck to analyze.  The reason for the
+blowup is that shrinking a list \ttp{[a]} produces a list \ttp{[[a]]}, the
+length of which is significantly longer than the original list.  Then, we take
+the cross-product of the generated lists.  For the example above, with some
+counterexamples, the shrinking stage may appear to loop for hours, consuming
+ever more memory, without returning a result.
+
+A programmer might try to control the complexity by truncating lists using the
 (\ttp{take n}) function that returns the first $n$ elements of a list.  The
@@
 %
-%\begin{figure}[ht]
-\begin{code}
-len = 10
-
+\begin{code}
 shrink (T w i0 i1 i2 i3) =
@@
                 , d <- tk i2, e <- tk i3 ]
-  where tk x = take len (shrink x)
-\end{code}
-%%   \caption{Truncated shrinking function for data type \ttp{B}.}
-%%   \label{lst:newshrink}
-%% \end{figure}
-
-%% \end{lstlisting}
-%% \end{minipage}
-%
-\noindent
-let us call this version ``truncated shrink''.  Truncating the controls the
-blowup of the input-space; the number of possible new values is $10^5$ in this
-case.  While truncation controls the blowup, the downside is that potentially
-smaller counterexamples may be omitted.
+  where tk x = take 10 (shrink x)
+\end{code}
+%
+\noindent
+Call this version ``truncated shrink''.  Truncation controls the blowup of the
+input-space; the maximum number of possible new values is $10^5$ in this case.
+While truncation controls the blowup, the downside is that potentially smaller
+counterexamples may be omitted.
+
+A more clever programmer that understands the semantics of QuickCheck's
+\ttp{shrink} implementation,\footnote{The following approach was suggested to
+  the author by John Hughes.} defines a \ttp{shrink} instance as follows:
+%
+\begin{code}
+shrink (T w i0 i1 i2 i3) =
+  let xs = shrink (w, i0, i1, i2, i3) in
+  let go (w', i0', i1', i2', i3') =
+    T w' i0' i1' i2' i3' in
+  map go xs
+\end{code}
+%
+\noindent
+This ``tuple shrink'' definition does not suffer the same shortcomings: it
+shrinks a value even if it contains an empty list, and the combinatorial blowup
+of shrink candidates is avoided, since a pair \ttp{(a,b)} is shrunk by
+attempting to shrink \ttp{a} while holding \ttp{b} constant, then attempting to
+shrink \ttp{b} while holding \ttp{a} constant, and the approach is generalized
+to larger tuples.  Still, with this definition, in some cases, shrinking takes
+over 20 seconds and returns a result containing over 60
+constructors.\footnote{Results are with a GHC-compiled program, using \ttp{-O2},
+  on a 4-core 2.7GHz Intel~i7 processor.}
 
@@
 \hline
- & QC (none) & QC (10) & QC (20) & SmartCheck \\
+         & QC (none)  & QC (trunc) & QC (tuple) & SmartCheck\\
 \hline \hline
-Max.  & 0.15s & 21.51s & 125.37s & 1.96s\\
+Maximum  & 0.40s, 124 & 139.15s, 90  & 97.53s, 53 & 0.s, 13\\
 \hline
-Mean  & 0.07s & 1.21s & 3.80s & 0.30s\\
+Mean     & 0.08s, 69  & 0.73s, 35  & 2.26s, 13  & 0.07s, 6\\
 \hline
-Median & 0.07s & 0.48s & 0.52s & 0.24s\\
-\hline
-Mean size & 70 & 34 & 32 & 5\\
+Median   & 0.07s, 68  & 0.18s, 32  & 1.90s, 13  & 0.07s, 5\\
 \hline
@@
   \end{center}
-  \caption{Results for 100 tests of the program in Figure~\ref{fig:initial}
+  \caption{Results for 1000 tests of the program in Figure~\ref{fig:initial}
     using interpreted Haskell.}
@@
 \end{table}
+\todo{table generated with InitEx2 on Feb. 3 --- do a final test with 1000}
 
 \paragraph{SmartCheck}
-Motivated by these limitations of QuickCheck and SmallCheck, we have developed
+Motivated by these limitations of QuickCheck and SmallCheck, we developed
 \emph{SmartCheck}.  SmartCheck extends QuickCheck to efficiently and generically
@@
 SmartCheck on the program shown in Figure~\ref{fig:initial} (we omit SmallCheck,
-since as we noted, it is not even feasible for an example like this).  In the
-first three columns, the results of using QuickCheck to generate counterexamples
-are shown.  The first column contains the results of using QuickCheck without
-shrinking to provide a baseline.  The next two columns show the result of using
-QuickCheck with definitions of truncated \ttp{shrink} that limit the number of
-smaller values to 10 and 20 elements, respectively.  (With the original
-definition for shrink from Figure~\ref{fig:initial}, some counterexamples cause
-QuickCheck to not return during the shrinking phase after hours of computation,
-so we omit these.)
-
-In each row, we show the maximum, mean, and median number of seconds required by
-QuickCheck to generate a counterexample over 100 runs.  In the last row, we show
-the mean size of the final value returned, measured by the taking the cumulative
-length of the lists of the fields of \ttp{B}.  Notice that while typically
-counterexamples are returned in a fraction of a second, there are outliers
-taking 21 seconds and 125 seconds for the two tests.  Both tests roughly cut the
-size of the counterexample in half; increasing the number of potential
-counterexamples can result in a significant performance penalty for this example
-without much benefit.
-
-In the last column are the results for SmartCheck.  Not only does it execute
-faster than QuickCheck with shrinking, it has significantly smaller outliers.
-The most striking difference, however, is the size of the counterexample
-returned, reducing on average the size of a counterexample found by QuickCheck
-without shrinking by a factor of 14.
-
-The purpose of this paper is to explain first how to efficiently and generically
-generate small counterexamples for algebraic data types and second how to
-generalize the counterexamples.
-
-\todo{check paper against this section again, especially generalization sec.}
+since as noted, it is not even feasible for the example).  We show the maximum,
+mean, and median amount of time required to generate a counterexample and to
+shrink it (and print it to standard out) as well as the corresponding number of
+constructors in the final result.
+
+In the first three columns, the results of using QuickCheck to generate
+counterexamples are shown.  The first column contains the results of using
+QuickCheck without shrinking to provide a baseline.  The next column shows the
+result of using QuickCheck with truncating shrink, which each list is truncated
+to 10 elements.  Next, we show the results of the ``tuple shrink'' instance.  In
+the last column are the results for SmartCheck, which includes the time required
+by QuickCheck to generate the original counterexample.  Not only does SmartCheck
+execute faster than the other strategies, it has smaller outliers.  In addition,
+the size of the counterexamples are significantly smaller.  And we stress that
+SmartCheck achieves these results \emph{generically}.
+
+\todo{change max to std dev, look for max info above.}
 
@@
 In Section~\ref{sec:experiments}, we discuss some of our experiences with using
-SmartCheck, including checking properties from Xmonad.
+SmartCheck, including checking properties from Xmonad, and related work and
+concluding remarks are made in Sections~\ref{sec:related}
+and~\ref{sec:conclusions}, respectively.
 
@@
 
-\todo{make clear that preconditions shouldn't be considered logical
-  preconditions}
-
 We focus on \emph{algebraic data types}~\cite{}.  Algebraic data types are
@@
 To present the algorithms in the following sections, we provide some definitions
-here.  Our presentation simplifies the implementation
-(Section~\ref{sec:implementation}) somewhat to improve the presentation.
-Consider the following types and type class.
+in the form of methods of a type class.  SmartCheck requires instances for data
+being analyzed.  In Section~\ref{sec:implementation}, we describe how instances
+for the type class are derived automatically.  Consider the following types and
+type class.
 %
@@
   baseType :: a -> Bool
+  subVals  :: a -> Tree SubVal
 \end{code}
@@
   We discuss the purpose of base types in Section~\ref{sec:base}.
+\item \ttp{subVals} returns a tree of all non base-type sub-values.  A tree is a
+  value with the type
+%
+  \begin{code}
+data Tree a = Node \{ rootLabel :: a
+                   , subForest :: [Tree a] \}
+  \end{code}
 \end{itemize}
@@
 \begin{code}
-data Tree = L | B Tree Tree
+data T = L | B T T
 \end{code}
@@
 baseType L          = False
+
+subVals (B (B L\sub{0} L\sub{1}) L\sub{2}) =
+  Tree (B L\sub{0} L\sub{1}) [Tree L\sub{0} [], Tree L\sub{1} []]
+       (Tree L\sub{1} [])
 \end{code}
@@
 For simplicity, we omit here Generics-specific super-class constraints on the
-\ttp{SubTypes} class.
+\ttp{SubTypes} class.  Moreover, our presentation simplifies the implementation
+(Section~\ref{sec:implementation}) somewhat to improve the presentation.
 
@@
 
-newVals :: SubVal -> Size -> Int -> IO [SubVal]
-newVals (SubVal a) sz tries =
-  replicateM tries s
+newVals :: Size -> Int -> SubVal -> IO [SubVal]
+newVals sz tries (SubVal a) =
+  replicateM tries s where
+  s  = liftM SubVal (sizedArbitrary sz a)
+
+reduce :: SubTypes a
+  => ScArgs -> (a -> Property) -> a -> IO a
+reduce args prop cex = reduce' 1
   where
-  s  = liftM SubVal (sizedArbitrary sz a)
-
-reduce :: SubTypes a
-       => a -> (a -> Property) -> IO a
-reduce cex prop = reduce' 1
-  where
-  reduce' idx =
-    case index cex idx of
-      Nothing -> return cex
-      Just v  -> do
-        vs <- newVals v (getSize v) scMaxReduce
-        case test cex idx vs prop of
-          Nothing   -> reduce' (idx+1)
-          Just cex' -> reduce cex' prop
+  reduce' idx
+    | Just v <- index cex idx
+    = do vs <- newVals (getSize v)
+                 (scMaxReduce args) v
+         maybe (reduce' (idx+1)) (reduce args prop)
+               (test cex idx vs prop)
+    | otherwise = return cex
 
@@
 Figure~\ref{fig:reduction} shows the reduction algorithm.  The function
-\ttp{reduce} takes a counterexample \ttp{cex} and the property \ttp{prop}.  At
-the beginning of execution, we assume that a counterexample has been obtained
-somehow; our implementation relies on QuickCheck as a ``front-end'' to generate
-a counterexample that the algorithm automatically shrinks.  The reduction begins
+\ttp{reduce} takes arguments to customize the algorithm's behavior, a
+counterexample \ttp{cex} and the property \ttp{prop}.  At the beginning of
+execution, we assume that a counterexample has been obtained somehow; our
+implementation relies on QuickCheck as a ``front-end'' to generate a
+counterexample that the algorithm automatically shrinks.  The reduction begins
 at the first proper sub-value of \ttp{cex}; call it \ttp{v}.  When the index
@@
 replace a sub-value.
-
-\todo{explain getSize}
 
@@
 increases or the size of the counterexample being evaluated decreases.  The
-complexity of the algorithm is $\Omega(n^2)$, where $n$ is the number of
+complexity of the algorithm is $\mathcal{O}(n^2)$, where $n$ is the number of
 constructors in the counterexample.  The analysis assumes generating new
@@
 \begin{code}
-reduce0 :: forall a . SubTypes a
-        => a -> (a -> Property) -> IO a
-reduce0 cex prop = reduce' 1
+reduceOpt :: forall a . SubTypes a
+  => ScArgs -> (a -> Property) -> a -> IO a
+reduceOpt args prop cex = reduce' 1
   where
-  reduce' idx =
-    case index cex idx of
-      Nothing -> return cex
-      Just v  -> case testHole v of
-                   Just v' -> reduce0 v' prop
-                   Nothing -> test' v idx
+  reduce' idx
+    | Just v <- index cex idx
+    = maybe (test' v idx) (reduceOpt args prop)
+            (testHole v)
+    | otherwise = return cex
 
   test' v idx = do
-    vs <- newVals v (getSize v) scMaxReduce
-    case test cex idx vs prop of
-      Nothing   -> reduce' (idx+1)
-      Just cex' -> reduce0 cex' prop
-
-  testHole (SubVal a) =
-    case cast a :: Maybe a of
-      Nothing -> Nothing
-      Just a' ->
-        if pass prop a' then Nothing
-          else Just a'
+    vs <- newVals (getSize v) (scMaxReduce args) v
+    maybe (reduce' (idx+1)) (reduceOpt args prop)
+          (test cex idx vs prop)
+
+  testHole (SubVal a) = do
+    a' <- cast a :: Maybe a
+    if pass prop a' then Nothing else Just a'
 \end{code}
@@
 directly, and rerun the reduction algorithm on \ttp{cex'}.  In
-Figure~\ref{fig:reduce0}, we show an updated reduction algorithm, \ttp{reduce0}
+Figure~\ref{fig:reduce0}, we show an updated reduction algorithm, \ttp{reduceOpt}
 that implements this optimization.  The function \ttp{testHole} tests the
@@
 extrapolate :: SubTypes a
-            => a -> (a -> Property) -> IO [Idx]
-extrapolate cex prop = extrapolate' 1 []
+  => ScArgs -> a -> (a -> Property) -> IO [Idx]
+extrapolate args cex prop = extrapolate' 1 []
   where
@@
     = extrapolate' (idx+1) idxs
-    | otherwise
-    = case index cex idx of
-        Nothing -> return idxs
-        Just v  -> do
-          vs <- newVals v scMaxSize scMaxExtrap
+    | Just v <- index cex idx = mkNewVals v
+    | otherwise = return idxs
+    where
+    mkNewVals v = do
+      vs <- newVals (scMaxSize args)
+                    (scMaxExtrap args) v
       extrapolate' (idx+1)
-            (if allFail cex idx vs prop
+        (if allFail args cex idx vs prop
            then idx:idxs else idxs)
 
-allFail :: SubTypes a
-        => a -> Idx -> [SubVal]
-        -> (a -> Property) -> Bool
-allFail cex idx vs prop =
-  length res >= scMinExtrap && and res
+allFail :: SubTypes a => ScArgs -> a -> Idx
+  -> [SubVal] -> (a -> Property) -> Bool
+allFail args cex idx vs prop =
+  length res >= scMinExtrap args && and res
   where
@@
 
-\subsection{Automated Precondition Strengthening}\label{sec:precondition}
-The extrapolation algorithm generalizes a counterexample, but in the
-``neighborhood'' of he original counterexample.  In particular, all
-generalizations are from the same variant as the original counterexample.  To
-help the programmer in the generalization step, we would also like a way to test
-the property again, ensuring we get counterexamples (if they exist) outside of
-the neighborhood of the original one.
-
-\begin{figure}[ht!]
-  \begin{center}
-    \includegraphics[scale=0.49]{Figs/cex-gen}
-   \end{center}
-  \caption{Counterexample generalization.}
-  \label{fig:cex-gen}
-\end{figure}
-
-Figure~\ref{fig:cex-gen} illustrates the situation in which we have some values
-and a property in the case that some values fail the property of the form
-(\ttp{pre ==> prop}).  Our goal is to find additional counterexamples in the
-shaded region, in which \ttp{pre} is satisfied, \ttp{prop} is not satisfied, and
-the counterexample is not already characterized by a universal sub-value
-generalization formula.  Iterating, we increasingly cover the shaded region
-with new formulas until no more uncovered counterexamples exist or the testing
-tool fails to find new counterexamples.
-
-\begin{figure}
-  \begin{code}
-matchesShapes :: SubTypes a => a -> [(a,([Idx],[Idx]))] -> Bool
-matchesShapes d = any (matchesShape d)
-
-matchesShape :: SubTypes a => a -> (a, ([Idx],[Idx])) -> Bool
-matchesShape a (b, (idxVals, idxConstrs))
-  | constr a /= constr b = False
-  | Just a' <- aRepl         = let x = subForest (subVals a') in
-                               let y = subForest (subVals b) in
-                               all foldEqConstrs (zip x y)
-  | otherwise                = False
-
-  where
-  foldEqConstrs :: (Tree SubVal, Tree SubVal) -> Bool
-  foldEqConstrs (Node (SubVal l0) sts0, Node (SubVal l1) sts1)
-    | constr l0 == constr l1 = next
-    | otherwise              = False
-    where next = all foldEqConstrs (zip sts0 sts1)
-
-  updateA idx d = fmap (replace d idx) (index b idx)
-  aRepl = foldl go (Just a) (idxVals ++ idxConstrs)
-    where go ma idx = maybe Nothing (updateA idx) ma
-  \end{code}
-  \caption{Shape matching algorithm.}
-  \label{fig:matches}
-\end{figure}
-
-\todo{talk about matches shape algorithm, real-eval-print}
-
 \subsection{Existential Sub-Value Generalization}
@@
 
-sumTest :: SubTypes a
-          => a -> (a -> Property) -> [Idx] -> IO [Idx]
-sumTest cex prop exIdxs = sumTest' 1 []
+sumTest :: SubTypes a => ScArgs -> a
+  -> (a -> Property) -> [Idx] -> IO [Idx]
+sumTest args cex prop exIdxs = sumTest' 1 []
   where
@@
     = sumTest' (idx+1) idxs
-    | otherwise
-    = case index cex idx of
-        Nothing -> return idxs
-        Just v  -> do
-          vs <- newVals v scMaxSize scConstrMax
+    | Just v <- index cex idx = fromSumTest v
+    | otherwise = return idxs
+    where
+    fromSumTest v = do
+      vs <- newVals (scMaxSize args)
+              (scConstrMax args) v
       sumTest' (idx+1)
@@
   constrFail' cons vs'
-    | length cons == length allCons
-    = True
-    | null vs'
-    = False
+    | length cons == length allCons = True
+    | null vs'                      = False
     | go v == Just True
@@
     = constrFail' cons (tail vs')
-
     where
@@
 
+\subsection{Automated Precondition Strengthening}\label{sec:precondition}
+The extrapolation algorithm generalizes a counterexample, but in the
+``neighborhood'' of he original counterexample.  In particular, all
+generalizations are from the same variant as the original counterexample.  To
+help the programmer in the generalization step, we would also like a way to test
+the property again, ensuring we get counterexamples (if they exist) outside of
+the neighborhood of the original one.
+
+\begin{figure}[ht!]
+  \begin{center}
+    \includegraphics[scale=0.49]{Figs/cex-gen}
+   \end{center}
+  \caption{Counterexample generalization.}
+  \label{fig:cex-gen}
+\end{figure}
+
+Figure~\ref{fig:cex-gen} illustrates the situation in which we have some values
+and a property in the case that some values fail the property of the form
+(\ttp{pre ==> prop}).  Our goal is to find additional counterexamples in the
+shaded region, in which \ttp{pre} is satisfied, \ttp{prop} is not satisfied, and
+the counterexample is not already characterized by a universal sub-value
+generalization formula.  Iterating, we increasingly cover the shaded region
+with new formulas until no more uncovered counterexamples exist or the testing
+tool fails to find new counterexamples.
+
+\begin{figure}
+  \begin{code}
+matchesShapes :: SubTypes a
+  => a -> [(a,[Idx])] -> Bool
+matchesShapes d = any (matchesShape d)
+
+matchesShape :: SubTypes a
+  => a -> (a, [Idx]) -> Bool
+matchesShape a (b, idxs)
+  | constr a /= constr b = False
+  | Just a' <- aRepl
+  = let x = subForest (subVals a') in
+    let y = subForest (subVals b)  in
+    all foldEqConstrs (zip x y)
+  | otherwise = False
+  where
+  updateA idx d =
+    fmap (replace d idx) (index b idx)
+  aRepl = foldl go (Just a) idxs where
+    go ma idx | Just x <- ma = updateA idx x
+              | otherwise    = Nothing
+  foldEqConstrs ( Node (SubVal l0) sts0
+                , Node (SubVal l1) sts1 )
+    | constr l0 == constr l1 = next
+    | otherwise              = False
+    where next = all foldEqConstrs (zip sts0 sts1)
+  \end{code}
+  \caption{Shape matching algorithm.}
+  \label{fig:matches}
+\end{figure}
+
+\todo{talk about matches shape algorithm, real-eval-print}
+
 \subsection{Generalization Details}
@@
 
+\todo{Instances for \ttp{shrink} are user-defined for complex data types (although a motivation for
+one of Haskell's first generics package was to provide generic
+shrinking~\cite{syb}).}
+
 \todo{avoid complexities \url{http://stackoverflow.com/questions/14006005/idiomatic-way-to-shrink-a-record-in-quickcheck}%
@@
 
-\todo{how to display counterexamples}
-\todo{Use Bayes' rule to do extrapolation?}
-
-\todo{future work: higher-kinded data, functions}
-
-\todo{add lazy testing---proof value is unused}
+We have presented new approaches for generically shrinking and generalizing
+counterexamples over algebraic data.  Shrinking has thus far been as much an art
+as a science, requiring experience with QuickCheck and deep understanding about
+the property being tested.  In many cases, SmartCheck automates the laborious
+task of shrinking.
+
+We envision a number of potential extensions and improvements to SmartCheck.
+First, we have considered only the simplest kind of data, algebraic data types,
+but we would like to extend these results to shrinking functions, too.  As noted
+in Section~\ref{sec:implementation}, SmartCheck does not work with GADTs
+currently, due to limitations with GHC Generics.
+
+Lazy SmallCheck can test partially-defined inputs by detecting the evaluation of
+undefined values~\cite{sc}.  This capability is useful in shrinking, too.  For
+example, the universal sub-value generalization algorithm
+(Section~\ref{sec:universal}) could be extended to shortcut testing and
+generalize a sub-value if it is not evaluated in testing the property.  Not only
+does this shortcut the generalization phase, but it gives a proof that the
+sub-value can be generalized.
+
+SmartCheck displays (generalized) counterexamples in a form similar to default
+\ttp{Show} instances or in a tree form, which can be helpful to parse the
+components of the value.  Better approaches for showing large data types are
+needed.
+
+Finally, we believe there are interesting connections between counterexample
+discovery and probabilistic reasoning.  Often, when a bug in a program is fixed,
+the fix is not correct, or it introduces new bugs.  Such knowledge might be
+useful in guiding the distribution of test cases or shrinking.
+
+\todo{final summary}
 

### 2013-02-10 examples/Div0.hs

@@ -38,9 +38,9 @@ instance Arbitrary Exp where
                   , liftM2 Div mkM' mkM' ]
       where mkM' = mkM =<< choose (0,n-1)
 
-  shrink (C i)       = map C (shrink i)
-  shrink (Add e0 e1) = [e0, e1]
-  shrink (Div e0 e1) = [e0, e1]
+  -- shrink (C i)       = map C (shrink i)
+  -- shrink (Add e0 e1) = [e0, e1]
+  -- shrink (Div e0 e1) = [e0, e1]
 
 -- property: so long as 0 isn't in the divisor, we won't try to divide by 0.
 -- It's false: something might evaluate to 0 still.

### 2013-02-10 paper/Example/InitEx2.hs

@@ -84,7 +84,14 @@ instance Arbitrary T where
 
 #if defined(qcNone) || defined(sc) || defined(feat)
   shrink _ = []
-#else
+#endif
+#ifdef qcjh
+  shrink (T w i0 i1 i2 i3) =
+    let xs = shrink (w, i0, i1, i2, i3) in
+    let go (w', i0', i1', i2', i3') = T w' i0' i1' i2' i3' in
+    map go xs
+#endif
+#if defined(qc10) || defined(qc20)
   shrink (T w i0 i1 i2 i3) =
     [ T a b c d e | a <- tk w
                   , b <- tk i0, c <- tk i1
@@ -135,18 +142,21 @@ test rnds run = do
   let app str = appendFile logFile (str ++ "\n")
   let avg vals = sum vals / fromIntegral rnds'
   let med vals = sort vals !! (rnds' `div` 2)
+  let stdDev vals = sqrt (avg distances)
+        where
+        distances = map (\x -> (x - m)^2) vals
+        m = avg vals
 
   app "***************"
   print res
   let times = fst $ unzip res'
   let szs :: [Double]
       szs = map fromIntegral (snd $ unzip res')
-  app $ "Num  : " ++ show rnds'
-  app $ "Max  : " ++ show (maximum times) ++ ", " ++ show (maximum szs)
-  app $ "Avg  : " ++ show (avg times) ++ ", " ++ show (avg szs)
-  app $ "Med  : " ++ show (med times) ++ ", " ++ show (med szs)
-  -- app $ "Size : " ++ show (fromIntegral (sum szs) /
-  --                             fromIntegral rnds' :: Double)
+  app $ "Num     : " ++ show rnds'
+  app $ "std dev : " ++ show (stdDev $ map (fromRational . toRational) times)
+                     ++ ", " ++ show (stdDev szs)
+  app $ "Avg     : " ++ show (avg times) ++ ", " ++ show (avg szs)
+  app $ "Med     : " ++ show (med times) ++ ", " ++ show (med szs)
 
   app ""
   app ""
@@ -189,7 +199,7 @@ main = do
 #ifdef sc
   test rnds runSC
 #endif
-#if defined(qcNone) || defined(qc10) || defined(qc20)
+#if defined(qcNone) || defined(qc10) || defined(qc20) || defined(qcjh)
   test rnds (runQC stdArgs prop)
 #endif
 

### 2013-02-10 paper/Paper.hs

@@ -62,7 +62,7 @@ class Arbitrary a => SubTypes a where
   subVals  :: a -> Tree SubVal
   constr   :: a -> String
   constrs  :: a -> [String]
---  baseType :: a -> Bool
+  baseType :: a -> Bool
 
 --------------------------------------------------------------------------------
 
@@ -218,13 +218,12 @@ matchesShapes :: SubTypes a
   => a -> [(a,[Idx])] -> Bool
 matchesShapes d = any (matchesShape d)
 
---------------------------------------------------------------------------------
-
 -- | At each index that we generalize (either value generalization or
 -- constructor generalization), we replace that value from b into a.  At this
 -- point, we check for constructor equality between the two values, decending
 -- their structures.
-matchesShape :: SubTypes a => a -> (a, [Idx]) -> Bool
+matchesShape :: SubTypes a
+  => a -> (a, [Idx]) -> Bool
 matchesShape a (b, idxs)
   | constr a /= constr b = False
   | Just a' <- aRepl

### 2013-02-21 examples/Div0.hs
renaming of arguments.

@@ -73,7 +73,7 @@ divTest = smartCheck args prop_div
                                 -- { maxSuccess = 1000
                                 -- , maxSize    = 20  }
                    , format  = PrintString
-                   , extrap  = True
+                   , runForall  = True
                    }
 
 -- Get the minimal offending sub-value.

### 2013-02-21 examples/Large.hs

@@ -65,7 +65,7 @@ prop_p :: B -> Property
 prop_p p = pre p ==> test p
 
 main :: IO ()
-main = smartCheck scStdArgs { extrap = False, constrGen = False } prop_p
+main = smartCheck scStdArgs { runForall = False, constrGen = False } prop_p
   -- smartCheck scStdArgs { format = PrintString 
   --                      , scMaxDepth = Just 20
   --                      } 

### 2013-02-21 src/Test/SmartCheck.hs

@@ -50,7 +50,7 @@ smartCheck args propT = do
       -- Run the smart reduction algorithm.
       d   <- smartRun args r prop'
       -- If we asked to extrapolate values, do so.
-      valIdxs <- if extrap args
+      valIdxs <- if runForall args
                    then -- Extrapolate with the original property to see if we
                         -- get a previously-visited value back.
                         extrapolate args d prop
@@ -58,7 +58,7 @@ smartCheck args propT = do
 
       -- If we asked to extrapolate constructors, do so, again with the original
       -- property.
-      csIdxs <- if constrGen args
+      csIdxs <- if runExists args
                   then constrsGen args d prop valIdxs
                   else return []
 

### 2013-02-21 src/Test/SmartCheck/Args.hs

@@ -38,21 +38,21 @@ data ScArgs =
                                      --   for failure in the reduction stage.
                                      --------------
          -- Extrapolation
-         , extrap       :: Bool      -- ^ Should we extrapolate?
+         , runForall    :: Bool      -- ^ Should we extrapolate?
                                      --------------
-         , scMaxExtrap  :: Int       -- ^ How hard (number of rounds) to look
+         , scMaxForall  :: Int       -- ^ How hard (number of rounds) to look
                                      --   for failures during the extrapolation
                                      --   and constructor generalization stages.
                                      --------------
-         , scMinExtrap  :: Int       -- ^ Minimum number of times a property's
+         , scMinForall  :: Int       -- ^ Minimum number of times a property's
                                      -- precondition must be passed to
                                      -- generalize it.
                                      --------------
          -- Constructor generalization
-         , constrGen    :: Bool      -- ^ Should we try to generalize
+         , runExists    :: Bool      -- ^ Should we try to generalize
                                      --   constructors?
                                      --------------
-         , scConstrMax    :: Int     -- ^ How hard (number of rounds) to look
+         , scMaxExists  :: Int     -- ^ How hard (number of rounds) to look
                                      -- for failing values with each
                                      -- constructor.  For "wide" sum types, this
                                      -- value should be increased.
@@ -70,12 +70,12 @@ scStdArgs = ScArgs { format       = PrintTree
                    ---------------------
                    , scMaxReduce  = 100
                    ---------------------
-                   , extrap       = True
-                   , scMaxExtrap  = 20
-                   , scMinExtrap  = 10
+                   , runForall    = True
+                   , scMaxForall  = 20
+                   , scMinForall  = 10
                    ---------------------
-                   , constrGen    = True
-                   , scConstrMax  = 20
+                   , runExists    = True
+                   , scMaxExists  = 20
                    }
 
 --------------------------------------------------------------------------------

### 2013-02-21 src/Test/SmartCheck/ConstructorGen.hs

@@ -86,7 +86,7 @@ arbSubset :: (SubTypes a, Generic a, ConNames (Rep a))
           -> S.Set String -> IO (Result a)
 arbSubset args a idx prop constrs =
   liftM snd $ iterateArbIdx a (idx, scMaxDepth args)
-                (scConstrMax args) (scMaxSize args) prop'
+                (scMaxExists args) (scMaxSize args) prop'
   where
   prop' b = newConstr b Q.==> prop b
   -- Make sure b's constructor is a new one.

### 2013-02-21 src/Test/SmartCheck/Extrapolate.hs

@@ -43,7 +43,7 @@ extrapolate args d origProp = do
   -- least one test passes a precondition, and for every test in which the
   -- precondition is passed, it fails.  We test values of all possible sizes, up
   -- to Q.maxSize.
-  test _ idx = iterateArbIdx d (idx, scMaxDepth args) (scMaxExtrap args)
+  test _ idx = iterateArbIdx d (idx, scMaxDepth args) (scMaxForall args)
                  (scMaxSize args) origProp
 
   -- Control-flow.
@@ -51,7 +51,7 @@ extrapolate args d origProp = do
   -- None of the tries satisfy prop (but something passed the precondition).
   -- Prevent recurring down this tree, since we can generalize.
   next _ (i, FailedProp) forest' idx idxs
-    | scMinExtrap args < i =
+    | scMinForall args < i =
         nextIter (forestReplaceChildren forest' idx False) idx (idx : idxs)
   next _ _ forest' idx idxs = nextIter forest' idx idxs
 

### 2013-02-21 paper/paper.tex
paper updates.

@@
 user to write \emph{shrink} functions to attempt to reduce the size of
-counterexamples discovered by QuickCheck.  Implementations of shrink often
-require significant boilerplate and usually just generate smaller tests than the
-discovered counterexample.  Therefore, shrinking can be inefficient and poor at
-discovering new, smaller counterexamples.  In this paper, we introduce
+counterexamples discovered.  Hand-written implementations of shrink can be
+difficult to write: they can complex and consist of significant boilerplate
+code.  Without a good understanding of QuickCheck and the property being tested,
+they can be very inefficient, generating poor new candidates, and they may fail
+to shrink a counterexample significantly.  In this paper, we introduce
 \emph{SmartCheck}, an automatic efficient counterexample reduction algorithm.
 SmartCheck reduces algebraic data using generic search heuristics to efficiently
-find smaller counterexamples.  In addition to shrinking, SmartCheck generalizes
-counterexamples by returning a formula that abstracts over individual
-counterexamples and make it easier for the programmer to uncover multiple
-counterexamples at once.  SmartCheck has been implemented for Haskell and is
-freely available.
+find smaller counterexamples independent of the program or property being
+tested.  In addition to shrinking, SmartCheck generalizes counterexamples by
+returning a formula that abstracts over individual counterexamples and make it
+easier for the programmer to uncover multiple counterexamples at once.
+SmartCheck has been implemented for Haskell and is freely available.
 \end{abstract}
@@
 
-\todo{complexity of each algorithm}
-\todo{termination for each algorithm}
-\todo{show quantification in intro}
 \todo{benchmark feat -- send to author}
@@
 
-In industrial uses, shrinking is essential.  Hughes, the inventor of QuickCheck
+In industrial uses, shrinking is essential.  Hughes, an inventor of QuickCheck
 and the CEO of Quiviq, a company that commercializes the technology, has noted
@@
 
-Typically, effective shrinking on complex data requires a good understanding of
+Typically, effectively shrinking complex data requires a good understanding of
 how shrinking in QuickCheck works and the semantics of the property and program
 being evaluated.  Bad definitions can be so inefficient that they are unusable.
-Shrinking presents a programmer's dilemma: shrinking is used to better
-understand a counterexample to understand why a property failed, but to
-implement an efficient shrink function in general requires knowing how to
-generate counterexamples to the property.
+Shrinking presents a programmer's paradox: shrinking is used to generate a
+smaller counterexample to better understand why a property failed, but to
+implement an efficient shrink function often requires a deep enough
+understanding of the property to efficiently generate smaller counterexamples to
+it!
 
@@
 To illustrate, consider the example in Figure~\ref{fig:initial}.\footnote{All
-  examples and algorithms in this paper are presented in Haskell.  We strive to
-  use idiomatic Haskell, and only rely on the language extensions of existential
-  type quantification and scoped type variables.~\cite{haskell98}.}  Data type
-\ttp{T} is a product type containing four lists of \ttp{Int16}s and a single
-\ttp{Word8}.  For our purposes, it is not important what \ttp{T} models, but for
-concreteness, suppose it models the values in four linked lists in an operating
-system, together with an 8-bit status register.
+  examples and algorithms in this paper are presented in
+  Haskell~\cite{haskell98}.  We strive to use idiomatic Haskell, and only rely
+  on the language extensions of existential type quantification and scoped type
+  variables.}  Data type \ttp{T} is a product type containing four lists of
+\ttp{Int16}s and a single \ttp{Word8}.  For our purposes, it is not important
+what \ttp{T} models, but for concreteness, suppose it models the values in four
+linked lists in an operating system, together with an 8-bit status register.
 
@@
 \noindent
-fails \ttp{prop} (\ttp{==>} is implication from the QuickCheck library).
+satisfies \ttp{pre} but fails \ttp{post} (\ttp{==>} is implication from the
+QuickCheck library).
 
 Despite the simplicity of the example, a typical counterexample returned by
-QuickCheck is not quite so simple.  With standard settings, the median
+QuickCheck can be large.  With standard settings and no shrinking, the median
 counterexample discovered contains around 70 \ttp{N} constructors, and
-counterexamples can contain over 100 constructors.  Thus, it pays to define
+counterexamples can contain over 100 \ttp{N} constructors.  Thus, it pays to define
 \ttp{shrink}!
 
-We might first naively try to shrink counterexamples for a data type like \ttp{T}
-by taking the cross-product of shrunk values over the arguments to the
+We might first naively try to shrink counterexamples for a data type like
+\ttp{T} by taking the cross-product of shrunk values over the arguments to the
 constructor \ttp{T}.  This can be expressed using Haskell's list-comprehension
@@
 \begin{code}
-shrink (T w i0 i1 i2 i3) =
-  let xs = shrink (w, i0, i1, i2, i3) in
-  let go (w', i0', i1', i2', i3') =
-    T w' i0' i1' i2' i3' in
-  map go xs
+shrink (T w i0 i1 i2 i3) = map go xs
+  where xs = shrink (w, i0, i1, i2, i3)
+        go (w', i0', i1', i2', i3') =
+          T w' i0' i1' i2' i3'
 \end{code}
@@
 attempting to shrink \ttp{a} while holding \ttp{b} constant, then attempting to
-shrink \ttp{b} while holding \ttp{a} constant, and the approach is generalized
-to larger tuples.  Still, with this definition, in some cases, shrinking takes
-over 20 seconds and returns a result containing over 60
-constructors.\footnote{Results are with a GHC-compiled program, using \ttp{-O2},
-  on a 4-core 2.7GHz Intel~i7 processor.}
+shrink \ttp{b} while holding \ttp{a} constant (and is generalized to larger
+tuples).  Still, using this definition, from some initial counterexamples,
+shrinking may take over 15 seconds and result in a final counterexample
+containing nearly 60 \ttp{N} constructors.\footnote{All results reported in the paper
+  are with a GHC-compiled program, using \ttp{-O2}, on a 4-core 2.7GHz Intel~i7
+  processor.}  We would like to do better, particularly since the data type and
+property are so simple.
 
@@
 
-With SmallCheck, a property is checked up to a depth set by the user that
-measures how large values are (typically, the depth is a number of constructors
-in the value).  For a product type, SmallCheck must check values produced by
-taking the cross-product of the type's fields at a given depth.  SmallCheck is
-unable to enumerate all possible values of type \ttp{T} to discover a
-counterexample, due to combinatorial explosion.  (A related library named Feat
-combines some aspects of SmallCheck and QuickCheck~\cite{feat}; we discuss it
-more in the related work (Section~\ref{sec:related}).
+For a product type, SmallCheck must check values produced by taking the
+cross-product of the type's fields at a given depth.  Unfortunately, SmallCheck
+fails long before it reaches a depth required to discover a counterexample, due
+to combinatorial explosion.  (A related library named Feat combines some aspects
+of SmallCheck and QuickCheck~\cite{feat}; we discuss it more in the related work
+(Section~\ref{sec:related}).
 
@@
 \noindent
-Such a formula has the added benefit of replacing a potentially large sub-term
-of a composite data type value with a quantifier, reducing the sheer size of the term.
+denoting that for any \ttp{Word8}, the counterexample holds.  Such a formula has
+the added benefit of replacing a potentially large sub-term of a composite data
+type value with a quantifier, reducing the sheer size of the term.
 
   \item Second, during testing, it is sometimes helpful to cover the input space
-    and by discovering different counterexamples to \ttp{prop}.  In the example
-    in Figure~\ref{fig:initial}, one counterexample is fairly representative of
+    by discovering different counterexamples to \ttp{prop}.  In the example in
+    Figure~\ref{fig:initial}, one counterexample is fairly representative of
     others, but this is not true in general (e.g., consider properties about
@@
 \hline \hline
-Maximum  & 0.40s, 124 & 139.15s, 90  & 97.53s, 53 & 0.s, 13\\
+Mean       & 0.08s, 69  & 0.53s, 35  & 2.24s, 12  & 0.10s, 7\\
 \hline
-Mean     & 0.08s, 69  & 0.73s, 35  & 2.26s, 13  & 0.07s, 6\\
+Median     & 0.07s, 69  & 0.17s, 33  & 1.81s, 13  & 0.10s, 6\\
+%% \hline
+%% MAD        & 0.02s, 13  & 0.09s, 8   & 0.62s, 3   & 0.02s, 2\\
 \hline
-Median   & 0.07s, 68  & 0.18s, 32  & 1.90s, 13  & 0.07s, 5\\
+95\%       & 0.14s, 100 & 1.76s, 65  & 4.06s, 19  & 0.18s, 12\\
 \hline
@@
   \end{center}
-  \caption{Results for 1000 tests of the program in Figure~\ref{fig:initial}
-    using interpreted Haskell.}
+  \caption{Results for 1000 tests of the program in Figure~\ref{fig:initial}.}
   \label{table:results}
 \end{table}
-\todo{table generated with InitEx2 on Feb. 3 --- do a final test with 1000}
+
+\todo{get 99\% }
 
@@
 SmartCheck on the program shown in Figure~\ref{fig:initial} (we omit SmallCheck,
-since as noted, it is not even feasible for the example).  We show the maximum,
-mean, and median amount of time required to generate a counterexample and to
-shrink it (and print it to standard out) as well as the corresponding number of
-constructors in the final result.
+since as noted, it is not even feasible for the example).  Each entry is a pair
+containing the number of seconds required for counterexample discovery and
+shrinking and the size of the final value obtained (by counting \ttp{N}
+constructors).  We show the mean, median, median, and the values at the 95th
+percentile values.  (The curves produced by these values are not normal, so we
+omit standard the deviation.)
 
@@
 result of using QuickCheck with truncating shrink, which each list is truncated
-to 10 elements.  Next, we show the results of the ``tuple shrink'' instance.  In
-the last column are the results for SmartCheck, which includes the time required
-by QuickCheck to generate the original counterexample.  Not only does SmartCheck
-execute faster than the other strategies, it has smaller outliers.  In addition,
-the size of the counterexamples are significantly smaller.  And we stress that
-SmartCheck achieves these results \emph{generically}.
-
-\todo{change max to std dev, look for max info above.}
+to 10 elements.  Compared with no shrinking, this approximately halves the size
+of the median counterexample, but doubles the time taken.  There can be large
+outliers, however, in the time taken during shrinking.  Next, we show the
+results of the ``tuple shrink'' instance.  While the median amount of time
+required is much higher, the size of the resulting value is approximately halved
+again.  The reason for the increase in time is because shrinking is very
+productive and produces a nonempty list for most values.
+
+In the last column are the results for SmartCheck, which includes the time
+required by QuickCheck to generate the original counterexample.  The median time
+required by SmartCheck to shrink is 0.02 seconds (SmartCheck's execution time
+minus the execution time for QuickCheck without shrinking), and the median size
+of a resulting value is six \ttp{N} constructors.  Moreover, SmartCheck has a
+very small variance---note the mean and median are nearly identical, meaning
+there are very few outliers.  And we stress that SmartCheck achieves these
+results \emph{generically}, requiring no shrink definitions to be provided by
+the user.
 
@@
 To present the algorithms in the following sections, we provide some definitions
-in the form of methods of a type class.  SmartCheck requires instances for data
-being analyzed.  In Section~\ref{sec:implementation}, we describe how instances
-for the type class are derived automatically.  Consider the following types and
-type class.
+in the form of methods of a \ttp{SubTypes} type class.  SmartCheck requires
+instances for data being analyzed.  In Section~\ref{sec:implementation}, we
+describe how instances for the type class are derived automatically.
 %
@@
 \item \ttp{baseType} returns whether the type of the value is an ``interesting
-  type'', which generally means it has structure (e.g., a sum or product type).
+  type'', which generally means it is user-defined and has positive arity.
   We discuss the purpose of base types in Section~\ref{sec:base}.
@@
   \end{code}
+%
 \end{itemize}
@@
 \noindent
-For example, consider a binary tree type:
+To demonstrate typical evaluations of the methods, consider a binary tree type:
 %
@@
 Here are example applications of \ttp{SubTypes} methods; in the following, we
-show the indexes with respect to the \ttp{tree} value:
+show the indexes with respect to the value \ttp{tree}:
 %
@@
 baseType (3 :: Int) = True
+baseType tree       = False
 baseType L          = False
@@
 Figure~\ref{fig:reduction} shows the reduction algorithm.  The function
-\ttp{reduce} takes arguments to customize the algorithm's behavior, a
-counterexample \ttp{cex} and the property \ttp{prop}.  At the beginning of
+\ttp{reduce} takes flags to customize the algorithm's behavior, a
+counterexample \ttp{cex}, and the property \ttp{prop}.  At the beginning of
 execution, we assume that a counterexample has been obtained somehow; our
@@
 
-\ttp{sizedArbitrary} returns a value inside the \ttp{IO} monad since it depends
-on a random seed for generation (we could pass in the generator seed, but for
-simplicity, let us assume one is generated locally).  The constant
-\ttp{scMaxReduce} is the maximum number of tries to discover a new
+The flag \ttp{scMaxReduce} is the maximum number of tries to discover a new
 counterexample by replacing \ttp{v} in \ttp{cex} and testing it.  The result of
@@
 counterexample \ttp{cex'} is found, we start a new breadth-first traversal of
-\ttp{cex'}, attempting to shrink it further.  The algorithm is guaranteed to
-terminate: intuitively, the measure for the function is that either the index
-increases or the size of the counterexample being evaluated decreases.  The
-complexity of the algorithm is $\mathcal{O}(n^2)$, where $n$ is the number of
-constructors in the counterexample.  The analysis assumes generating new
-sub-values and testing them is done in constant time.
+\ttp{cex'}, attempting to shrink it further.
+
+The algorithm is guaranteed to terminate: intuitively, the measure for the
+function is that either the index increases or the size of the counterexample
+being evaluated decreases.  The algorithm's complexity is
+$\mathcal{O}(n^2)$, where $n$ is the number of constructors in the
+counterexample, assuming that generating new sub-values and testing them is done
+in constant time.
 
@@
 
-\todo{correspond to dimension 0 from\cite{agata}}
 The number of constructors in an algebraic data value primarily determines how
-large and understandable it is.  In our experience, the size of fields that come
-from primitive types, like \ttp{Int} or \ttp{Char}, have less of a role in
-understanding a counterexample.  However, a testing algorithm might spend
-significant time testing new values in which primitive types are reduces.
+large it is.  In our experience, the size of fields that come from primitive
+types, like \ttp{Int} or \ttp{Char}, have less of a role in understanding a
+counterexample, so in general, we want to prevent our reduction algorithm from
+spending time shrinking values from these types.
 
@@
 directly, and rerun the reduction algorithm on \ttp{cex'}.  In
-Figure~\ref{fig:reduce0}, we show an updated reduction algorithm, \ttp{reduceOpt}
+Figure~\ref{fig:reduce0}, we show an updated reduction algorithm, \ttp{reduceOpt},
 that implements this optimization.  The function \ttp{testHole} tests the
@@
 algorithm.  Thus, new sub-values must be strictly smaller than the minimum of
-\ttp{scMaxSize} and the size of the sub-value being replaced.  (The size of the
-value is with respect to the \ttp{size} method in the \ttp{SubTypes} class.)
-Finally, \ttp{scMaxDepth} determines the maximum depth in the counterexample the
+\ttp{scMaxSize} and the size of the sub-value being replaced.  Finally,
+\ttp{scMaxDepth} determines the maximum depth in the counterexample the
 reduction algorithm should analyze.  A value of \ttp{Nothing} means that the
@@
 depth L         = 0
-depth (B t0 t1) =
-  1 + max (depth t0) (depth t1)
+depth (B t0 t1) = 1 + max (depth t0) (depth t1)
 \end{code}
@@
 counterexample to a set of counterexamples for which the program and property
-fails.  For example, in the introduction, we realized that the specific
-counterexample
-%
-\begin{code}
-B [A (-32769)] [] [] []
-\end{code}
-%
-\noindent
-is representative of the set of counterexamples in which underflow causes the
-sum of the fields to be greater or equal to 64.
-
-The generalization step is an important under-appreciated step in the debugging
-process.  For counterexamples that are large, deeply-nested values, the
-generalization step can be difficult.  One way we might help the programmer is
-to provide her not with a single counterexample but with a quantified formula
-that characterizes a set of counterexamples.  In particular, a formula reduces
-the noise in favor of the signal by abstracting away portions of large
-counterexample that are irrelevant to why it violates the property.
+fails.  The generalization step is an important yet under-appreciated step in
+the debugging process.  A characterizing formula reduces the noise in favor of
+the signal by abstracting away portions of large counterexample that are
+irrelevant to why it violates the property.
 
@@
 \subsection{Universal Sub-Value Generalization}\label{sec:universal}
-An example best explains the first kind of quantified formulas we generate.
-Recall the calculator language from Section~\ref{sec:subval}.  Recall the
-culprit buried within a counterexample, causing it to result in a
-division-by-zero error:
+Recall from the introduction the quantified formula
+%
+\begin{code}
+forall x . T x [N (-20000)] [N (-20000)] [] []
+\end{code}
+%
+\noindent
+characterizing a set of counterexamples.  Another example using the calculator
+language from Section~\ref{sec:subval} is to generalize a counterexample like
 %
@@
 \noindent
-Notice that any value substituted in the dividend results in an error.  We can
-quantify the error then as follows:
+by
 %
@@
 \noindent
-Not only does the generalization assist the programmer's insight, but it reduces
-the sheer size of the counterexample, and it can do so significantly.  We call
-the kind of formula just shown \emph{universal sub-value generalization} and is
-implemented in SmartCheck.
+since any dividend results in divide-by-zero.  Not only do the generalizations
+assist the programmer's insight, but they reduce the sheer size of the
+counterexample.  We call the kind of formula just shown \emph{universal
+  sub-value generalization} and is implemented in SmartCheck.
 
@@
       vs <- newVals (scMaxSize args)
-                    (scMaxExtrap args) v
+                    (scMaxForall args) v
       extrapolate' (idx+1)
@@
 allFail args cex idx vs prop =
-  length res >= scMinExtrap args && and res
+  length res >= scMinForall args && and res
   where
@@
 \end{figure}
+
+\todo{explain scMaxForall}
 
@@
 is true if in \ttp{cex}, the value at index \ttp{idx0}, is a child of index
-\ttp{idx1}, in a tree representation of the data.  This prevents the algorithm
-from trying to generalize sub-values that are abstracted away already since
-their parents have been generalized.
+\ttp{idx1}, in a tree representation of \ttp{cex} (i.e., \ttp{subVals cex}).  This
+prevents the algorithm from trying to generalize sub-values that are abstracted
+away already since their parents have been generalized.
 
@@
 Figure~\ref{fig:reduction}.  Here, we bound the size of values to generate by
-the constant \ttp{scMaxSize}, which is independent of the size of the particular
+the flag \ttp{scMaxSize}, which is independent of the size of the particular
 sub-value.
@@
 \noindent
-is the roughly the dual of \ttp{pass} in the reduction algorithm: (\ttp{fail
+is roughly the dual of \ttp{pass} in the reduction algorithm: (\ttp{fail
   prop cex}) returns (\ttp{Just True}) if \ttp{cex} passes \ttp{prop}'s
@@
 non-trivially satisfies \ttp{prop}, and \ttp{Nothing} if \ttp{cex} fails
-\ttp{prop}'s precondition.  The constant \ttp{scMinExtrap} is the minimum number
+\ttp{prop}'s precondition.  The flag \ttp{scMinForall} is the minimum number
 of \ttp{Just False} results required from \ttp{fail} to extrapolate from failed
@@
 tuning the parameters, the risk of an unsound generalization is reduced.
+
+The algorithm's complexity is $\mathcal{O}(n)$, where $n$ is the number of
+constructors in the counterexample.  Again, we assume that the cost for each
+generating random values and testing them at each index is constant.
 
@@
 Because there are only three variants in the type, we have found counterexamples
-built from each of them in the divisor.  We therefore return a formula like the
-following:
+built from each of them in the divisor.  We therefore can claim the following
+formula holds:
 %
@@
 forall values x .
-  forall constructors C\(\sb{0}\) .
+  forall constructors C .
     there exist arguments \(\stackrel{\rightarrow}{y}\) .
-      such that Div x (C\(\sb{0} \stackrel{\rightarrow}{y}\))
-\end{code}
-%
+      such that Div x (C\(\stackrel{\rightarrow}{y}\))
+\end{code}
+%
+\noindent
+We therefore present an existential sub-value generalization algorithm that
+performs constructor generalization.  Like with the other algorithms, this
+algorithm also performs a breadth-first search over a counterexample.
 
@@
       vs <- newVals (scMaxSize args)
-              (scConstrMax args) v
+              (scMaxExists args) v
       sumTest' (idx+1)
@@
 
-The function \ttp{sumTest} takes a counterexample, a property, and a list of
-indexes that have already been generalized---perhaps by the extrapolation
-algorithm in Figure~\ref{fig:universal}.  The list of course may be empty if no
-sub-values have been previously extrapolated.  Like with the other algorithms,
-this algorithm also performs a breadth-first search over a counterexample.  In a
-call to \ttp{subTrees}, discussed in Section~\ref{sec:universal}, the guard
-prevents constructor generalization if the current index is a sub-value of a
-previously generalized value.  Otherwise, a list of well-typed new values are
-generated by a call to \ttp{newVals}, as shown in Figure~\ref{fig:reduction}.
-In the arguments to \ttp{newVals}, we bound the size of values generated with
-\ttp{scMaxSize} as before, and bound the number of values generated with the
-constant \ttp{scConstrMax}.  Because values are randomly generated, for ``wide''
-sum-types (i.e., with a large number of constructors), \ttp{scConstrMax} should
-be large enough to ensure with high probability that each variant is generated.
+We show the algorithm in Figure~\ref{fig:constrs}.  The function \ttp{sumTest}
+takes a set of flags, a counterexample, a property, and a list of indexes that
+have already been generalized---perhaps by the extrapolation algorithm in
+Figure~\ref{fig:universal}.  The list of course may be empty if no sub-values
+have been previously extrapolated.  In a call to \ttp{subTrees}, discussed in
+Section~\ref{sec:universal}, the guard prevents constructor generalization if
+the current index is a sub-value of a previously generalized value.  Otherwise,
+a list of well-typed new values are generated by a call to \ttp{newVals}, as
+shown in Figure~\ref{fig:reduction}.  In the arguments to \ttp{newVals}, we
+bound the size of values generated with \ttp{scMaxSize} as before, and bound the
+number of values generated with the flag \ttp{scMaxExists}.  Because values are
+randomly generated, for ``wide'' sum-types (i.e., with a large number of
+constructors), \ttp{scMaxExists} should be large enough to ensure with high
+probability that each variant is generated.
 
@@
 current constructor is not an element of constructors already seen.  Thus,
-\ttp{go v} returns \ttp{Just True} if \ttp{replace cex idx v} passes this
+(\ttp{go v}) returns (\ttp{Just True}) if (\ttp{replace cex idx v}) passes this
 precondition (and any other preconditions of \ttp{prop}), but fails the
@@
 least one counterexample, so inductive reasoning is not required.
+
+This algorithm's complexity is also $\mathcal{O}(n)$, where $n$ is the number of
+constructors in the counterexample.
 
@@
                 , Node (SubVal l1) sts1 )
-    | constr l0 == constr l1 = next
+    | constr l0 == constr l1 =
+      all foldEqConstrs (zip sts0 sts1)
     | otherwise              = False
-    where next = all foldEqConstrs (zip sts0 sts1)
   \end{code}
@@
 
-\todo{talk about matches shape algorithm, real-eval-print}
-
-\subsection{Generalization Details}
-
-\todo{flags, optimizations}
+Figure~\ref{fig:matches} show the shape-matching algorithm used for precondition
+strengthening.  The function takes a candidate counterexample and a list of
+previous counterexamples, together with the indexes at which they have been
+generalized.  The basic idea of the algorithm is to determine whether two values
+have the same ``shape''.  Intuitively, we consider two values to have the same
+shape if in their respective tree representations, their constructors at each
+node in the tree match, ignoring all base types (Section~\ref{sec:base}).
+That is, intuitively, for values \ttp{a} and \ttp{b},
+%
+\begin{code}
+subVals a == subVals b
+\end{code}
+%
+\noindent
+Furthermore, indexes that have been universally generalized (and all their
+children) match any value, since a universally generalized index indicates that
+counterexamples have been found for any possible constructor.  In our
+implementation, we also consider existentially generalized indexes to match any
+value.  Doing so is a design choice that is more aggressive about covering the
+space of counterexamples at the risk of omitting some.
+
+To understand the shape matching algorithm, consider a few values of type
+\ttp{Exp}, defined in Section~\ref{sec:subval}:
+%
+\begin{code}
+e0 = Div (C 1) (C 2)
+e1 = Div (C 1) (C 3)
+e2 = Div (Add (C 1) (C 2)) (C 7)
+e3 = Div (Div (C 8) (C 2)) (C 7)
+\end{code}
+%
+\noindent
+Then the following hold:
+%
+\begin{code}
+matchesShape e0 (e1, [])  == True
+matchesShape e1 (e2, [])  == False
+matchesShape e1 (e2, [0]) == True
+matchesShape e3 (e2, [0]) == True
+\end{code}
+%
+The first equation holds because we ignore base types.  The second equation
+fails because at index 0, \ttp{e1} contains the constructor \ttp{C} and \ttp{e2}
+contains the constructor \ttp{Add}.  However, index 0 has been generalized, then
+we ignore the sub-value at index 0, and they match (the third equation).
+The same reasoning holds for the fourth equation.
 
@@
 
-\todo{shrink paper rehger pointed to ``Automatic isolation of compiler errors''
-  from 1994}
-\todo{\url{http://www.cs.utah.edu/~regehr/papers/pldi12-preprint.pdf}}
-\url{http://blog.regehr.org/archives/697} --- check out C-reduce (Regehr) and Delta%
-  debugging (\url{http://www.st.cs.uni-saarland.de/dd/}).  C-reduce is in PLDI.%
-
-Related to SmartCheck is the recent Haskell library Feat~\cite{feat}.  Feat
-provides automated enumerations of algebraic data types in Haskell, allowing for
-fast access to very large indexes.  For example, from the enumeration of
+Counterexample reduction in general is a relatively unexplored topic, but there
+has been some treatment of the topic.  Zeller and Hildebrandt describe an
+application of greedy search to shrink counterexamples they call
+``delta-debugging'' (DD)~\cite{dd}.  The authors apply their work shrinking HTML
+inputs to crash Mozilla and shrinking C programs to trigger a bug in GCC.
+Subsequent generalizations are reported by Misherghi and Su in which they
+perform greedy search on tree-structured data; they call their approach
+hierarchical delta-debugging (HDD)~\cite{hdd}.
+
+HDD is most similar to SmartCheck's reduction algorithm, with an important
+difference: HDD (and DD) is deterministic, so the algorithm only succeeds in
+reducing the counterexample only if a new counterexample can be constructed from
+the original one.  Our approach combines the speed of delta debugging with the
+power of QuickCheck to \emph{randomly} discover structurally smaller
+counterexamples.  Furthermore, neither paper explores the idea of counterexample
+generalization, however.
+
+Within the functional programming community, one of the few treatments of
+generic shrinking is as a motivation for generic programming in Haskell's
+``Scrap your boilerplate'' generic programming library~\cite{syb}.
+
+Besides QuickCheck and SmallCheck, another testing framework related to
+SmartCheck is the recent Haskell library Feat~\cite{feat}.  Feat provides
+automated enumerations of algebraic data types in Haskell, allowing for fast
+access to very large indexes.  For example, from the enumeration of
 (\ttp{[Bool]})
@@
 
-combines some of the advantages of SmallCheck and QuickCheck:
-
-Feat provides a functional representation of enumerations,
-allowing for fast indexing of large values in an enumeration.
-
-\todo{Instances for \ttp{shrink} are user-defined for complex data types (although a motivation for
-one of Haskell's first generics package was to provide generic
-shrinking~\cite{syb}).}
-
-\todo{avoid complexities \url{http://stackoverflow.com/questions/14006005/idiomatic-way-to-shrink-a-record-in-quickcheck}%
-%
-The SYB paper:%
-  \url{http://research.microsoft.com/en-us/um/people/simonpj/Papers/hmap/index.htm}%
-  Has shrink example%
-%
-* QuickSpec:%
-  \url{http://www.cse.chalmers.se/~nicsma/quickspec.pdf}%
-%
-%
-* Also check out \url{http://www.whyprogramsfail.com/}%
-%
-* GenCheck \url{https://github.com/JacquesCarette/GenCheck}%
-%
-* Dart-check: \url{http://code.google.com/p/dart-immutable/} (and G+ post by Paul%
-  Brauner)%
-}
-%
-\url{http://www.haskell.org/haskellwiki/Research_papers/Testing_and_correctness}
+Feat finds small counterexamples effectively, but a limitation compared to
+QuickCheck is that value generation is applicative rather than monadic.
+Consequently, defining custom generators (i.e., instances for \ttp{arbitrary})
+is highly constrained.  On the other hand, generators can be derived
+automatically with Feat.
+
+We have benchmarked Feat as well on the example in the introduction
+(Figure~\ref{fig:initial}).  While it performs quite competitively to SmartCheck
+in terms the size of counterexamples discovered: the mean and median number of
+\ttp{N} constructors in the counterexamples returned by Feat are both seven.
+However, Feat performs more poorly with respect to how long counterexample
+discovery takes: the mean time taken by Feat is 5.60 seconds, with a long tail:
+five percent of runs take over 14 seconds (recall that the mean time taken by
+SmartCheck is 0.12 seconds, and 99\% of runs take less than 0.30 seconds).
+
+SmartCheck bears some similarity to QuickSpec, a testing-based library that
+infers equational properties about programs~\cite{qs} insofar as they both
+attempt to generalize counterexamples based on specific inputs.
+
+\todo{shrink paper rehger pointed to ``Automatic isolation of compiler errors''
+  from 1994}
+\todo{\url{http://www.cs.utah.edu/~regehr/papers/pldi12-preprint.pdf}}
+
+%% \todo{avoid complexities \url{http://stackoverflow.com/questions/14006005/idiomatic-way-to-shrink-a-record-in-quickcheck}%
+%% %
+%% * Also check out \url{http://www.whyprogramsfail.com/}%
+%% %
+%% \url{http://www.haskell.org/haskellwiki/Research_papers/Testing_and_correctness}%}
+
+%% \todo{Not relevant really: GenCheck
+%%   \url{https://github.com/JacquesCarette/GenCheck}%
+\url{http://blog.regehr.org/archives/697} --- check out C-reduce (Regehr)
+%% }
 
@@
 
+\todo{difficult to compare performance against arbitrary hand-written shrinking}
 \todo{final summary}

### 2013-02-21 paper/Paper.hs

@@ -35,14 +35,14 @@ data ScArgs = ScArgs {
                               -- types.
   , scMaxReduce  :: Int       -- ^ How hard (number of rounds) to look for
                               -- failure in the reduction stage.
-  , extrap       :: Bool      -- ^ Should we extrapolate?
-  , scMaxExtrap  :: Int       -- ^ How hard (number of rounds) to look for
+  , runForall    :: Bool      -- ^ Should we extrapolate?
+  , scMaxForall  :: Int       -- ^ How hard (number of rounds) to look for
                               -- failures during the extrapolation and
                               -- constructor generalization stages.
-  , scMinExtrap  :: Int       -- ^ Minimum number of times a property's
+  , scMinForall  :: Int       -- ^ Minimum number of times a property's
                               -- precondition must be passed to generalize it.
-  , constrGen    :: Bool      -- ^ Should we try to generalize constructors?
-  , scConstrMax  :: Int       -- ^ How hard (number of rounds) to look for
+  , runExists    :: Bool      -- ^ Should we try to generalize constructors?
+  , scMaxExists  :: Int       -- ^ How hard (number of rounds) to look for
                               -- failing values with each constructor.  For
                               -- "wide" sum types, this value should be
                               -- increased.
@@ -155,7 +155,7 @@ extrapolate args cex prop = extrapolate' 1 []
     where
     mkNewVals v = do
       vs <- newVals (scMaxSize args)
-                    (scMaxExtrap args) v
+                    (scMaxForall args) v
       extrapolate' (idx+1)
         (if allFail args cex idx vs prop
            then idx:idxs else idxs)
@@ -163,7 +163,7 @@ extrapolate args cex prop = extrapolate' 1 []
 allFail :: SubTypes a => ScArgs -> a -> Idx
   -> [SubVal] -> (a -> Property) -> Bool
 allFail args cex idx vs prop =
-  length res >= scMinExtrap args && and res
+  length res >= scMinForall args && and res
   where
   res  = mapMaybe go vs
   go   = fail prop . replace cex idx
@@ -188,7 +188,7 @@ sumTest args cex prop exIdxs = sumTest' 1 []
     where
     fromSumTest v = do
       vs <- newVals (scMaxSize args)
-              (scConstrMax args) v
+              (scMaxExists args) v
       sumTest' (idx+1)
         (if constrFail cex idx vs prop
[…]

### 2013-02-21 paper/Example/InitEx2.hs
Example updates.

@@ -86,10 +86,10 @@ instance Arbitrary T where
   shrink _ = []
 #endif
 #ifdef qcjh
-  shrink (T w i0 i1 i2 i3) =
-    let xs = shrink (w, i0, i1, i2, i3) in
-    let go (w', i0', i1', i2', i3') = T w' i0' i1' i2' i3' in
-    map go xs
+  shrink (T w i0 i1 i2 i3) = map go xs
+    where xs = shrink (w, i0, i1, i2, i3)
+          go (w', i0', i1', i2', i3') =
+            T w' i0' i1' i2' i3'
 #endif
 #if defined(qc10) || defined(qc20)
   shrink (T w i0 i1 i2 i3) =
@@ -134,34 +134,67 @@ prop_sc t = pre t S.==> post t
 -- Testing
 --------------------------------------------------------------------------------
 
-test :: Int -> IO (Maybe T) -> IO ()
-test rnds run = do
+test :: FilePath -> Int -> IO (Maybe T) -> IO ()
+test f rnds run = do
   res <- replicateM rnds (test' run)
-  let res' = catMaybes res
-  let rnds' = length res'
-  let app str = appendFile logFile (str ++ "\n")
-  let avg vals = sum vals / fromIntegral rnds'
-  let med vals = sort vals !! (rnds' `div` 2)
-  let stdDev vals = sqrt (avg distances)
-        where
-        distances = map (\x -> (x - m)^2) vals
-        m = avg vals
-
-  app "***************"
-  print res
+  let res'        = catMaybes res
+  let rnds'       = length res'
+  let app str     = appendFile logFile (str ++ "\n")
+  let avg vals    = sum vals / fromIntegral rnds'
+  let med vals    = sort vals !! (rnds' `div` 2)
   let times = fst $ unzip res'
   let szs :: [Double]
       szs = map fromIntegral (snd $ unzip res')
+  let stdDev vals = sqrt (avg distances)
+        where
+        distances = map (\x -> (x - m)^(2::Integer)) vals
+        m = avg vals
+  let percentile n vals = sort vals !! ((rnds' `div` 100) * n)
+  -- http://en.wikipedia.org/wiki/Median_absolute_deviation
+  let medAbsDev vals = med (map dist vals)
+        where dist v = abs (v - median)
+              median = med vals
+
+  app "***************"
[…]

### 2013-02-22 paper/paper.tex
updates for paper.

@@
 % can be 9pt
-\documentclass[10pt]{sigplanconf}
+\documentclass{sigplanconf}
 \usepackage{url}
@@
 
-\newenvironment{code}{\begin{alltt}\small}{\end{alltt}}
+\newenvironment{code}{\begin{alltt}}{\end{alltt}}
 %% \newenvironment{cols}{\begin{tabular}{m{3.6cm}|m{3.6cm}}{Haskell} &
@@
 
-In industrial uses, shrinking is essential.  Hughes, an inventor of QuickCheck
-and the CEO of Quiviq, a company that commercializes the technology, has noted
-that ``without it [shrinking], randomly generated failing cases would often be
-so large as to be almost useless.''~\cite{qcjh}.  Hughes~\emph{et al.} give an
-extended example in which shrinking is essential in debugging telecom
-software~\cite{telecom}.
+In industrial uses, shrinking is essential.  Hughes, in describing industrial
+uses of QuickCheck, has noted that ``without it [shrinking], randomly generated
+failing cases would often be so large as to be almost useless.''~\cite{qcjh}.
+Hughes~\emph{et al.} give an extended example in which shrinking is essential in
+debugging telecom software~\cite{telecom}.
 
@@
 \end{itemize}
+
+\paragraph{SmartCheck}
+
+\begin{figure}[ht]
+\includegraphics[scale=0.6]{Graphs/data}
+\includegraphics[scale=0.6]{Graphs/time-big}
+\includegraphics[scale=0.6]{Graphs/time-small}
+  \caption{Results for 2000 tests of the program in Figure~\ref{fig:initial}.}
+  \label{fig:graphs}
+\end{figure}
 
@@
   \end{center}
-  \caption{Results for 1000 tests of the program in Figure~\ref{fig:initial}.}
+  \caption{Summarizing data for the graphs in Figure~\ref{fig:graphs}. Entries
+    contain execution time (in seconds) and counterexample sizes (counting
+    \ttp{N} constructors).}
   \label{table:results}
@@
 
-\todo{get 99\% }
-
-\paragraph{SmartCheck}
 Motivated by these limitations of QuickCheck and SmallCheck, we developed
@@
 find small counterexamples as well as generalize them as described in the
-preceding paragraph.  To motivate the benefit of SmartCheck, consider
-Table~\ref{table:results}, giving the results for using QuickCheck and
-SmartCheck on the program shown in Figure~\ref{fig:initial} (we omit SmallCheck,
-since as noted, it is not even feasible for the example).  Each entry is a pair
-containing the number of seconds required for counterexample discovery and
-shrinking and the size of the final value obtained (by counting \ttp{N}
-constructors).  We show the mean, median, median, and the values at the 95th
-percentile values.  (The curves produced by these values are not normal, so we
-omit standard the deviation.)
-
-In the first three columns, the results of using QuickCheck to generate
-counterexamples are shown.  The first column contains the results of using
-QuickCheck without shrinking to provide a baseline.  The next column shows the
-result of using QuickCheck with truncating shrink, which each list is truncated
-to 10 elements.  Compared with no shrinking, this approximately halves the size
-of the median counterexample, but doubles the time taken.  There can be large
-outliers, however, in the time taken during shrinking.  Next, we show the
-results of the ``tuple shrink'' instance.  While the median amount of time
-required is much higher, the size of the resulting value is approximately halved
-again.  The reason for the increase in time is because shrinking is very
-productive and produces a nonempty list for most values.
+preceding paragraph.  SmartCheck takes a counterexample produced by some
+oracle---in these experiments, we use QuickCheck---and then shrinks the
+counterexample.
+
+To motivate the benefit of SmartCheck, consider the graphs in
+Figure~\ref{fig:graphs} and the Table~\ref{table:results}, giving the results
+for using QuickCheck and SmartCheck on the program shown in
+Figure~\ref{fig:initial} (we omit SmallCheck, since as noted, it is not even
+feasible for the example).  The graphs show the number of results, out of 2000
+test runs, with the specified size of the final counterexample (by counting
+\ttp{N} constructors) or requiring the specified execution time.  We show two
+execution time graphs, since the scale of time scales differ dramatically.  In
+the table, we show the mean, median, median, and the results at the 95th
+percentile values.  (The curves produced by these values are not necessarily
+Gaussian, so we omit standard the deviation.)
+
+The experiments show three approaches to shrinking with QuickCheck: no
+shrinking, to provide a baseline, truncated shrinking, as described above, in
+which each list is truncated to 10 elements, and tuple shrinking, also described
+above.  Note that SmartCheck produces smaller counterexamples than the other
+approaches, with a very small tail.  SmartCheck has a very small variance---note
+the mean and median are nearly identical, meaning there are very few outliers.
+
+More dramatic, however, are the differences
+in execution time.  The execution times for SmartCheck include the initial
+counterexample generation time required by QuickCheck; thus the time required by
+SmartCheck to shrink a counterexample is on average just 0.2 seconds (0.10 -
+0.08).  As can be seen in the graphs, SmartCheck's execution time is a linear
+function (like QuickCheck, without shrinking).  Truncation and tuple shrinking
+produce very long tails, with some counterexamples taking over a minute to
+shrink.  SmartCheck is largely insensitive to the size of the original
+counterexample provided.
 
@@
 minus the execution time for QuickCheck without shrinking), and the median size
-of a resulting value is six \ttp{N} constructors.  Moreover, SmartCheck has a
-very small variance---note the mean and median are nearly identical, meaning
-there are very few outliers.  And we stress that SmartCheck achieves these
-results \emph{generically}, requiring no shrink definitions to be provided by
-the user.
+of a resulting value is six \ttp{N} constructors.
+
+And we stress that SmartCheck achieves these results \emph{generically},
+requiring no shrink definitions to be provided by the user.
 
@@
 \begin{sloppypar}
+\small
 \noindent%
@@
 \medskip%
+\noindent
 In the counterexample, the culprit is a buried sub-value,
@@
 
-\todo{xmonad}
-\todo{use SC to test itself?}
-\todo{experiments? scale vs. broad data, linear data?}
+\subsection{Xmonad}
+Xmonad is a popular X11 window manager written in Haskell~\cite{xmonad}.  Xmonad
+is a large software project with many contributors, so consequently, a
+QuickCheck test harness is included to help ensure new commits do not introduce
+bugs.  At the heart of Xmonad is a \ttp{StackSet} data type that encodes the
+relationship between windows, work spaces, and which window has the focus.
+Xmonad contains properties to ensure the correct manipulation of
+\ttp{StackSet}s.
+
+Xmonad passes all of its QuickCheck tests, but let us see what might happen to a
+new contributor if things go awry.  Suppose a developer defines an insertion
+function that inserts a new focus into the \ttp{StackSet}, except she forgets to
+place the value inside.  The correct insertion function from Xmonad is
+%
+\begin{code}
+insertUp a s = if member a s then s else insert
+  where
+  insert = modify (Just $ Stack a [] [])
+                  (\textbackslash(Stack t l r) ->
+                      Just $ Stack a l (t:r)) s
+\end{code}
+%
+\noindent
+So suppose the developer forgets to cons \ttp{t} onto \ttp{r}.  Fortunately,
+there are QuickCheck tests that depend on \ttp{insertUp}.  One such test is
+\ttp{prop\_shift\_reversible}, which tests if the \ttp{shift} function, which
+checks a property about shifting focus back and forth between windows.
+
+With the incorrect version of \ttp{insertUp}, QuickCheck returns a wall of text
+to the user, showing the fields of a \ttp{StackSet}:
+
+\medskip%
+\begin{sloppypar}
+\footnotesize
+\noindent%
+\ttp{StackSet \{current = Screen \{workspace = Workspace \{tag = NonNegative
+  \{getNonNegative = 5\}, layout = -2, stack = Just (Stack \{focus = 'I', up =
+  ", down = "\})\}, screen = 5, screenDetail = 0\}, visible = [Screen
+    \{workspace = Workspace \{tag = NonNegative \{getNonNegative = 0\}, layout =
+    -2, stack = Just (Stack \{focus = '\char`\\161', up = \", down =
+    \"\})\}, screen = 1, screenDetail = -2\},Screen \{workspace = Workspace
+    \{tag = NonNegative \{getNonNegative = 2\}, layout = -2, stack = Just (Stack
+    \{focus = '3', up = \", down = "\})\}, screen = 2, screenDetail =
+    -1\},Screen \{workspace = Workspace \{tag = NonNegative \{getNonNegative =
+    3\}, layout = -2, stack = Just (Stack \{focus = 'T', up = \", down =
+    \"\})\}, screen = 3, screenDetail ) = 2\},Screen \{workspace = Workspace
+    \{tag = NonNegative \{getNonNegative = 4\}, layout = -2, stack = Nothing\},
+    screen = 4, screenDetail = -1\},Screen \{workspace = Workspace \{tag =
+    NonNegative \{getNonNegative = 6\}, layout = -2, stack = Just (Stack \{focus
+    = 'm', up = \", down = \"\})\}, screen = 6, screenDetail = 2\},Screen
+    \{workspace = Workspace \{tag = NonNegative \{getNonNegative = 7\}, layout =
+    -2, stack = Nothing\}, screen = 0, screenDetail = 0\}], hidden = [Workspace
+    \{tag = NonNegative \{getNonNegative = 1\}, layout = -2, stack = Just (Stack
+    \{focus = '\char`\\DC3', up = \", down = \"\})\},Workspace \{tag = NonNegative \{getNonNegative = 8\}, layout = -2, stack = Nothing\},Workspace \{tag = NonNegative \{getNonNegative = 9\}, layout = -2, stack = Just (Stack \{focus = '\char`\\ACK', up = \", down = \"\})\}], floating = fromList []\} }
+\end{sloppypar}
+\medskip%
+\noindent
+This is a wall of text, which provides little useful information about what went
+wrong.  SmartCheck, however, takes the value and returns the following:
+%
+\begin{code}
+Extrapolated value:
+forall x0 x1 x2:
+  StackSet x0 x1 x2 (fromList [])
+\end{code}
+%
+\noindent
+The programmer immediately sees that the first three fields in the record are
+irrelevant,  The last field is a dictionary (i.e., an ordered map), which we
+declare to be a base type, so it is not analyzed by SmartCheck.  The upshot is
+that any \ttp{StackSet} fails because insertion is ill-defined.  Such an
+insight, while obvious in hind-sight, can take hours to arrive at as the
+programmer manually tries to write her own shrink functions or manually
+deconstruct a large record.
+
+\todo{xmonad} \todo{experiments? scale vs. broad data, linear data?}
 
@@
 power of QuickCheck to \emph{randomly} discover structurally smaller
-counterexamples.  Furthermore, neither paper explores the idea of counterexample
-generalization, however.
+counterexamples; in short, this is the first treatment we know of that
+integrates counterexample generation with counterexample reduction.
+Furthermore, neither paper explores the idea of counterexample generalization.
 
@@
 discovery takes: the mean time taken by Feat is 5.60 seconds, with a long tail:
-five percent of runs take over 14 seconds (recall that the mean time taken by
+five percent of runs take over 14 seconds (the mean time taken by
 SmartCheck is 0.12 seconds, and 99\% of runs take less than 0.30 seconds).
@@
 attempt to generalize counterexamples based on specific inputs.
-
-\todo{shrink paper rehger pointed to ``Automatic isolation of compiler errors''
-  from 1994}
-\todo{\url{http://www.cs.utah.edu/~regehr/papers/pldi12-preprint.pdf}}
 
@@
 %% \url{http://www.haskell.org/haskellwiki/Research_papers/Testing_and_correctness}%}
+%% \todo{shrink paper rehger pointed to ``Automatic isolation of compiler errors''
+%%   from 1994}
 
@@
 %%   \url{https://github.com/JacquesCarette/GenCheck}%
-\url{http://blog.regehr.org/archives/697} --- check out C-reduce (Regehr)
+%% \url{http://blog.regehr.org/archives/697} --- check out C-reduce (Regehr)
+%% \todo{\url{http://www.cs.utah.edu/~regehr/papers/pldi12-preprint.pdf}}
 %% }
-
-\todo{stuff from SC repo}
 
@@
 currently, due to limitations with GHC Generics.
+
+We have not profiled the performance of our implementation of SmartCheck, so
+additional performance gains might be possible.  In particular, we perform
+significant computations over tree-structured data, and a zipper~\cite{zipper},
+which we do not currently implement, could improve performance.
 

### 2013-02-22 paper/Example/InitEx.hs

@@ -1,172 +1,0 @@
-{-# LANGUAGE CPP #-}
-{-# LANGUAGE TemplateHaskell #-}
-{-# LANGUAGE DeriveDataTypeable #-}
-{-# LANGUAGE DeriveGeneric #-}
-
--- | Try to generate a very large counterexample.
-
-module Main where
-
-import Test.SmartCheck
-import Test.SmartCheck.Reduce
-import Test.QuickCheck
-import Test.LazySmallCheck hiding (Property, test, (==>))
-import qualified Test.LazySmallCheck as S
-
-import GHC.Generics hiding (P, C)
-import Data.Typeable
-import Control.Monad
-
-import Data.Int
-import Data.List
-import Data.Maybe
-import Data.Time
-import System.Environment
-
-import Test.Feat
-
------------------------------------------------------------------
-
--- Let's try to generate a product type of long lists when all we need is a
--- single element to have a long list.
-
--- Container so that we don't have base types.
-
------------------------------------------------------------------
-data A = A Int16
-  deriving (Read, Show, Typeable, Generic)
-
-data B = B [A] [A] [A] [A]
-  deriving (Read, Show, Typeable, Generic)
-
-instance Serial Int16 where
-  series d = drawnFrom [(-d')..d']
-    where d' = fromIntegral d
-
-instance SubTypes A
-
-instance Arbitrary A where
-  arbitrary = liftM A arbitrary
-  shrink (A i) = map A (shrink i)
-
-instance Serial A where
-  series = cons1 A
-
-instance SubTypes B
-
--- qc/shrink takes over 1m seconds
-instance Arbitrary B where
-#ifdef feat
-  arbitrary = sized uniform
-#else
-  arbitrary = liftM4 B arbitrary arbitrary
-                       arbitrary arbitrary
-#endif
-
-#if defined(qcNone) || defined(sc) || defined(feat)
-  shrink (B a b c d) = []
-#else
-  shrink (B a b c d) = [ B w x y z |
-                         w <- tk a
-                       , x <- tk b
-                       , y <- tk c
-                       , z <- tk d ]
-    where
-#ifdef qc10
-    sz = 10
-#endif
-#ifdef qc20
-    sz = 20
-#endif
-    tk x = take sz (shrink x)
-#endif
-
-deriveEnumerable ''A
-deriveEnumerable ''B
-
-instance Serial B where
-  series = cons4 B
-
[…]

### 2013-02-22 paper/Example/InitEx2.hs

@@ -169,9 +169,9 @@ test f rnds run = do
   app ""
   app ""
 
-  appendFile (f ++ "_time.csv") (mkCSV $ plot 50 times)
+  appendFile (f ++ "_time.csv") (mkCSV $ plot 200 times)
 --  appendFile (f ++ "_timelog.csv") (mkCSV $ plot rnds (map (log . (+1)) times))
-  appendFile (f ++ "_vals.csv") (mkCSV $ plot 200 szs)
+  appendFile (f ++ "_vals.csv") (mkCSV $ plot 50 szs)
 --  appendFile (f ++ "_valslog.csv") (mkCSV $ plot rnds (map (log . (+100)) szs))
 
 -- For gnuplot ---------------------------------------
@@ -180,7 +180,7 @@ mkCSV [] = "\n"
 mkCSV ((x,y):rst) = show x ++ ", " ++ show y ++ "\n" ++ mkCSV rst
 
 -- Make 100 compartments to put data in.
-plot :: Int -> [Double] -> [(Double,Double)]
+plot :: Double -> [Double] -> [(Double,Double)]
 plot comparts vals = filter (\(_,n) -> n /= 0.0) $ cz vs (min' + compartSz, 0)
   where
   vs          = sort vals

### 2013-02-22 src/Test/SmartCheck.hs

@@ -6,7 +6,7 @@
 module Test.SmartCheck
   ( smartCheck
   , runQC
-  , SubTypes
+  , SubTypes(..)
   , module Test.SmartCheck.Args
   ) where
 

### 2013-03-03 examples/Heap_Program.hs
more examples.

@@ -46,41 +46,13 @@ instance (SC.SubTypes a, Arbitrary a, Generic a)
 instance (SC.SubTypes a, Ord a, Arbitrary a, Generic a)
          => SC.SubTypes (HeapPP a)
 
--- instance (Ord a, Arbitrary a, Data a, Show a) => SC.SubTypes (Heap a) where
---   subTypes Nil = []
---   subTypes (Node x h0 h1) =
---     [ T.Node (SC.subT x) []
---     , T.Node (SC.subT h0) (SC.subTypes h0)
---     , T.Node (SC.subT h1) (SC.subTypes h1)
---     ]
-
--- instance (Arbitrary a, Data a, Show a) => SC.SubTypes (HeapP a) where
---   subTypes Empty = []
---   subTypes (Unit a) = [ T.Node (SC.subT a) [] ]
---   subTypes (Insert i h)      =
---     [ T.Node (SC.subT i) []
---     , T.Node (SC.subT h) (SC.subTypes h)
---     ]
---   subTypes (SafeRemoveMin h) =
---     [ T.Node (SC.subT h) (SC.subTypes h) ]
---   subTypes (Merge h0 h1)     =
---     [ T.Node (SC.subT h0) (SC.subTypes h0)
---     , T.Node (SC.subT h1) (SC.subTypes h1)
---     ]
---   subTypes (FromList a) = [ T.Node (SC.subT a) [] ]
-
--- instance (Ord a, Arbitrary a, Data a, Show a) => SC.SubTypes (HeapPP a) where
---   subTypes (HeapPP hp h) =
---     [ T.Node (SC.subT hp) (SC.subTypes hp)
---     , T.Node (SC.subT h) (SC.subTypes h)
---     ]
-
 instance (Ord a, Arbitrary a) => Arbitrary (Heap a) where
   arbitrary = do p <- arbitrary :: Gen (HeapP a)
                  return $ heap p
 
 --------------------------------------------------------------------------
 -- skew heaps
+-- Smallest values on top.
 
 data Heap a
   = Node a (Heap a) (Heap a)
@@ -204,8 +176,8 @@ instance (Ord a, Arbitrary a) => Arbitrary (HeapPP a) where
     do p <- arbitrary
        return (HeapPP p (heap p))
 
-  shrink (HeapPP p _) =
-    [ HeapPP p' (heap p') | p' <- shrink p ]
+  -- shrink (HeapPP p _) =
[…]

### 2013-03-03 examples/Lambda.hs

@@ -1,0 +1,443 @@
+-- From QuickCheck test suite.
+
+{-# LANGUAGE ScopedTypeVariables, TemplateHaskell #-}
+{-# LANGUAGE DeriveGeneric #-}
+{-# LANGUAGE DeriveDataTypeable #-}
+module Main where
+
+--------------------------------------------------------------------------
+-- imports
+
+import Data.Functor ((<$>))
+import Data.Typeable
+import GHC.Generics
+import Test.SmartCheck
+
+import Test.QuickCheck
+
+import Data.Maybe
+
+import Control.Monad
+  ( liftM
+  , liftM2
+  )
+
+import Data.Char
+  ( toUpper
+  )
+
+import Data.Set (Set)
+import qualified Data.Set as Set
+
+import Debug.Trace
+
+--------------------------------------------------------------------------
+-- types for lambda expressions
+
+-- variables
+
+newtype Var = MkVar String
+  deriving ( Eq, Ord, Show, Read, Typeable, Generic )
+
+instance SubTypes Var
+
+-- instance Show Var where
+--   show (MkVar s) = s
+
+varList :: [Var]
+varList = [ MkVar s
+          | let vs = [ c:v | v <- "" : vs, c <- ['a'..'z'] ]
+          , s <- vs
+          ]
+
+instance Arbitrary Var where
+  arbitrary = growingElements [ MkVar [c] | c <- ['a'..'z'] ]
+
+-- constants
+
+newtype Con = MkCon String
+  deriving ( Eq, Ord, Show, Read, Typeable, Generic )
+
+instance SubTypes Con
+
+-- instance Show Con where
+--   show (MkCon s) = s
+
+instance Arbitrary Con where
+  arbitrary = growingElements [ MkCon [c] | c <- ['A'..'Z'] ]
+
+-- expressions
+
+data Exp
+  = Lam Var Exp
+  | App Exp Exp
+  | Var Var
+  | Con Con
+ deriving ( Eq, Ord, Show, Read, Typeable, Generic )
+
+instance SubTypes Exp
+
+-- instance Show Exp where
+--   showsPrec n (Lam x t) = showParen (n>0) (showString "\\" . shows x . showString "." . shows t)
+--   showsPrec n (App s t) = showParen (n>1) (showsPrec 1 s . showString " " . showsPrec 2 t)
+--   showsPrec _ (Var x)   = shows x
+--   showsPrec _ (Con c)   = shows c
+
+instance Arbitrary Exp where
[…]

### 2013-03-03 paper/paper.tex
paper updates.

@@
 counterexamples discovered.  Hand-written implementations of shrink can be
-difficult to write: they can complex and consist of significant boilerplate
-code.  Without a good understanding of QuickCheck and the property being tested,
-they can be very inefficient, generating poor new candidates, and they may fail
-to shrink a counterexample significantly.  In this paper, we introduce
-\emph{SmartCheck}, an automatic efficient counterexample reduction algorithm.
-SmartCheck reduces algebraic data using generic search heuristics to efficiently
-find smaller counterexamples independent of the program or property being
-tested.  In addition to shrinking, SmartCheck generalizes counterexamples by
-returning a formula that abstracts over individual counterexamples and make it
-easier for the programmer to uncover multiple counterexamples at once.
-SmartCheck has been implemented for Haskell and is freely available.
+difficult to write: they may be complex, inefficient, and consist of significant
+boilerplate code.  We introduce \emph{SmartCheck}, an automatic efficient
+counterexample reduction algorithm.  SmartCheck reduces algebraic data using
+generic search heuristics to efficiently find smaller counterexamples
+independent of the program or property being tested.  In addition to shrinking,
+SmartCheck generalizes counterexamples by returning a \emph{formula} that
+abstracts over individual counterexamples and make it easier for the programmer
+to uncover multiple counterexamples at once.  SmartCheck has been implemented
+for Haskell and is freely available.
 \end{abstract}
@@
 %% embedded domain-specific language, compiler, verification
-
-\todo{benchmark feat -- send to author}
 
@@
 
-In industrial uses, shrinking is essential.  Hughes, in describing industrial
-uses of QuickCheck, has noted that ``without it [shrinking], randomly generated
+In industrial uses, shrinking is essential.  Hughes, in describing commercial applications of QuickCheck, has noted that ``without it [shrinking], randomly generated
 failing cases would often be so large as to be almost useless.''~\cite{qcjh}.
@@
 being evaluated.  Bad definitions can be so inefficient that they are unusable.
-Shrinking presents a programmer's paradox: shrinking is used to generate a
-smaller counterexample to better understand why a property failed, but to
-implement an efficient shrink function often requires a deep enough
-understanding of the property to efficiently generate smaller counterexamples to
-it!
 
@@
   on the language extensions of existential type quantification and scoped type
-  variables.}  Data type \ttp{T} is a product type containing four lists of
-\ttp{Int16}s and a single \ttp{Word8}.  For our purposes, it is not important
+  variables.}  Data type \ttp{T} is a product type containing four lists of wrapped
+\ttp{Int16}s and a single wrapped \ttp{Word8}.  For our purposes, it is not important
 what \ttp{T} models, but for concreteness, suppose it models the values in four
@@
 First, the result of (\ttp{shrink t}) is null if any list contained in \ttp{t}
-is null, then \ttp{t} will not be shrunk.  More troublesome is that with
+is null, so \ttp{t} will not be shrunk.  More troublesome is that with
 QuickCheck's default \ttp{Arbitrary} instances, the length of potential
@@
 the cross-product of the generated lists.  For the example above, with some
-counterexamples, the shrinking stage may appear to loop for hours, consuming
+counterexamples, the shrinking stage may appear to execute for hours, consuming
 ever more memory, without returning a result.
@@
 cross-product of the type's fields at a given depth.  Unfortunately, SmallCheck
-fails long before it reaches a depth required to discover a counterexample, due
-to combinatorial explosion.  (A related library named Feat combines some aspects
-of SmallCheck and QuickCheck~\cite{feat}; we discuss it more in the related work
-(Section~\ref{sec:related}).
+fails long before it reaches a depth required to discover a counterexample.  (A
+related library named Feat combines some aspects of SmallCheck and
+QuickCheck~\cite{feat}; we discuss it in Section~\ref{sec:related}).
 
@@
 Motivated by these limitations of QuickCheck and SmallCheck, we developed
-\emph{SmartCheck}.  SmartCheck extends QuickCheck to efficiently and generically
-find small counterexamples as well as generalize them as described in the
-preceding paragraph.  SmartCheck takes a counterexample produced by some
-oracle---in these experiments, we use QuickCheck---and then shrinks the
-counterexample.
-
-To motivate the benefit of SmartCheck, consider the graphs in
+\emph{SmartCheck}.  SmartCheck takes a counterexample produced by some
+oracle---in these experiments, we use QuickCheck---and generically minimizes and
+generalizes the counterexample, as described in the preceding bullet points.
+
+To motivate the benefits of SmartCheck, consider the graphs in
 Figure~\ref{fig:graphs} and the Table~\ref{table:results}, giving the results
@@
 Figure~\ref{fig:initial} (we omit SmallCheck, since as noted, it is not even
-feasible for the example).  The graphs show the number of results, out of 2000
-test runs, with the specified size of the final counterexample (by counting
-\ttp{N} constructors) or requiring the specified execution time.  We show two
-execution time graphs, since the scale of time scales differ dramatically.  In
-the table, we show the mean, median, median, and the results at the 95th
-percentile values.  (The curves produced by these values are not necessarily
-Gaussian, so we omit standard the deviation.)
+feasible for the example).  We graph counterexample size and execution time.
+Out of 2000 runs, we show the number of runs resulting in a shrunk
+counterexample of a given size (by counting \ttp{N} constructors), and we show
+the number of runs executing within a given time limit (in seconds).  Note that
+we provide two execution-time graphs, since the scale of time scales differ
+dramatically.  In the table, we show the mean, median, median, and the results
+at the 95th percentile values.  (The curves produced by these values are not
+necessarily Gaussian, so we omit standard the deviation.)
 
@@
 above.  Note that SmartCheck produces smaller counterexamples than the other
-approaches, with a very small tail.  SmartCheck has a very small variance---note
-the mean and median are nearly identical, meaning there are very few outliers.
-
-More dramatic, however, are the differences
-in execution time.  The execution times for SmartCheck include the initial
-counterexample generation time required by QuickCheck; thus the time required by
-SmartCheck to shrink a counterexample is on average just 0.2 seconds (0.10 -
-0.08).  As can be seen in the graphs, SmartCheck's execution time is a linear
-function (like QuickCheck, without shrinking).  Truncation and tuple shrinking
-produce very long tails, with some counterexamples taking over a minute to
-shrink.  SmartCheck is largely insensitive to the size of the original
+approaches, with a very small tail.
+
+More dramatic, however, are the differences in execution time.  The execution
+times for SmartCheck include the initial counterexample generation time required
+by QuickCheck; thus the time required by SmartCheck to shrink a counterexample
+is on average just 0.2 seconds (0.10 - 0.08).  As can be seen in the graphs,
+SmartCheck's execution time scales linearly (like QuickCheck, without
+shrinking).  Although it is difficult to tell from the graphs, truncation and
+tuple shrinking produce very long tails, with some counterexamples taking over a
+minute to shrink.  SmartCheck is largely insensitive to the size of the original
 counterexample provided.
 
-In the last column are the results for SmartCheck, which includes the time
-required by QuickCheck to generate the original counterexample.  The median time
-required by SmartCheck to shrink is 0.02 seconds (SmartCheck's execution time
-minus the execution time for QuickCheck without shrinking), and the median size
-of a resulting value is six \ttp{N} constructors.
-
 And we stress that SmartCheck achieves these results \emph{generically},
-requiring no shrink definitions to be provided by the user.
+requiring no shrink definitions to be provided by the user.  Indeed, there is
+nothing particularly special about the example given.
 
@@
 
-We focus on \emph{algebraic data types}~\cite{}.  Algebraic data types are
+We focus on \emph{algebraic data types}.  Algebraic data types are
 represented as ``sums of products''.  We sometimes refer to the elements of a
@@
 \begin{itemize}
-\item \ttp{size} returns the size of a value,
+\item \ttp{size} returns the size of a value---intuitively, the number of
+  constructors contained within it,
 \item \ttp{index} returns a \emph{sub-value} at an breadth-first index in a
@@
 \section{Shrinking Data}\label{sec:shrinking}
-In this section, we describe how to efficiently shrink data values.  Before we
-begin, recall the basic idea behind the \ttp{shrink} method of the
-\ttp{Arbitrary} class: generate a list of values, each of which is smaller than
-the current counterexample.  Each of the new values generated typically bears no
-relationship to the original counterexample other than being smaller.
+In this section, we describe how to efficiently and generically shrink algebraic data
+values.  Before we begin, recall the basic idea behind the \ttp{shrink} method
+of the \ttp{Arbitrary} class: generate a list of values, each of which is
+smaller than the current counterexample.  Each of the new values generated
+typically bears no relationship to the original counterexample other than being
+smaller.
 
@@
 
-The approach taken by \ttp{shrink} is the most naive graph search algorithm: it
-simply generates a random set of potential vertexes to check in the graph.
-However, it does not exploit that the original counterexample \ttp{cex} is
-already in a neighborhood of other likely counterexamples.  So rather than
-choose random vertexes, it may be more efficient to test other values that are
-closely related to \ttp{cex}.  By restricting the search to \ttp{cex}'s
-neighborhood, a smaller counterexample might be missed, and only a local minimum
-is returned.  However, with a large set of values, finding a global minimum by
-random sampling vertexes in the graph is possible but too inefficient, as we saw
-from the example in the introduction.
+The approach taken by typical \ttp{shrink} implementations is the most naive
+graph search algorithm: it simply generates a random set of potential vertexes
+to check in the graph.  However, it does not exploit that the original
+counterexample \ttp{cex} is already in a neighborhood of other likely
+counterexamples.  So rather than choose random vertexes, it may be more
+efficient to test other values that are closely related to \ttp{cex}.  By
+restricting the search to \ttp{cex}'s neighborhood, a smaller counterexample
+might be missed, and only a local minimum is returned.  However, with a large
+set of values, finding a global minimum by random sampling vertexes in the graph
+is possible but too inefficient, as we saw from the example in the introduction.
 
@@
 
-The algorithm is guaranteed to terminate: intuitively, the measure for the
+The algorithm is guaranteed to terminate: informally, the measure for the
 function is that either the index increases or the size of the counterexample
@@
 \subsection{Reduction Algorithm Details}
-Having describe the reduction algorithm, there are two important details about
+Having described the reduction algorithm, there are two important details about
 its design we describe below.
@@
 \subsubsection{Variant Counterexample Hypothesis}
-A motivation for the design of the reduction algorithm is something we cal the
+A motivation for the design of the reduction algorithm is something we call the
 \emph{variant counterexample hypothesis}.  Recall that the algorithm begins at
@@
 specially: lists of base types are considered to be base types themselves.  This
-prevents the algorithm from analyzing, for example, strings.  A list of a
-non-base-types is shrunk by the algorithm.
+prevents the algorithm from analyzing, for example, strings.  A list of
+non-base-types, however, is shrunk by the algorithm.
 
@@
 efficiency, particularly for large product types with significant ``fan out''.
-The number of sub-values of a value product type can grow exponentially with
+The number of sub-values of a product type value can grow exponentially with
 respect to the depth.  Furthermore, note that as the reduction algorithm
@@
 \noindent
-since any dividend results in divide-by-zero.  Not only do the generalizations
-assist the programmer's insight, but they reduce the sheer size of the
-counterexample.  We call the kind of formula just shown \emph{universal
-  sub-value generalization} and is implemented in SmartCheck.
+since any dividend results in divide-by-zero for the given divisor.  Not only do
+the generalizations assist the programmer's insight, but they reduce the sheer
+size of the counterexample.  We call the kind of formula just shown
+\emph{universal sub-value generalization} and it is implemented in SmartCheck.
 
@@
 
-\todo{explain scMaxForall}
-
 The extrapolation algorithm is shown in Figure~\ref{fig:universal}.  The
@@
 New sub-values are generated by \ttp{newVals}, shown in
-Figure~\ref{fig:reduction}.  Here, we bound the size of values to generate by
-the flag \ttp{scMaxSize}, which is independent of the size of the particular
-sub-value.
+Figure~\ref{fig:reduction}.  Like in the reduction algorithm, we bound the size
+of values to generate by the flag \ttp{scMaxSize}, which is independent of the
+size of the particular sub-value.  The flag \ttp{scMaxForall} is the analogue of
+the \ttp{scMaxReduce} flag, determining the number of values generated in trying
+to generalize a value.
 
@@
 
-The extrapolation algorithm is unsound insofar as it extrapolations from a set
-of counterexamples to a universal claim, similar to QuickSpec~\cite{qs}.  By
-tuning the parameters, the risk of an unsound generalization is reduced.
-
 The algorithm's complexity is $\mathcal{O}(n)$, where $n$ is the number of
@@
 
+\paragraph{Soundness}
+The extrapolation algorithm is unsound in two ways.  First, it extrapolates from a set
+of counterexamples to a universal claim, similar to QuickSpec~\cite{qs}.  By
+tuning the parameters, the risk of an unsound generalization is reduced.
+Second, note that each index is generalized with respect to the concrete
+sub-values that exist in the counterexample.  So in some cases, a formula may be
+returned that is overly general.  For example, consider the counterexample in
+which both fields of \ttp{Add} contain values causing the failure:
+%
+\begin{code}
+Add (Div (C 1) (C (-2)) (C 2))
+    (Div (C 0) (C (-1)) (C 1))
+\end{code}
+%
+\noindent
+The universal generalization algorithm might return the formula
+\noindent
+%
+\begin{code}
+forall values x0 x1 . Add x0 x1
+\end{code}
+%
+\noindent
+The algorithm could be modified to so that when attempting to generalize a
+sub-value at an index, all previously-generalized indexes are re-generalized
+simultaneously, by taking the cross-product of random values for the current
+index and for the previously-generalized indexes.  While this reduces the
+potential for over-generalizing a counterexample, it is exponential in the
+number of indexes generalized.  For universally quantified formulas, the reader
+should read it as shorthand for quantifying each variable independently and
+taking the conjunction of formulas.  For example, instead of
+%
+\begin{code}
+forall values x0 x1 . Add x0 x1
+\end{code}
+%
+\noindent
+one should read
+%
+%
+\begin{code}
+      forall values x0 .
+        Add x0 (Div (C 0) (C (-1)) (C 1))
+  and forall values x1 .
+        Add (Div (C 1) (C (-2)) (C 2)) x1
+\end{code}
+
 \subsection{Existential Sub-Value Generalization}
@@
 fails because there is a bug for some of the variants but not others.  For
-example, recall the calculator language from Section~\ref{sec:subval}.  The
-property \ttp{prop\_div} fails only for values that contain a variant tagged
+example, recall again the calculator language from Section~\ref{sec:subval}.  The
+no-division-by-zero property fails only for values that contain a variant tagged
 with the \ttp{Div} constructor.  Recall again the generalized counterexample
@@
 containing every possible variant as a sub-value, then we abstract it.  For
-example, the following all characterize counterexamples:
+example, the following are all counterexamples:
 %
@@
 \ttp{idx} in counterexample \ttp{cex} with every possible variant given the type
-and construct a counterexample to the property.  There are three guards to the
+and construct a counterexample to the property.  There are four guards to the
 recursive function \ttp{constrFail'}: the first guard holds if the list of
@@
 variants.  Third, for a specific sub-value \ttp{v}, we test whether it fails the
-property.  If so, we add its constructor to the list of constructors.  Note in
-the definition of \ttp{prop'}, we add an additional precondition that the
-current constructor is not an element of constructors already seen.  Thus,
-(\ttp{go v}) returns (\ttp{Just True}) if (\ttp{replace cex idx v}) passes this
-precondition (and any other preconditions of \ttp{prop}), but fails the
-property.
+property.  If so, we add its constructor to the list of constructors.
+Otherwise, we simply recurse.  Note in the definition of \ttp{prop'}, we add an
+additional precondition that the current constructor is not an element of
+constructors already seen.  Thus, (\ttp{go v}) returns (\ttp{Just True}) if
+(\ttp{replace cex idx v}) passes this precondition (and any other preconditions
+of \ttp{prop}), but fails the property.
 
@@
 \subsection{Automated Precondition Strengthening}\label{sec:precondition}
-The extrapolation algorithm generalizes a counterexample, but in the
-``neighborhood'' of he original counterexample.  In particular, all
-generalizations are from the same variant as the original counterexample.  To
-help the programmer in the generalization step, we would also like a way to test
-the property again, ensuring we get counterexamples (if they exist) outside of
-the neighborhood of the original one.
+The universal and existential generalization algorithms generalize a
+counterexample, but in the ``neighborhood'' of the original counterexample.  In
+particular, all generalizations are from the same variant as the original
+counterexample.  To help the programmer in the generalization step, we would
+also like a way to test the property again, ensuring we get counterexamples (if
+they exist) outside of the neighborhood of the original one.
 
@@
 and a property in the case that some values fail the property of the form
-(\ttp{pre ==> prop}).  Our goal is to find additional counterexamples in the
-shaded region, in which \ttp{pre} is satisfied, \ttp{prop} is not satisfied, and
+(\ttp{pre ==> post}).  Our goal is to find additional counterexamples in the
+shaded region, in which \ttp{pre} is satisfied, \ttp{post} is not satisfied, and
 the counterexample is not already characterized by a universal sub-value
@@
 unknown at compile-time whether the value to be replaced and replacing value
-have the same types.
+have the same types.  (We only require \ttp{Read} because standard QuickCheck
+does not return a counterexample; we read the shown value.)
 
@@
 %
+Default instances are provided for common Prelude types and some additional
+ones, including all types for which QuickCheck provides instances.
 
@@
 method can be used to discover an initial counterexample, SmartCheck by default
-uses QuickCheck to discover an initial counterexample.  If QuickCheck is used,
-instances for \ttp{Arbitrary} must be provided.
+uses QuickCheck to discover an initial counterexample.
 
@@
 one large structure together with smaller inputs.  While multiple inputs could
-be analyzed simultaneously by SmartCheck, it is not clear how to clearly
+be analyzed simultaneously by SmartCheck, it is not obvious how to clearly
 represent multiple large counterexamples to the user---the implementation
-currently prints to standard out.  (In future work, a web-based interface could help.)
+currently prints to standard out.  (In future work, a web-based interface could
+help.)
 
@@
 being tested.  The arguments is a record type that includes the values described
-above to control SmartCheck's behavior.
-
-\todo{default instances for all common types, including those provided by QC}
+above to control SmartCheck's behavior.  The standard arguments are contained in
+the value \ttp{scStdArgs}.
+
+Counterexamples can be optionally shown in a tree format by setting the
+\ttp{format} field of the arguments to be equal to \ttp{PrintTree}.  for
+example, the tree format shows a counterexample like
+%
+\begin{code}
+Div (C 1) (Add (C 0) (C 2))
+\end{code}
+%
+\noindent
+as
+%
+\begin{code}
+\scriptsize
+Div
+|
++- C 1
+|
+`- Add
+   |
+   +- C 0
+   |
+   `- C 2
+\end{code}
+%
+\noindent
+We find that for very large and nested data structures, a tree representation
+aids in visually parsing the value.
 
 \section{Experiments}\label{sec:experiments}
+Here we describe three experiments using SmartCheck, including an Xmonad
+property and a property about a natural language processing library.
+
+%% \subsection{Lambda Calculus}
+%% Included in QuickCheck's test suite is an untyped lambda calculus module.  One
+%% of the properties, \ttp{prop\_EvalProducesValue}, claims that a normal form
+%% results from repeatedly applying the $\beta$-reduction rule to a lambda term.
+%% But the property is just the negation of the theorem that the untyped lambda
+%% calculus is not (weakly) normalizing, so we should expect failure.  In testing,
+%% though, we do not know how many $\beta$-reduction steps might be required for a
+%% given term, so a finite limit on reduction is introduced, either through a
+%% timeout or after a number of steps.  If we limit $\beta$-reduction to 10,000
+%% steps, a counterexample is returned by QuickCheck:
+%% %
+%% \medskip%
+%% \begin{sloppypar}
+%% \scriptsize
+%% \noindent%
+%% \ttp{Closed (App (App (App (Lam (MkVar "l") (App (Lam (MkVar "u") (App (Var (MkVar "l")) (Var (MkVar "u")))) (App (Lam (MkVar "u") (Var (MkVar "l"))) (Var (MkVar "l"))))) (Lam (MkVar "s") (App (App (App (Var (MkVar "s")) (Var (MkVar "s"))) (App (Var (MkVar "s")) (Var (MkVar "s")))) (App (Lam (MkVar "a") (Var (MkVar "a"))) (Lam (MkVar "w") (Var (MkVar "s"))))))) (App (Lam (MkVar "y") (App (Var (MkVar "y")) (Var (MkVar "y")))) (Lam (MkVar "q") (App (App (Lam (MkVar "a") (Var (MkVar "q"))) (App (Var (MkVar "q")) (Var (MkVar "q")))) (App (Var (MkVar "q")) (App (Con (MkCon "H")) (Var (MkVar "q")))))))) (App (App (App (App (Lam (MkVar "s") (Var (MkVar "s"))) (App (Lam (MkVar "b") (Con (MkCon "J"))) (Lam (MkVar "v") (Var (MkVar "v"))))) (Lam (MkVar "s") (Lam (MkVar "z") (Var (MkVar "z"))))) (App (Con (MkCon "Z")) (App (App (Lam (MkVar "l") (Con (MkCon "V"))) (Lam (MkVar "e") (Var (MkVar "e")))) (App (Lam (MkVar "u") (Var (MkVar "u"))) (Lam (MkVar "x") (Var (MkVar "x"))))))) (Lam (MkVar "a") (App (App (Lam (MkVar "i") (App (Var (MkVar "i")) (Var (MkVar "i")))) (Var (MkVar "a"))) (App (Lam (MkVar "t") (Lam (MkVar "u") (Var (MkVar "a")))) (App (App (Con (MkCon "P")) (Con (MkCon "M"))) (Var (MkVar "a"))))))))}
+%% \end{sloppypar}
+%% \medskip%
+%% \noindent
+%% SmartCheck reduces the counterexample automatically to a more manageable size:
+%% %
+%% \medskip%
+%% \begin{sloppypar}
+%% \scriptsize
+%% \noindent%
+%% \ttp{Closed (App (App (App (Lam (MkVar "l") (App (Var (MkVar "l")) (App (Lam (MkVar "u") (Var (MkVar "l"))) (Var (MkVar "l"))))) (Lam (MkVar "s") (App (App (Con (MkCon "D")) (App (Var (MkVar "s")) (Var (MkVar "s")))) (Var (MkVar "c"))))) (Con (MkCon "D"))) (Var (MkVar "b")))}
+%% \end{sloppypar}
+%% \medskip%
+%% \noindent
+%% and then gne:eralizes it to the following formula:
+%% %
+%% \begin{code}
+%% forall x0 x1 x2: Closed (App (App x0 x1) x2)
+%% \end{code}
+%% %
 
@@
 
-With the incorrect version of \ttp{insertUp}, QuickCheck returns a wall of text
-to the user, showing the fields of a \ttp{StackSet}:
+With the incorrect version of \ttp{insertUp}, an example returned by QuickCheck
+would typically include a \ttp{StackSet} value as follows:
 
@@
 \begin{sloppypar}
-\footnotesize
+\scriptsize
 \noindent%
@@
     \{workspace = Workspace \{tag = NonNegative \{getNonNegative = 0\}, layout =
-    -2, stack = Just (Stack \{focus = '\char`\\161', up = \", down =
-    \"\})\}, screen = 1, screenDetail = -2\},Screen \{workspace = Workspace
+    -2, stack = Just (Stack \{focus = `\char`\\161', up = ", down =
+    "\})\}, screen = 1, screenDetail = -2\},Screen \{workspace = Workspace
     \{tag = NonNegative \{getNonNegative = 2\}, layout = -2, stack = Just (Stack
-    \{focus = '3', up = \", down = "\})\}, screen = 2, screenDetail =
+    \{focus = '3', up = ", down = "\})\}, screen = 2, screenDetail =
     -1\},Screen \{workspace = Workspace \{tag = NonNegative \{getNonNegative =
-    3\}, layout = -2, stack = Just (Stack \{focus = 'T', up = \", down =
-    \"\})\}, screen = 3, screenDetail ) = 2\},Screen \{workspace = Workspace
+    3\}, layout = -2, stack = Just (Stack \{focus = 'T', up = ", down =
+    "\})\}, screen = 3, screenDetail ) = 2\},Screen \{workspace = Workspace
     \{tag = NonNegative \{getNonNegative = 4\}, layout = -2, stack = Nothing\},
@@
     NonNegative \{getNonNegative = 6\}, layout = -2, stack = Just (Stack \{focus
-    = 'm', up = \", down = \"\})\}, screen = 6, screenDetail = 2\},Screen
+    = 'm', up = ", down = "\})\}, screen = 6, screenDetail = 2\},Screen
     \{workspace = Workspace \{tag = NonNegative \{getNonNegative = 7\}, layout =
@@
     \{tag = NonNegative \{getNonNegative = 1\}, layout = -2, stack = Just (Stack
-    \{focus = '\char`\\DC3', up = \", down = \"\})\},Workspace \{tag = NonNegative \{getNonNegative = 8\}, layout = -2, stack = Nothing\},Workspace \{tag = NonNegative \{getNonNegative = 9\}, layout = -2, stack = Just (Stack \{focus = '\char`\\ACK', up = \", down = \"\})\}], floating = fromList []\} }
+    \{focus = `\char`\\DC3', up = ", down = "\})\},Workspace \{tag = NonNegative \{getNonNegative = 8\}, layout = -2, stack = Nothing\},Workspace \{tag = NonNegative \{getNonNegative = 9\}, layout = -2, stack = Just (Stack \{focus = `\char`\\ACK', up = ", down = "\})\}], floating = fromList []\} }
 \end{sloppypar}
@@
 \noindent
+\todo{double check}
 The programmer immediately sees that the first three fields in the record are
-irrelevant,  The last field is a dictionary (i.e., an ordered map), which we
-declare to be a base type, so it is not analyzed by SmartCheck.  The upshot is
-that any \ttp{StackSet} fails because insertion is ill-defined.  Such an
-insight, while obvious in hind-sight, can take hours to arrive at as the
-programmer manually tries to write her own shrink functions or manually
-deconstruct a large record.
-
-\todo{xmonad} \todo{experiments? scale vs. broad data, linear data?}
+irrelevant.  The upshot is that any \ttp{StackSet} fails because insertion is
+ill-defined.  Such an insight, while obvious in hind-sight, can take hours to
+arrive at as the programmer manually tries to write her own shrink functions or
+manually deconstruct a large record.
+
+\subsection{Natural Language Processing}
+In 2012, a question was posted on the programming message board
+\emph{stackoverflow} asking about how to shrinking large nested data
+types.\footnote{\url{http://stackoverflow.com/questions/8788542/how-do-i-get-good-small-shrinks-out-of-quickcheck}}
+The poster writes:
+\begin{quote}
+$\ldots$ I tend to get an incomprehensible page full of output. $\ldots$
+  Implementing the shrink function for each of my types seems to help a little,
+  but not as much as I'd like.  $\ldots$ If I try to tune my shrink
+  implementations, I also find that QC starts taking a very long time.
+\end{quote}
+\noindent
+The question relates to the Geni natural language processing (NLP) package
+implemented in Haskell~\cite{kow}.  Specifically, counterexamples to a property
+attempting to show that a macro expansion function is its own inverse are
+enormous, requiring between 150 to 200 console output lines. \todo{redo tests}
+
+Using SmartCheck, we are able to reduce counterexamples to around 25 lines of
+output.  Most of the savings in the counterexample size were due to universal
+generalization, like in the Xmonad case: entire record fields are abstracted
+away.
+
+``Retro-fitting'' SmartCheck to GenI, which is a large software system, requires
+some work.  Fortunately, because the program already uses QuickCheck,
+\ttp{arbitrary} instances are already defined, but we had to ensure all relevant
+types derived the \ttp{Read}, \ttp{Show}, \ttp{Typeable}, and \ttp{Generic}
+type classes as well, together with \ttp{SubTypes}.  The work does not require
+understanding the program, but it does require hunting for definitions and
+multiple rounds of recompilation.
 
@@
 We have presented new approaches for generically shrinking and generalizing
-counterexamples over algebraic data.  Shrinking has thus far been as much an art
-as a science, requiring experience with QuickCheck and deep understanding about
-the property being tested.  In many cases, SmartCheck automates the laborious
-task of shrinking.
+counterexamples over algebraic data.  SmartCheck automates the laborious
+task of shrinking, and extrapolating from counterexamples, and in our
+experience, performs better and faster than hand-written shrink functions.
 
 We envision a number of potential extensions and improvements to SmartCheck.
-First, we have considered only the simplest kind of data, algebraic data types,
-but we would like to extend these results to shrinking functions, too.  As noted
-in Section~\ref{sec:implementation}, SmartCheck does not work with GADTs
-currently, due to limitations with GHC Generics.
-
-We have not profiled the performance of our implementation of SmartCheck, so
-additional performance gains might be possible.  In particular, we perform
-significant computations over tree-structured data, and a zipper~\cite{zipper},
-which we do not currently implement, could improve performance.
+First, we have considered only the simplest kind of data, algebraic data types.
+As noted in Section~\ref{sec:implementation}, SmartCheck does not work with
+GADTs currently, due to limitations with GHC Generics.  It would be interesting
+to see if the approaches described here could be extended to function types as
+well---we are particularly motivated by Claessen's recent work in shrinking and
+showing functions~\cite{claessen}.
 
@@
 generalize a sub-value if it is not evaluated in testing the property.  Not only
-does this shortcut the generalization phase, but it gives a proof that the
+does this shortcut the generalization phase, but it gives a \emph{proof} that the
 sub-value can be generalized.
@@
 components of the value.  Better approaches for showing large data types are
-needed.
-
-Finally, we believe there are interesting connections between counterexample
+needed.  In particular, an interactive web-based viewer with hyperlinks to close
+or expand sub-values would be particularly useful.
+
+Another aspect of displaying large counterexamples that we have not explored is
+to exploit sharing.  Constructs might be repeated that can be abstracted out.
+For example, instead of a counterexample like
+%
+\begin{code}
+Add (Div (C 1) (C (-2)) (C 2))
+    (Div (C 1) (C (-1)) (C 1))
+\end{code}
+%
+\noindent
+we might instead return
+%
+\begin{code}
+Add (div (-2) 2) (div (-1) 1)
+  where div x y = Div (C 1) (C x) (C y)
+\end{code}
+%
+\noindent
+Discovering and exploiting sharing is future work.
+
+We believe there are interesting connections between counterexample
 discovery and probabilistic reasoning.  Often, when a bug in a program is fixed,
@@
 
-\todo{difficult to compare performance against arbitrary hand-written shrinking}
-\todo{final summary}
-
-\section*{Acknowledgements}
-\todo{Thank Rehger for comments/pointing to paper.}
+Regarding our implementation, we have not profiled the performance of our
+implementation of SmartCheck, so additional performance gains might be possible.
+In particular, we perform significant computations over tree-structured data,
+and a generic zipper~\cite{zipper}, which we do not currently implement, could
+improve performance.
+
+Debugging is a difficult task.  Functional programming has been at the forefront
+of testing research, with tools like QuickCheck and SmallCheck.  We were
+motivated to build a tool like SmartCheck just because of how effective
+QuickCheck is at discovering counterexamples automatically---there would be no
+such problem of having very large counterexamples if inputs were written by
+hand.  We hope SmartCheck and the ideas in this paper continue the tradition of
+highly-automated testing and debugging in the functional programming community.
+
+\section*{Acknowledgments}
 

### 2013-03-03 paper/Paper.hs

@@ -37,8 +37,7 @@ data ScArgs = ScArgs {
                               -- failure in the reduction stage.
   , runForall    :: Bool      -- ^ Should we extrapolate?
   , scMaxForall  :: Int       -- ^ How hard (number of rounds) to look for
-                              -- failures during the extrapolation and
-                              -- constructor generalization stages.
+                              -- failures during the extrapolation stage.
   , scMinForall  :: Int       -- ^ Minimum number of times a property's
                               -- precondition must be passed to generalize it.
   , runExists    :: Bool      -- ^ Should we try to generalize constructors?

### 2013-03-03 src/Test/SmartCheck.hs
smartcheck wibbles.

@@ -8,6 +8,12 @@ module Test.SmartCheck
   , runQC
   , SubTypes(..)
   , module Test.SmartCheck.Args
+  -- ** For constructing new instances of `SubTypes`
+  , gst
+  , grc
+  , gtc
+  , gsf
+  , gsz
   ) where
 
 import Test.SmartCheck.Args
@@ -65,7 +71,7 @@ smartCheck args propT = do
       let replIdxs = Replace valIdxs csIdxs
 
       -- If either kind of extrapolation pass yielded fruit, prettyprint it.
-      if not $ null (valIdxs ++ csIdxs)
+      if (runForall args || runExists args) && (not $ null (valIdxs ++ csIdxs))
         then output d replIdxs
         else smartPrtLn "Could not extrapolate a new value."
 

### 2013-03-03 src/Test/SmartCheck/Args.hs

@@ -42,7 +42,7 @@ data ScArgs =
                                      --------------
          , scMaxForall  :: Int       -- ^ How hard (number of rounds) to look
                                      --   for failures during the extrapolation
-                                     --   and constructor generalization stages.
+                                     --   stage.
                                      --------------
          , scMinForall  :: Int       -- ^ Minimum number of times a property's
                                      -- precondition must be passed to

### 2013-03-03 src/Test/SmartCheck/Types.hs

@@ -15,9 +15,12 @@ module Test.SmartCheck.Types
   , Subst(..)
   , Replace(..)
   , errorMsg
-  , replaceChild'
-  , toConstr'
-  , showForest'
+  -- ** For constructing new instances of `SubTypes`
+  , gst
+  , grc
+  , gtc
+  , gsf
+  , gsz
   ) where
 
 import GHC.Generics

### 2013-03-03 paper/paper.tex
paper updates.

@@
 
-\title{SmartCheck: Automatic and Efficient Counterexample Reduction and Generalization}
+\title{SmartCheck: Automatic and Efficient Counterexample Reduction and Generalization\\$\;$\\DRAFT}
 %\subtitle{Subtitle Text, if any}
@@
 In Section~\ref{sec:experiments}, we discuss some of our experiences with using
-SmartCheck, including checking properties from Xmonad, and related work and
+SmartCheck, including checking properties from XMonad, and related work and
 concluding remarks are made in Sections~\ref{sec:related}
@@
 \section{Experiments}\label{sec:experiments}
-Here we describe three experiments using SmartCheck, including an Xmonad
+Here we describe three experiments using SmartCheck, including an XMonad
 property and a property about a natural language processing library.
@@
 
-\subsection{Xmonad}
-Xmonad is a popular X11 window manager written in Haskell~\cite{xmonad}.  Xmonad
+\subsection{XMonad}
+XMonad is a popular X11 window manager written in Haskell~\cite{xmonad}.  XMonad
 is a large software project with many contributors, so consequently, a
 QuickCheck test harness is included to help ensure new commits do not introduce
-bugs.  At the heart of Xmonad is a \ttp{StackSet} data type that encodes the
+bugs.  At the heart of XMonad is a \ttp{StackSet} data type that encodes the
 relationship between windows, work spaces, and which window has the focus.
-Xmonad contains properties to ensure the correct manipulation of
+XMonad contains properties to ensure the correct manipulation of
 \ttp{StackSet}s.
 
-Xmonad passes all of its QuickCheck tests, but let us see what might happen to a
-new contributor if things go awry.  Suppose a developer defines an insertion
-function that inserts a new focus into the \ttp{StackSet}, except she forgets to
-place the value inside.  The correct insertion function from Xmonad is
-%
-\begin{code}
-insertUp a s = if member a s then s else insert
-  where
-  insert = modify (Just $ Stack a [] [])
-                  (\textbackslash(Stack t l r) ->
-                      Just $ Stack a l (t:r)) s
-\end{code}
-%
-\noindent
-So suppose the developer forgets to cons \ttp{t} onto \ttp{r}.  Fortunately,
-there are QuickCheck tests that depend on \ttp{insertUp}.  One such test is
-\ttp{prop\_shift\_reversible}, which tests if the \ttp{shift} function, which
-checks a property about shifting focus back and forth between windows.
-
-With the incorrect version of \ttp{insertUp}, an example returned by QuickCheck
-would typically include a \ttp{StackSet} value as follows:
-
+XMonad passes all of its QuickCheck tests, but let us see what might happen to a
+new contributor if things go awry.  Suppose a developer defines a deletion
+function to delete a window, if it exists.  There is a deletion function that
+exists in XMonad, which is quite complex, given the amount of state that is managed
+by \ttp{StackSet}.  However, one function used in deletion is to filter the
+stack of windows associated with each workspace defined:
+%
+\begin{code}
+removeFromWorkspace ws =
+  ws { stack = stack ws >>= filter (/=w) }
+\end{code}
+%
+\noindent
+Now, suppose the programmer makes a simple typo (or perhaps has a more serious
+misunderstanding) and instead writes
+%
+\begin{code}
+removeFromWorkspace ws =
+  ws { stack = stack ws >>= filter (==w) }
+\end{code}
+%
+\noindent
+When testing the property \ttp{prop\_delete}, which says that deleting the
+focused window of the current stack removes it from the \ttp{StackSet} \ttp{x}.
+%
+\begin{code}
+prop_delete x =
+    case peek x of
+        Nothing -> True
+        Just i  -> not (member i (delete i x))
+\end{code}
+%
+\noindent
+QuickCheck returns
+%
 \medskip%
@@
 \ttp{StackSet \{current = Screen \{workspace = Workspace \{tag = NonNegative
-  \{getNonNegative = 5\}, layout = -2, stack = Just (Stack \{focus = 'I', up =
-  ", down = "\})\}, screen = 5, screenDetail = 0\}, visible = [Screen
-    \{workspace = Workspace \{tag = NonNegative \{getNonNegative = 0\}, layout =
-    -2, stack = Just (Stack \{focus = `\char`\\161', up = ", down =
-    "\})\}, screen = 1, screenDetail = -2\},Screen \{workspace = Workspace
-    \{tag = NonNegative \{getNonNegative = 2\}, layout = -2, stack = Just (Stack
-    \{focus = '3', up = ", down = "\})\}, screen = 2, screenDetail =
-    -1\},Screen \{workspace = Workspace \{tag = NonNegative \{getNonNegative =
-    3\}, layout = -2, stack = Just (Stack \{focus = 'T', up = ", down =
-    "\})\}, screen = 3, screenDetail ) = 2\},Screen \{workspace = Workspace
-    \{tag = NonNegative \{getNonNegative = 4\}, layout = -2, stack = Nothing\},
-    screen = 4, screenDetail = -1\},Screen \{workspace = Workspace \{tag =
-    NonNegative \{getNonNegative = 6\}, layout = -2, stack = Just (Stack \{focus
-    = 'm', up = ", down = "\})\}, screen = 6, screenDetail = 2\},Screen
-    \{workspace = Workspace \{tag = NonNegative \{getNonNegative = 7\}, layout =
-    -2, stack = Nothing\}, screen = 0, screenDetail = 0\}], hidden = [Workspace
-    \{tag = NonNegative \{getNonNegative = 1\}, layout = -2, stack = Just (Stack
-    \{focus = `\char`\\DC3', up = ", down = "\})\},Workspace \{tag = NonNegative \{getNonNegative = 8\}, layout = -2, stack = Nothing\},Workspace \{tag = NonNegative \{getNonNegative = 9\}, layout = -2, stack = Just (Stack \{focus = `\char`\\ACK', up = ", down = "\})\}], floating = fromList []\} }
+      \{getNonNegative = 0\}, layout = -1, stack = Just (Stack \{focus = 'S', up =
+        "", down = ""\})\}, screen = 1, screenDetail = 1\}, visible = [Screen
+    \{workspace = Workspace \{tag = NonNegative \{getNonNegative = 2\}, layout = -1,
+        stack = Nothing\}, screen = 2, screenDetail = -1\},Screen \{workspace =
+      Workspace \{tag = NonNegative \{getNonNegative = 3\}, layout = -1, stack =
+        Nothing\}, screen = 0, screenDetail = -1\}], hidden = [Workspace \{tag =
+      NonNegative \{getNonNegative = 1\}, layout = -1, stack = Just (Stack \{focus
+        = `\char`\\NUL', up = "", down = ""\})\},Workspace \{tag = NonNegative
+      \{getNonNegative = 4\}, layout = -1, stack = Just (Stack \{focus = 'I', up =
+        "", down = ""\})\}], floating = fromList []\} }
 \end{sloppypar}
@@
 \noindent
-This is a wall of text, which provides little useful information about what went
-wrong.  SmartCheck, however, takes the value and returns the following:
-%
-\begin{code}
-Extrapolated value:
-forall x0 x1 x2:
-  StackSet x0 x1 x2 (fromList [])
-\end{code}
-%
-\noindent
-\todo{double check}
-The programmer immediately sees that the first three fields in the record are
-irrelevant.  The upshot is that any \ttp{StackSet} fails because insertion is
-ill-defined.  Such an insight, while obvious in hind-sight, can take hours to
-arrive at as the programmer manually tries to write her own shrink functions or
-manually deconstruct a large record.
+This is a relatively small counterexample, but even the smallest \ttp{StackSet}
+values are somewhat large due to the number of fields within it.  But let us see
+if we can do better.
+
+SmartCheck returns a smaller counterexample, and then generalizes it:
+%
+\begin{code}
+forall values x0 x1 x2 x3:
+  StackSet
+    (Screen (Workspace x0 (-1) (Just x1)) 1 1)
+    x2 x3 (fromList [])
+\end{code}
+%
+\noindent
+We can look at what was generalized and what was not.  First, the last element
+of \ttp{StackSet} is a map, and we treated it as a base type, so it is not
+analyzed at all.  But it is an empty map.  The second and third fields of
+\ttp{StackSet} are generalized, but the first one is not.  So there is something
+particular about it.  So the culprit is one of the small constants (1 and -1) or
+having a \ttp{Just} value rather than a \ttp{Nothing}: it turns out that what
+matters is having a \ttp{Just} value, which is the stack field that deletion
+works on.
+
+%% n insertion
+%% function that inserts a new focus into the \ttp{StackSet}, except she forgets to
+%% place the value inside.  The correct insertion function from XMonad is
+%% %
+%% \begin{code}
+%% insertUp a s = if member a s then s else insert
+%%   where
+%%   insert = modify (Just $ Stack a [] [])
+%%                   (\textbackslash(Stack t l r) ->
+%%                       Just $ Stack a l (t:r)) s
+%% \end{code}
+%% %
+%% \noindent
+%% So suppose the developer forgets to cons \ttp{t} onto \ttp{r}.  Fortunately,
+%% there are QuickCheck tests that depend on \ttp{insertUp}.  One such test is
+%% \ttp{prop\_shift\_reversible}, which tests if the \ttp{shift} function, which
+%% checks a property about shifting focus back and forth between windows.
+
+%% With the incorrect version of \ttp{insertUp}, an example returned by QuickCheck
+%% would typically include a \ttp{StackSet} value as follows:
+
+%% \medskip%
+%% \begin{sloppypar}
+%% \scriptsize
+%% \noindent%
+%% \ttp{StackSet \{current = Screen \{workspace = Workspace \{tag = NonNegative
+%%   \{getNonNegative = 5\}, layout = -2, stack = Just (Stack \{focus = 'I', up =
+%%   ", down = "\})\}, screen = 5, screenDetail = 0\}, visible = [Screen
+%%     \{workspace = Workspace \{tag = NonNegative \{getNonNegative = 0\}, layout =
+%%     -2, stack = Just (Stack \{focus = `\char`\\161', up = ", down =
+%%     "\})\}, screen = 1, screenDetail = -2\},Screen \{workspace = Workspace
+%%     \{tag = NonNegative \{getNonNegative = 2\}, layout = -2, stack = Just (Stack
+%%     \{focus = '3', up = ", down = "\})\}, screen = 2, screenDetail =
+%%     -1\},Screen \{workspace = Workspace \{tag = NonNegative \{getNonNegative =
+%%     3\}, layout = -2, stack = Just (Stack \{focus = 'T', up = ", down =
+%%     "\})\}, screen = 3, screenDetail ) = 2\},Screen \{workspace = Workspace
+%%     \{tag = NonNegative \{getNonNegative = 4\}, layout = -2, stack = Nothing\},
+%%     screen = 4, screenDetail = -1\},Screen \{workspace = Workspace \{tag =
+%%     NonNegative \{getNonNegative = 6\}, layout = -2, stack = Just (Stack \{focus
+%%     = 'm', up = ", down = "\})\}, screen = 6, screenDetail = 2\},Screen
+%%     \{workspace = Workspace \{tag = NonNegative \{getNonNegative = 7\}, layout =
+%%     -2, stack = Nothing\}, screen = 0, screenDetail = 0\}], hidden = [Workspace
+%%     \{tag = NonNegative \{getNonNegative = 1\}, layout = -2, stack = Just (Stack
+%%     \{focus = `\char`\\DC3', up = ", down = "\})\},Workspace \{tag = NonNegative \{getNonNegative = 8\}, layout = -2, stack = Nothing\},Workspace \{tag = NonNegative \{getNonNegative = 9\}, layout = -2, stack = Just (Stack \{focus = `\char`\\ACK', up = ", down = "\})\}], floating = fromList []\} }
+%% \end{sloppypar}
+%% \medskip%
+%% \noindent
+%% This is a wall of text, which provides little useful information about what went
+%% wrong.  SmartCheck, however, takes the value and returns the following:
+%% %
+%% \begin{code}
+%% Extrapolated value:
+%% forall values x0 x1 x2:
+%%   StackSet x0 x1 x2 (fromList [])
+%% \end{code}
+%% %
+%% \noindent
+%% (The last field of \ttp{StackSet} is a map, and we treat it as a base type.)
+%% The programmer immediately sees that failing instances were found when
+%% substituting arbitrary well-typed values into the first three fields of
+%% \ttp{StackSet}.  Therefore, there is likely nothing specific in the
+%% counterexample causing the failures.
+
+%% the first three fields in the record are
+%% irrelevant.  The upshot is that any \ttp{StackSet} fails because insertion is
+%% ill-defined.  Such an insight, while obvious in hind-sight, can take hours to
+%% arrive at as the programmer manually tries to write her own shrink functions or
+%% manually deconstruct a large record.
 
@@
 In 2012, a question was posted on the programming message board
-\emph{stackoverflow} asking about how to shrinking large nested data
+Stack Overflow asking about how to shrinking large nested data
 types.\footnote{\url{http://stackoverflow.com/questions/8788542/how-do-i-get-good-small-shrinks-out-of-quickcheck}}
@@
 attempting to show that a macro expansion function is its own inverse are
-enormous, requiring between 150 to 200 console output lines. \todo{redo tests}
-
-Using SmartCheck, we are able to reduce counterexamples to around 25 lines of
-output.  Most of the savings in the counterexample size were due to universal
-generalization, like in the Xmonad case: entire record fields are abstracted
-away.
-
-``Retro-fitting'' SmartCheck to GenI, which is a large software system, requires
+enormous, requiring 200-300 80-character lines to print.
+
+Using SmartCheck, we are able to reduce counterexamples to around 25
+80-character lines of output.  Most of the savings in the counterexample size
+were due to universal generalization, like in the XMonad case: entire record
+fields are abstracted away.
+
+``Retro-fitting'' SmartCheck to GenI, which is a large software system, required
 some work.  Fortunately, because the program already uses QuickCheck,

### 2013-03-18 examples/Large.hs
examples: update API usage.

@@ -65,7 +65,7 @@ prop_p :: B -> Property
 prop_p p = pre p ==> test p
 
 main :: IO ()
-main = smartCheck scStdArgs { runForall = False, constrGen = False } prop_p
+main = smartCheck scStdArgs { runForall = False, runExists = False } prop_p
   -- smartCheck scStdArgs { format = PrintString 
   --                      , scMaxDepth = Just 20
   --                      } 

### 2013-03-18 paper/paper.tex
paper: updated figure, wibbles.

@@
 
-In industrial uses, shrinking is essential.  Hughes, in describing commercial applications of QuickCheck, has noted that ``without it [shrinking], randomly generated
-failing cases would often be so large as to be almost useless.''~\cite{qcjh}.
-Hughes~\emph{et al.} give an extended example in which shrinking is essential in
-debugging telecom software~\cite{telecom}.
+In industrial uses, shrinking is essential.  In describing commercial
+applications of QuickCheck, Hughes has noted that ``without it [shrinking],
+randomly generated failing cases would often be so large as to be almost
+useless.''~\cite{qcjh}.  Hughes~\emph{et al.} also give an extended example in
+which shrinking is essential in debugging telecom software~\cite{telecom}.
 
@@
 dramatically.  In the table, we show the mean, median, median, and the results
-at the 95th percentile values.  (The curves produced by these values are not
-necessarily Gaussian, so we omit standard the deviation.)
+at the 95th percentile values.  (The distributions produced by these values are
+not necessarily Gaussian, so we omit standard the deviation.)
 
@@
 
-We focus on \emph{algebraic data types}.  Algebraic data types are
-represented as ``sums of products''.  We sometimes refer to the elements of a
-sum type as a \emph{variant} that is \emph{tagged} by a constructor.  For
-example, the type
+We focus on algebraic data types represented as ``sums of products''.  We
+sometimes refer to the elements of a sum type as a \emph{variant} that is
+\emph{tagged} by a constructor.  For example, the type
 %
@@
   value's type,
-\item \ttp{baseType} returns whether the type of the value is an ``interesting
+\item \ttp{baseType} is true when the type of the value is an ``uninteresting
   type'', which generally means it is user-defined and has positive arity.
@@
 \noindent
-To demonstrate typical evaluations of the methods, consider a binary tree type:
+To illustrate typical evaluations of the methods, consider a binary tree type:
 %
@@
 subVals (B (B L\sub{0} L\sub{1}) L\sub{2}) =
-  Tree (B L\sub{0} L\sub{1}) [Tree L\sub{0} [], Tree L\sub{1} []]
-       (Tree L\sub{1} [])
+  Node (B L\sub{0} L\sub{1}) [Node L\sub{0} [], Node L\sub{1} []]
+       (Node L\sub{1} [])
 \end{code}
@@
 Figure~\ref{fig:reduction} shows the reduction algorithm.  The function
-\ttp{reduce} takes flags to customize the algorithm's behavior, a
-counterexample \ttp{cex}, and the property \ttp{prop}.  At the beginning of
-execution, we assume that a counterexample has been obtained somehow; our
-implementation relies on QuickCheck as a ``front-end'' to generate a
-counterexample that the algorithm automatically shrinks.  The reduction begins
-at the first proper sub-value of \ttp{cex}; call it \ttp{v}.  When the index
-\ttp{idx} becomes out-of-bounds and returns \ttp{Nothing}, the algorithm
-terminates.  Otherwise, a list of new random values are generated.
+\ttp{reduce} takes flags to customize the algorithm's behavior, a counterexample
+\ttp{cex}, and the property \ttp{prop}.  At the beginning of execution, we
+assume that a counterexample has been obtained somehow; our implementation
+relies on QuickCheck to generate an initial counterexample and to generate
+arbitrary well-typed random values (by the \ttp{arbitrary} method).  The
+reduction begins at the first proper sub-value of \ttp{cex}; call it \ttp{v}.
+When the index \ttp{idx} becomes out-of-bounds and returns \ttp{Nothing}, the
+algorithm terminates.  Otherwise, a list of new random values are generated.
 %
@@
 like QuickCheck's \ttp{arbitrary} method, \ttp{sizedArbitrary} generates
-successfully larger counterexamples when generating new values with which to
+successively larger counterexamples when generating new values with which to
 replace a sub-value.
@@
 For example, consider a simple calculator language containing constants,
-addition, and division:
+addition, and division, together with an evaluator that checks if the divisor is
+0 and returning \ttp{Nothing} in that case:
 %
@@
          | Div Exp Exp
-\end{code}
-%
-\noindent
-Consider a program that takes an \ttp{Exp} as an argument that throws an error
-if an evaluation of an \ttp{Exp} value contains a division-by-zero.  Testing for
-the occurrence of the error, we might have a counterexample like the following:
+
+eval :: Exp -> Maybe Int
+eval (C i) = Just i
+eval (Add e0 e1) =
+  liftM2 (+) (eval e0) (eval e1)
+eval (Div e0 e1) =
+  let e = eval e1 in
+  if e == Just 0 then Nothing
+    else liftM2 div (eval e0) e
+\end{code}
+%
+\noindent
+Now consider the property \ttp{prop\_div}, claiming that if \ttp{divSubTerms}
+holds on an expression, then the evaluator returns \ttp{Just} a value:
+%
+\begin{code}
+divSubTerms :: Exp -> Bool
+divSubTerms (C _)         = True
+divSubTerms (Div _ (C 0)) = False
+divSubTerms (Add e0 e1)   =  divSubTerms e0
+                          && divSubTerms e1
+divSubTerms (Div e0 e1)   = divSubTerms e0
+                          && divSubTerms e1
+
+prop_div e = divSubTerms e ==> eval e /= Nothing
+\end{code}
+%
+Testing \ttp{prop\_div}, we might have a counterexample like the following:
 %
@@
 \paragraph{Soundness}
-The extrapolation algorithm is unsound in two ways.  First, it extrapolates from a set
-of counterexamples to a universal claim, similar to QuickSpec~\cite{qs}.  By
-tuning the parameters, the risk of an unsound generalization is reduced.
+The extrapolation algorithm is unsound in two ways.  First, it extrapolates from
+a set of counterexamples to a universal claim, similar to QuickSpec~\cite{qs}.
+By tuning the parameters, the risk of an unsound generalization is reduced by
+requiring more or larger values to fail the property.
+
 Second, note that each index is generalized with respect to the concrete
@@
 \noindent
-The universal generalization algorithm might return the formula
+since no matter what random value the first field of \ttp{Add} is replaced with,
+the property fails by the second field and vice versa for the second.  The
+universal generalization algorithm might return the formula
 \noindent
@@
 \noindent
-The algorithm could be modified to so that when attempting to generalize a
+The algorithm could be modified so that when attempting to generalize a
 sub-value at an index, all previously-generalized indexes are re-generalized
@@
 forall values x .
-  forall constructors C .
+  forall constructors c .
     there exist arguments \(\stackrel{\rightarrow}{y}\) .
-      such that Div x (C\(\stackrel{\rightarrow}{y}\))
+      such that Div x (c\(\stackrel{\rightarrow}{y}\))
 \end{code}
@@
 
-Figure~\ref{fig:cex-gen} illustrates the situation in which we have some values
+Figure~\ref{fig:cex-gen} illustrates a property of the form (\ttp{pre ==>
+  post}).  Points are specific counterexamples
+
+ the situation in which we have some values
 and a property in the case that some values fail the property of the form
-(\ttp{pre ==> post}).  Our goal is to find additional counterexamples in the
+.  Our goal is to find additional counterexamples in the
 shaded region, in which \ttp{pre} is satisfied, \ttp{post} is not satisfied, and
@@
 
-Figure~\ref{fig:matches} show the shape-matching algorithm used for precondition
+Figure~\ref{fig:matches} shows the shape-matching algorithm used for precondition
 strengthening.  The function takes a candidate counterexample and a list of
@@
 
-SmartCheck does not implement a counterexample discovery algorithm.  While any
-method can be used to discover an initial counterexample, SmartCheck by default
-uses QuickCheck to discover an initial counterexample.
+SmartCheck does not implement a counterexample discovery algorithm or an
+\ttp{arbitrary} method.  Any initial counterexample can be passed in or
+generated, and the algorithm depends on \ttp{arbitrary} to generate new
+well-typed values.
 
@@
 \noindent
-We can look at what was generalized and what was not.  First, the last element
+Let us examine what was generalized and what was not.  First, the last element
 of \ttp{StackSet} is a map, and we treated it as a base type, so it is not
@@
 In 2012, a question was posted on the programming message board
-Stack Overflow asking about how to shrinking large nested data
+Stack Overflow asking about how to shrink large nested data
 types.\footnote{\url{http://stackoverflow.com/questions/8788542/how-do-i-get-good-small-shrinks-out-of-quickcheck}}
@@
 
-Counterexample reduction in general is a relatively unexplored topic, but there
-has been some treatment of the topic.  Zeller and Hildebrandt describe an
-application of greedy search to shrink counterexamples they call
-``delta-debugging'' (DD)~\cite{dd}.  The authors apply their work shrinking HTML
-inputs to crash Mozilla and shrinking C programs to trigger a bug in GCC.
-Subsequent generalizations are reported by Misherghi and Su in which they
-perform greedy search on tree-structured data; they call their approach
-hierarchical delta-debugging (HDD)~\cite{hdd}.
+A few other researchers have also investigated counterexample shrinking.  Zeller
+and Hildebrandt describe an application of greedy search to shrink
+counterexamples they call ``delta-debugging'' (DD)~\cite{dd}.  The authors apply
+their work shrinking HTML inputs to crash Mozilla and shrinking C programs to
+trigger a bug in GCC.  Subsequent generalizations are reported by Misherghi and
+Su in which they perform greedy search on tree-structured data; they call their
+approach hierarchical delta-debugging (HDD)~\cite{hdd}.
 
@@
 
-SmartCheck bears some similarity to QuickSpec, a testing-based library that
-infers equational properties about programs~\cite{qs} insofar as they both
+Finally, SmartCheck bears some similarity to QuickSpec, a testing-based library
+that infers equational properties about programs~\cite{qs} insofar as they both
 attempt to generalize counterexamples based on specific inputs.
@@
 Add (div (-2) 2) (div (-1) 1)
-  where div x y = Div (C 1) (C x) (C y)
+  where div x y = Div (C x) (C y)
 \end{code}

### 2013-03-22 src/Test/SmartCheck.hs
update to SmartCheck for the newest version of QuickCheck.

@@ -182,8 +182,8 @@ lenChk chk ls = if chk ls then errorMsg "No value to SmartCheck!"
 getOut :: Q.Result -> Maybe [String]
 getOut res =
   case res of
-    Q.Failure _ _ _ _ _ _ out -> Just $ lines out
-    _                         -> Nothing
+    Q.Failure _ _ _ _ _ _ _ out -> Just $ lines out
+    _                           -> Nothing
 
 genProp :: (Show a, Q.Testable prop, Q.Arbitrary a)
         => (a -> prop) -> Q.Property

### 2013-03-22 examples/Comp.hs
comp examples update.

@@ -25,7 +25,7 @@ instance Ord I where
 
 data X = X0 I
        | X1 I
-  deriving (Eq, Read, Show, Typeable, Generic)
+  deriving (Read, Show, Typeable, Generic)
 
 instance Arbitrary X where
   arbitrary = oneof [ liftM X0 arbitrary
@@ -38,6 +38,12 @@ instance Ord X where
   compare (X0 i) (X1 j) = compare i j
   compare (X1 i) (X0 j) = compare i j
   compare (X1 i) (X1 j) = compare i j
+
+instance Eq X where
+  (X0 i) == (X0 j) = i == j
+  (X0 i) == (X1 j) = i == j
+  (X1 i) == (X0 j) = i == j
+  (X1 i) == (X1 j) = i == j
 
 data A = A0 Int
        | A1 Int
@@ -72,7 +78,7 @@ instance Arbitrary B where
 instance SubTypes B
 
 comp :: X -> X -> A -> Bool
-comp a0 a1 _ = a0 < a1
+comp a0 a1 _ = a0 /= a1
 
 comp2 :: X -> Bool
 comp2 a0 = a0 < a0

### 2013-03-22 paper/paper.tex
paper updates.

@@
 \begin{abstract}
-QuickCheck, developed by Klassen and Hughes, is a powerful library for automatic
-test-case generation.  Because QuickCheck performs random testing, it can
-discover very large counterexamples.  QuickCheck provides an interface for the
-user to write \emph{shrink} functions to attempt to reduce the size of
+QuickCheck, developed by Claessen and Hughes, is a powerful library for automatic
+test-case generation.  Because QuickCheck performs random testing, some of the
+counterexamples discovered are very large.  QuickCheck provides an interface for
+the user to write \emph{shrink} functions to attempt to reduce the size of
 counterexamples discovered.  Hand-written implementations of shrink can be
@@
 
-Typically, effectively shrinking complex data requires a good understanding of
-how shrinking in QuickCheck works and the semantics of the property and program
-being evaluated.  Bad definitions can be so inefficient that they are unusable.
+Defining an efficient and effective shrink method requires a good understanding
+of how shrinking in QuickCheck works and the semantics of the property and
+program being evaluated.  Bad definitions can be so inefficient that they are
+unusable.
+
+In addition, shrinking is one side of the coin when it comes to making counterexamples more
+understandable.  The other side is generalization.  For example, the following
+is a relatively small counterexample returned when QuickChecking an (incorrect)
+implementation in Xmonad, a popular X11 window manager written in
+Haskell~\cite{xmonad}:
+%
+\medskip%
+\begin{sloppypar}
+%\footnotesize
+\noindent%
+\ttp{StackSet \{current = Screen \{workspace = Workspace \{tag = NonNegative
+      \{getNonNegative = 0\}, layout = -1, stack = Just (Stack \{focus = 'S', up =
+        "", down = ""\})\}, screen = 1, screenDetail = 1\}, visible = [Screen
+    \{workspace = Workspace \{tag = NonNegative \{getNonNegative = 2\}, layout = -1,
+        stack = Nothing\}, screen = 2, screenDetail = -1\},Screen \{workspace =
+      Workspace \{tag = NonNegative \{getNonNegative = 3\}, layout = -1, stack =
+        Nothing\}, screen = 0, screenDetail = -1\}], hidden = [Workspace \{tag =
+      NonNegative \{getNonNegative = 1\}, layout = -1, stack = Just (Stack \{focus
+        = `\char`\\NUL', up = "", down = ""\})\},Workspace \{tag = NonNegative
+      \{getNonNegative = 4\}, layout = -1, stack = Just (Stack \{focus = 'I', up =
+        "", down = ""\})\}], floating = fromList []\} }
+\end{sloppypar}
+\medskip%
+\noindent
+Functional programmers should be familiar with having to debug a ``wall of
+text'' when showing data structures.  What if instead a counterexample like the
+following were returned by the testing framework:
+%
+\begin{code}
+forall values x0 x1 x2 x3:
+  StackSet
+    (Screen (Workspace x0 (-1) (Just x1)) 1 1)
+    x2 x3 (fromList [])
+\end{code}
+%
+\noindent
+The formula quantifies away all the irrelevant portions of the data structure
+with respect to the counterexample, meaning that for any well-typed values that
+replace \ttp{x0}, \ttp{x1}, \ttp{x2}, and \ttp{x3}, the counterexample still
+holds.
+
+\paragraph{SmartCheck}
+
+Motivated by the problems of reducing and generalizing large counterexamples, we
+have developed \emph{SmartCheck}.  SmartCheck takes a counterexample produced by
+some oracle---in these experiments, we use QuickCheck---and generically
+minimizes and generalizes the counterexample.
+
+As demonstrated by a motivating example in Section~\ref{sec:example}, SmartCheck
+can shrink more than an order-of-magnitude faster than efficient hand-written
+shrink implementations for some properties and programs.  More importantly, the
+time required to shrink scales linearly with the size of the
+counterexample---other common shrink implementations often scale in polynomial
+time or worse.  And the counterexamples returned are two to five times smaller.
+
+After presenting some preliminary definitions in
+Section~\ref{sec:preliminaries}, in Section~\ref{sec:shrinking}, we describe the
+generic shrinking algorithm implemented.  The algorithm is particularly novel in
+that it generates new random sub-values during the reduction phase (see related
+work in Section~\ref{sec:related}).
+
+%%  than the tuple or list-truncation approaches
+%% described earlier, and the size of resulting counterexamples are two-five times
+%% smaller (more detailed performance numbers are given in
+%% Section~\ref{sec:performance}).  Furthermore, other shrink approaches tend to
+%% have ``long tails,'' meaning that for some original counterexamples, that are
+%% not shrunk much or take a very long time to shrink.
+%% Motivated by these limitations of QuickCheck and SmallCheck, we developed
+%% \emph{SmartCheck}.
+
+SmartCheck implements three novel approaches to automatically generalize
+counterexamples, which are described in Section~\ref{sec:generalization}.  The
+first algorithm universally quantifies sub-values that always fail in tests, as
+briefly described earlier in the introduction.  The second algorithm
+existentially quantifies sub-values for types in which every possible variant
+fails the property.  For example, finding counterexamples (\ttp{Left 2}) and
+(\ttp{Right True}) for the type (\ttp{Either Int Bool}) means there exists a
+counterexample regardless of the variant chosen.
+
+The third algorithm automatically strengthens properties by omitting ``similar''
+counterexamples to the ones previously observed.  The algorithm is motivated by
+noting that there are often multiple ways in which a property may fail; for
+example, a property that parsing a pretty-printed abstract syntax tree (AST)
+results in the original AST may fail due to multiple bugs.  During testing, it
+is useful to discover all these possibilities in one go.  In practice, the
+problem is solved by discovering a counterexample \ttp{cex}, abstracting it, and
+then adding a new precondition to the property that informally says ``omit
+counterexamples of form \ttp{cex}.''  Adding preconditions manually is laborious
+and may cause the programmer to make premature fixes to the program, if she
+believes she has isolated the error.
+
+We describe our implementation based on generic programming in
+Section~\ref{sec:implementation}; the implementation is open-source.  In
+Section~\ref{sec:experiments}, we discuss some of our experiences with using
+SmartCheck, including checking properties from XMonad and a natural language
+processing library.
+
+\section{A Motivating Example}\label{sec:example}
 
@@
 
-To illustrate, consider the example in Figure~\ref{fig:initial}.\footnote{All
-  examples and algorithms in this paper are presented in
-  Haskell~\cite{haskell98}.  We strive to use idiomatic Haskell, and only rely
-  on the language extensions of existential type quantification and scoped type
-  variables.}  Data type \ttp{T} is a product type containing four lists of wrapped
-\ttp{Int16}s and a single wrapped \ttp{Word8}.  For our purposes, it is not important
-what \ttp{T} models, but for concreteness, suppose it models the values in four
-linked lists in an operating system, together with an 8-bit status register.
+In this section, we motivate in more detail the challenges in shrinking
+counterexamples (we focus on shrinking and only touch on generalization here).
+Consider the example in Figure~\ref{fig:initial}.\footnote{All examples and
+  algorithms in this paper are presented in Haskell 2010~\cite{haskell2010} plus
+  the language extensions of existential type quantification and scoped type
+  variables.}  Data type \ttp{T} is a product type containing four lists of
+wrapped \ttp{Int16}s and a single wrapped \ttp{Word8}.  For our purposes, it is
+not important what \ttp{T} models, but for concreteness, suppose it models the
+values in four linked lists in an operating system, together with an 8-bit
+status register.
 
@@
 \noindent
-satisfies \ttp{pre} but fails \ttp{post} (\ttp{==>} is implication from the
-QuickCheck library).
+satisfies \ttp{pre} but fails \ttp{post} (the \ttp{==>}) operator in the figure
+implements implication from the QuickCheck library).
 
@@
 shrink \ttp{b} while holding \ttp{a} constant (and is generalized to larger
-tuples).  Still, using this definition, from some initial counterexamples,
-shrinking may take over 15 seconds and result in a final counterexample
-containing nearly 60 \ttp{N} constructors.\footnote{All results reported in the paper
-  are with a GHC-compiled program, using \ttp{-O2}, on a 4-core 2.7GHz Intel~i7
-  processor.}  We would like to do better, particularly since the data type and
-property are so simple.
+tuples).  Thus, each argument to \ttp{T} is independently shrunk.
+
+Still, using this definition, from some initial counterexamples, shrinking may
+take over 15 seconds and result in a final counterexample containing nearly 60
+\ttp{N} constructors.\footnote{All results reported in the paper are with a
+  GHC-compiled program, using \ttp{-O2}, on a 4-core 2.7GHz Intel~i7 processor.}
+We would like to do better, particularly since the data type and property are so
+simple.
 
@@
 
-Finally, even if we could overcome the problems described above with QuickCheck
-and SmallCheck, we are left with two others.
-\begin{itemize}
-  \item A counterexample is usually representative of other counterexamples, but
-    that generalization is left to the user to discover for herself.  For
-    example, consider the counterexample presented earlier:
-%
-\begin{code}
-T (N 10) [N (-20000)] [N (-20000)] [] []
-\end{code}
-%
-\noindent
-For any \ttp{Word8}, the counterexample still holds.  Thus, we can be more
-informative and return a quantified formula rather than a single counterexample:
-%
-\begin{code}
-forall x . T x [N (-20000)] [N (-20000)] [] []
-\end{code}
-%
-\noindent
-denoting that for any \ttp{Word8}, the counterexample holds.  Such a formula has
-the added benefit of replacing a potentially large sub-term of a composite data
-type value with a quantifier, reducing the sheer size of the term.
-
-  \item Second, during testing, it is sometimes helpful to cover the input space
-    by discovering different counterexamples to \ttp{prop}.  In the example in
-    Figure~\ref{fig:initial}, one counterexample is fairly representative of
-    others, but this is not true in general (e.g., consider properties about
-    abstract syntax trees).  Ideally, the testing framework would automatically
-    add a predicate characterizing a counterexample as the precondition to
-    \ttp{prop} that strengthens the property.  That is, for a specific
-    counterexample \ttp{ex}, we have a predicate \ttp{predEx} that holds if its
-    argument is another counterexample that is similar to (i.e., some
-    generalization of) \ttp{ex}.  Then we test against a new property
-%
-\begin{code}
-prop' a = not (cex a) ==> prop a
-\end{code}
+%% Finally, even if we could overcome the problems described above with QuickCheck
+%% and SmallCheck, we are left with two others.
+%
+
+%% Second, during testing, it is sometimes helpful to cover the input space
+%%     by discovering different counterexamples to \ttp{prop}.  In the example in
+%%     Figure~\ref{fig:initial}, one counterexample is fairly representative of
+%%     others, but this is not true in general (e.g., consider properties about
+%%     abstract syntax trees).  Ideally, the testing framework would automatically
+%%     add a predicate characterizing a counterexample as the precondition to
+%%     \ttp{prop} that strengthens the property.  That is, for a specific
+%%     counterexample \ttp{ex}, we have a predicate \ttp{predEx} that holds if its
+%%     argument is another counterexample that is similar to (i.e., some
+%%     generalization of) \ttp{ex}.  Then we test against a new property
+%% %
+%% \begin{code}
+%% prop' a = not (cex a) ==> prop a
+%% \end{code}
 %
@@
 %% violate the precondition.
-\end{itemize}
-
-\paragraph{SmartCheck}
+%\end{itemize}
+
+Here we provide some illustrative comparisons between SmartCheck's shrinking
+against other shrinking approaches.  As a representative example, we use the
+example presented in Section~\ref{sec:intro}.  The results for SmartCheck are
+fairly generalizable to other types and properties since SmartCheck's reduction
+algorithm has no dependencies on the specifics of either; however, user-defined
+shrink methods may vary significantly based on the property and counterexample
+type.
 
@@
 
-Motivated by these limitations of QuickCheck and SmallCheck, we developed
-\emph{SmartCheck}.  SmartCheck takes a counterexample produced by some
-oracle---in these experiments, we use QuickCheck---and generically minimizes and
-generalizes the counterexample, as described in the preceding bullet points.
-
 To motivate the benefits of SmartCheck, consider the graphs in
@@
 
-And we stress that SmartCheck achieves these results \emph{generically},
-requiring no shrink definitions to be provided by the user.  Indeed, there is
-nothing particularly special about the example given.
-
-\paragraph{Contributions}
-This paper, and the library it describes, makes the following contributions:
-
-\begin{itemize}
-
-\item an efficient counterexample reduction strategy for algebraic
-  data types (Section~\ref{sec:shrinking});
-
-\item an approach to counterexample generalization, to present the user with a
-  formula describing a set of counterexamples to a property
-  (Section~\ref{sec:generalization});
-
-\item an approach to generically strengthen a property with a precondition that
-  characterizes an already-discovered counterexample for the purpose of finding
-  new, dissimilar counterexamples (Section~\ref{sec:generalization});
-
-\item an implementation based on generic programming that is not only more
-  efficient that hand-written \ttp{shrink} instances in many cases but that
-  generates the instances automatically for the user
-  (Section~\ref{sec:implementation}).
+Finally, note that after shrinking the counterexample, SmartCheck will also
+generalize it.  Note that for any \ttp{Word8} value in the first argument of
+\ttp{T}, the counterexample still holds.  Thus, we can be more informative and
+return a quantified formula rather than a single counterexample:
+%
+\begin{code}
+forall x . T x [N (-20000)] [N (-20000)] [] []
+\end{code}
+
+%% \begin{itemize}
+
+%% \item an efficient counterexample reduction strategy for algebraic
+%%   data types (Section~\ref{sec:shrinking});
+
+%\item an approach to counterexample generalization, to present the user with a
+%%   formula describing a set of counterexamples to a property
+%%   (Section~\ref{sec:generalization});
+
+%% \item an approach to generically strengthen a property with a precondition that
+%%   characterizes an already-discovered counterexample for the purpose of finding
+%%   new, dissimilar counterexamples (Section~\ref{sec:generalization});
+
+%\item
 
@@
 %%   functional program testing.
-\end{itemize}
-
-\noindent
-In Section~\ref{sec:experiments}, we discuss some of our experiences with using
-SmartCheck, including checking properties from XMonad, and related work and
-concluding remarks are made in Sections~\ref{sec:related}
-and~\ref{sec:conclusions}, respectively.
+% \end{itemize}
 
@@
 
-\subsection{Existential Sub-Value Generalization}
+\subsection{Existential Sub-Value Generalization}\label{sec:existential}
 
@@
 removeFromWorkspace ws =
-  ws { stack = stack ws >>= filter (/=w) }
+  ws \{ stack = stack ws >>= filter (/=w) \}
 \end{code}
@@
 removeFromWorkspace ws =
-  ws { stack = stack ws >>= filter (==w) }
+  ws \{ stack = stack ws >>= filter (==w) \}
 \end{code}
@@
 \noindent
-QuickCheck returns
-%
-\medskip%
-\begin{sloppypar}
-\scriptsize
-\noindent%
-\ttp{StackSet \{current = Screen \{workspace = Workspace \{tag = NonNegative
-      \{getNonNegative = 0\}, layout = -1, stack = Just (Stack \{focus = 'S', up =
-        "", down = ""\})\}, screen = 1, screenDetail = 1\}, visible = [Screen
-    \{workspace = Workspace \{tag = NonNegative \{getNonNegative = 2\}, layout = -1,
-        stack = Nothing\}, screen = 2, screenDetail = -1\},Screen \{workspace =
-      Workspace \{tag = NonNegative \{getNonNegative = 3\}, layout = -1, stack =
-        Nothing\}, screen = 0, screenDetail = -1\}], hidden = [Workspace \{tag =
-      NonNegative \{getNonNegative = 1\}, layout = -1, stack = Just (Stack \{focus
-        = `\char`\\NUL', up = "", down = ""\})\},Workspace \{tag = NonNegative
-      \{getNonNegative = 4\}, layout = -1, stack = Just (Stack \{focus = 'I', up =
-        "", down = ""\})\}], floating = fromList []\} }
-\end{sloppypar}
-\medskip%
-\noindent
+%QuickCheck returns
+\todo{fill in}
+
 This is a relatively small counterexample, but even the smallest \ttp{StackSet}
@@
 \section*{Acknowledgments}
+I thank Joe Hurd, John Launchbury, and John Hughes for insightful comments on
+earlier drafts of this paper.  In particular, John Hughes discovered errors and
+omissions that we have fixed.
 

### 2013-03-22 examples/Div0.hs
example tweaks.

@@ -44,7 +44,7 @@ instance Arbitrary Exp where
 
 -- property: so long as 0 isn't in the divisor, we won't try to divide by 0.
 -- It's false: something might evaluate to 0 still.
-prop_div :: Exp -> ScProp
+prop_div :: Exp -> ScProperty
 prop_div e = divSubTerms e --> eval e /= Nothing
 -- prop_div e = property $ case x of
 --                           Nothing -> True

### 2013-03-22 examples/LambdaCalc.hs

@@ -113,7 +113,7 @@ instance Arbitrary Pr where
 -- prop0 (Pr (e0, e1)) = alphaEq e0 e1 ==> betaEq e0 e1
 
 -- if you do a beta reduction to nf
--- prop1 :: Pr -> ScProp
+-- prop1 :: Pr -> ScProperty
 -- prop1 (Pr e0 e1) = -- Timeout due to possible non-termination
 --   within 1000 $ alphaEq e0 e1 --> betaEq e0 (substVar "x" "y" e1)
 

### 2013-03-22 paper/paper.tex
paper: major updates.

@@
 
-\title{SmartCheck: Automatic and Efficient Counterexample Reduction and Generalization\\$\;$\\DRAFT}
+\title{SmartCheck: Automatic and Efficient Counterexample Reduction and Generalization}
 %\subtitle{Subtitle Text, if any}
@@
 of how shrinking in QuickCheck works and the semantics of the property and
-program being evaluated.  Bad definitions can be so inefficient that they are
-unusable.
-
-In addition, shrinking is one side of the coin when it comes to making counterexamples more
-understandable.  The other side is generalization.  For example, the following
-is a relatively small counterexample returned when QuickChecking an (incorrect)
-implementation in Xmonad, a popular X11 window manager written in
-Haskell~\cite{xmonad}:
+program being evaluated.  Bad definitions can be so slow or so ineffective at
+shrinking that they are unusable.
+
+In addition, shrinking is one side of the coin when it comes to making
+counterexamples more understandable: the other side is generalization.  For
+example, the following is a relatively small counterexample returned when
+QuickChecking a property in Xmonad (with a bug inserted), a popular X11 window
+manager written in Haskell~\cite{xmonad}:
 %
@@
 \noindent
-Functional programmers should be familiar with having to debug a ``wall of
-text'' when showing data structures.  What if instead a counterexample like the
-following were returned by the testing framework:
+(This counterexample uses Haskell's default \ttp{Show} instances, which uses
+record syntax.)  Functional programmers should be familiar with having to debug
+a ``wall of text'' when debugging data structures.  What if instead a
+counterexample like the following were returned by the testing framework:
 %
@@
 The formula quantifies away all the irrelevant portions of the data structure
-with respect to the counterexample, meaning that for any well-typed values that
+with respect to the property, meaning that for any well-typed values that
 replace \ttp{x0}, \ttp{x1}, \ttp{x2}, and \ttp{x3}, the counterexample still
@@
 counterexample---other common shrink implementations often scale in polynomial
-time or worse.  And the counterexamples returned are two to five times smaller.
+time or worse.  (Shrinking time is important in real-world uses; see
+Section~\ref{sec:experiments} for a real-world example.)  And the
+counterexamples returned are two to five times smaller.
 
@@
 \end{code}
-  \caption{QuickCheck and SmallCheck for a product-type input.}
+  \caption{Example program and property.}
   \label{fig:initial}
@@
 %
+\begin{samepage}
 \begin{code}
@@
 \end{code}
+\end{samepage}
 %
@@
 
-The programmer might ask herself: ``What if we omit the need for shrinking
-counterexamples altogether?''  SmallCheck is another testing framework for
-Haskell that does just this.  SmallCheck is guaranteed to return a smallest
-counterexample, if one exists~\cite{sc}.  SmallCheck does this by enumerating
-all possible inputs, ordered from smallest to largest, up to some user-defined
-bound.  While SmallCheck is effective for testing many programs and properties
-(in accordance with the \emph{small scope hypothesis}~\cite{jackson}),
-counterexamples to even relatively simple properties may be infeasible to
-discover due to input-space explosion.
+SmallCheck is another testing framework for Haskell for which shrinking is
+irrelevant.  SmallCheck is guaranteed to return a smallest counterexample, if
+one exists~\cite{sc}.  SmallCheck does this by enumerating all possible inputs,
+ordered from smallest to largest, up to some user-defined bound.  While
+SmallCheck is effective for testing many programs and properties (in accordance
+with the \emph{small scope hypothesis}~\cite{jackson}), counterexamples to even
+relatively simple properties may be infeasible to discover due to input-space
+explosion.
 
@@
 For a product type, SmallCheck must check values produced by taking the
-cross-product of the type's fields at a given depth.  Unfortunately, SmallCheck
-fails long before it reaches a depth required to discover a counterexample.  (A
-related library named Feat combines some aspects of SmallCheck and
-QuickCheck~\cite{feat}; we discuss it in Section~\ref{sec:related}).
+cross-product of the type's fields at a given depth.  Unfortunately, using
+default instances, SmallCheck fails long before it reaches a depth required to
+discover a counterexample.  (A related library named Feat combines some aspects
+of SmallCheck and QuickCheck~\cite{feat}; we discuss it in
+Section~\ref{sec:related}).
 
@@
 %\end{itemize}
-
-Here we provide some illustrative comparisons between SmartCheck's shrinking
-against other shrinking approaches.  As a representative example, we use the
-example presented in Section~\ref{sec:intro}.  The results for SmartCheck are
-fairly generalizable to other types and properties since SmartCheck's reduction
-algorithm has no dependencies on the specifics of either; however, user-defined
-shrink methods may vary significantly based on the property and counterexample
-type.
 
@@
 
-To motivate the benefits of SmartCheck, consider the graphs in
-Figure~\ref{fig:graphs} and the Table~\ref{table:results}, giving the results
+To motivate the benefits of SmartCheck reduction algorithm, consider the graphs
+in Figure~\ref{fig:graphs} and the Table~\ref{table:results}, giving the results
 for using QuickCheck and SmartCheck on the program shown in
@@
 at the 95th percentile values.  (The distributions produced by these values are
-not necessarily Gaussian, so we omit standard the deviation.)
+not necessarily Gaussian, so we omit the standard deviation.)
 
@@
 by QuickCheck; thus the time required by SmartCheck to shrink a counterexample
-is on average just 0.2 seconds (0.10 - 0.08).  As can be seen in the graphs,
-SmartCheck's execution time scales linearly (like QuickCheck, without
-shrinking).  Although it is difficult to tell from the graphs, truncation and
-tuple shrinking produce very long tails, with some counterexamples taking over a
-minute to shrink.  SmartCheck is largely insensitive to the size of the original
-counterexample provided.
+is on average just two hundredths of a second (0.10 - 0.08).  As can be seen in
+the graphs, SmartCheck's execution time scales linearly (like QuickCheck,
+without shrinking).  Truncation and tuple shrinking produce very long tails; the
+longest execution time recorded for truncation shrinking and tuple shrinking are
+88 seconds and 110 seconds, respectively.  SmartCheck is largely insensitive to
+the size of the original counterexample provided.
+
+The results for SmartCheck are fairly generalizable to other types and
+properties since SmartCheck's reduction algorithm has no dependencies on the
+specifics of either; however, user-defined shrink methods may vary significantly
+based on the property and counterexample type.
 
@@
 
-We focus on algebraic data types represented as ``sums of products''.  We
-sometimes refer to the elements of a sum type as a \emph{variant} that is
-\emph{tagged} by a constructor.  For example, the type
+SmartCheck focuses on algebraic data types represented as ``sums of products''
+including recursive and mutually recursive types.  We sometimes refer to the
+elements of a sum type as a \emph{variant} that is \emph{tagged} by a
+constructor.  For example, the type
 %
@@
 To present the algorithms in the following sections, we provide some definitions
-in the form of methods of a \ttp{SubTypes} type class.  SmartCheck requires
-instances for data being analyzed.  In Section~\ref{sec:implementation}, we
-describe how instances for the type class are derived automatically.
+in the form of methods of a \ttp{SubTypes} type class.  (SmartCheck requires
+instances to be defined for data being analyzed; in
+Section~\ref{sec:implementation}, we describe how instances for the type class
+are derived automatically.)
 %
@@
   constrs :: a -> [String]
-  baseType :: a -> Bool
+  opaque  :: a -> Bool
   subVals :: a -> Tree SubVal
@@
 \item \ttp{index} returns a \emph{sub-value} at an breadth-first index in a
-  value,
+  value,\footnote{Rather than indexes, an alternative representation is to use a
+    zipper data structure~\cite{zipper}; see Section~\ref{sec:conclusions} for a
+    discussion.}
 \item \ttp{replace} replaces a sub-value at a particular focus (returning the
@@
   value's type,
-\item \ttp{baseType} is true when the type of the value is an ``uninteresting
+\item \ttp{opaque} is true when the type of the value is an ``uninteresting
   type'', which generally means it is user-defined and has positive arity.
-  We discuss the purpose of base types in Section~\ref{sec:base}.
-\item \ttp{subVals} returns a tree of all non base-type sub-values.  A tree is a
+  We discuss the purpose of opaque types in Section~\ref{sec:base}.
+\item \ttp{subVals} returns a tree of all non opaque-type sub-values.  A tree is a
   value with the type
@@
 
-baseType (3 :: Int) = True
-baseType tree       = False
-baseType L          = False
+opaque (3 :: Int) = True
+opaque tree       = False
+opaque L          = False
 
@@
 
-%% Finally, \ttp{baseType} is true for our tree type but false for a
+%% Finally, \ttp{opaque} is true for our tree type but false for a
 %% structureless type, such as \ttp{Int}.
@@
 graph search algorithm: it simply generates a random set of potential vertexes
-to check in the graph.  However, it does not exploit that the original
-counterexample \ttp{cex} is already in a neighborhood of other likely
-counterexamples.  So rather than choose random vertexes, it may be more
-efficient to test other values that are closely related to \ttp{cex}.  By
-restricting the search to \ttp{cex}'s neighborhood, a smaller counterexample
-might be missed, and only a local minimum is returned.  However, with a large
-set of values, finding a global minimum by random sampling vertexes in the graph
-is possible but too inefficient, as we saw from the example in the introduction.
+that are strictly smaller according to the given measure to check in the graph.
+However, it does not exploit that the original counterexample \ttp{cex} is
+already in a neighborhood of other likely counterexamples.  So rather than
+choose random vertexes, it may be more efficient to test other values that are
+closely related to \ttp{cex}.  By restricting the search to \ttp{cex}'s
+neighborhood, a smaller counterexample might be missed, and only a local minimum
+is returned.  However, with a large set of values, finding a global minimum by
+random sampling vertexes in the graph is possible but too inefficient, as we saw
+from the example in the introduction.
 
@@
 The algorithm we present for efficiently searching for new counterexamples is an
-instance of greedy breadth-first search over a tree structures, representing a
+instance of greedy breadth-first search over a tree structure that represents a
 value.  At each node, during the traversal, we generate arbitrary
@@
 \ttp{reduce} takes flags to customize the algorithm's behavior, a counterexample
-\ttp{cex}, and the property \ttp{prop}.  At the beginning of execution, we
-assume that a counterexample has been obtained somehow; our implementation
-relies on QuickCheck to generate an initial counterexample and to generate
-arbitrary well-typed random values (by the \ttp{arbitrary} method).  The
-reduction begins at the first proper sub-value of \ttp{cex}; call it \ttp{v}.
-When the index \ttp{idx} becomes out-of-bounds and returns \ttp{Nothing}, the
-algorithm terminates.  Otherwise, a list of new random values are generated.
-%
-\begin{code}
-sizedArbitrary :: SubTypes a => a -> IO a
+\ttp{cex}, and the property \ttp{prop}.  The reduction begins at the first
+proper sub-value of \ttp{cex}; call it \ttp{v}.  When the index \ttp{idx}
+becomes out-of-bounds and returns \ttp{Nothing}, the algorithm terminates.
+Otherwise, a list of new random values are generated.
+%
+\begin{code}
+sizedArbitrary :: SubTypes a => Size -> a -> IO a
 \end{code}
@@
 A motivation for the design of the reduction algorithm is something we call the
-\emph{variant counterexample hypothesis}.  Recall that the algorithm begins at
-the \emph{first} sub-value of the counterexample rather than the 0th sub-value.
-The 0th sub-value of \ttp{cex} is \ttp{cex} itself.  No invariant of the
-algorithm would be violated by beginning with the 0th sub-value, and in
-particular, the algorithm would still terminate.  However, we begin with a
-strict sub-value because of the hypothesis.  The hypothesis is that in the
-search space of possible values from a given type \ttp{T}, starting with a
-particular counterexample \ttp{cex}, we are more likely to find other
-counterexamples from the same variant as \ttp{cex}.  For example, consider a
-property about unintended variable capture over a language's parse tree
-represented by a sum type with constructors for module imports, function
-definitions, and global-variable assignments, respectively.  A function
-definition counterexample can only be reduced to smaller function definition
-counterexamples, the only construct in which variable capture is possible.
+\emph{variant counterexample hypothesis}: in the search space of possible values
+from a given type \ttp{T}, if a known counterexample \ttp{cex} is a variant
+\ttp{V} of \ttp{T}, then it is most probable that other counterexamples are also
+from variant \ttp{V}.  Recall that the algorithm begins at the \emph{first} sub-value of
+the counterexample rather than the 0th sub-value.  The 0th sub-value of
+\ttp{cex} is \ttp{cex} itself.  No invariant of the algorithm would be violated
+by beginning with the 0th sub-value, and in particular, the algorithm would
+still terminate.  However, we begin with a strict sub-value because of the
+hypothesis.  For example, consider a property about unintended variable capture
+over a language's parse tree represented by a sum type with constructors for
+module imports, function definitions, and global-variable assignments,
+respectively.  A function definition counterexample can only be reduced to
+smaller function definition counterexamples, the only construct in which
+variable capture is possible.
 
@@
 
-\subsubsection{Base Types}\label{sec:base}
+\subsubsection{Opaque Types}\label{sec:base}
 
 The number of constructors in an algebraic data value primarily determines how
-large it is.  In our experience, the size of fields that come from primitive
+large it is.  In some cases, the size of fields that come from primitive
 types, like \ttp{Int} or \ttp{Char}, have less of a role in understanding a
@@
 
-Consequently, we distinguish a class of types referred to as \emph{base types}.
-If the reduction algorithm encounters a base type, it is ignored.  By default,
-this includes numeric types (e.g., \ttp{Int}, \ttp{Word}, \ttp{Float}, etc.,
-\ttp{Bool}, and \ttp{Char}.  We also treat recursive container types like lists
-specially: lists of base types are considered to be base types themselves.  This
-prevents the algorithm from analyzing, for example, strings.  A list of
-non-base-types, however, is shrunk by the algorithm.
+Consequently, we distinguish a class of types referred to as \emph{opaque
+  types}.  If the reduction algorithm encounters a opaque type, it is ignored.
+By default, this includes types without constructors (e.g., \ttp{Int}, \ttp{Word},
+\ttp{Float}, \ttp{Bool}, and \ttp{Char}).  We also treat recursive container
+types like lists specially: lists of opaque types are considered to be opaque
+types themselves.  This prevents the algorithm from analyzing, for example,
+strings.  A list of non-opaque-types, however, is shrunk by the algorithm.
 
 By providing custom instances, the user can declare any substructure in a data
-type to be a base type.  Doing so effectively treats values from that type as
+type to be a opaque type.  Doing so effectively treats values from that type as
 ``black boxes'' for SmartCheck.  Doing so can help make SmartCheck more
@@
 or is irrelevant to the property.
+
+Conversely, if the user wishes for types that are treated as opaque by default
+by SmartCheck to be shrunk, the easiest approach is to define shrink instances
+that operate on the opaque types while leaving the rest of the value constant.
+These types are shrunk relatively efficiently by conventional QuickCheck's
+pre-defined instances.
 
@@
 \noindent
-is true if in \ttp{cex}, the value at index \ttp{idx0}, is a child of index
-\ttp{idx1}, in a tree representation of \ttp{cex} (i.e., \ttp{subVals cex}).  This
-prevents the algorithm from trying to generalize sub-values that are abstracted
-away already since their parents have been generalized.
+is true if in \ttp{cex}, the value at index \ttp{idx0} is a child of index
+\ttp{idx1} in a tree representation of \ttp{cex} (i.e., \ttp{subVals cex}).  The
+check needs only to compare indexes, so it is efficient.  The \ttp{subTrees}
+guard prevents the algorithm from trying to generalize sub-values that are
+abstracted away already since their parents have been generalized.
 
@@
 The extrapolation algorithm is unsound in two ways.  First, it extrapolates from
-a set of counterexamples to a universal claim, similar to QuickSpec~\cite{qs}.
-By tuning the parameters, the risk of an unsound generalization is reduced by
-requiring more or larger values to fail the property.
+a set of counterexamples to a universal claim, similar to QuickSpec or
+Daikon~\cite{qs,daikon}.  By tuning the parameters, the risk of an unsound
+generalization is reduced by requiring more or larger values to fail the
+property.
 
@@
 \noindent
-The algorithm could be modified so that when attempting to generalize a
-sub-value at an index, all previously-generalized indexes are re-generalized
-simultaneously, by taking the cross-product of random values for the current
-index and for the previously-generalized indexes.  While this reduces the
-potential for over-generalizing a counterexample, it is exponential in the
-number of indexes generalized.  For universally quantified formulas, the reader
-should read it as shorthand for quantifying each variable independently and
-taking the conjunction of formulas.  For example, instead of
+An over-generalized formula is only a problem when generalizing without
+minimizing the counterexample first, as pointed out to this author by John
+Hughes, so in practice with SmartCheck, this form of unsoundness does not arise.
+However, if it does, for universally quantified formulas, the reader should read
+it as shorthand for quantifying each variable independently and taking the
+conjunction of formulas.  For example, instead of
 %
@@
 We therefore introduce another kind of generalization we call \emph{existential
-  sub-value generalization}.  In this generalization, if there a counterexample
-containing every possible variant as a sub-value, then we abstract it.  For
-example, the following are all counterexamples:
-%
-\begin{code}
+  sub-value generalization}.  In this generalization, if there is a
+counterexample containing every possible variant as a sub-value, then we
+abstract it.  For example, suppose that \ttp{prop\_div} had no precondition
+\ttp{divSubTerms} testing for the occurrence of a constant (\ttp{C 0}) in the
+divisor.  Then the following would all counterexamples:
+%
+\begin{code}
+Div (Add (C 3) (C 2)) (C 0)
 Div (C 7) (Add (C (-5)) (C 5))
-Div (Add (C 3) (C 2)) (C 0)
 Div (C (-2)) (Div (C 0) (C 1))
@@
 Figure~\ref{fig:cex-gen} illustrates a property of the form (\ttp{pre ==>
-  post}).  Points are specific counterexamples
-
- the situation in which we have some values
-and a property in the case that some values fail the property of the form
-.  Our goal is to find additional counterexamples in the
-shaded region, in which \ttp{pre} is satisfied, \ttp{post} is not satisfied, and
-the counterexample is not already characterized by a universal sub-value
-generalization formula.  Iterating, we increasingly cover the shaded region
-with new formulas until no more uncovered counterexamples exist or the testing
-tool fails to find new counterexamples.
+  post}).  Points are specific counterexamples that satisfy the precondition but
+fail the post-condition, and the enclosing ring represents the generalization of
+counterexamples resulting from either universal or existential generalization.
+Our goal is to find additional counterexamples in the shaded region.  As new
+counterexamples are discovered in the shaded region (and generalized), the
+counterexample space becomes covered until no more classes of counterexamples
+exist or it becomes too difficult for the testing framework to discover them.
 
@@
 shape if in their respective tree representations, their constructors at each
-node in the tree match, ignoring all base types (Section~\ref{sec:base}).
+node in the tree match, ignoring all opaque types (Section~\ref{sec:base}).
 That is, intuitively, for values \ttp{a} and \ttp{b},
@@
 %
-The first equation holds because we ignore base types.  The second equation
+The first equation holds because we ignore opaque types.  The second equation
 fails because at index 0, \ttp{e1} contains the constructor \ttp{C} and \ttp{e2}
@@
 have the same types.  (We only require \ttp{Read} because standard QuickCheck
-does not return a counterexample; we read the shown value.)
+does not return a counterexample; we read the shown values.)
 
@@
 \ttp{arbitrary} method.  Any initial counterexample can be passed in or
-generated, and the algorithm depends on \ttp{arbitrary} to generate new
+generated, and the algorithm depends on QuickCheck's \ttp{arbitrary} to generate new
 well-typed values.
@@
 operates on a single input during testing, which is assumed to be the first
-argument to the function.  Our experience is that often a program operates over
-one large structure together with smaller inputs.  While multiple inputs could
-be analyzed simultaneously by SmartCheck, it is not obvious how to clearly
-represent multiple large counterexamples to the user---the implementation
-currently prints to standard out.  (In future work, a web-based interface could
-help.)
+argument to the function.  During all stages of SmartCheck's execution, the
+remaining arguments are held constant.  For example, for a property of type
+%
+\begin{code}
+prop :: Foo -> Int -> Bool
+\end{code}
+%
+\noindent
+and a counterexample
+%
+\begin{code}
+(SomeFoo ...) 3
+\end{code}
+%
+\noindent
+returned by QuickCheck, SmartCheck reduces and generalizes \ttp{SomeFoo} with
+respect to the property
+%
+\begin{code}
+prop' :: Foo -> Bool
+prop' foo = prop foo 3
+\end{code}
+
+Our experience is that often a program operates over one large structure
+together with smaller inputs.  While multiple inputs could be analyzed
+simultaneously by SmartCheck, it is not obvious how to clearly represent
+multiple large counterexamples to the user---the implementation currently prints
+to standard out.  (In future work, a web-based interface could help.)
 
@@
 \noindent
-where \ttp{args} are the arguments passed in, and \ttp{prop} is the property
-being tested.  The arguments is a record type that includes the values described
-above to control SmartCheck's behavior.  The standard arguments are contained in
-the value \ttp{scStdArgs}.
+where \ttp{args} (the arguments) are passed in, and \ttp{prop} is the property
+being tested.
+
+The interface types and functions for SmartCheck with analogous behavior to
+QuickCheck's are prefixed with an \ttp{sc} to avoid name space collisions with
+QuickCheck.  So the return type of property must be either \ttp{Bool} or
+\ttp{ScProperty}, a data type that corresponds to QuickCheck's \ttp{Property}
+type.  A function that returns \ttp{Bool} can by turned into one that returns
+\ttp{ScProperty} using the \ttp{scProperty} function (corresponding to
+QuickCheck's \ttp{property} function.  SmartCheck's property implication uses an
+arrow \ttp{-->} corresponding to QuickCheck's \ttp{==>}.  The reason that we
+cannot just use QuickCheck properties directly is that we must do some
+``pre-processing'' to extract the arguments from a counterexample.
+
+The arguments is a record type that includes the values described above to
+control SmartCheck's behavior.  The standard arguments are contained in the
+value \ttp{scStdArgs}.
 
@@
 \subsection{XMonad}
-XMonad is a popular X11 window manager written in Haskell~\cite{xmonad}.  XMonad
-is a large software project with many contributors, so consequently, a
-QuickCheck test harness is included to help ensure new commits do not introduce
-bugs.  At the heart of XMonad is a \ttp{StackSet} data type that encodes the
-relationship between windows, work spaces, and which window has the focus.
-XMonad contains properties to ensure the correct manipulation of
-\ttp{StackSet}s.
+Recall from the introduction the XMonad example.  The XMonad window manager is a
+large software project with many contributors, so consequently, a QuickCheck
+test harness is included to help ensure new commits do not introduce bugs.  At
+the heart of XMonad is a \ttp{StackSet} data type that encodes the relationship
+between windows, work spaces, and which window has the focus.  XMonad contains
+properties to ensure the correct manipulation of \ttp{StackSet}s.
 
@@
 \noindent
-%QuickCheck returns
-\todo{fill in}
-
-This is a relatively small counterexample, but even the smallest \ttp{StackSet}
-values are somewhat large due to the number of fields within it.  But let us see
-if we can do better.
-
-SmartCheck returns a smaller counterexample, and then generalizes it:
+QuickCheck returns the large value shown in the introduction.  That value is a
+relatively small counterexample, but even the smallest \ttp{StackSet} values
+are somewhat large due to the number of fields within it.  Recall the value
+returned by SmartCheck after generalization:
 %
@@
 Let us examine what was generalized and what was not.  First, the last element
-of \ttp{StackSet} is a map, and we treated it as a base type, so it is not
+of \ttp{StackSet} is a map, and we treated it as a opaque type, so it is not
 analyzed at all.  But it is an empty map.  The second and third fields of
@@
 We have benchmarked Feat as well on the example in the introduction
-(Figure~\ref{fig:initial}).  While it performs quite competitively to SmartCheck
+(Figure~\ref{fig:initial}).  It performs quite competitively to SmartCheck
 in terms the size of counterexamples discovered: the mean and median number of
 \ttp{N} constructors in the counterexamples returned by Feat are both seven.
-However, Feat performs more poorly with respect to how long counterexample
+However, Feat performs less well with respect to how long counterexample
 discovery takes: the mean time taken by Feat is 5.60 seconds, with a long tail:
@@
 that infers equational properties about programs~\cite{qs} insofar as they both
-attempt to generalize counterexamples based on specific inputs.
+attempt to generalize counterexamples based on specific inputs.  QuickSpec
+attempts to infer equational properties of programs through random testing.
+Daikon infers assertions for C, C++, Java, and Perl by observing relationships
+between variables in executions of a program~\cite{daikon}.  SmartCheck does not
+attempt to infer properties like these tools do, so it is in one sense less
+general.  However, SmartCheck's three kinds counterexample generalization are novel.
 
@@
 \begin{code}
-Add (Div (C 1) (C (-2)) (C 2))
-    (Div (C 1) (C (-1)) (C 1))
+Add (Div (C 1) (Add (C (-2)) (C 2)))
+    (Div (C 1) (Add (C (-1)) (C 1)))
 \end{code}
@@
 Add (div (-2) 2) (div (-1) 1)
-  where div x y = Div (C x) (C y)
+  where div x y = Div (C 1) (Add (C x) (C y))
 \end{code}
@@
 hand.  We hope SmartCheck and the ideas in this paper continue the tradition of
-highly-automated testing and debugging in the functional programming community.
+highly-automated testing and debugging in the functional programming community
+and beyond!
 
 \section*{Acknowledgments}
-I thank Joe Hurd, John Launchbury, and John Hughes for insightful comments on
-earlier drafts of this paper.  In particular, John Hughes discovered errors and
-omissions that we have fixed.
-
-%% \balancecolumns
+I thank Joe Hurd, John Launchbury, Simon Winwood, and especially John Hughes for
+reading an earlier draft of this paper.  Their deep insights and bug catches
+dramatically improved both the form and content.
 
@@
 
+\balancecolumns
+
 \end{document}

### 2013-03-22 paper/Paper.hs

@@ -61,7 +61,7 @@ class Arbitrary a => SubTypes a where
   subVals  :: a -> Tree SubVal
   constr   :: a -> String
   constrs  :: a -> [String]
-  baseType :: a -> Bool
+  opaque :: a -> Bool
 
 --------------------------------------------------------------------------------
 
@@ -238,7 +238,7 @@ matchesShape a (b, idxs)
               | otherwise    = Nothing
   foldEqConstrs ( Node (SubVal l0) sts0
                 , Node (SubVal l1) sts1 )
-    -- Don't need a baseType test, since they don't ever appear in subTypes.
+    -- Don't need a opaque test, since they don't ever appear in subTypes.
     | constr l0 == constr l1 =
       all foldEqConstrs (zip sts0 sts1)
     | otherwise              = False

### 2013-03-22 src/Test/SmartCheck.hs
smartCheck: change type name.

@@ -9,7 +9,7 @@ module Test.SmartCheck
     smartCheck
 
   -- ** Type of SmartCheck properties.
-  , ScProp()
+  , ScProperty()
   -- ** Implication for SmartCheck properties.
   , (-->)
 
@@ -48,7 +48,7 @@ import Generics.Deriving
 smartCheck :: forall a prop.
   ( Read a, Q.Arbitrary a, SubTypes a
   , Generic a, ConNames (Rep a)
-  , ScProperty prop, Q.Testable prop
+  , ScProp prop, Q.Testable prop
   ) => ScArgs -> (a -> prop) -> IO ()
 smartCheck args scProp = do
   -- Run standard QuickCheck or read in value.
@@ -150,7 +150,7 @@ runAgainMsg = putStrLn $
 
 -- | Run QuickCheck initially, to get counterexamples for each argument,
 -- includding the one we want to focus on for SmartCheck, plus a `Property`.
-runQCInit :: (Show a, Read a, Q.Arbitrary a, ScProperty prop, Q.Testable prop)
+runQCInit :: (Show a, Read a, Q.Arbitrary a, ScProp prop, Q.Testable prop)
           => Q.Args -> (a -> prop) -> IO (Maybe a, a -> Q.Property)
 runQCInit args scProp = do
   res <- Q.quickCheckWithResult args (genProp $ propify scProp)
@@ -193,11 +193,11 @@ genProp prop = Q.forAllShrink Q.arbitrary Q.shrink prop
 
 -- | Type for SmartCheck properties.  Moral equivalent of QuickCheck's
 -- `Property` type.
-data ScProp = Implies (Bool, Bool)
-            | Simple  Bool
+data ScProperty = Implies (Bool, Bool)
+                | Simple  Bool
   deriving (Show, Read, Eq)
 
-instance Q.Testable ScProp where
+instance Q.Testable ScProperty where
   property (Simple prop)         = Q.property prop
   property (Implies prop)        = Q.property (toQCImp prop)
   exhaustive (Simple prop)       = Q.exhaustive prop
@@ -206,7 +206,7 @@ instance Q.Testable ScProp where
 -- same as ==>
 infixr 0 -->
 -- | Moral equivalent of QuickCheck's `==>` operator.
-(-->) :: Bool -> Bool -> ScProp
+(-->) :: Bool -> Bool -> ScProperty
 pre --> post = Implies (pre, post)
 
[…]

### 2013-03-24 paper/paper.tex
paper updates.

@@
 \begin{abstract}
-QuickCheck, developed by Claessen and Hughes, is a powerful library for automatic
-test-case generation.  Because QuickCheck performs random testing, some of the
-counterexamples discovered are very large.  QuickCheck provides an interface for
-the user to write \emph{shrink} functions to attempt to reduce the size of
-counterexamples discovered.  Hand-written implementations of shrink can be
-difficult to write: they may be complex, inefficient, and consist of significant
-boilerplate code.  We introduce \emph{SmartCheck}, an automatic efficient
-counterexample reduction algorithm.  SmartCheck reduces algebraic data using
-generic search heuristics to efficiently find smaller counterexamples
-independent of the program or property being tested.  In addition to shrinking,
-SmartCheck generalizes counterexamples by returning a \emph{formula} that
-abstracts over individual counterexamples and make it easier for the programmer
-to uncover multiple counterexamples at once.  SmartCheck has been implemented
-for Haskell and is freely available.
+QuickCheck, developed by Claessen and Hughes, is a powerful library for
+automatic test-case generation.  Because QuickCheck performs random testing,
+some of the counterexamples discovered are very large.  QuickCheck provides an
+interface for the user to write \emph{shrink} functions to attempt to reduce the
+size of counterexamples discovered.  Hand-written implementations of shrink can
+be complex, inefficient, and consist of significant boilerplate code.  We
+introduce \emph{SmartCheck}, an automatic efficient counterexample reduction
+algorithm.  SmartCheck reduces algebraic data using generic search heuristics to
+efficiently find smaller counterexamples independent of the program or property
+being tested.  In addition to shrinking, SmartCheck generalizes counterexamples
+by returning a \emph{formula} that abstracts over individual counterexamples and
+make it easier for the programmer to uncover multiple counterexamples at once.
+SmartCheck has been implemented for Haskell and is freely available.
 \end{abstract}
@@
 \noindent
-The purpose of \ttp{smaller} is to generate strictly smaller values (according
-to some measure) from a given counterexample.  These new values are then tested
+The purpose of \ttp{smaller} is to generate strictly smaller values, according
+to some measure, from a given counterexample.  These new values are then tested
 to attempt to find a smaller counterexample.  Today, \ttp{smaller} is called
@@
 example, the following is a relatively small counterexample returned when
-QuickChecking a property in Xmonad (with a bug inserted), a popular X11 window
+QuickChecking a property in (a buggy version of) Xmonad, a popular X11 window
 manager written in Haskell~\cite{xmonad}:
@@
 a ``wall of text'' when debugging data structures.  What if instead a
-counterexample like the following were returned by the testing framework:
+counterexample like the following were returned, stating that for any well-typed
+values $x_0$, $x_1$, $x_2$, and $x_3$, a counterexample is returned:
 %
@@
 The formula quantifies away all the irrelevant portions of the data structure
-with respect to the property, meaning that for any well-typed values that
-replace \ttp{x0}, \ttp{x1}, \ttp{x2}, and \ttp{x3}, the counterexample still
-holds.
+with respect to the property, so that the user can focus on the heart of the
+problem in the counterexample.
 
 \paragraph{SmartCheck}
-
 Motivated by the problems of reducing and generalizing large counterexamples, we
-have developed \emph{SmartCheck}.  SmartCheck takes a counterexample produced by
-some oracle---in these experiments, we use QuickCheck---and generically
-minimizes and generalizes the counterexample.
+developed \emph{SmartCheck}.  SmartCheck takes a counterexample produced by some
+oracle---we use QuickCheck---and generically minimizes and generalizes the
+counterexample.
 
@@
 time or worse.  (Shrinking time is important in real-world uses; see
-Section~\ref{sec:experiments} for a real-world example.)  And the
-counterexamples returned are two to five times smaller.
+Section~\ref{sec:experiments} for examples.)  And the counterexamples returned
+are two to five times smaller.
 
@@
 fails the property.  For example, finding counterexamples (\ttp{Left 2}) and
-(\ttp{Right True}) for the type (\ttp{Either Int Bool}) means there exists a
-counterexample regardless of the variant chosen.
+(\ttp{Right True}) for the type
+%
+\begin{code}
+\ttp{Either Int Bool}
+\end{code}
+%
+\noindent
+means there exists a counterexample regardless of the variant chosen.
 
@@
 results in the original AST may fail due to multiple bugs.  During testing, it
-is useful to discover all these possibilities in one go.  In practice, the
-problem is solved by discovering a counterexample \ttp{cex}, abstracting it, and
-then adding a new precondition to the property that informally says ``omit
-counterexamples of form \ttp{cex}.''  Adding preconditions manually is laborious
-and may cause the programmer to make premature fixes to the program, if she
-believes she has isolated the error.
+is useful to discover counterexamples arising from each bug in one go.  In
+practice, the problem is solved by discovering a counterexample \ttp{cex},
+abstracting it, and then adding a new precondition to the property that
+informally says ``omit counterexamples of form \ttp{cex}.''  Adding
+preconditions manually is laborious and may cause the programmer to make
+premature fixes to the program, if she believes she has isolated the error.
 
@@
   GHC-compiled program, using \ttp{-O2}, on a 4-core 2.7GHz Intel~i7 processor.}
-We would like to do better, particularly since the data type and property are so
-simple.
 
@@
 \hline
-           & QC (none)  & QC (trunc) & QC (tuple) & SmartCheck\\
+           & QC none    & QC trunc   & QC tuple   & SmartCheck\\
 \hline \hline
@@
 
-To motivate the benefits of SmartCheck reduction algorithm, consider the graphs
-in Figure~\ref{fig:graphs} and the Table~\ref{table:results}, giving the results
-for using QuickCheck and SmartCheck on the program shown in
-Figure~\ref{fig:initial} (we omit SmallCheck, since as noted, it is not even
-feasible for the example).  We graph counterexample size and execution time.
-Out of 2000 runs, we show the number of runs resulting in a shrunk
-counterexample of a given size (by counting \ttp{N} constructors), and we show
-the number of runs executing within a given time limit (in seconds).  Note that
-we provide two execution-time graphs, since the scale of time scales differ
-dramatically.  In the table, we show the mean, median, median, and the results
-at the 95th percentile values.  (The distributions produced by these values are
-not necessarily Gaussian, so we omit the standard deviation.)
-
-The experiments show three approaches to shrinking with QuickCheck: no
-shrinking, to provide a baseline, truncated shrinking, as described above, in
-which each list is truncated to 10 elements, and tuple shrinking, also described
-above.  Note that SmartCheck produces smaller counterexamples than the other
-approaches, with a very small tail.
+To motivate the benefits of SmartCheck reduction algorithm, consider
+Table~\ref{table:results} and the corresponding graphs in
+Figure~\ref{fig:graphs}, giving the results for using QuickCheck and SmartCheck
+on the program shown in Figure~\ref{fig:initial} (we omit SmallCheck, since as
+noted, it is not even feasible for the example).  We graph the final
+counterexample size and execution time.  Out of 2000 runs, we show the number of
+runs resulting in a shrunk counterexample of a given size (by counting \ttp{N}
+constructors), and we show the number of runs executing within a given time
+limit (in seconds).  Note that we provide two execution-time graphs, since time
+scales differ dramatically (we use a logrithmic $y$ axis in one graph).  In the
+table, we show the mean, median, median, and the results at the 95th percentile
+values.  (The distributions produced by these values are not necessarily
+Gaussian, so we omit the standard deviation.)
+
+The experiments show three approaches to shrinking with QuickCheck: no shrinking
+(QC none), to provide a baseline, truncated shrinking (QC trunc), as described
+above, in which each list is truncated to 10 elements, and tuple shrinking (QC
+tuple), also described above.  SmartCheck produces smaller counterexamples than
+the other approaches, with a very small tail.
 
@@
 times for SmartCheck include the initial counterexample generation time required
-by QuickCheck; thus the time required by SmartCheck to shrink a counterexample
-is on average just two hundredths of a second (0.10 - 0.08).  As can be seen in
-the graphs, SmartCheck's execution time scales linearly (like QuickCheck,
-without shrinking).  Truncation and tuple shrinking produce very long tails; the
-longest execution time recorded for truncation shrinking and tuple shrinking are
-88 seconds and 110 seconds, respectively.  SmartCheck is largely insensitive to
-the size of the original counterexample provided.
+by QuickCheck (QC none); thus the time required by SmartCheck to shrink a
+counterexample is on average just two hundredths of a second (0.10 - 0.08).  As
+can be seen in the graphs, SmartCheck's execution time scales linearly (like
+QuickCheck without shrinking).  Truncation and tuple shrinking produce very
+long tails; the longest execution time recorded for truncation shrinking and
+tuple shrinking are 88 seconds and 110 seconds, respectively.  SmartCheck is
+largely insensitive to the size of the original counterexample provided.
 
@@
 generalize it.  Note that for any \ttp{Word8} value in the first argument of
-\ttp{T}, the counterexample still holds.  Thus, we can be more informative and
-return a quantified formula rather than a single counterexample:
+\ttp{T}, the counterexample still holds.  For example, SmartCheck might return a
+counterexample of the form
 %
@@
   value's type,
-\item \ttp{opaque} is true when the type of the value is an ``uninteresting
-  type'', which generally means it is user-defined and has positive arity.
-  We discuss the purpose of opaque types in Section~\ref{sec:base}.
+\item \ttp{opaque} is false when the type of the value is an ``interesting
+  type''; informally, this is a type other than a primitive type like \ttp{Int}
+  or \ttp{Char}.  See Section~\ref{sec:base} for a full discussion.
 \item \ttp{subVals} returns a tree of all non opaque-type sub-values.  A tree is a
@@
 graph search algorithm: it simply generates a random set of potential vertexes
-that are strictly smaller according to the given measure to check in the graph.
-However, it does not exploit that the original counterexample \ttp{cex} is
-already in a neighborhood of other likely counterexamples.  So rather than
-choose random vertexes, it may be more efficient to test other values that are
-closely related to \ttp{cex}.  By restricting the search to \ttp{cex}'s
-neighborhood, a smaller counterexample might be missed, and only a local minimum
-is returned.  However, with a large set of values, finding a global minimum by
-random sampling vertexes in the graph is possible but too inefficient, as we saw
-from the example in the introduction.
+that are strictly smaller according to the given measure to check the property
+against.  However, it does not exploit that the original counterexample
+\ttp{cex} is already in a neighborhood of other likely counterexamples.  So
+rather than choose random vertexes, it may be more efficient to test other
+values that are closely related to \ttp{cex}.  By restricting the search to
+\ttp{cex}'s neighborhood, a smaller counterexample might be missed, and only a
+local minimum is returned.  However, with a large set of values, finding a
+global minimum by random sampling vertexes in the graph is possible but too
+inefficient.
 
@@
 value.  At each node, during the traversal, we generate arbitrary
-\emph{structurally} smaller sub-values and build a new value from that, leaving
+structurally smaller sub-values and build a new value from that, leaving
 the remainder of the tree unchanged.  Intuitively, by a \emph{structurally
@@
 from a given type \ttp{T}, if a known counterexample \ttp{cex} is a variant
-\ttp{V} of \ttp{T}, then it is most probable that other counterexamples are also
-from variant \ttp{V}.  Recall that the algorithm begins at the \emph{first} sub-value of
-the counterexample rather than the 0th sub-value.  The 0th sub-value of
-\ttp{cex} is \ttp{cex} itself.  No invariant of the algorithm would be violated
-by beginning with the 0th sub-value, and in particular, the algorithm would
-still terminate.  However, we begin with a strict sub-value because of the
-hypothesis.  For example, consider a property about unintended variable capture
-over a language's parse tree represented by a sum type with constructors for
-module imports, function definitions, and global-variable assignments,
-respectively.  A function definition counterexample can only be reduced to
-smaller function definition counterexamples, the only construct in which
-variable capture is possible.
+\ttp{v} of \ttp{T}, then it is most probable that other counterexamples are also
+from variant \ttp{v}.  Recall that the algorithm begins at the \emph{first}
+sub-value of the counterexample rather than the 0th sub-value.  The 0th
+sub-value of \ttp{cex} is \ttp{cex} itself.  No invariant of the algorithm would
+be violated by beginning with the 0th sub-value, and in particular, the
+algorithm would still terminate.  However, we begin with a strict sub-value
+because of the hypothesis.  As an example supporting the hypothesis, consider a
+property about unintended variable capture over a language's parse tree
+represented by a sum type with constructors for module imports, function
+definitions, and global-variable assignments, respectively.  A function
+definition counterexample can only be reduced to smaller function definition
+counterexamples, the only construct in which variable capture is possible.
 
@@
 
-Conversely, if the user wishes for types that are treated as opaque by default
-by SmartCheck to be shrunk, the easiest approach is to define shrink instances
-that operate on the opaque types while leaving the rest of the value constant.
-These types are shrunk relatively efficiently by conventional QuickCheck's
-pre-defined instances.
+Conversely, if the user wishes to shrink types that are treated as opaque by
+default, she can sometimes do so before handing the counterexample to
+SmartCheck, say by using QuickCheck's shrink instances to shrink base types
+while leaving the remaining structure for SmartCheck to shrink.  In the future,
+depending on user experiences, default instances for types can be removed so
+that SmartCheck's reduction algorithm does not ignore values from those types.
 
@@
 \noindent
-In the counterexample, the culprit is a buried sub-value,
+In the counterexample, the culprit is a buried sub-value:
 %
@@
 hand, the property itself is a formula that over-approximates the failing
-inputs.  We describe two kinds of formula that fall between these two extremes.
+inputs.  In this section, we describe two kinds of formula that fall between
+these two extremes that we call universal and existential sub-value
+generalization, respectively.  We then describe a third approach to
+generalization to automatically strengthen a property's precondition to obtain
+new counterexamples.
 
 \subsection{Universal Sub-Value Generalization}\label{sec:universal}
-Recall from the introduction the quantified formula
+Recall from the end of Section~\ref{sec:example} the quantified formula
 %
@@
 \ttp{prop}, a breadth-first search over the sub-values of the \ttp{cex} is
-performed.  For each sub-value, the algorithm generates new sub-values, replaces
+performed.  For each sub-value, the algorithm generates new sub-values and replaces
 them in \ttp{cex} to create a list of new potential counterexamples.  If enough
@@
 original counterexample.  In the recursive function \ttp{extrapolate'}, there is
-a function guard in which a call
+a function guard with a call
 %
@@
 \ttp{idx1} in a tree representation of \ttp{cex} (i.e., \ttp{subVals cex}).  The
-check needs only to compare indexes, so it is efficient.  The \ttp{subTrees}
-guard prevents the algorithm from trying to generalize sub-values that are
-abstracted away already since their parents have been generalized.
+\ttp{subTrees} guard prevents the algorithm from trying to generalize sub-values
+that are abstracted away already since their parents have been generalized.
 
@@
 returned that is overly general.  For example, consider the counterexample in
-which both fields of \ttp{Add} contain values causing the failure:
-%
-\begin{code}
-Add (Div (C 1) (C (-2)) (C 2))
-    (Div (C 0) (C (-1)) (C 1))
-\end{code}
-%
-\noindent
-since no matter what random value the first field of \ttp{Add} is replaced with,
+which both arguments of the outermost \ttp{Add} constructor contain values
+causing the failure:
+%
+\begin{code}
+Add (Div (C 1) (Add (C (-2)) (C 2)))
+    (Div (C 0) (Add (C (-1)) (C 1)))
+\end{code}
+%
+\noindent
+Since no matter what random value the first field of \ttp{Add} is replaced with,
 the property fails by the second field and vice versa for the second.  The
@@
 An over-generalized formula is only a problem when generalizing without
-minimizing the counterexample first, as pointed out to this author by John
+minimizing the counterexample first, as pointed out to the author by John
 Hughes, so in practice with SmartCheck, this form of unsoundness does not arise.
-However, if it does, for universally quantified formulas, the reader should read
-it as shorthand for quantifying each variable independently and taking the
-conjunction of formulas.  For example, instead of
+However, if one were to generalize without shrinking first, the reader should
+read a universally quantified formula as shorthand for quantifying each variable
+independently and taking the conjunction of formulas.  For example, instead of
 %
@@
 node in the tree match, ignoring all opaque types (Section~\ref{sec:base}).
-That is, intuitively, for values \ttp{a} and \ttp{b},
+That is, for values \ttp{a} and \ttp{b},
 %
@@
 %
-The first equation holds because we ignore opaque types.  The second equation
-fails because at index 0, \ttp{e1} contains the constructor \ttp{C} and \ttp{e2}
-contains the constructor \ttp{Add}.  However, index 0 has been generalized, then
-we ignore the sub-value at index 0, and they match (the third equation).
-The same reasoning holds for the fourth equation.
+The first equation holds because we ignore opaque types, so the arguments to
+\ttp{C} are considered equal.  The second equation fails because at index 0,
+\ttp{e1} contains the constructor \ttp{C} and \ttp{e2} contains the constructor
+\ttp{Add}.  However, index 0 has been generalized, then we ignore the sub-value
+at index 0, and they match (the third equation).  The same reasoning holds for
+the fourth equation.
 
@@
 \noindent
-and a counterexample
+and a pair of arguments for which \ttp{prop} returns \ttp{False}
 %
@@
 \ttp{ScProperty} using the \ttp{scProperty} function (corresponding to
-QuickCheck's \ttp{property} function.  SmartCheck's property implication uses an
-arrow \ttp{-->} corresponding to QuickCheck's \ttp{==>}.  The reason that we
-cannot just use QuickCheck properties directly is that we must do some
-``pre-processing'' to extract the arguments from a counterexample.
+QuickCheck's \ttp{property} function).  SmartCheck's property implication uses
+an arrow \ttp{-->} corresponding to QuickCheck's \ttp{==>}.
 
@@
 \section{Experiments}\label{sec:experiments}
-Here we describe three experiments using SmartCheck, including an XMonad
+Here we describe two experiments using SmartCheck, including an XMonad
 property and a property about a natural language processing library.
@@
 \noindent
-Now, suppose the programmer makes a simple typo (or perhaps has a more serious
-misunderstanding) and instead writes
+Now, suppose the programmer makes a simple typo and instead writes
 %
@@
 some work.  Fortunately, because the program already uses QuickCheck,
-\ttp{arbitrary} instances are already defined, but we had to ensure all relevant
-types derived the \ttp{Read}, \ttp{Show}, \ttp{Typeable}, and \ttp{Generic}
-type classes as well, together with \ttp{SubTypes}.  The work does not require
-understanding the program, but it does require hunting for definitions and
-multiple rounds of recompilation.
+\ttp{arbitrary} instances are already defined, but we had to hunt down and
+define default instances for a large number of definitions.  Fortunately,
+however, it required no understanding of the program or property.  We were able
+to send a substantially reduced and generalized counterexample to the author,
+making the cause of the bug more obvious.
 
@@
 attempts to infer equational properties of programs through random testing.
-Daikon infers assertions for C, C++, Java, and Perl by observing relationships
-between variables in executions of a program~\cite{daikon}.  SmartCheck does not
-attempt to infer properties like these tools do, so it is in one sense less
-general.  However, SmartCheck's three kinds counterexample generalization are novel.
+Similarly, Daikon infers assertions for C, C++, Java, and Perl by observing
+relationships between variables in executions of a program~\cite{daikon}.
+SmartCheck does not attempt to infer properties like these tools do, so it is in
+one sense less general.  However, SmartCheck's three kinds counterexample
+generalization are novel.
 
@@
 hand.  We hope SmartCheck and the ideas in this paper continue the tradition of
-highly-automated testing and debugging in the functional programming community
+highly-automated testing and debugging in the functional programming community,
 and beyond!

### 2013-03-29 paper/paper.tex
final paper updates.

@@
 \begin{abstract}
-QuickCheck, developed by Claessen and Hughes, is a powerful library for
-automatic test-case generation.  Because QuickCheck performs random testing,
-some of the counterexamples discovered are very large.  QuickCheck provides an
-interface for the user to write \emph{shrink} functions to attempt to reduce the
-size of counterexamples discovered.  Hand-written implementations of shrink can
-be complex, inefficient, and consist of significant boilerplate code.  We
-introduce \emph{SmartCheck}, an automatic efficient counterexample reduction
-algorithm.  SmartCheck reduces algebraic data using generic search heuristics to
-efficiently find smaller counterexamples independent of the program or property
-being tested.  In addition to shrinking, SmartCheck generalizes counterexamples
-by returning a \emph{formula} that abstracts over individual counterexamples and
-make it easier for the programmer to uncover multiple counterexamples at once.
+QuickCheck, developed by Claessen and Hughes, is a powerful library for automatic test-case generation.
+Because QuickCheck performs random testing, some of the counterexamples discovered are very large.
+QuickCheck provides an interface for the user to write shrink functions to attempt to reduce the size of counterexamples discovered.
+Hand-written implementations of shrink can be complex, inefficient, and consist of significant boilerplate code.
+Furthermore, shrinking is only one aspect in debugging: counterexample generalization is the process of extrapolating from individual counterexamples to a class of counterexamples, often requiring a flash of insight from the programmer.
+To improve counterexample reduction and generalization, we introduce \emph{SmartCheck}.
+SmartCheck is a debugging tool that reduces algebraic data using generic search heuristics to efficiently find smaller counterexamples independent of the program or property being tested.
+In addition to shrinking, SmartCheck also automatically generalizes counterexamples to formulas representing them.
 SmartCheck has been implemented for Haskell and is freely available.
@@
 The QuickCheck testing framework was a revolutionary step-forward in
-type-directed random testing~\cite{qc}.  Originally designed for Haskell,
+type-directed random testing~\cite{qc,monadic}.  Originally designed for Haskell,
 QuickCheck has been ported to other languages and is a now a widely-used testing
@@
 In addition, shrinking is one side of the coin when it comes to making
-counterexamples more understandable: the other side is generalization.  For
-example, the following is a relatively small counterexample returned when
-QuickChecking a property in (a buggy version of) Xmonad, a popular X11 window
-manager written in Haskell~\cite{xmonad}:
+counterexamples more understandable: the other side is extrapolation from
+individual counterexamples to a class of counterexamples.  This leap of
+abstraction is often implicitly made by the programmer in determining the reason
+why counterexamples fail the property.  For example, the following is a
+relatively small counterexample returned when QuickChecking a property in (a
+buggy version of) Xmonad, a popular X11 window manager written in
+Haskell~\cite{xmonad}:
 %
@@
 a ``wall of text'' when debugging data structures.  What if instead a
-counterexample like the following were returned, stating that for any well-typed
-values $x_0$, $x_1$, $x_2$, and $x_3$, a counterexample is returned:
-%
+formula like the following were returned, stating that for any well-typed
+values $x_0$, $x_1$, $x_2$, and $x_3$, tested, a counterexample is found?
+
+%
+\begin{samepage}
 \begin{code}
@@
 with respect to the property, so that the user can focus on the heart of the
-problem in the counterexample.
+problem in a class of counterexamples.
 
@@
 developed \emph{SmartCheck}.  SmartCheck takes a counterexample produced by some
-oracle---we use QuickCheck---and generically minimizes and generalizes the
+oracle (we use QuickCheck) and generically minimizes and generalizes the
 counterexample.
@@
 can shrink more than an order-of-magnitude faster than efficient hand-written
-shrink implementations for some properties and programs.  More importantly, the
-time required to shrink scales linearly with the size of the
-counterexample---other common shrink implementations often scale in polynomial
-time or worse.  (Shrinking time is important in real-world uses; see
+shrink implementations for some properties and programs.  As well, the time
+required to shrink scales linearly with the size of the counterexample---other
+common shrink implementations often scale in polynomial time or worse.
+(Shrinking time is important in real-world uses; see
 Section~\ref{sec:experiments} for examples.)  And the counterexamples returned
@@
 means there exists a counterexample regardless of the variant chosen.
+Existential generalization is important for large sum types, as found in
+abstract syntax tree (AST) definitions, for example.
 
@@
 noting that there are often multiple ways in which a property may fail; for
-example, a property that parsing a pretty-printed abstract syntax tree (AST)
-results in the original AST may fail due to multiple bugs.  During testing, it
-is useful to discover counterexamples arising from each bug in one go.  In
-practice, the problem is solved by discovering a counterexample \ttp{cex},
-abstracting it, and then adding a new precondition to the property that
-informally says ``omit counterexamples of form \ttp{cex}.''  Adding
-preconditions manually is laborious and may cause the programmer to make
-premature fixes to the program, if she believes she has isolated the error.
+example, a property that parsing a pretty-printed AST results in the original
+AST may fail due to multiple bugs.  During testing, it is useful to discover
+counterexamples arising from each bug in one go.  In practice, the problem is
+solved by discovering a counterexample \ttp{cex}, abstracting it, and then
+adding a new precondition to the property that informally says ``omit
+counterexamples of form \ttp{cex}.''  Adding preconditions manually is laborious
+and may cause the programmer to make premature fixes to the program, if she
+believes she has isolated the error before she actually does.
 
@@
   the language extensions of existential type quantification and scoped type
-  variables.}  Data type \ttp{T} is a product type containing four lists of
-wrapped \ttp{Int16}s and a single wrapped \ttp{Word8}.  For our purposes, it is
-not important what \ttp{T} models, but for concreteness, suppose it models the
-values in four linked lists in an operating system, together with an 8-bit
-status register.
+  variables.  We use Haskell \ttp{Prelude} functions, \ttp{mapMaybe} from
+  \ttp{Data.Maybe}, and \ttp{liftM} and \ttp{replicateM} from
+  \ttp{Control.Monad}.}  Data type \ttp{T} is a product type containing four
+lists of wrapped \ttp{Int16}s and a single wrapped \ttp{Word8}.  For our
+purposes, it is not important what \ttp{T} models, but for concreteness, suppose
+it models the values in four linked lists in an operating system, together with
+an 8-bit status register.
 
@@
 satisfies \ttp{pre} but fails \ttp{post} (the \ttp{==>}) operator in the figure
-implements implication from the QuickCheck library).
+is implication from the QuickCheck library).
 
@@
 First, the result of (\ttp{shrink t}) is null if any list contained in \ttp{t}
-is null, so \ttp{t} will not be shrunk.  More troublesome is that with
+is null, in which case \ttp{t} will not be shrunk.  More troublesome is that with
 QuickCheck's default \ttp{Arbitrary} instances, the length of potential
@@
   where xs = shrink (w, i0, i1, i2, i3)
-        go (w', i0', i1', i2', i3') =
-          T w' i0' i1' i2' i3'
+        go (w', i0', i1', i2', i3')
+           = T w' i0' i1' i2' i3'
 \end{code}
@@
 Still, using this definition, from some initial counterexamples, shrinking may
-take over 15 seconds and result in a final counterexample containing nearly 60
-\ttp{N} constructors.\footnote{All results reported in the paper are with a
-  GHC-compiled program, using \ttp{-O2}, on a 4-core 2.7GHz Intel~i7 processor.}
+take over one minute and result in a final counterexample containing just under
+100 \ttp{N} constructors.\footnote{All results reported in the paper are with a
+  GHC-7.6.1-compiled program, using \ttp{-O2}, on a 4-core 2.7GHz Intel~i7
+  processor.}
 
 SmallCheck is another testing framework for Haskell for which shrinking is
-irrelevant.  SmallCheck is guaranteed to return a smallest counterexample, if
+irrelevant: SmallCheck is guaranteed to return a smallest counterexample, if
 one exists~\cite{sc}.  SmallCheck does this by enumerating all possible inputs,
@@
 on the program shown in Figure~\ref{fig:initial} (we omit SmallCheck, since as
-noted, it is not even feasible for the example).  We graph the final
-counterexample size and execution time.  Out of 2000 runs, we show the number of
-runs resulting in a shrunk counterexample of a given size (by counting \ttp{N}
-constructors), and we show the number of runs executing within a given time
-limit (in seconds).  Note that we provide two execution-time graphs, since time
-scales differ dramatically (we use a logrithmic $y$ axis in one graph).  In the
-table, we show the mean, median, median, and the results at the 95th percentile
-values.  (The distributions produced by these values are not necessarily
-Gaussian, so we omit the standard deviation.)
+noted, it is not even feasible for the example with default instances).  We
+graph the final counterexample size and execution time.  Out of 2000 runs, we
+show the number of runs resulting in a shrunk counterexample of a given size (by
+counting \ttp{N} constructors), and we show the number of runs executing within
+a given time limit (in seconds).  Note that we provide two execution-time
+graphs, since time scales differ dramatically (we use a logrithmic scale for the
+$y$ axis in one graph and a linear scale in the other).  In the table, we show
+the mean, median, median, and the results at the 95th percentile values.  (The
+distributions produced by these values are not necessarily Gaussian, so we omit
+the standard deviation.)
 
@@
 can be seen in the graphs, SmartCheck's execution time scales linearly (like
-QuickCheck without shrinking).  Truncation and tuple shrinking produce very
-long tails; the longest execution time recorded for truncation shrinking and
-tuple shrinking are 88 seconds and 110 seconds, respectively.  SmartCheck is
-largely insensitive to the size of the original counterexample provided.
-
-The results for SmartCheck are fairly generalizable to other types and
-properties since SmartCheck's reduction algorithm has no dependencies on the
-specifics of either; however, user-defined shrink methods may vary significantly
-based on the property and counterexample type.
-
-Finally, note that after shrinking the counterexample, SmartCheck will also
-generalize it.  Note that for any \ttp{Word8} value in the first argument of
-\ttp{T}, the counterexample still holds.  For example, SmartCheck might return a
-counterexample of the form
+QuickCheck without shrinking).  Truncation and tuple shrinking produce very long
+tails; the longest execution time recorded for truncation shrinking and tuple
+shrinking are 88 seconds and 110 seconds, respectively, although these outliers
+can vary wildly.  SmartCheck is largely insensitive to the size of the original
+counterexample provided.
+
+In regards to generalization, after shrinking the counterexample, SmartCheck
+will also generalize it.  Note that for any \ttp{Word8} value in the first
+argument of \ttp{T}, the counterexample still holds.  For example, SmartCheck
+might return a counterexample of the form
 %
@@
 \end{code}
+
+Finally, these performance results for SmartCheck are fairly generalizable to
+other types and properties, since the algorithm is not dependent on the specific
+program or property being checked.  However, the results are particularly
+pronounced for counterexamples that are particularly ``deep'' or ``broad'' in
+terms of their tree structure.  For programs and properties for which small
+counterexamples are easily found, QuickCheck's shrinking is perfectly
+acceptable.  Furthermore, user-defined shrink methods may vary significantly
+based on the property and counterexample type, so it is not always clear which
+implementation to compare against.
 
@@
 \section{Shrinking Data}\label{sec:shrinking}
-In this section, we describe how to efficiently and generically shrink algebraic data
-values.  Before we begin, recall the basic idea behind the \ttp{shrink} method
-of the \ttp{Arbitrary} class: generate a list of values, each of which is
+In this section, we describe how to efficiently and generically shrink algebraic
+data values.  Before we begin, recall the basic idea behind the \ttp{shrink}
+method of the \ttp{Arbitrary} class: generate a list of values, each of which is
 smaller than the current counterexample.  Each of the new values generated
-typically bears no relationship to the original counterexample other than being
-smaller.
-
-Consider the problem of finding a small counterexample as a graph search
-problem.  The set of vertexes are all possible well-typed values.  There is an
-edge from vertex $v$ to $v'$ if $v'$ is smaller than $v$ by some measure.  The
-goal is to traverse the graph, starting from a counterexample, to find a vertex
-from the set of smallest counterexamples in the graph.
-
-The approach taken by typical \ttp{shrink} implementations is the most naive
-graph search algorithm: it simply generates a random set of potential vertexes
-that are strictly smaller according to the given measure to check the property
-against.  However, it does not exploit that the original counterexample
-\ttp{cex} is already in a neighborhood of other likely counterexamples.  So
-rather than choose random vertexes, it may be more efficient to test other
-values that are closely related to \ttp{cex}.  By restricting the search to
-\ttp{cex}'s neighborhood, a smaller counterexample might be missed, and only a
-local minimum is returned.  However, with a large set of values, finding a
-global minimum by random sampling vertexes in the graph is possible but too
-inefficient.
+may not bear any relationship to the original counterexample other than being
+smaller.  SmartCheck pursues an approach that searches for smaller but
+structurally similar counterexamples, as we make precise below.
+
+We describe the algorithm in Section~\ref{sec:reduct} and then describe
+algorithmic details in Section~\ref{sec:details}.  Some optimizations to the
+reduction algorithm are described in Section~\ref{sec:optimizations}.
+
+%% Consider the problem of finding a small counterexample as a graph search
+%% problem.  The set of vertexes are all possible well-typed values.  Let $<$ be
+%% a partial ordering on values relating them by some size measure.  Let there
+%% be an edge from vertex $v$ to $v'$ if $v < v'$ and there exists no $v''$ such
+%% that $v < v'' < v'$.  The goal is to traverse the graph, starting from a
+%% counterexample, to find a vertex from the set of smallest counterexamples in
+%% the graph.
+
+%% The approach taken by typical \ttp{shrink} implementations is the most naive
+%% graph search algorithm: it simply generates a random set of potential vertexes
+%% that are strictly smaller according to the given measure to check the property
+%% against.  However, it does not exploit that the original counterexample
+%% \ttp{cex} is already in a neighborhood of other likely counterexamples.  So
+%% rather than choose random vertexes, it may be more efficient to test other
+%% values that are closely related to \ttp{cex}.  By restricting the search to
+%% \ttp{cex}'s neighborhood, a smaller counterexample might be missed, and only a
+%% local minimum is returned.  However, with a large set of values, finding a
+%% global minimum by random sampling vertexes in the graph is possible but too
+%% inefficient.
 
@@
 
-\subsection{Reduction Algorithm Details}
+\subsection{Reduction Algorithm Details}\label{sec:details}
 Having described the reduction algorithm, there are two important details about
@@
   types}.  If the reduction algorithm encounters a opaque type, it is ignored.
-By default, this includes types without constructors (e.g., \ttp{Int}, \ttp{Word},
+By default, this includes ``primitive'' types (e.g., \ttp{Int}, \ttp{Word},
 \ttp{Float}, \ttp{Bool}, and \ttp{Char}).  We also treat recursive container
@@
 type to be a opaque type.  Doing so effectively treats values from that type as
-``black boxes'' for SmartCheck.  Doing so can help make SmartCheck more
-efficient if the user knows that some portion of the structure cannot be shrunk
-or is irrelevant to the property.
+``black boxes'' for SmartCheck, making SmartCheck more efficient if the user
+knows that some portion of the structure cannot be shrunk or is irrelevant to
+the property.
 
@@
 
-\subsection{Reduction Algorithm Optimizations}
+\subsection{Reduction Algorithm Optimizations}\label{sec:optimizations}
 The reduction algorithm description above omits some details and optimizations
@@
 divSubTerms (C _)         = True
-divSubTerms (Div _ (C 0)) = False
+odivSubTerms (Div _ (C 0)) = False
 divSubTerms (Add e0 e1)   =  divSubTerms e0
@@
 \medskip%
-\noindent
-In the counterexample, the culprit is a buried sub-value:
+%
+\noindent
+The cause is that \ttp{divSubTerms} fails to check whether the
+divisor evaluates to zero.  In the counterexample, the culprit is a buried
+sub-value:
 %
@@
 %
+\noindent
 Thus, when attempting to shrink an \ttp{Exp} value, it pays to test whether a
@@
 \ttp{prop}, a breadth-first search over the sub-values of the \ttp{cex} is
-performed.  For each sub-value, the algorithm generates new sub-values and replaces
-them in \ttp{cex} to create a list of new potential counterexamples.  If enough
-new counterexamples non-trivially fail \ttp{prop} (i.e., fail the property's
-consequent but satisfy its antecendent), then we extrapolate, claiming that for
+performed.  For each sub-value, the algorithm generates new sub-values and
+replaces them in \ttp{cex} to create a list of new potential counterexamples.
+If no new value satisfies the property, then we extrapolate, claiming that for
 \emph{any} new value replacing that sub-value in \ttp{cex}, the property will
@@
 The function \ttp{allFail} takes a counterexample \ttp{cex}, an index into
-\ttp{cex}, a list of new sub-values, and a property.  It returns true if enough
-new values pass the property's precondition and fail the property.  The function
+\ttp{cex}, a list of new sub-values, and a property.  It returns true if no new
+values satisfy the property.  The function
 %
@@
 
-Second, note that each index is generalized with respect to the concrete
-sub-values that exist in the counterexample.  So in some cases, a formula may be
-returned that is overly general.  For example, consider the counterexample in
-which both arguments of the outermost \ttp{Add} constructor contain values
-causing the failure:
+In some cases, a formula may be returned that is overly general.  For example,
+consider the counterexample in which both arguments of the outermost \ttp{Add}
+constructor contain values causing the failure:
 %
@@
 counterexample containing every possible variant as a sub-value, then we
-abstract it.  For example, suppose that \ttp{prop\_div} had no precondition
-\ttp{divSubTerms} testing for the occurrence of a constant (\ttp{C 0}) in the
-divisor.  Then the following would all counterexamples:
+abstract it.  For example, suppose that \ttp{divSubTerms} had no equation
+%
+\begin{code}
+divSubTerms (Div _ (C 0)) = False
+\end{code}
+%
+\noindent
+Then the following would all counterexamples:
 %
@@
 additional precondition that the current constructor is not an element of
-constructors already seen.  Thus, (\ttp{go v}) returns (\ttp{Just True}) if
-(\ttp{replace cex idx v}) passes this precondition (and any other preconditions
-of \ttp{prop}), but fails the property.
+constructors already seen.  Thus, (\ttp{go v}) returns
+%
+\begin{code}
+\ttp{Just True}
+\end{code}
+%
+\noindent
+if
+%
+\begin{code}
+\ttp{replace cex idx v}
+\end{code}
+%
+passes this precondition (and any other preconditions of \ttp{prop}), but fails
+the property.
 
@@
   post}).  Points are specific counterexamples that satisfy the precondition but
-fail the post-condition, and the enclosing ring represents the generalization of
+fail the post-condition, and the enclosing oval represents the generalization of
 counterexamples resulting from either universal or existential generalization.
@@
 The data type to be tested must derive the \ttp{Read}, \ttp{Show},
-\ttp{Typeable}, and \ttp{Generic} type classes.  Deriving \ttp{Typeable} and
-\ttp{Generic} require using the GHC language extensions \ttp{DeriveDataTypeable}
-and \ttp{DeriveGeneric}, respectively.  \ttp{Typeable} is used for dynamic
-typing in defining the \ttp{replace} method of \ttp{SubTypes} since it is
-unknown at compile-time whether the value to be replaced and replacing value
-have the same types.  (We only require \ttp{Read} because standard QuickCheck
-does not return a counterexample; we read the shown values.)
+\ttp{Typeable}, and \ttp{Generic} type classes.  (We only require \ttp{Read}
+because standard QuickCheck does not return a counterexample; we read the shown
+values.)  Deriving \ttp{Typeable} and \ttp{Generic} require using the GHC
+language extensions \ttp{DeriveDataTypeable} and \ttp{DeriveGeneric},
+respectively.  \ttp{Typeable} is used for dynamic typing in defining the
+\ttp{replace} method of \ttp{SubTypes} since it is unknown at compile-time
+whether the value to be replaced and replacing value have the same types.
+However, through SmartCheck's interface, run-time failures due to
+type-mismatches will not occur.
 
@@
 
-Our experience is that often a program operates over one large structure
+Our experience is that often a program operates over one large data structure
 together with smaller inputs.  While multiple inputs could be analyzed
@@
 %
+%\scriptsize
 \begin{samepage}
 \begin{code}
-\scriptsize
 Div
@@
 were due to universal generalization, like in the XMonad case: entire record
-fields are abstracted away.
+fields are abstracted away.  From that, we (syntactically) shrunk the
+counterexample by-hand further by naming common sub-expressions.
+
+We were able to send a substantially reduced and generalized counterexample to
+the message poster, making the cause of the bug more obvious.  He responded:
+``$\ldots$While your improved shrinking may not have gone `all'
+the way to the bottom, it got me a \emph{huge} chunk of the way there!''  Recall
+that through the entire process, we never had to learn how GenI works, what the
+property meant, or how to write a custom shrink method!
 
@@
 some work.  Fortunately, because the program already uses QuickCheck,
-\ttp{arbitrary} instances are already defined, but we had to hunt down and
-define default instances for a large number of definitions.  Fortunately,
-however, it required no understanding of the program or property.  We were able
-to send a substantially reduced and generalized counterexample to the author,
-making the cause of the bug more obvious.
+\ttp{arbitrary} instances were already defined, but we had to hunt down and
+define default instances for a large number of definitions.
 
@@
 
-A few other researchers have also investigated counterexample shrinking.  Zeller
+Other researchers have also investigated counterexample shrinking.  Zeller
 and Hildebrandt describe an application of greedy search to shrink
 counterexamples they call ``delta-debugging'' (DD)~\cite{dd}.  The authors apply
-their work shrinking HTML inputs to crash Mozilla and shrinking C programs to
+their work to shrinking HTML inputs to crash Mozilla and shrinking C programs to
 trigger a bug in GCC.  Subsequent generalizations are reported by Misherghi and
@@
 generic shrinking is as a motivation for generic programming in Haskell's
-``Scrap your boilerplate'' generic programming library~\cite{syb}.
+``Scrap your boilerplate'' generic programming library~\cite{syb}.  There, the
+motivation was not to design new approaches to counterexample reduction, but
+simply to derive instances for the \ttp{shrink} method.
 
@@
 Consequently, defining custom generators (i.e., instances for \ttp{arbitrary})
-is highly constrained.  On the other hand, generators can be derived
+is constrained.  On the other hand, generators can be derived
 automatically with Feat.
@@
 \noindent
-Discovering and exploiting sharing is future work.
-
-We believe there are interesting connections between counterexample
-discovery and probabilistic reasoning.  Often, when a bug in a program is fixed,
-the fix is not correct, or it introduces new bugs.  Such knowledge might be
-useful in guiding the distribution of test cases or shrinking.
+Discovering and exploiting sharing automatically is future work.
+
+%% We believe there are interesting connections between counterexample
+%% discovery and probabilistic reasoning.  Often, when a bug in a program is fixed,
+%% the fix is not correct, or it introduces new bugs.  Such knowledge might be
+%% useful in guiding the distribution of test cases or shrinking.
 
@@
 \section*{Acknowledgments}
-I thank Joe Hurd, John Launchbury, Simon Winwood, and especially John Hughes for
-reading an earlier draft of this paper.  Their deep insights and bug catches
-dramatically improved both the form and content.
+I thank Andy Gill, Joe Hurd, John Launchbury, Simon Winwood, and especially John
+Hughes for reading an earlier draft of this paper.  Their deep insights and bug
+catches dramatically improved both the form and content.
 

### 2014-03-14 src/Test/SmartCheck.hs
sc: wibbles.

@@ -61,7 +61,7 @@ smartCheckInput :: forall a prop.
 smartCheckInput args scProp = do
   smartPrtLn "Input value to SmartCheck:"
   mcex <- fmap Just (readLn :: IO a)
-  smartCheckRun args (mcex, propify scProp)
+  smartCheckRun args (mcex, Q.property . scProp)
 
 smartCheckRun :: forall a.
   ( SubTypes a
@@ -69,7 +69,7 @@ smartCheckRun :: forall a.
   ) => ScArgs -> (Maybe a, a -> Q.Property) -> IO ()
 smartCheckRun args (origMcex, origProp) = do
   smartPrtLn $
-    "(If any stage takes too long, try modifying the standard "
+    "(If any stage takes too long, try modifying SmartCheck's standard "
       ++ "arguments (see Args.hs).)"
   smartCheck' [] origMcex origProp
   where
@@ -102,7 +102,6 @@ smartCheckRun args (origMcex, origProp) = do
                 let matchesProp a =
                             not (matchesShapes a oldVals)
                       Q.==> prop a
---                mcex' <- runQC (qcArgs args) matchesProp
                 (mcex', _) <- runQC (qcArgs args) (Q.noShrinking . matchesProp)
                 smartCheck' oldVals mcex' matchesProp
         else smartPrtLn "Done."
@@ -152,10 +151,13 @@ runAgainMsg = putStrLn $
 
 --------------------------------------------------------------------------------
 
--- | Run QuickCheck initially, to get counterexamples for each argument,
--- includding the one we want to focus on for SmartCheck, plus a 'Property'.  We
--- allow QuickCheck to attempt to shrink all arguments except the first here.
-runQC :: (Show a, Q.Arbitrary a, Q.Testable prop)
+-- | Run QuickCheck, to get a counterexamples for each argument, including the
+-- one we want to focus on for SmartCheck, which is the first argument.  That
+-- argument is never shrunk by QuickCheck, but others may be shrunk by
+-- QuickCheck.  Returns the value (if it exists) and a 'Property' (by applying
+-- the 'property' method to the 'Testable' value).  In each iteration of
+-- 'runQC', non-SmartCheck arguments are not necessarily held constant
+runQC :: forall a prop . (Show a, Q.Arbitrary a, Q.Testable prop)
           => Q.Args -> (a -> prop) -> IO (Maybe a, a -> Q.Property)
 runQC args scProp = do
   (mCex, res) <- scQuickCheckWithResult args scProp
[…]

### 2014-03-14 src/Test/SmartCheck/Test.hs

@@ -35,8 +35,10 @@ OF THIS SOFTWARE, EVEN IF ADVISED OF THE POSSIBILITY OF SUCH DAMAGE.
 
 {-# LANGUAGE ScopedTypeVariables #-}
 {-# LANGUAGE NamedFieldPuns #-}
+{-# LANGUAGE ExistentialQuantification #-}
 
--- | The main test loop.
+-- | SmartCheck's interface to QuickCheck.
+
 module Test.SmartCheck.Test
   ( scQuickCheckWithResult
   , stdArgs
@@ -164,12 +166,17 @@ giveUp st _f =
        )
      success st
      theOutput <- terminalOutput (terminal st)
-     return (Nothing, GaveUp{ numTests = numSuccessTests st
-                            , labels   = summary st
-                            , output   = theOutput
-                            })
+     return ( Nothing
+            , GaveUp{ numTests = numSuccessTests st
+                    , labels   = summary st
+                    , output   = theOutput
+                    }
+            )
 
-runATest :: forall a. (Arbitrary a) => State -> (QCGen -> Int -> (a -> Prop)) -> IO (Maybe a, Result)
+runATest :: forall a. (Arbitrary a)
+         => State
+         -> (QCGen -> Int -> (a -> Prop))
+         -> IO (Maybe a, Result)
 runATest st f =
   do -- CALLBACK before_test
      putTemp (terminal st)
@@ -189,7 +196,7 @@ runATest st f =
          genA = unGen arbitrary
      let rndA = genA rnd1 size
 
-     let mkRes a = return (Just rndA, a)
+     let mkRes res = return (Just rndA, res)
 
      MkRose res ts <- protectRose (reduceRose (unProp (p rndA)))
      callbackPostTest st res
@@ -226,7 +233,8 @@ runATest st f =
                               numTests = numSuccessTests st+1,
                               output = theOutput }
              else
-              mkRes  Failure{ usedSeed       = randomSeed st -- correct! (this will be split first)
+              mkRes  Failure{ -- correct! (this will be split first)
+                              usedSeed       = randomSeed st
                             , usedSize       = size
                             , numTests       = numSuccessTests st+1
                             , numShrinks     = numShrinks

### 2014-03-14 examples/Comp.hs
examples: remove unnec. derives.

@@ -13,7 +13,7 @@ import Data.Typeable
 -----------------------------------------------------------------
 
 data I = I Int
-  deriving (Eq, Read, Show, Typeable, Generic)
+  deriving (Eq, Show, Typeable, Generic)
 
 instance Arbitrary I where
   arbitrary = liftM I arbitrary
@@ -25,7 +25,7 @@ instance Ord I where
 
 data X = X0 I
        | X1 I
-  deriving (Read, Show, Typeable, Generic)
+  deriving (Show, Typeable, Generic)
 
 instance Arbitrary X where
   arbitrary = oneof [ liftM X0 arbitrary
@@ -47,7 +47,7 @@ instance Eq X where
 
 data A = A0 Int
        | A1 Int
-  deriving (Eq, Read, Show, Typeable, Generic)
+  deriving (Eq, Show, Typeable, Generic)
 
 instance Ord A where
   compare (A0 i) (A0 j) = compare i j
@@ -62,7 +62,7 @@ instance Arbitrary A where
 instance SubTypes A
 
 data L = L [Int]
-  deriving (Read, Show, Typeable, Generic)
+  deriving (Show, Typeable, Generic)
 
 instance Arbitrary L where
   arbitrary = liftM L arbitrary
@@ -70,7 +70,7 @@ instance Arbitrary L where
 instance SubTypes L
 
 data B = B L X
-  deriving (Read, Show, Typeable, Generic)
+  deriving (Show, Typeable, Generic)
 
 instance Arbitrary B where
   arbitrary = liftM2 B arbitrary arbitrary

### 2014-03-14 paper/paper.tex
grammar.
paper: tweak, commit .pdf so folks can read it.

@@
 relatively small counterexample returned when QuickChecking a property in (a
-buggy version of) Xmonad, a popular X11 window manager written in
+bug-injected version of) Xmonad, a popular X11 window manager written in
 Haskell~\cite{xmonad}:
@@
 (This counterexample uses Haskell's default \ttp{Show} instances, which uses
-record syntax.)  Functional programmers should be familiar with having to debug
-a ``wall of text'' when debugging data structures.  What if instead a
+record syntax.)  Programmers should be familiar with having to debug
+a ``wall of text'' as shown above.  What if instead a
 formula like the following were returned, stating that for any well-typed
@@
 developed \emph{SmartCheck}.  SmartCheck takes a counterexample produced by some
-oracle (we use QuickCheck) and generically minimizes and generalizes the
-counterexample.
-
-As demonstrated by a motivating example in Section~\ref{sec:example}, SmartCheck
+oracle and generically minimizes and generalizes the counterexample.
+
+As shown in the motivating example in Section~\ref{sec:example}, SmartCheck
 can shrink more than an order-of-magnitude faster than efficient hand-written
@@
 (Shrinking time is important in real-world uses; see
-Section~\ref{sec:experiments} for examples.)  And the counterexamples returned
-are two to five times smaller.
+Section~\ref{sec:experiments} for examples.)
 
@@
 counterexamples, which are described in Section~\ref{sec:generalization}.  The
-first algorithm universally quantifies sub-values that always fail in tests, as
-briefly described earlier in the introduction.  The second algorithm
-existentially quantifies sub-values for types in which every possible variant
-fails the property.  For example, finding counterexamples (\ttp{Left 2}) and
-(\ttp{Right True}) for the type
+first algorithm universally quantifies sub-values that always fail in tests.
+The second algorithm existentially quantifies sub-values for types in which
+every possible variant fails the property.  For example, finding counterexamples
+(\ttp{Left 2}) and (\ttp{Right True}) for the type
 %
@@
 means there exists a counterexample regardless of the variant chosen.
-Existential generalization is important for large sum types, as found in
+Existential generalization is useful for large sum types, as found in
 abstract syntax tree (AST) definitions, for example.
@@
 noting that there are often multiple ways in which a property may fail; for
-example, a property that parsing a pretty-printed AST results in the original
-AST may fail due to multiple bugs.  During testing, it is useful to discover
-counterexamples arising from each bug in one go.  In practice, the problem is
-solved by discovering a counterexample \ttp{cex}, abstracting it, and then
-adding a new precondition to the property that informally says ``omit
-counterexamples of form \ttp{cex}.''  Adding preconditions manually is laborious
-and may cause the programmer to make premature fixes to the program, if she
-believes she has isolated the error before she actually does.
+example, a property stating that pretty-printing an AST and then parsing it
+results in the original AST may fail due to multiple bugs.  During testing, it
+is useful to discover counterexamples arising from each bug in one go.  In
+practice, the problem is solved by discovering a counterexample \ttp{cex},
+abstracting it, and then adding a new precondition to the property that
+informally says ``omit counterexamples of form \ttp{cex}.''  Adding
+preconditions manually is laborious and may cause the programmer to make
+premature fixes to the program, if she believes she has isolated the error
+before she actually does.
 
@@
 \begin{code}
-data N a = N a
-type I   = [N Int16]
-data T   = T (N Word8) I I I I
+type I   = [Int16]
+data T   = T (Word8) I I I I
 
@@
 toList (T w i0 i1 i2 i3) =
-  [go w] : (map . map) go [i0, i1, i2, i3]
-  where go (N i) = fromIntegral i
+  [fromIntegral w] : (map . map) fromIntegral rst
+  where
+  rst = [i0, i1, i2, i3]
 
@@
 
+\todo{redo analysis, check text}
 In this section, we motivate in more detail the challenges in shrinking
@@
 \begin{code}
-T (N 10) [N (-20000)] [N (-20000)] [] []
-\end{code}
-%
-\noindent
-satisfies \ttp{pre} but fails \ttp{post} (the \ttp{==>}) operator in the figure
+T 10 [-20000] [-20000] [] []
+\end{code}
+%
+\noindent
+satisfies \ttp{pre} but fails \ttp{post} (the \ttp{==>} operator in the figure
 is implication from the QuickCheck library).
@@
 QuickCheck can be large.  With standard settings and no shrinking, the median
-counterexample discovered contains around 70 \ttp{N} constructors, and
-counterexamples can contain over 100 \ttp{N} constructors.  Thus, it pays to define
-\ttp{shrink}!
+counterexample discovered contains around 70 \ttp{Int16} values, and
+counterexamples can contain over 100 values.  \todo{recheck} Thus, it pays to
+define \ttp{shrink}!
 
@@
 take over one minute and result in a final counterexample containing just under
-100 \ttp{N} constructors.\footnote{All results reported in the paper are with a
-  GHC-7.6.1-compiled program, using \ttp{-O2}, on a 4-core 2.7GHz Intel~i7
+100 \ttp{Int16} values.\footnote{All results reported in the paper are with a
+  GHC-7.6.3-compiled program, using \ttp{-O2}, on a 4-core 2.7GHz Intel~i7
   processor.}
@@
 \begin{code}
-forall x . T x [N (-20000)] [N (-20000)] [] []
+forall x . T x [-20000] [-20000] [] []
 \end{code}
@@
 
-The data type to be tested must derive the \ttp{Read}, \ttp{Show},
-\ttp{Typeable}, and \ttp{Generic} type classes.  (We only require \ttp{Read}
-because standard QuickCheck does not return a counterexample; we read the shown
-values.)  Deriving \ttp{Typeable} and \ttp{Generic} require using the GHC
+The data type to be tested must derive the \ttp{Typeable}, and \ttp{Generic}
+type classes.  Deriving \ttp{Typeable} and \ttp{Generic} require using the GHC
 language extensions \ttp{DeriveDataTypeable} and \ttp{DeriveGeneric},
@@
 However, through SmartCheck's interface, run-time failures due to
-type-mismatches will not occur.
+type-mismatches will not occur.  Additionally, like with QuickCheck, deriving
+\ttp{Show} is required to print counterexamples discovered.
 
@@
 
-A property may be a function taking any number of arguments, but SmartCheck only
-operates on a single input during testing, which is assumed to be the first
-argument to the function.  During all stages of SmartCheck's execution, the
-remaining arguments are held constant.  For example, for a property of type
-%
-\begin{code}
-prop :: Foo -> Int -> Bool
-\end{code}
-%
-\noindent
-and a pair of arguments for which \ttp{prop} returns \ttp{False}
-%
-\begin{code}
-(SomeFoo ...) 3
-\end{code}
-%
-\noindent
-returned by QuickCheck, SmartCheck reduces and generalizes \ttp{SomeFoo} with
-respect to the property
-%
-\begin{code}
-prop' :: Foo -> Bool
-prop' foo = prop foo 3
-\end{code}
-
-Our experience is that often a program operates over one large data structure
-together with smaller inputs.  While multiple inputs could be analyzed
-simultaneously by SmartCheck, it is not obvious how to clearly represent
-multiple large counterexamples to the user---the implementation currently prints
-to standard out.  (In future work, a web-based interface could help.)
+The kinds of programs SmartCheck is specialized for are ones that operate over a
+large data structure together with smaller inputs.  Therefore, properties
+provided to SmartCheck are expected to be of the form
+%
+\begin{code}
+Testable prop => a -> prop
+\end{code}
+%
+where \ttp{a} is the type of the value for SmartCheck to analyze, and \ttp{prop}
+is a testable property, as defined by QuickCheck; morally, these are functions
+(or degenerately, values) that evaluate to a Boolean value.
+
+If QuickCheck is used to discover a counterexample, all arguments except the
+first are shrunk, if their types have \ttp{shrink} methods defined for them.
+The first argument is returned to SmartCheck to be shrunk or generalized
+according to the algorithms described earlier.
+
+%% \todo{is still subject to QC's maxsize constraints}
+%% During all stages of SmartCheck's execution, the
+%% remaining arguments are held constant.  For example, for a property of type
+%% %
+%% \begin{code}
+%% prop :: Foo -> Int -> Bool
+%% \end{code}
+%% %
+%% \noindent
+%% and a pair of arguments for which \ttp{prop} returns \ttp{False}
+%% %
+%% \begin{code}
+%% (SomeFoo ...) 3
+%% \end{code}
+%% %
+%% \noindent
+%% returned by QuickCheck, SmartCheck reduces and generalizes \ttp{SomeFoo} with
+%% respect to the property
+%% %
+%% \begin{code}
+%% prop' :: Foo -> Bool
+%% prop' foo = prop foo 3
+%% \end{code}
 
@@
 QuickCheck's are prefixed with an \ttp{sc} to avoid name space collisions with
-QuickCheck.  So the return type of property must be either \ttp{Bool} or
-\ttp{ScProperty}, a data type that corresponds to QuickCheck's \ttp{Property}
-type.  A function that returns \ttp{Bool} can by turned into one that returns
-\ttp{ScProperty} using the \ttp{scProperty} function (corresponding to
-QuickCheck's \ttp{property} function).  SmartCheck's property implication uses
-an arrow \ttp{-->} corresponding to QuickCheck's \ttp{==>}.
-
-The arguments is a record type that includes the values described above to
-control SmartCheck's behavior.  The standard arguments are contained in the
-value \ttp{scStdArgs}.
+QuickCheck.  Others are specialized for SmartCheck; e.g., enabling or disabling
+universal or existential extrapolated, number of extrapolation rounds, and
+limits on the depth and size of the values to generate.
 
@@
 \section{Experiments}\label{sec:experiments}
-Here we describe two experiments using SmartCheck, including an XMonad
-property and a property about a natural language processing library.
+We describe two experiments using SmartCheck, including an XMonad property and a
+property about a natural language processing library.
 
@@
 Recall from the introduction the XMonad example.  The XMonad window manager is a
-large software project with many contributors, so consequently, a QuickCheck
+large software project with many contributors, so naturally, a QuickCheck
 test harness is included to help ensure new commits do not introduce bugs.  At
@@
 between windows, work spaces, and which window has the focus.  XMonad contains
-properties to ensure the correct manipulation of \ttp{StackSet}s.
+properties to ensure the correct manipulation of \ttp{StackSet}s.  Due to having
+one large data-structure that is essential to the entire program, XMonad is a
+perfect candidate for SmartCheck.
 
@@
 QuickCheck returns the large value shown in the introduction.  That value is a
-relatively small counterexample, but even the smallest \ttp{StackSet} values
-are somewhat large due to the number of fields within it.  Recall the value
-returned by SmartCheck after generalization:
+relatively small counterexample, but even the smallest \ttp{StackSet} values are
+somewhat visually overwhelming due to the number of fields within it.  Recall
+the value returned by SmartCheck after generalization:
 %
@@
 \noindent
-Let us examine what was generalized and what was not.  First, the last element
-of \ttp{StackSet} is a map, and we treated it as a opaque type, so it is not
-analyzed at all.  But it is an empty map.  The second and third fields of
+Let us examine what was generalized.  In our test run, we chose to treat data
+maps as opaque, so the forth element of \ttp{StackSet} is not generalized, but
+is simply the empty map, so looks uninteresting.  The second and third fields of
 \ttp{StackSet} are generalized, but the first one is not.  So there is something
@@
 matters is having a \ttp{Just} value, which is the stack field that deletion
-works on.
+works on!
 
@@
 \end{quote}
+
+\todo{how long unanswered?}
+
 \noindent
@@
 We were able to send a substantially reduced and generalized counterexample to
-the message poster, making the cause of the bug more obvious.  He responded:
-``$\ldots$While your improved shrinking may not have gone `all'
-the way to the bottom, it got me a \emph{huge} chunk of the way there!''  Recall
-that through the entire process, we never had to learn how GenI works, what the
-property meant, or how to write a custom shrink method!
+the message poster, making the cause of the bug more obvious.  The author responded:
+\begin{quote}
+$\ldots$While your improved shrinking may not have gone `all'
+the way to the bottom, it got me a \emph{huge} chunk of the way there!
+\end{quote}
+
+\noindent
+Recall that through the entire process, we never had to learn how GenI works,
+what the property meant, or how to write a custom shrink method!
 
 ``Retro-fitting'' SmartCheck to GenI, which is a large software system, required
-some work.  Fortunately, because the program already uses QuickCheck,
-\ttp{arbitrary} instances were already defined, but we had to hunt down and
-define default instances for a large number of definitions.
+some boilerplate work.  Fortunately, because the program already uses
+QuickCheck, \ttp{arbitrary} instances were already defined, but we had to hunt
+down and define default instances for a large number of definitions (indeed, the
+job could have largely been done with Sed and Awk).
 
@@
 in terms the size of counterexamples discovered: the mean and median number of
-\ttp{N} constructors in the counterexamples returned by Feat are both seven.
+\ttp{N} constructors in the counterexamples returned by Feat are both seven. \todo{check}
 However, Feat performs less well with respect to how long counterexample
@@
 SmartCheck does not attempt to infer properties like these tools do, so it is in
-one sense less general.  However, SmartCheck's three kinds counterexample
+one sense less general.  However, SmartCheck's three kinds of counterexample
 generalization are novel.
@@
 
+\todo{mention memoization (generalized tries)}
+
+\todo{tail-recursive algorithm is fast and easy to port to imperative languages}
+
 \section*{Acknowledgments}
@@
 Hughes for reading an earlier draft of this paper.  Their deep insights and bug
-catches dramatically improved both the form and content.
+catches dramatically improved both the form and content.  This paper was
+submitted (and rejected) from ICFP'13; the anonymous reviewers' suggestions
+improved the presentation and especially the implementation.
 

### 2014-03-14 src/Test/SmartCheck.hs
sc: wibble.

@@ -165,7 +165,6 @@ runQC args scProp = do
              then (mCex, propify scProp)
              else (Nothing, propify scProp)
 
-
 -- | Returns 'True' if a counterexample is returned and 'False' otherwise.
 failureRes :: Q.Result -> Bool
 failureRes res =

### 2014-03-14 src/Test/SmartCheck/Types.hs
types: smart shrink list containers.

@@ -340,23 +340,26 @@ instance SubTypes ()      where baseType _    = True
 -- --  toConstrAndBase = toConstrAndBase'
 --  showForest    = showForest'
 
--- For container types like list, if it's over a baseType, we don't want to
--- evaluate the container either.  The intuition is that, e.g., for [Int], it'll
--- be shrunk enough by QuickCheck and doesn't really have "interesting
--- structure".
+-- For example, this makes String a baseType automatically.
+-- instance (Q.Arbitrary a, SubTypes a, Typeable a) => SubTypes [a] where
+--   subTypes      = if baseType (undefined :: a) then \_ -> []
+--                     else gst . from
+--   baseType _    = baseType (undefined :: a)
+--   replaceChild x forest y = if baseType (undefined :: a)
+--                               then replaceChild' x forest y
+--                               else fmap to $ grc (from x) forest y
+--   toConstr      = if baseType (undefined :: a) then toConstr'
+--                     else gtc . from
+--   showForest    = if baseType (undefined :: a) then showForest'
+--                     else gsf . from
 
 -- For example, this makes String a baseType automatically.
 instance (Q.Arbitrary a, SubTypes a, Typeable a) => SubTypes [a] where
-  subTypes      = if baseType (undefined :: a) then \_ -> []
-                    else gst . from
-  baseType _    = baseType (undefined :: a)
-  replaceChild x forest y = if baseType (undefined :: a)
-                              then replaceChild' x forest y
-                              else fmap to $ grc (from x) forest y
-  toConstr      = if baseType (undefined :: a) then toConstr'
-                    else gtc . from
-  showForest    = if baseType (undefined :: a) then showForest'
-                    else gsf . from
+  subTypes      = gst . from
+  baseType _    = False
+  replaceChild x forest y = fmap to $ grc (from x) forest y
+  toConstr      = gtc . from
+  showForest    = gsf . from
 
 instance (Integral a, Q.Arbitrary a, SubTypes a, Typeable a)
   => SubTypes (Ratio a) where

### 2014-03-14 src/Test/SmartCheck.hs
sc: wibbles.

@@ -162,8 +162,8 @@ runQC :: forall a prop . (Show a, Q.Arbitrary a, Q.Testable prop)
 runQC args scProp = do
   (mCex, res) <- scQuickCheckWithResult args scProp
   return $ if failureRes res
-             then (mCex, propify scProp)
-             else (Nothing, propify scProp)
+             then (mCex,    Q.property . scProp)
+             else (Nothing, Q.property . scProp)
 
 -- | Returns 'True' if a counterexample is returned and 'False' otherwise.
 failureRes :: Q.Result -> Bool

### 2014-03-14 src/Test/SmartCheck/Reduce.hs

@@ -106,11 +106,11 @@ resultToMaybe res =
 -- | Get the maximum depth of d's subforest at idx.  Intuitively, it's the
 -- maximum number of constructors you have *below* the constructor at idx.  So
 -- for a unary constructor C, the value [C, C, C]
-
+--
 -- (:) C
 --   (:) C
 --     (:) C []
-
+--
 -- At (Idx 0 0) in v, we're at C, so subValSize v (Idx 0 0) == 0.
 -- At (Idx 0 1) in v, we're at (C : C : []), so subValSize v (Idx 0 1) == 2, since
 -- we have the constructors :, C (or :, []) in the longest path underneath.

### 2014-03-14 src/Test/SmartCheck/Render.hs

@@ -83,7 +83,7 @@ replaceWithVars format d idxs vars =
   f :: Tree VarRepl -> (String, Idx) -> Tree VarRepl
   f tree (var, idx) = Node (rootLabel tree) $
     case getIdxForest sf idx of
-      Nothing                 -> errorMsg "replaceWithVars"
+      Nothing                 -> errorMsg "replaceWithVars1"
       Just (Node (Right _) _) -> sf -- Don't replace anything
       Just (Node (Left  _) _) -> forestReplaceChildren sf idx (Right var)
 
@@ -94,7 +94,7 @@ replaceWithVars format d idxs vars =
   -- data.  showForest is one of our generic methods.
   t :: Tree VarRepl
   t = let forest = showForest d in
-      if null forest then errorMsg "replaceWithVars"
+      if null forest then errorMsg "replaceWithVars2"
          else fmap Left (head forest) -- Should be a singleton
 
   -- Note: we put value idxs before constrs, since they take precedence.

### 2014-03-14 src/Test/SmartCheck/SmartGen.hs

@@ -75,8 +75,8 @@ iterateArb d ext idx tries max prop = do
               Result x      -> return (i+1, Result x)
     where
     (size, g0) = randomR (0, currMax) g
-    s = sample ext g size
     sample SubT { unSubT = v } = newVal v
+    s = sample ext g size
     rec res' =
       iterateArb' res' g0 (try + 1)
         -- XXX what ratio is right to increase size of values?  This gives us

### 2014-03-14 src/Test/SmartCheck/Test.hs

@@ -50,7 +50,6 @@ module Test.SmartCheck.Test
 import Prelude hiding (break)
 
 import Test.QuickCheck
-import Test.QuickCheck.Arbitrary
 import Test.QuickCheck.Gen
 import Test.QuickCheck.Property hiding ( Result( reason, theException) )
 import qualified Test.QuickCheck.Property as P

### 2014-03-14 src/Test/SmartCheck/Types.hs

@@ -109,10 +109,7 @@ instance Show SubT where
   show (SubT t) = show t
 
 -- | This class covers algebraic datatypes that can be transformed into Trees.
--- subTypes is the main method, placing values into trees.  For types that can't
--- be put into a *structural* order (e.g., Int), we don't want SmartCheck to
--- touch them, so that aren't placed in the tree (the baseType method tells
--- subTypes which types have this property).
+-- subTypes is the main method, placing values into trees.
 --
 -- for a datatype with constructors A and C,
 --
@@ -243,7 +240,9 @@ instance GST a => GST (M1 i k a) where
   gsz (M1 a) = gsz a
 
 instance (Show a, Q.Arbitrary a, SubTypes a, Typeable a) => GST (K1 i a) where
-  gst (K1 a) = if baseType a then [] else [ Node (subT a) (subTypes a) ]
+  gst (K1 a) = if baseType a
+                 then []
+                 else [ Node (subT a) (subTypes a) ]
 
   grc (K1 a) forest c =
     case forest of

### 2014-03-15 paper/paper.tex
paper wibbles.

@@
 
+\todo{need to test with arbitrary for ints turned off.  (Use newtype)}
+
 \begin{figure}[ht]
@@
   constructors contained within it,
-\item \ttp{index} returns a \emph{sub-value} at an breadth-first index in a
+\item \ttp{index} returns a \emph{sub-value} at a breadth-first index in a
   value,\footnote{Rather than indexes, an alternative representation is to use a
-    zipper data structure~\cite{zipper}; see Section~\ref{sec:conclusions} for a
-    discussion.}
+    zipper data structure~\cite{zipper} to traverse data.  We have chosen
+    explicit indexes to write simple tail-recursive algorithms that can easily
+    be transcribed to imperative languages.}
 \item \ttp{replace} replaces a sub-value at a particular focus (returning the
@@
 
-The number of constructors in an algebraic data value primarily determines how
-large it is.  In some cases, the size of fields that come from primitive
-types, like \ttp{Int} or \ttp{Char}, have less of a role in understanding a
-counterexample, so in general, we want to prevent our reduction algorithm from
-spending time shrinking values from these types.
-
-Consequently, we distinguish a class of types referred to as \emph{opaque
-  types}.  If the reduction algorithm encounters a opaque type, it is ignored.
-By default, this includes ``primitive'' types (e.g., \ttp{Int}, \ttp{Word},
-\ttp{Float}, \ttp{Bool}, and \ttp{Char}).  We also treat recursive container
-types like lists specially: lists of opaque types are considered to be opaque
-types themselves.  This prevents the algorithm from analyzing, for example,
-strings.  A list of non-opaque-types, however, is shrunk by the algorithm.
-
-By providing custom instances, the user can declare any substructure in a data
-type to be a opaque type.  Doing so effectively treats values from that type as
-``black boxes'' for SmartCheck, making SmartCheck more efficient if the user
-knows that some portion of the structure cannot be shrunk or is irrelevant to
-the property.
-
-Conversely, if the user wishes to shrink types that are treated as opaque by
-default, she can sometimes do so before handing the counterexample to
-SmartCheck, say by using QuickCheck's shrink instances to shrink base types
-while leaving the remaining structure for SmartCheck to shrink.  In the future,
-depending on user experiences, default instances for types can be removed so
-that SmartCheck's reduction algorithm does not ignore values from those types.
+SmartCheck focuses on efficiently shrinking and generalizing large data
+structures.  It is not intended as a general replacement for QuickCheck's
+\ttp{shrink} method.  Consequently, SmartCheck ignores ``primitive'' types
+without value constructors, such as \ttp{Char}, \ttp{Int}, and \ttp{Word16}.
+Our experience is that for the kinds of properties with counterexamples that
+contain massive data structures, shrinking primitive types does not
+significantly help in understanding them.  Furthermore, by ignoring these types
+by fiat, shrinking time is dependent only on the size of a data structure as
+measured by the number of constructors.
+
+We generalize the idea of ignoring primitive types by introducing the concept of
+\emph{opaque types}.  If the reduction algorithm encounters a opaque type, it is
+ignored.  Opaque types include the primitive types mentioned above, but in the
+user can declare any substructure in a data type to be a opaque type by
+providing custom instances, .  Doing so effectively treats values from that type
+as ``black boxes'', making SmartCheck more efficient if the user knows that some
+portion of the structure cannot be shrunk or is irrelevant to the property.
+
+%% The number of constructors in an algebraic data value primarily determines how
+%% large it is.  In some cases, the size of fields that come from primitive
+%% types, like \ttp{Int} or \ttp{Char}, have less of a role in understanding a
+%% counterexample, so in general, we want to prevent our reduction algorithm from
+%% spending time shrinking values from these types.
+
+%% By default, this includes ``primitive'' types (e.g., \ttp{Int}, \ttp{Word},
+%% \ttp{Float}, \ttp{Bool}, and \ttp{Char}).  We also treat recursive container
+%% types like lists specially: lists of opaque types are considered to be opaque
+%% types themselves.  This prevents the algorithm from analyzing, for example,
+%% strings.  A list of non-opaque-types, however, is shrunk by the algorithm.
+
+%% Conversely, if the user wishes to shrink types that are treated as opaque by
+%% default, she can sometimes do so before handing the counterexample to
+%% SmartCheck, say by using QuickCheck's shrink instances to shrink base types
+%% while leaving the remaining structure for SmartCheck to shrink.  In the future,
+%% depending on user experiences, default instances for types can be removed so
+%% that SmartCheck's reduction algorithm does not ignore values from those types.
 
@@
 removeFromWorkspace ws =
-  ws \{ stack = stack ws >>= filter (/=w) \}
+  ws \{ stack = stack ws >>= filter (/= w) \}
 \end{code}
@@
 removeFromWorkspace ws =
-  ws \{ stack = stack ws >>= filter (==w) \}
+  ws \{ stack = stack ws >>= filter (== w) \}
 \end{code}

### 2014-03-19 paper/paper.tex
paper updates.

@@
 \usepackage{paralist}
-
+\usepackage{multirow}
 \usepackage{pgf}
@@
 
-\newenvironment{code}{\begin{alltt}}{\end{alltt}}
+\newenvironment{code}{\begin{alltt}\footnotesize}{\end{alltt}}
 %% \newenvironment{cols}{\begin{tabular}{m{3.6cm}|m{3.6cm}}{Haskell} &
@@
 user can declare any substructure in a data type to be a opaque type by
-providing custom instances, .  Doing so effectively treats values from that type
+providing custom instances.  Doing so effectively treats values from that type
 as ``black boxes'', making SmartCheck more efficient if the user knows that some
 portion of the structure cannot be shrunk or is irrelevant to the property.
+
+Opaque types can be conditional.  For example, the user may want lists to be
+shrunk in general, unless the element of the list are opaque themselves.  Such a
+definition is possible.
+
+Opaque types are defined by providing the method for \ttp{opaque} in the
+\ttp{SubTypes} type class.
 
@@
 
-\subsection{Natural Language Processing}
+\subsection{Natural Language Processing} \label{sec:nlp}
 In 2012, a question was posted on the programming message board
@@
 QuickCheck, \ttp{arbitrary} instances were already defined, but we had to hunt
-down and define default instances for a large number of definitions (indeed, the
-job could have largely been done with Sed and Awk).
+down the data types used to write \ttp{instance Subtypes D}, for each data type
+\ttp{D} (indeed, the job could have largely been done with Sed and Awk).
 
 %-------------------------------------------------------------------------------
+
+\subsection{Benchmarks}
+
+No set of testing benchmarks exists over which to compare different test-case
+generation and minimization approaches.  Therefore, we have collected a small
+number of benchmarks, in addition to the more involved case-studies described
+earlier in this section.
+
+The benchmarks presented, in addition to the motivating example presented in
+Section~\ref{sec:example}, are
+
+\begin{itemize}
+  \item  \emph{Reverse}, with the false property
+\begin{code}
+prop_rev :: [a] -> Bool
+prop_rev ls = ls == reverse ls
+\end{code}
+\noindent
+(the example appears in the original QuickCheck documentation);
+
+  \item \emph{Div0}, a division-by-zero property for a simple calculator
+    language (introduced in Section~\ref{sec:subval});
+
+  \item \emph{Heap}, an example from the QuickCheck test suite, in which an
+    incorrect ``to sorted list'' function is checked.
+\end{itemize}
+
+\noindent
+All benchmarks can be found at
+\url{https://github.com/leepike/SmartCheck/tree/master/regression}.  Benchmarks
+are against QuickCheck~2.7.   The benchmarks only compare the size of shrunk
+values and the time required to generate them using standard arguments to
+QuickCheck and SmartCheck.  We do not benchmark the generalization algorithms,
+since there is no analog in QuickCheck  or other testing frameworks.
+
+\todo{include std dev.}
+\todo{fill in benchmarks!}
+
+\begin{table}[ht]
+\footnotesize
+  \begin{center}
+    \begin{tabular}{|r|r||c|c|}
+\hline
+\multicolumn{2}{|c||}{} & QuickCheck & SmartCheck \\
+\hline \hline
+
+\multirow{4}{*}{Reverse}
+  & Mean       & 0.08s, 69  & 0.53s, 35  \\
+  \cline{2-4}
+  & Median     & 0.07s, 69  & 0.17s, 33  \\
+  \cline{2-4}
+  & Std. dev.  &            &            \\
+  \cline{2-4}
+  & 95\%       & 0.14s, 100 & 1.76s, 65  \\
+  \hline
+
+\multirow{4}{*}{Div0}
+  & Mean       & 0.08s, 69  & 0.53s, 35  \\
+  \cline{2-4}
+  & Median     & 0.07s, 69  & 0.17s, 33  \\
+  \cline{2-4}
+  & Std. dev.  &            &            \\
+  \cline{2-4}
+  & 95\%       & 0.14s, 100 & 1.76s, 65  \\
+  \hline
+
+\multirow{4}{*}{Heap}
+  & Mean       & 0.08s, 69  & 0.53s, 35  \\
+  \cline{2-4}
+  & Median     & 0.07s, 69  & 0.17s, 33  \\
+  \cline{2-4}
+  & Std. dev.  &            &            \\
+  \cline{2-4}
+  & 95\%       & 0.14s, 100 & 1.76s, 65  \\
+  \hline
+    \end{tabular}
+
+  \end{center}
+  \caption{Summarizing data for the graphs in Figure~\ref{fig:graphs}. Entries
+    contain execution time (in seconds) and counterexample sizes (counting
+    constructors).}
+  \label{table:results}
+\end{table}
+
+The \emph{Reverse} benchmark essentially provides a lower-bound on the benefit
+of using SmartCheck to discover small counterexamples, since a list of length
+two falsifies the property (SmartCheck is still useful in generalizing the
+counterexample, however!).  Perhaps surprisingly, SmartCheck still slightly
+outperforms QuickCheck.  The other two benchmarks are slightly larger examples,
+where SmartCheck's benefits are more pronounced.  But again, we emphasize the
+most pronounced benefit comes in much larger examples, like described in
+Section~\ref{sec:nlp}.
 
@@
 Hughes for reading an earlier draft of this paper.  Their deep insights and bug
-catches dramatically improved both the form and content.  This paper was
-submitted (and rejected) from ICFP'13; the anonymous reviewers' suggestions
-improved the presentation and especially the implementation.
+catches dramatically improved both the form and content.  This paper was and
+rejected from ICFP'13; the anonymous reviewers' suggestions improved the
+presentation additional benchmarking.
 

### 2014-03-20 paper/paper.tex
paper wibbles.

@@
 
+\newcommand{\undef}[1]{{\bf #1}}
+
 \begin{document}
@@
 \begin{abstract}
-QuickCheck, developed by Claessen and Hughes, is a powerful library for automatic test-case generation.
-Because QuickCheck performs random testing, some of the counterexamples discovered are very large.
-QuickCheck provides an interface for the user to write shrink functions to attempt to reduce the size of counterexamples discovered.
-Hand-written implementations of shrink can be complex, inefficient, and consist of significant boilerplate code.
-Furthermore, shrinking is only one aspect in debugging: counterexample generalization is the process of extrapolating from individual counterexamples to a class of counterexamples, often requiring a flash of insight from the programmer.
-To improve counterexample reduction and generalization, we introduce \emph{SmartCheck}.
-SmartCheck is a debugging tool that reduces algebraic data using generic search heuristics to efficiently find smaller counterexamples independent of the program or property being tested.
-In addition to shrinking, SmartCheck also automatically generalizes counterexamples to formulas representing them.
-SmartCheck has been implemented for Haskell and is freely available.
+QuickCheck, developed by Claessen and Hughes, is a powerful library for
+automatic test-case generation.  Because QuickCheck performs random testing,
+some of the counterexamples discovered are very large.  QuickCheck provides an
+interface for the user to write shrink functions to attempt to reduce the size
+of counterexamples.  Hand-written implementations of shrink can be complex,
+inefficient, and consist of significant boilerplate code.  Furthermore,
+shrinking is only one aspect in debugging: counterexample generalization is the
+process of extrapolating from individual counterexamples to a class of
+counterexamples, often requiring a flash of insight from the programmer.  To
+improve counterexample reduction and generalization, we introduce
+\emph{SmartCheck}.  SmartCheck is a debugging tool that reduces algebraic data
+using generic search heuristics to efficiently find smaller counterexamples.  In
+addition to shrinking, SmartCheck also automatically generalizes counterexamples
+to formulas representing classes of counterexamples.  SmartCheck has been implemented for Haskell and is freely available.
 \end{abstract}
@@
 (This counterexample uses Haskell's default \ttp{Show} instances, which uses
-record syntax.)  Programmers should be familiar with having to debug
+record syntax.)  Programmers may be familiar with having to debug
 a ``wall of text'' as shown above.  What if instead a
@@
 As shown in the motivating example in Section~\ref{sec:example}, SmartCheck
-can shrink more than an order-of-magnitude faster than efficient hand-written
+can shrink more than an order-of-magnitude faster than even the most efficient hand-written
 shrink implementations for some properties and programs.  As well, the time
@@
 A more clever programmer that understands the semantics of QuickCheck's
-\ttp{shrink} implementation,\footnote{The following approach was suggested to
+\ttp{shrink} implementation\footnote{The following approach was suggested to
   the author by John Hughes.} defines a \ttp{shrink} instance as follows:
@@
 %\end{itemize}
+
+\todo{QC without shrinking Ints}
 
@@
 $y$ axis in one graph and a linear scale in the other).  In the table, we show
-the mean, median, median, and the results at the 95th percentile values.  (The
-distributions produced by these values are not necessarily Gaussian, so we omit
-the standard deviation.)
+the mean, median, median, and the results at the 95th percentile values.  (We
+give standard deviations, but not that the plots are not necessarily Gaussian.)
 
@@
 
-In regards to generalization, after shrinking the counterexample, SmartCheck
-will also generalize it.  Note that for any \ttp{Word8} value in the first
-argument of \ttp{T}, the counterexample still holds.  For example, SmartCheck
-might return a counterexample of the form
-%
-\begin{code}
-forall x . T x [-20000] [-20000] [] []
-\end{code}
+%% In regards to generalization, after shrinking the counterexample, SmartCheck
+%% will also generalize it.  Note that for any \ttp{Word8} value in the first
+%% argument of \ttp{T}, the counterexample still holds.  For example, SmartCheck
+%% might return a counterexample of the form
+%% %
+%% \begin{code}
+%% forall x . T x [-20000] [-20000] [] []
+%% \end{code}
 
@@
 pronounced for counterexamples that are particularly ``deep'' or ``broad'' in
-terms of their tree structure.  For programs and properties for which small
-counterexamples are easily found, QuickCheck's shrinking is perfectly
-acceptable.  Furthermore, user-defined shrink methods may vary significantly
-based on the property and counterexample type, so it is not always clear which
-implementation to compare against.
+terms of their tree structure.  %% For programs and properties for which small
+%% counterexamples are easily found, QuickCheck's shrinking is perfectly
+%% acceptable.  Furthermore, user-defined shrink methods may vary significantly
+%% based on the property and counterexample type, so it is not always clear which
+%% implementation to compare against.
+We now describe in detail SmartCheck's algorithm for generic shrinking.
 
@@
 \item \ttp{index} returns a \emph{sub-value} at a breadth-first index in a
-  value,\footnote{Rather than indexes, an alternative representation is to use a
-    zipper data structure~\cite{zipper} to traverse data.  We have chosen
-    explicit indexes to write simple tail-recursive algorithms that can easily
-    be transcribed to imperative languages.}
+  value,
 \item \ttp{replace} replaces a sub-value at a particular focus (returning the
@@
 We use this invariant as a termination case in recursive algorithms over the
-sub-values of a value.
+sub-values of a value.  (Rather than indexes into a data-structure, an
+alternative representation is to use a zipper data structure~\cite{zipper} to
+traverse data.  We have chosen explicit indexes to write simple tail-recursive
+algorithms that can easily be transcribed to imperative languages.)
 
@@
 In this section, we describe how to efficiently and generically shrink algebraic
-data values.  Before we begin, recall the basic idea behind the \ttp{shrink}
+data values.  Recall the basic idea behind the \ttp{shrink}
 method of the \ttp{Arbitrary} class: generate a list of values, each of which is
@@
   replicateM tries s where
-  s = liftM SubVal (sizedArbitrary sz a)
+  s = liftM SubVal (\undef{sizedArbitrary} sz a)
 
@@
     let cex' = replace cex idx v in
-    if pass prop cex' then go vs'
+    if \undef{pass} prop cex' then go vs'
       else Just cex'
@@
 
-Figure~\ref{fig:reduction} shows the reduction algorithm.  The function
+Figure~\ref{fig:reduction} shows the reduction algorithm.  In this algorithm and
+subsequent algorithms in the paper, functions in {\bf bold font} are left
+undefined but their implementation is described in the text.  The function
 \ttp{reduce} takes flags to customize the algorithm's behavior, a counterexample
@@
 \ttp{v} of \ttp{T}, then it is most probable that other counterexamples are also
-from variant \ttp{v}.  Recall that the algorithm begins at the \emph{first}
-sub-value of the counterexample rather than the 0th sub-value.  The 0th
-sub-value of \ttp{cex} is \ttp{cex} itself.  No invariant of the algorithm would
-be violated by beginning with the 0th sub-value, and in particular, the
-algorithm would still terminate.  However, we begin with a strict sub-value
-because of the hypothesis.  As an example supporting the hypothesis, consider a
+from variant \ttp{v}.  As an example supporting the hypothesis, consider a
 property about unintended variable capture over a language's parse tree
@@
 counterexamples, the only construct in which variable capture is possible.
+
+Recall that the algorithm begins at the \emph{first} sub-value of the
+counterexample rather than the 0th sub-value.  The 0th sub-value of \ttp{cex} is
+\ttp{cex} itself.  No invariant of the algorithm would be violated by beginning
+with the 0th sub-value, and in particular, the algorithm would still terminate.
+However, we begin with a strict sub-value because of the hypothesis.
+
+The hypothesis may be incorrect for some properties, in which case SmartCheck
+may potentially fail to discover a smaller counterexample.  However, in
+Sections~\ref{sec:existential} and~\ref{sec:precondition}, we describe an
+approaches to generalize counterexamples based on discovering new counterexample
+variants.  These generalization techniques are executed in an (optional)
+generalization phase, run after the reduction phase, in which this hypothesis is
+implemented.
 
@@
 \ttp{SubTypes} type class.
+
+\todo{reference into with QC, no int shrinking}
 
@@
   testHole (SubVal a) = do
-    a' <- cast a :: Maybe a
-    if pass prop a' then Nothing else Just a'
+    a' <- \undef{cast} a :: Maybe a
+    if \undef{pass} prop a' then Nothing else Just a'
 \end{code}
@@
 \subsection{Universal Sub-Value Generalization}\label{sec:universal}
-Recall from the end of Section~\ref{sec:example} the quantified formula
-%
-\begin{code}
-forall x . T x [-20000] [-20000] [] []
-\end{code}
-%
-\noindent
-characterizing a set of counterexamples.  Another example using the calculator
-language from Section~\ref{sec:subval} is to generalize a counterexample like
+Consider again the calculator language from Section~\ref{sec:subval}.  The
+property \ttp{prop\_div} is violated for any numerator, so we might generalize a
+counterexample like
 %
@@
 \noindent
-by
+by the formula
 %
@@
 subTrees :: SubTypes a => a -> Idx -> [Idx] -> Bool
-subTrees cex idx = any (subTree cex idx)
+subTrees cex idx = any (\undef{subTree} cex idx)
 
@@
   res  = mapMaybe go vs
-  go   = fail prop . replace cex idx
+  go   = \undef{fail} prop . replace cex idx
   \end{code}
@@
 
-The extrapolation algorithm is shown in Figure~\ref{fig:universal}.  The
-algorithm is similar to the reduction algorithm in Figure~\ref{fig:reduction},
-and in the implementation, the algorithms are generalized and combined.  The
-function \ttp{extrapolate} returns a list of indexes to be generalized in the
-original counterexample.  In the recursive function \ttp{extrapolate'}, there is
-a function guard with a call
+The extrapolation algorithm is shown in Figure~\ref{fig:universal}; let sketch
+its specification.  The algorithm is similar to the reduction algorithm in
+Figure~\ref{fig:reduction} (and in the implementation, the algorithms are
+generalized and combined).  The function \ttp{extrapolate} returns a list of
+indexes to be generalized in the original counterexample.  In the recursive
+function \ttp{extrapolate'}, there is a function guard with a call
 %
@@
 \noindent
+The value
+%
+\begin{code}
+subTree cex idx0 idx1
+\end{code}
+%
 is true if in \ttp{cex}, the value at index \ttp{idx0} is a child of index
@@
 that are abstracted away already since their parents have been generalized.
-
 New sub-values are generated by \ttp{newVals}, shown in
-Figure~\ref{fig:reduction}.  Like in the reduction algorithm, we bound the size
-of values to generate by the flag \ttp{scMaxSize}, which is independent of the
-size of the particular sub-value.  The flag \ttp{scMaxForall} is the analogue of
-the \ttp{scMaxReduce} flag, determining the number of values generated in trying
-to generalize a value.
+Figure~\ref{fig:reduction}.
 
@@
 non-trivially satisfies \ttp{prop}, and \ttp{Nothing} if \ttp{cex} fails
-\ttp{prop}'s precondition.  The flag \ttp{scMinForall} is the minimum number
-of \ttp{Just False} results required from \ttp{fail} to extrapolate from failed
-tests to a universal claim.
+\ttp{prop}'s precondition.
+
+Like in the reduction algorithm user-specified flags bound the behavior of the
+algorithm.  We bound the size of values to generate by the flag \ttp{scMaxSize},
+which is independent of the size of the particular sub-value.  The flag
+\ttp{scMaxForall} is the analogue of the \ttp{scMaxReduce} flag, determining the
+number of values generated in trying to generalize a value.  The flag
+\ttp{scMinForall} is the minimum number of \ttp{Just False} results required
+from \ttp{fail} to extrapolate from failed tests to a universal claim.  So, for
+example, if \ttp{scMaxForall} is set to 30 and \ttp{scMinForall} is set to 20,
+we generate 30 new values, 20 of which must pass the precondition but fail the
+property to claim the counterexample can be generalized.
 
@@
 
-Regarding our implementation, we have not profiled the performance of our
-implementation of SmartCheck, so additional performance gains might be possible.
-In particular, we perform significant computations over tree-structured data,
-and a generic zipper~\cite{zipper}, which we do not currently implement, could
-improve performance.
+%% Regarding our implementation, we have not profiled the performance of our
+%% implementation of SmartCheck, so additional performance gains might be possible.
+%% In particular, we perform significant computations over tree-structured data,
+%% and a generic zipper~\cite{zipper}, which we do not currently implement, could
+%% improve performance.
 

### 2014-04-29 paper/paper.tex
paper: add problem def.

@@
 \end{quote}
-%
-\noindent
-QuickCheck defines a type class \ttp{Arbitrary} that presents a method
-\ttp{arbitrary} for generating random values of a given type.  Gill added another
-method to the type class:
+
+That is, automated testing helps find a counterexample, but it may be a very
+large counterexample.  We are left facing the \emph{shrinking problem}: assume
+there exist a set of inputs $i_0, \ i_1, \ldots, \i_n$, such that for program
+$P$ (modeled as a function mapping inputs to output) and property $\phi$,
+$\phi(P(i_0, \ i_1, \ldots, \i_n))$ is false.  The shrinking problem is to find
+the smallest set of inputs $i'_0, \ i'_1, \ldots, \i'_n$, according to some
+measure, such that $\phi(P(i\_0, \ i\_1, \ldots, \i\_n))$ is also false.
+
+In general, the problem too expensive to solve, since it requires showing that
+for a set of failing inputs, there exist no strictly smaller inputs that also
+cause the property to fail.  Such a demonstration requires exhaustive testing or
+proof.  Solutions to the shrinking problem are therefore approximated to
+efficiently discover sufficiently small inputs causing the input to fail.
+Usually, the shrunk inputs are derived algorithmically (but perhaps
+nondeterministically) from a known larger set of inputs resulting in failure,
+and the inputs may not be minimal.
+
+To address the shrinking problem, QuickCheck defines a type class
+\ttp{Arbitrary} that presents a method \ttp{arbitrary} for generating random
+values of a given type.  Gill added another method to the type class:
 %

### 2014-04-29 paper/paper.tex
paper: draft notice.
paper: wibbles, grab latest graphs.

@@
 \title{SmartCheck: Automatic and Efficient Counterexample Reduction and Generalization}
-%\subtitle{Subtitle Text, if any}
+\subtitle{UNPUBLISHED DRAFT}
 
@@
 \begin{figure}[ht]
-\includegraphics[scale=0.6]{Graphs/data}
-\includegraphics[scale=0.6]{Graphs/time-big}
-\includegraphics[scale=0.6]{Graphs/time-small}
+\includegraphics[scale=0.6]{../regression/PaperExample1/out/data}
+\includegraphics[scale=0.6]{../regression/PaperExample1/out/time-big}
+\includegraphics[scale=0.6]{../regression/PaperExample1/out/time-small}
   \caption{Results for 2000 tests of the program in Figure~\ref{fig:initial}.}
@@
     incorrect ``to sorted list'' function is checked.
+
+  \item \emph{Parser}, a parser/pretty-printer for a toy imperative language
+    containing a parser bug that switches the arguments of an binary operator.
 \end{itemize}
@@
 \url{https://github.com/leepike/SmartCheck/tree/master/regression}.  Benchmarks
-are against QuickCheck~2.7.   The benchmarks only compare the size of shrunk
-values and the time required to generate them using standard arguments to
-QuickCheck and SmartCheck.  We do not benchmark the generalization algorithms,
-since there is no analog in QuickCheck  or other testing frameworks.
+are against
+QuickCheck~2.7.3.\footnote{\url{http://hackage.haskell.org/package/QuickCheck-2.7.3}}
+The benchmarks only compare the size of shrunk values and the time required to
+generate them using standard arguments to QuickCheck and SmartCheck.  We do not
+benchmark the generalization algorithms, since there is no analog in QuickCheck
+or other testing frameworks.
 

### 2014-04-29 regression/PaperExample1/PaperExample1.hs
regression: build wibbles, don't shrink Int16s at all.

@@ -31,18 +31,13 @@ import Test.Feat
 
 -----------------------------------------------------------------
 
--- Let's try to generate a product type of long lists when all we need is a
--- single element to have a long list.
-
------------------------------------------------------------------
-
-#ifdef qcjhint
+#if defined(qcjhint) || defined(qcNone) || defined(qc10) || defined(qc20)
 -- So that Int16s aren't shrunk by default arbitrary instances.
 newtype J = J { getInt :: Int16 } deriving Show
-#endif
-
-#ifdef qcjhint
 type I = [J]
+instance Arbitrary J where
+  arbitrary = fmap J arbitrary
+
 #else
 type I = [Int16]
 #endif
@@ -91,10 +86,6 @@ instance Arbitrary T where
     where xs = shrink (w, i0, i1, i2, i3)
           go (w', i0', i1', i2', i3') = T w' i0' i1' i2' i3'
 #endif
-#ifdef qcjhint
-instance Arbitrary J where
-  arbitrary = fmap J arbitrary
-#endif
 #if defined(qc10) || defined(qc20)
   shrink (T w i0 i1 i2 i3) =
     [ T a b c d e | a <- tk w
@@ -118,7 +109,7 @@ deriveEnumerable ''T
 
 toList :: T -> [[Int16]]
 toList (T w i0 i1 i2 i3) =
-#ifdef qcjhint
+#if defined(qcjhint) || defined(qcNone) || defined(qc10) || defined(qc20)
   [fromIntegral w] : (map . map) (fromIntegral . getInt) [i0, i1, i2, i3]
 #else
   [fromIntegral w] : (map . map) fromIntegral [i0, i1, i2, i3]

### 2014-05-03 paper/paper.tex
paper: with updated data.

@@
 
+\newcommand{\sci}[2]{\ensuremath{#1 \times 10^{#2}}}
 \begin{document}
@@
 
-\todo{need to test with arbitrary for ints turned off.  (Use newtype)}
-
 \begin{figure}[ht]
@@
 type I   = [Int16]
-data T   = T (Word8) I I I I
+data T   = T I I I I I
 
 toList :: T -> [[Int16]]
-toList (T w i0 i1 i2 i3) =
-  [fromIntegral w] : (map . map) fromIntegral rst
-  where
-  rst = [i0, i1, i2, i3]
+toList (T i0 i1 i2 i3 i4) =
+  (map . map) fromIntegral [i0, i1, i2, i3, i4]
 
@@
 
-\todo{redo analysis, check text}
 In this section, we motivate in more detail the challenges in shrinking
-counterexamples (we focus on shrinking and only touch on generalization here).
+counterexamples by comparing manual approaches using QuickCheck to SmartCheck.
+(We focus on shrinking rather than generalization here since counterexample
+generalization is unique to SmartCheck.)  We will show how a small data type and
+simple property can result in large counterexamples without any shrinking.  Then
+we show the difficulty in designing an efficient shrink implementation manually,
+which is required when using QuickCheck; we will walk through a couple of poor
+designs before arriving at a ``canonical'' manual solution.  However, even the
+canonical solution produces significantly larger counterexamples and takes
+significantly longer to produce them than SmartCheck, even though SmartCheck's
+shrinking strategy is completely automatic and generic.
+
 Consider the example in Figure~\ref{fig:initial}.\footnote{All examples and
@@
   the language extensions of existential type quantification and scoped type
-  variables.  We use Haskell \ttp{Prelude} functions, \ttp{mapMaybe} from
-  \ttp{Data.Maybe}, and \ttp{liftM} and \ttp{replicateM} from
-  \ttp{Control.Monad}.}  Data type \ttp{T} is a product type containing four
-lists of wrapped \ttp{Int16}s and a single wrapped \ttp{Word8}.  For our
-purposes, it is not important what \ttp{T} models, but for concreteness, suppose
-it models the values in four linked lists in an operating system, together with
-an 8-bit status register.
+  variables.}  Data type \ttp{T} is a product type containing five lists of
+signed 8-bit integers.  For our purposes, it is not important what \ttp{T}
+models, but for concreteness, suppose it models the values in five linked lists.
 
@@
 The input to the program satisfies the invariant \ttp{pre}, that the sum of
-values in each list of \ttp{Int16}s is less than or equal to 256 (and by
-definition, the \ttp{Word8} value is less than or equal to 256).  Assuming this,
+values in each list of \ttp{Int16}s is less than or equal to 256.  Assuming this,
 we want to show \ttp{post} holds, that the sum of all the values from \ttp{T} is
@@
 \begin{code}
-T 10 [-20000] [-20000] [] []
+T [-20000] [-20000] [] [] []
 \end{code}
@@
 QuickCheck can be large.  With standard settings and no shrinking, the median
-counterexample discovered contains around 70 \ttp{Int16} values, and
-counterexamples can contain over 100 values.  \todo{recheck} Thus, it pays to
-define \ttp{shrink}!
+counterexample discovered contains just under 70 \ttp{Int16} values, and
+counterexamples can contain over 100 values.  Thus, it pays to define
+\ttp{shrink}!
 
@@
 \noindent
-While the above definition appears reasonable, there are two problems with it.
-First, the result of (\ttp{shrink t}) is null if any list contained in \ttp{t}
-is null, in which case \ttp{t} will not be shrunk.  More troublesome is that with
-QuickCheck's default \ttp{Arbitrary} instances, the length of potential
-counterexamples returned by \ttp{shrink} can easily exceed $10^{10}$, which is
-an intractable number of tests for QuickCheck to analyze.  The reason for the
-blowup is that shrinking a list \ttp{[a]} produces a list \ttp{[[a]]}, the
-length of which is significantly longer than the original list.  Then, we take
-the cross-product of the generated lists.  For the example above, with some
-counterexamples, the shrinking stage may appear to execute for hours, consuming
-ever more memory, without returning a result.
-
-A programmer might try to control the complexity by truncating lists using the
-(\ttp{take n}) function that returns the first $n$ elements of a list.  The
-trade-off is quicker shrinking with a lower probability of finding a smaller
-counterexample.  For example, we might redefine shrink as follows:
+While the above definition appears reasonable on first glance, there are two
+problems with it.  First, the result of (\ttp{shrink t}) is null if any list
+contained in \ttp{t} is null, in which case \ttp{t} will not be shrunk.  More
+troublesome is that with QuickCheck's default \ttp{Arbitrary} instances, the
+length of potential counterexamples returned by \ttp{shrink} can easily exceed
+$10^{10}$, which is an intractable number of tests for QuickCheck to analyze.
+The reason for the blowup is that shrinking a list \ttp{[a]} produces a list
+\ttp{[[a]]}, a list of lists; each list is a new shrinking of the original list.
+Then, we take the cross-product of the generated lists.  For the example above,
+with some counterexamples, the shrinking stage may appear to execute for hours,
+consuming ever more memory, without returning a result.
+
+The first way we will control the complexity is focusing on structurally
+shrinking the data type rather than shrinking the base types.  We prevent
+QuickCheck from using its default shrink instance for \ttp{Int16} by defining a
+newtype:
+%
+\begin{code}
+newtype T = T { getInt :: Int16 }
+\end{code}
+%
+\noindent
+Without shrinking the \ttp{Int16}s, the performance is dramatically improved.
+
+Next, a programmer might try to control the complexity by truncating lists using
+truncation where
+%
+\begin{code}
+take n ls
+\end{code}
+%
+returns the first $n$ elements \ttp{ls}.  The trade-off is quicker shrinking
+with a lower probability of finding a smaller counterexample.  For example, we
+might redefine shrink as follows:
 %
@@
 Call this version ``truncated shrink''.  Truncation controls the blowup of the
-input-space; the maximum number of possible new values is $10^5$ in this case.
-While truncation controls the blowup, the downside is that potentially smaller
-counterexamples may be omitted.
-
-A more clever programmer that understands the semantics of QuickCheck's
-\ttp{shrink} implementation\footnote{The following approach was suggested to
-  the author by John Hughes.} defines a \ttp{shrink} instance as follows:
+input-space, but the downside is that potentially smaller counterexamples may be
+omitted.
+
+Someone who understands the semantics of QuickCheck's \ttp{shrink}
+implementation defines a \ttp{shrink} instance as follows:\footnote{The
+  following approach was suggested to the author by John Hughes.}
 %
@@
 attempting to shrink \ttp{a} while holding \ttp{b} constant, then attempting to
-shrink \ttp{b} while holding \ttp{a} constant (and is generalized to larger
-tuples).  Thus, each argument to \ttp{T} is independently shrunk.
-
+shrink \ttp{b} while holding \ttp{a} constant (and is generalized from pairs to
+arbitrary tuples).  Thus, each argument to \ttp{T} is independently shrunk.
 Still, using this definition, from some initial counterexamples, shrinking may
-take over one minute and result in a final counterexample containing just under
-100 \ttp{Int16} values.\footnote{All results reported in the paper are with a
-  GHC-7.6.3-compiled program, using \ttp{-O2}, on a 4-core 2.7GHz Intel~i7
-  processor.}
+result in a final counterexample containing just over 100 \ttp{Int16} values.
 
@@
 
-\todo{QC without shrinking Ints}
-
 \begin{figure}[ht]
 \includegraphics[scale=0.6]{../regression/PaperExample1/out/data}
-\includegraphics[scale=0.6]{../regression/PaperExample1/out/time-big}
-\includegraphics[scale=0.6]{../regression/PaperExample1/out/time-small}
-  \caption{Results for 2000 tests of the program in Figure~\ref{fig:initial}.}
+%% \includegraphics[scale=0.6]{../regression/PaperExample1/out/time-big}
+%% \includegraphics[scale=0.6]{../regression/PaperExample1/out/time-small}
+  \caption{Results for 1000 tests of the program in Figure~\ref{fig:initial}.}
   \label{fig:graphs}
@@
 \hline \hline
-Mean       & 0.08s, 69  & 0.53s, 35  & 2.24s, 12  & 0.10s, 7\\
+Mean       & 81, 0.013s  & 40, 0.231s   & 13, 0.228s  & 6, 0.022s\\
 \hline
-Median     & 0.07s, 69  & 0.17s, 33  & 1.81s, 13  & 0.10s, 6\\
+Std. dev.  & 21, 0.001   & 21, 1.742    & 7,  0.962   & 4, 0.001s\\
+% Median     & 0.009s, 67  & 0.17s, 33  & 1.81s, 13  & 0.10s, 6\\
 %% \hline
@@
 \hline
-95\%       & 0.14s, 100 & 1.76s, 65  & 4.06s, 19  & 0.18s, 12\\
+95\%       & 116, 0.025s  & 84, 0.283s  & 26, 0.501s  & 13, 0.039s\\
 \hline
@@
   \end{center}
-  \caption{Summarizing data for the graphs in Figure~\ref{fig:graphs}. Entries
+  \caption{Summarizing data for the graph in Figure~\ref{fig:graphs}. Entries
     contain execution time (in seconds) and counterexample sizes (counting
-    \ttp{N} constructors).}
+    \ttp{Int16} values).}
   \label{table:results}
@@
 on the program shown in Figure~\ref{fig:initial} (we omit SmallCheck, since as
-noted, it is not even feasible for the example with default instances).  We
-graph the final counterexample size and execution time.  Out of 2000 runs, we
-show the number of runs resulting in a shrunk counterexample of a given size (by
-counting \ttp{N} constructors), and we show the number of runs executing within
-a given time limit (in seconds).  Note that we provide two execution-time
-graphs, since time scales differ dramatically (we use a logrithmic scale for the
-$y$ axis in one graph and a linear scale in the other).  In the table, we show
-the mean, median, median, and the results at the 95th percentile values.  (We
-give standard deviations, but not that the plots are not necessarily Gaussian.)
+noted, it is not even feasible for the example with default
+instances).\footnote{All results reported in the paper are with a
+  GHC-7.6.3-compiled program, using \ttp{-O2}, on a 2.8 GHz 4 core i7 with 16GB
+  memory under OSX v. 10.9.2.}  We graph the final counterexample size and
+execution time.
 
@@
 (QC none), to provide a baseline, truncated shrinking (QC trunc), as described
-above, in which each list is truncated to 10 elements, and tuple shrinking (QC
-tuple), also described above.  SmartCheck produces smaller counterexamples than
-the other approaches, with a very small tail.
+above, in which each list is truncated to 10 elements and \ttp{Int16} values are
+not shrunk, and tuple shrinking (QC tuple), also described above, and
+\ttp{Int16} values are also not shrunk.
+
+With each shrinking strategy, we generate 1000 runs in which a random
+counterexample is discovered (by QuickCheck), and then shrunk.  We plot the
+number of runs resulting in a shrunk counterexample of a given size.
+
+In Table~\ref{table:results}, we summarize the graphed values and provide
+execution time in seconds as well.  For both execution time and final value
+size, we summarize the mean, median, standard deviation, and the results at the
+95th percentile.  (While we provide the standard deviations, note that the plots
+are not necessarily Gaussian.)
+
+The execution time includes the time to
+discover an initial counterexample as well as time to shrink it.
+
+SmartCheck produces smaller counterexamples than the other approaches, with a
+very small tail.
 
@@
 \multirow{4}{*}{Reverse}
-  & Mean       & 0.08s, 69  & 0.53s, 35  \\
+  & Mean       & 2, 0.002s  & 2, \sci{4.8}{4}s  \\
   \cline{2-4}
-  & Median     & 0.07s, 69  & 0.17s, 33  \\
+  & Std. dev.  & 0, 0.002   & 0, \sci{8.1}{4}  \\
   \cline{2-4}
-  & Std. dev.  &            &            \\
+  & 95\%       & 2, 0.007s  & 2, \sci{5.8}{4}s  \\
+  \hline
+
+\multirow{4}{*}{Div0}
+  & Mean       & 10, 0.001s  & 5, 0.001s  \\
   \cline{2-4}
-  & 95\%       & 0.14s, 100 & 1.76s, 65  \\
+  & Std. dev.  & 6, 0.002    & 0, 0.001   \\
+  \cline{2-4}
+  & 95\%       & 23, 0.002s  & 5, 0.002s \\
   \hline
 
-\multirow{4}{*}{Div0}
-  & Mean       & 0.08s, 69  & 0.53s, 35  \\
+\multirow{4}{*}{Heap}
+  & Mean       & 15, 0.016s  & 7, 0.007s  \\
   \cline{2-4}
-  & Median     & 0.07s, 69  & 0.17s, 33  \\
+  & Std. dev.  & 1, 0.025    & 2, 0.003   \\
   \cline{2-4}
-  & Std. dev.  &            &            \\
+  & 95\%       & 19, 0.045s  & 12, 0.015s  \\
+  \hline
+
+\multirow{4}{*}{Parser}
+  & Mean       & 191, 0.013s  & 16, 0.241s  \\
   \cline{2-4}
-  & 95\%       & 0.14s, 100 & 1.76s, 65  \\
+  & Std. dev.  & 195, 0.018   & 75, 0.800   \\
+  \cline{2-4}
+  & 95\%       & 577, 0.044s  & 14, 0.521s  \\
   \hline
 
-\multirow{4}{*}{Heap}
-  & Mean       & 0.08s, 69  & 0.53s, 35  \\
-  \cline{2-4}
-  & Median     & 0.07s, 69  & 0.17s, 33  \\
-  \cline{2-4}
-  & Std. dev.  &            &            \\
-  \cline{2-4}
-  & 95\%       & 0.14s, 100 & 1.76s, 65  \\
-  \hline
     \end{tabular}
@@
 \end{table}
+
+\todo{note that 95 is smaller than average for heap}
 
@@
 \section*{Acknowledgments}
-I thank Andy Gill, Joe Hurd, John Launchbury, Simon Winwood, and especially John
-Hughes for reading an earlier draft of this paper.  Their deep insights and bug
-catches dramatically improved both the form and content.  This paper was and
-rejected from ICFP'13; the anonymous reviewers' suggestions improved the
-presentation additional benchmarking.
+I thank Andy Gill, Joe Hurd, John Launchbury, Simon Winwood, the anonymous
+reviewers of ICFP'13 (from which an early draft of this paper was rejected), and
+especially John Hughes for reading an earlier draft of this paper.  Their deep
+insights and bug catches dramatically improved both the form and content.
 

### 2014-05-03 regression/Div0/Div0.hs
regression: fixed qc runs with shrinking.

@@ -20,7 +20,7 @@ import Data.Typeable
 data Exp = C Int
          | Add Exp Exp
          | Div Exp Exp
-  deriving (Show, Typeable, Generic)
+  deriving (Show, Read, Typeable, Generic)
 
 instance SubTypes Exp
 
@@ -70,7 +70,7 @@ findVal a@(Add e0 e1)
 findVal _                 = error "not possible"
 
 size :: Exp -> Int
-size exp = case exp of
+size e = case e of
   C _       -> 1
   Add e0 e1 -> 1 + size e0 + size e1
   Div e0 e1 -> 1 + size e0 + size e1
@@ -91,9 +91,9 @@ main = do
   let rnds = read rnds' :: Int
   let file  = read file' :: String
 #ifdef qc
-  test file rnds (runQC' stdArgs prop_div) size
+  test file rnds $ runQC' proxy stdArgs prop_div size
 #else
-  test file rnds (runSC scStdArgs prop_div) size
+  test file rnds $ runSC scStdArgs prop_div size
 #endif
 
 --------------------------------------------------------------------------------

### 2014-05-03 regression/Heap/Heap_Program.hs

@@ -22,7 +22,6 @@ import Test.QuickCheck.Poly
 
 import Data.List
   ( sort
-  , (\\)
   )
 import Data.Typeable
 
@@ -37,6 +36,12 @@ import qualified Test.SmartCheck as SC
 
 deriving instance Typeable OrdA
 deriving instance Generic OrdA
+
+instance Read OrdA where
+  readsPrec i = \s ->
+    let rd = readsPrec i s :: [(Integer,String)] in
+    let go (i',s') = (OrdA i', s') in
+    map go rd
 
 heapProgramTest :: IO ()
 heapProgramTest = SC.smartCheck SC.scStdArgs prop_ToSortedList
@@ -60,7 +65,7 @@ instance (Ord a, Arbitrary a) => Arbitrary (Heap a) where
 data Heap a
   = Node a (Heap a) (Heap a)
   | Nil
- deriving ( Eq, Ord, Show, Typeable, Generic )
+ deriving ( Eq, Ord, Show, Read, Typeable, Generic )
 
 empty :: Heap a
 empty = Nil
@@ -122,7 +127,7 @@ data HeapP a
   | SafeRemoveMin (HeapP a)
   | Merge (HeapP a) (HeapP a)
   | FromList [a]
- deriving ( Show, Typeable, Generic )
+ deriving ( Show, Read, Typeable, Generic )
 
 heap :: Ord a => HeapP a -> Heap a
 heap Empty             = empty
@@ -172,7 +177,7 @@ instance Arbitrary a => Arbitrary (HeapP a) where
   shrink _                 = []
 
 data HeapPP a = HeapPP (HeapP a) (Heap a)
- deriving ( Show, Typeable, Generic )
+ deriving ( Show, Read, Typeable, Generic )
 
 instance (Ord a, Arbitrary a) => Arbitrary (HeapPP a) where
   arbitrary =
@@ -198,7 +203,7 @@ sizePP :: HeapPP a -> Int
 sizePP (HeapPP h0 h1) = sizeP h0 + sizeH h1
 
 sizeP :: HeapP a -> Int
-sizeP heap = case heap of
+sizeP hp = case hp of
   Empty           -> 1
   Unit _          -> 1
   Insert _ h      -> 1 + sizeP h
@@ -207,9 +212,12 @@ sizeP heap = case heap of
   FromList ls     -> 1 + length ls
 
 sizeH :: Heap a -> Int
[…]

### 2014-05-03 regression/List/List.hs

@@ -9,11 +9,7 @@ module Main where
 import Test
 import Test.QuickCheck
 import Test.SmartCheck
-import Control.Monad
 import System.Environment
-
-import GHC.Generics
-import Data.Typeable
 
 -----------------------------------------------------------------
 
@@ -36,9 +32,9 @@ main = do
   let rnds = read rnds' :: Int
   let file  = read file' :: String
 #ifdef qc
-  test file rnds (runQC' stdArgs prop_rev :: IO (Maybe [Int])) length
+  test file rnds $ runQC' (proxy :: Proxy [Int]) stdArgs prop_rev length
 #else
-  test file rnds (runSC scStdArgs prop_rev) length
+  test file rnds $ runSC scStdArgs prop_rev length
 #endif
 
 --------------------------------------------------------------------------------

### 2014-05-03 regression/PaperExample1/PaperExample1.hs

@@ -10,7 +10,6 @@ module Main where
 
 import Test
 import Test.SmartCheck
-import Test.SmartCheck.Reduce
 import Test.QuickCheck
 #ifdef small
 import Test.LazySmallCheck hiding (Property, test, (==>))
@@ -21,7 +20,6 @@ import GHC.Generics hiding (P, C)
 import Data.Typeable
 
 import Data.Int
-import Data.Word
 import Control.Monad
 import System.Environment
 
@@ -31,19 +29,19 @@ import Test.Feat
 
 -----------------------------------------------------------------
 
-#if defined(qcjhint) || defined(qcNone) || defined(qc10) || defined(qc20)
+#if defined(qcjh) || defined(qcNone) || defined(qc10) || defined(qc20)
 -- So that Int16s aren't shrunk by default arbitrary instances.
-newtype J = J { getInt :: Int16 } deriving Show
+newtype J = J { getInt :: Int16 } deriving (Show, Read)
 type I = [J]
 instance Arbitrary J where
   arbitrary = fmap J arbitrary
-
+  shrink (J i) = map J (shrink i)
 #else
 type I = [Int16]
 #endif
 
-data T = T (Word8) I I I I
-  deriving (Show, Typeable, Generic)
+data T = T I I I I I
+  deriving (Read, Show, Typeable, Generic)
 
 -- SmallCheck --------------------------
 #ifdef small
@@ -81,16 +79,16 @@ instance Arbitrary T where
 #if defined(qcNone) || defined(feat)
   shrink _ = []
 #endif
-#if defined(qcjh) || defined(qcjhint)
-  shrink (T w i0 i1 i2 i3) = map go xs
-    where xs = shrink (w, i0, i1, i2, i3)
-          go (w', i0', i1', i2', i3') = T w' i0' i1' i2' i3'
+#if defined(qcjh)
+  shrink (T i0 i1 i2 i3 i4) = map go xs
+    where xs = shrink (i0, i1, i2, i3, i4)
+          go (i0', i1', i2', i3', i4') = T i0' i1' i2' i3' i4'
 #endif
 #if defined(qc10) || defined(qc20)
-  shrink (T w i0 i1 i2 i3) =
-    [ T a b c d e | a <- tk w
-                  , b <- tk i0, c <- tk i1
-                  , d <- tk i2, e <- tk i3 ]
+  shrink (T i0 i1 i2 i3 i4) =
[…]

### 2014-05-03 regression/Parser/Parser.hs

@@ -387,7 +387,7 @@ prop_parse :: Lang -> Bool
 prop_parse e = read' (show' e) == e
 
 scargs :: ScArgs
-scargs = scStdArgs { qcArgs  = theArgs
+scargs = scStdArgs { qcArgs  = stdArgs
                                 -- { maxSuccess = 1000
                                 -- , maxSize    = 20  }
                    , format  = PrintString
@@ -398,18 +398,15 @@ scargs = scStdArgs { qcArgs  = theArgs
                    , scMaxReduce = 10
                    }
 
-theArgs :: Args
-theArgs = stdArgs { maxSuccess = 100 }
-
 main :: IO ()
 main = do
   [file', rnds'] <- getArgs
   let rnds = read rnds' :: Int
   let file  = read file' :: String
 #ifdef qc
-  test file rnds (runQC' theArgs prop_parse) size
+  test file rnds $ runQC' proxy stdArgs prop_parse size
 #else
-  test file rnds (runSC scargs prop_parse) size
+  test file rnds $ runSC scargs prop_parse size
 #endif
 
 --------------------------------------------------------------------------------

### 2014-05-03 regression/Test.hs

@@ -13,15 +13,15 @@ import Test.SmartCheck.Reduce
 
 --------------------------------------------------------------------------------
 
-test :: FilePath -> Int -> IO (Maybe a) -> (a -> Int) -> IO ()
-test f rnds run size = do
-  res <- replicateM rnds (test' run size)
+test :: FilePath -> Int -> IO Res -> IO ()
+test f rnds run = do
+  res <- replicateM rnds run
   let res'        = catMaybes res
   let rnds'       = length res'
   let app str     = appendFile logFile (str ++ "\n")
   let avg vals    = sum vals / fromIntegral rnds'
   let med vals    = sort vals !! (rnds' `div` 2)
-  let times = fst $ unzip res'
+  let times       = fst $ unzip res'
   let szs :: [Double]
       szs = map fromIntegral (snd $ unzip res')
   let stdDev vals = sqrt (avg distances)
@@ -37,7 +37,7 @@ test f rnds run size = do
   app "***************"
   print res
   app $ "Num     : " ++ show rnds'
-  app $ "std dev : " ++ show (stdDev $ map (fromRational . toRational) times)
+  app $ "std dev : " ++ show (stdDev $ map (fromRational . toRational) times :: Double)
                      ++ ", " ++ show (stdDev szs)
   app $ "Avg     : " ++ show (avg times) ++ ", " ++ show (avg szs)
   app $ "Med     : " ++ show (med times) ++ ", " ++ show (med szs)
@@ -54,16 +54,15 @@ test f rnds run size = do
 
 type Res = Maybe (Double, Int)
 
-test' :: IO (Maybe a) -> (a -> Int) -> IO Res
-test' run size = do
-  start <- getCurrentTime
-  res <- run
-  stop <- getCurrentTime
-  let diff = diffUTCTime stop start
-  print diff
-  case res of
-    Nothing -> return Nothing
-    Just r  -> return $ Just (fromRational $ toRational diff, size r)
+-- test' :: IO (Maybe a) -> (a -> Int) -> IO Res
+-- test' run size = do
+--   start <- getCurrentTime
+--   res <- run
+--   stop <- getCurrentTime
+--   let diff = diffUTCTime stop start
+--   case res of
+--     Nothing -> return Nothing
+--     Just r  -> return $ Just (fromRational $ toRational diff, size r)
 
 -- For gnuplot ---------------------------------------
[…]

### 2014-05-03 src/Test/SmartCheck/Test.hs
smartcheck: comment wibble.

@@ -1,6 +1,6 @@
 {-
 
-The following is modified by Lee Pike 2014 and still retains the following
+The following is modified by Lee Pike (2014) and still retains the following
 license:
 
 Copyright (c) 2000-2012, Koen Claessen

### 2014-05-05 paper/paper.tex
paper: updates.

@@
 we show the difficulty in designing an efficient shrink implementation manually,
-which is required when using QuickCheck; we will walk through a couple of poor
-designs before arriving at a ``canonical'' manual solution.  However, even the
+which is required when using QuickCheck; we will show a poor
+design before arriving at a ``canonical'' manual solution.  However, even the
 canonical solution produces significantly larger counterexamples and takes
@@
   variables.}  Data type \ttp{T} is a product type containing five lists of
-signed 8-bit integers.  For our purposes, it is not important what \ttp{T}
-models, but for concreteness, suppose it models the values in five linked lists.
+signed 8-bit integers.  %% For our purposes, it is not important what \ttp{T}
+%% models, but for concreteness, suppose it models the values in five linked lists.
 
@@
 Despite the simplicity of the example, a typical counterexample returned by
-QuickCheck can be large.  With standard settings and no shrinking, the median
-counterexample discovered contains just under 70 \ttp{Int16} values, and
-counterexamples can contain over 100 values.  Thus, it pays to define
+QuickCheck can be large.  With standard settings and no shrinking, the average
+counterexample discovered contains just over 80 \ttp{Int16} values, and over
+five percent contain over 100 values.  Thus, it pays to define
 \ttp{shrink}!
@@
 \begin{code}
-newtype T = T { getInt :: Int16 }
-\end{code}
-%
-\noindent
-Without shrinking the \ttp{Int16}s, the performance is dramatically improved.
-
+newtype J = J { getInt :: Int16 }
+\end{code}
+%
+\noindent
 Next, a programmer might try to control the complexity by truncating lists using
@@
 arbitrary tuples).  Thus, each argument to \ttp{T} is independently shrunk.
-Still, using this definition, from some initial counterexamples, shrinking may
-result in a final counterexample containing just over 100 \ttp{Int16} values.
-
-SmallCheck is another testing framework for Haskell for which shrinking is
-irrelevant: SmallCheck is guaranteed to return a smallest counterexample, if
-one exists~\cite{sc}.  SmallCheck does this by enumerating all possible inputs,
-ordered from smallest to largest, up to some user-defined bound.  While
-SmallCheck is effective for testing many programs and properties (in accordance
-with the \emph{small scope hypothesis}~\cite{jackson}), counterexamples to even
-relatively simple properties may be infeasible to discover due to input-space
-explosion.
-
-%% \begin{figure}
-%%   \begin{code}
-%% instance Serial Int16 where
-%%   series d = drawnFrom [(-d')..d']
-%%     where d' = fromIntegral d
-
-%% instance Serial A where series = cons1 A
-%% instance Serial B where series = cons4 B
-%%   \end{code}
-%%   \caption{SmallCheck instances for the data types in
-%%     Figure~\ref{fig:initial}.}
-%%   \label{fig:smallshrink}
-%% \end{figure}
-
-For a product type, SmallCheck must check values produced by taking the
-cross-product of the type's fields at a given depth.  Unfortunately, using
-default instances, SmallCheck fails long before it reaches a depth required to
-discover a counterexample.  (A related library named Feat combines some aspects
-of SmallCheck and QuickCheck~\cite{feat}; we discuss it in
-Section~\ref{sec:related}).
+
+Finally, with this combination of shrinking tuples and disabling shrinking for
+\ttp{Int16} values, we have an efficient and effective shrinking approach.
 
@@
 \hline \hline
-Mean       & 81, 0.013s  & 40, 0.231s   & 13, 0.228s  & 6, 0.022s\\
+Mean       & 83, 0.013s  & 40, 0.064s   & 10, 0.031s  & 6, 0.021s\\
 \hline
-Std. dev.  & 21, 0.001   & 21, 1.742    & 7,  0.962   & 4, 0.001s\\
+Std. dev.  & 22, 0.014   & 20, 0.334    & 4,  0.039   & 4, 0.009\\
 % Median     & 0.009s, 67  & 0.17s, 33  & 1.81s, 13  & 0.10s, 6\\
@@
 \hline
-95\%       & 116, 0.025s  & 84, 0.283s  & 26, 0.501s  & 13, 0.039s\\
+95\%       & 119, 0.023s  & 84, 0.153s  & 16, 0.052s  & 13, 0.036s\\
 \hline
@@
 
-To motivate the benefits of SmartCheck reduction algorithm, consider
-Table~\ref{table:results} and the corresponding graphs in
-Figure~\ref{fig:graphs}, giving the results for using QuickCheck and SmartCheck
-on the program shown in Figure~\ref{fig:initial} (we omit SmallCheck, since as
-noted, it is not even feasible for the example with default
-instances).\footnote{All results reported in the paper are with a
-  GHC-7.6.3-compiled program, using \ttp{-O2}, on a 2.8 GHz 4 core i7 with 16GB
-  memory under OSX v. 10.9.2.}  We graph the final counterexample size and
-execution time.
-
-The experiments show three approaches to shrinking with QuickCheck: no shrinking
-(QC none), to provide a baseline, truncated shrinking (QC trunc), as described
-above, in which each list is truncated to 10 elements and \ttp{Int16} values are
-not shrunk, and tuple shrinking (QC tuple), also described above, and
-\ttp{Int16} values are also not shrunk.
-
-With each shrinking strategy, we generate 1000 runs in which a random
-counterexample is discovered (by QuickCheck), and then shrunk.  We plot the
-number of runs resulting in a shrunk counterexample of a given size.
+The results of our discussion above are summarized in Table~\ref{table:results}
+and the corresponding graph in Figure~\ref{fig:graphs}. They show three
+approaches to shrinking with QuickCheck: no shrinking (QC none), to provide a
+baseline, truncated shrinking (QC trunc) with \ttp{Int16} shrinking disabled, in
+which each list is truncated to 10 elements, and tuple shrinking (QC tuple),
+also with \ttp{Int16} shrinking disabled.  We omit SmallCheck, since as noted,
+it is not even feasible for the example with default instances.\footnote{All
+  results reported in the paper are with a GHC-7.6.3-compiled program, using
+  \ttp{-O2}, on a 2.8 GHz 4~core i7 with 16GB memory under OSX v. 10.9.2.}  We
+graph the final counterexample size, measured in the total number of \ttp{Int16}
+values, and execution time for counterexample discovery and shrinking.
+Figure~\ref{fig:graphs} shows how many of the 1000 runs resulted in a final
+value of a particular size, fitting the data with a Bezier curve.
 
@@
 
-The execution time includes the time to
-discover an initial counterexample as well as time to shrink it.
-
-SmartCheck produces smaller counterexamples than the other approaches, with a
-very small tail.
-
-More dramatic, however, are the differences in execution time.  The execution
-times for SmartCheck include the initial counterexample generation time required
-by QuickCheck (QC none); thus the time required by SmartCheck to shrink a
-counterexample is on average just two hundredths of a second (0.10 - 0.08).  As
-can be seen in the graphs, SmartCheck's execution time scales linearly (like
-QuickCheck without shrinking).  Truncation and tuple shrinking produce very long
-tails; the longest execution time recorded for truncation shrinking and tuple
-shrinking are 88 seconds and 110 seconds, respectively, although these outliers
-can vary wildly.  SmartCheck is largely insensitive to the size of the original
-counterexample provided.
+SmartCheck produces consistently small counterexamples and does so
+quickly---even smaller and faster than the best hand-tuned solution---all
+without writing any problem-specific code.  Note that the ``no shrinking''
+example provides a lower-bound on execution time (only including counterexample
+discovery time) and an upper-bound on counterexample size.  In this example, the
+hand-written ``tuple shrink'' algorithm is only slightly slower and produces
+slightly larger results.  In Section~\ref{sec:benchmarks}, we provide additional
+benchmarks, but now we turn to the design of SmartCheck.
+
+%% More dramatic, however, are the differences in execution time.  The execution
+%% times for SmartCheck include the initial counterexample generation time
+%% required by QuickCheck (QC none); thus the time required by SmartCheck to
+%% shrink a counterexample is on average just two hundredths of a second (0.10 -
+%% 0.08).  As can be seen in the graphs, SmartCheck's execution time scales
+%% linearly (like QuickCheck without shrinking).  Truncation and tuple shrinking
+%% produce very long tails; the longest execution time recorded for truncation
+%% shrinking and tuple shrinking are 88 seconds and 110 seconds, respectively,
+%% although these outliers can vary wildly.  SmartCheck is largely insensitive
+%% to the size of the original counterexample provided.
 
@@
 
-Finally, these performance results for SmartCheck are fairly generalizable to
-other types and properties, since the algorithm is not dependent on the specific
-program or property being checked.  However, the results are particularly
-pronounced for counterexamples that are particularly ``deep'' or ``broad'' in
-terms of their tree structure.  %% For programs and properties for which small
-%% counterexamples are easily found, QuickCheck's shrinking is perfectly
-%% acceptable.  Furthermore, user-defined shrink methods may vary significantly
-%% based on the property and counterexample type, so it is not always clear which
-%% implementation to compare against.
+%% Finally, these performance results for SmartCheck are fairly generalizable to
+%% other types and properties, since the algorithm is not dependent on the
+%% specific program or property being checked.  However, the results are
+%% particularly pronounced for counterexamples that are particularly ``deep'' or
+%% ``broad'' in terms of their tree structure.  %% For programs and properties
+%% for which small counterexamples are easily found, QuickCheck's shrinking is
+%% perfectly acceptable.  Furthermore, user-defined shrink methods may vary
+%% significantly based on the property and counterexample type, so it is not
+%% always clear which implementation to compare against.
 We now describe in detail SmartCheck's algorithm for generic shrinking.
@@
 
-\subsection{Benchmarks}
-
-No set of testing benchmarks exists over which to compare different test-case
-generation and minimization approaches.  Therefore, we have collected a small
-number of benchmarks, in addition to the more involved case-studies described
-earlier in this section.
+\subsection{Benchmarks}\label{sec:benchmarks}
+
+Unfortunately, no set of testing benchmarks exists over which to compare
+different test-case generation and minimization approaches.  Therefore, we have
+collected a small number of benchmarks, in addition to the more involved
+case-studies described earlier in this section.
 
 The benchmarks presented, in addition to the motivating example presented in
-Section~\ref{sec:example}, are
+Section~\ref{sec:example}, compare standard SmartCheck against hand-written
+shrink implementations.  The benchmarks are as follows:
 
@@
 \noindent
-All benchmarks can be found at
-\url{https://github.com/leepike/SmartCheck/tree/master/regression}.  Benchmarks
-are against
-QuickCheck~2.7.3.\footnote{\url{http://hackage.haskell.org/package/QuickCheck-2.7.3}}
-The benchmarks only compare the size of shrunk values and the time required to
-generate them using standard arguments to QuickCheck and SmartCheck.  We do not
-benchmark the generalization algorithms, since there is no analog in QuickCheck
-or other testing frameworks.
-
-\todo{include std dev.}
-\todo{fill in benchmarks!}
+All benchmarks can be found
+online.\footnote{\url{https://github.com/leepike/SmartCheck/tree/master/regression}}
+We compare the size of the final counterexample returned (by counting
+constructors) and the time required for counterexample generation and shrinking
+in seconds.  The results are presented in Table~\ref{table:results}.  Again, we
+summarize the mean, median, standard deviation, and the results at the 95th
+percentile.  (While we provide the standard deviations, note that the plots are
+not necessarily Gaussian.)
 
@@
 
-\todo{note that 95 is smaller than average for heap}
-
 The \emph{Reverse} benchmark essentially provides a lower-bound on the benefit
@@
 counterexample, however!).  Perhaps surprisingly, SmartCheck still slightly
-outperforms QuickCheck.  The other two benchmarks are slightly larger examples,
-where SmartCheck's benefits are more pronounced.  But again, we emphasize the
-most pronounced benefit comes in much larger examples, like described in
-Section~\ref{sec:nlp}.
+outperforms QuickCheck.  The other two benchmarks have much larger counterexamples,
+where SmartCheck's benefits are more pronounced.  For example, consider the
+parser example in which the initial counterexample to shrink may be in a large
+parse tree.  SmartCheck performs significantly better than optimal hand-written
+shrink instances, but even SmartCheck can sometimes return large
+counterexamples---this is why the mean counterexample size is larger than the
+size at the 95th percentile: there are a small number of large values not
+successfully shrunk by SmartCheck.
 
@@
 
+SmallCheck is another testing framework for Haskell for which shrinking is
+irrelevant: SmallCheck is guaranteed to return a smallest counterexample, if
+one exists~\cite{sc}.  SmallCheck does this by enumerating all possible inputs,
+ordered from smallest to largest, up to some user-defined bound.
+
+While SmallCheck is effective for testing many programs and properties (in
+accordance with the \emph{small scope hypothesis}~\cite{jackson}),
+counterexamples to even relatively simple properties may be practically
+infeasible to discover due to the size of the input space.  For example,
+SmallCheck does not find a counterexample to the example presented in
+Section~\ref{sec:example} after running it for several minutes.
+
 Besides QuickCheck and SmallCheck, another testing framework related to
@@
 
-Feat finds small counterexamples effectively, but a limitation compared to
-QuickCheck is that value generation is applicative rather than monadic.
-Consequently, defining custom generators (i.e., instances for \ttp{arbitrary})
-is constrained.  On the other hand, generators can be derived
-automatically with Feat.
-
-We have benchmarked Feat as well on the example in the introduction
-(Figure~\ref{fig:initial}).  It performs quite competitively to SmartCheck
-in terms the size of counterexamples discovered: the mean and median number of
-\ttp{N} constructors in the counterexamples returned by Feat are both seven. \todo{check}
-However, Feat performs less well with respect to how long counterexample
-discovery takes: the mean time taken by Feat is 5.60 seconds, with a long tail:
-five percent of runs take over 14 seconds (the mean time taken by
-SmartCheck is 0.12 seconds, and 99\% of runs take less than 0.30 seconds).
+Feat is used to discover counterexamples, not shrink them.  However, shrinking
+is less necessary with Feat, since discovered counterexamples are often small,
+if one is found.  For example, on the overflow example in
+Section~\ref{sec:example}, with a limit of 100 test cases, Feat finds a
+counterexample just two percent of the time, whereas QuickCheck finds one nearly
+100\%.  Even at a limit of 10000, a counterexample is found about 50\% of the
+time (with a correspondingly longer search time).  Sampling from a uniform
+distribution does not work so well here.  Feat does a better job of discovering
+counterexamples in the parser benchmark, but the size of the average
+counterexample contains 500 constructors, with a standard deviation of 500
+(compared with 16 and 75, respectively, for SmartCheck).  Still, Feat is
+powerful at what it does well and can be seamlessly used with SmartCheck, since
+it just defines the \ttp{arbitrary} method.
+
+%% When Feat finds a counterexample, it is a small one, rendering shrinking
+%%   A limitation compared to QuickCheck (and SmartCheck, since
+%% SmartCheck uses QuickCheck as a front-end for initial counterexample discovery)
+%% is that value generation is applicative rather than monadic.  Consequently,
+%% defining custom generators (i.e., instances for \ttp{arbitrary}) is constrained.
+%% On the other hand, generators can be derived automatically with Feat.  The main
+%% limitation, however
+
+%% We have benchmarked Feat as well on the example in the introduction
+%% (Figure~\ref{fig:initial}).  It performs quite competitively to SmartCheck in
+%% terms the size of counterexamples discovered: the mean and median number of
+%% \ttp{N} constructors in the counterexamples returned by Feat are both
+%% seven. \todo{check} However, Feat performs less well with respect to how long
+%% counterexample discovery takes: the mean time taken by Feat is 5.60 seconds,
+%% with a long tail: five percent of runs take over 14 seconds (the mean time taken
+%% by SmartCheck is 0.12 seconds, and 99\% of runs take less than 0.30 seconds).
 
@@
 
-\todo{mention memoization (generalized tries)}
-
-\todo{tail-recursive algorithm is fast and easy to port to imperative languages}
+%% \todo{mention memoization (generalized tries)}
+
+%% \todo{tail-recursive algorithm is fast and easy to port to imperative languages}
 

### 2014-05-05 regression/PaperExample1/PaperExample1.hs
regression/build: wibbles.

@@ -8,7 +8,7 @@
 
 module Main where
 
-import Test
+--import Test
 import Test.SmartCheck
 import Test.QuickCheck
 #ifdef small
@@ -35,7 +35,6 @@ newtype J = J { getInt :: Int16 } deriving (Show, Read)
 type I = [J]
 instance Arbitrary J where
   arbitrary = fmap J arbitrary
-  shrink (J i) = map J (shrink i)
 #else
 type I = [Int16]
 #endif
@@ -143,20 +142,20 @@ prop_small t = pre t S.==> post t
 size :: T -> Int
 size t = sum $ map length (toList t)
 
-main :: IO ()
-main = do
-  [file', rnds'] <- getArgs
-  let rnds = read rnds' :: Int
-  let file  = read file' :: String
-#ifdef feat
-  test file rnds $ runQC' proxy stdArgs prop size
-#endif
-#ifdef smart
-  test file rnds $ runSC scStdArgs prop size
-#endif
-#if defined(qcNone) || defined(qc10) || defined(qc20) || defined(qcjh) || defined(qcNaive)
-  test file rnds $ runQC' proxy stdArgs prop size
-#endif
+-- main :: IO ()
+-- main = do
+--   [file', rnds'] <- getArgs
+--   let rnds = read rnds' :: Int
+--   let file  = read file' :: String
+-- #ifdef feat
+--   test file rnds $ runQC' proxy stdArgs {maxSuccess = 10000} prop size
+-- #endif
+-- #ifdef smart
+--   test file rnds $ runSC scStdArgs prop size
+-- #endif
+-- #if defined(qcNone) || defined(qc10) || defined(qc20) || defined(qcjh) || defined(qcNaive)
+--   test file rnds $ runQC' proxy stdArgs prop size
+-- #endif
 
 #ifdef smart
 -- Tester (not part of the benchmark).

### 2014-05-05 regression/Parser/Parser.hs

@@ -1,4 +1,5 @@
 {-# LANGUAGE CPP #-}
+{-# LANGUAGE TemplateHaskell #-}
 {-# LANGUAGE DeriveDataTypeable #-}
 {-# LANGUAGE DeriveGeneric #-}
 
@@ -20,6 +21,10 @@ import Control.Applicative
 
 import Control.Monad.State
 import Data.Char
+
+#ifdef feat
+import Test.Feat
+#endif
 
 -----------------------------------------------------------------
 
@@ -78,6 +83,17 @@ data Exp = Int Int
 
 instance SubTypes Exp
 
+-- Feat --------------------------------
+#ifdef feat
+deriveEnumerable ''Lang
+deriveEnumerable ''Var
+deriveEnumerable ''Mod
+deriveEnumerable ''Func
+deriveEnumerable ''Stmt
+deriveEnumerable ''Exp
+#endif
+-- Feat --------------------------------
+
 --------------------------------------------------------------------------------
 
 nonEmpty :: Gen [a] -> Gen [a]
@@ -89,18 +105,24 @@ instance Arbitrary Var where
 
 instance Arbitrary Lang where
   arbitrary = Lang <$> nonEmpty arbitrary <*> nonEmpty arbitrary
+#ifndef feat
   shrink (Lang m f) = map go (shrink (m, f))
     where go (a,b) = Lang a b
+#endif
 
 instance Arbitrary Mod where
   arbitrary = Mod <$> nonEmpty arbitrary <*> nonEmpty arbitrary
+#ifndef feat
   shrink (Mod a b) = map go (shrink (a, b))
     where go (x,y) = Mod x y
+#endif
 
 instance Arbitrary Func where
   arbitrary = Func <$> arbitrary <*> nonEmpty arbitrary <*> nonEmpty arbitrary
+#ifndef feat
   shrink (Func f a st) = map go (shrink (a, st))
     where go (x, s) = Func f x s
+#endif
 
 instance Arbitrary Stmt where
   arbitrary = do
@@ -110,10 +132,12 @@ instance Arbitrary Stmt where
     let a1 = Alloc v e
     let a2 = Return e
     elements [a0, a1, a2]
+#ifndef feat
   shrink stmt = case stmt of
     Assign v e -> map (Assign v) (shrink e)
     Alloc v e  -> map (Alloc v) (shrink e)
     Return e   -> map Return (shrink e)
+#endif
 
[…]

### 2014-05-09 paper/paper.tex
regression: add generic QC shrink instances.

@@
 \begin{abstract}
-QuickCheck, developed by Claessen and Hughes, is a powerful library for
-automatic test-case generation.  Because QuickCheck performs random testing,
-some of the counterexamples discovered are very large.  QuickCheck provides an
-interface for the user to write shrink functions to attempt to reduce the size
-of counterexamples.  Hand-written implementations of shrink can be complex,
-inefficient, and consist of significant boilerplate code.  Furthermore,
-shrinking is only one aspect in debugging: counterexample generalization is the
-process of extrapolating from individual counterexamples to a class of
-counterexamples, often requiring a flash of insight from the programmer.  To
-improve counterexample reduction and generalization, we introduce
-\emph{SmartCheck}.  SmartCheck is a debugging tool that reduces algebraic data
-using generic search heuristics to efficiently find smaller counterexamples.  In
-addition to shrinking, SmartCheck also automatically generalizes counterexamples
-to formulas representing classes of counterexamples.  SmartCheck has been implemented for Haskell and is freely available.
+QuickCheck is a powerful library for automatic test-case generation.  Because
+QuickCheck performs random testing, some of the counterexamples discovered are
+very large.  QuickCheck provides an interface for the user to write shrink
+functions to attempt to reduce the size of counterexamples.  Hand-written
+implementations of shrink can be complex, inefficient, and consist of
+significant boilerplate code.  Furthermore, shrinking is only one aspect in
+debugging: counterexample generalization is the process of extrapolating from
+individual counterexamples to a class of counterexamples, often requiring a
+flash of insight from the programmer.  To improve counterexample reduction and
+generalization, we introduce \emph{SmartCheck}.  SmartCheck is a debugging tool
+that reduces algebraic data using generic search heuristics to efficiently find
+smaller counterexamples.  In addition to shrinking, SmartCheck also
+automatically generalizes counterexamples to formulas representing classes of
+counterexamples.  SmartCheck has been implemented for Haskell and is freely
+available.
 \end{abstract}
@@
 with respect to the property, so that the user can focus on the heart of the
-problem in a class of counterexamples.
+problem in a class of counterexamples.  Given a program and large
+counterexample, \emph{SmartCheck} returns such a formula.
 
@@
 Motivated by the problems of reducing and generalizing large counterexamples, we
-developed \emph{SmartCheck}.  SmartCheck takes a counterexample produced by some
+developed SmartCheck.  SmartCheck takes a counterexample produced by some
 oracle and generically minimizes and generalizes the counterexample.
@@
 After presenting some preliminary definitions in
-Section~\ref{sec:preliminaries}, in Section~\ref{sec:shrinking}, we describe the
-generic shrinking algorithm implemented.  The algorithm is particularly novel in
-that it generates new random sub-values during the reduction phase (see related
-work in Section~\ref{sec:related}).
+Section~\ref{sec:preliminaries}, in Section~\ref{sec:shrinking}, we describe
+SmartCheck's generic counterexample reduction algorithm.  The algorithm is
+particularly novel in that it generates new random sub-values during the
+reduction phase (see related work in Section~\ref{sec:related}).
 
@@
 toList :: T -> [[Int16]]
-toList (T i0 i1 i2 i3 i4) =
-  (map . map) fromIntegral [i0, i1, i2, i3, i4]
+toList (T i0 i1 i2 i3 i4) = [i0, i1, i2, i3, i4]
 
 pre :: T -> Bool
-pre t = all ((>) 256 . sum) (toList t)
+pre t = all ((< 256) . sum) (toList t)
 
@@
   variables.}  Data type \ttp{T} is a product type containing five lists of
-signed 8-bit integers.  %% For our purposes, it is not important what \ttp{T}
+signed 16-bit integers.  %% For our purposes, it is not important what \ttp{T}
 %% models, but for concreteness, suppose it models the values in five linked lists.
@@
 \begin{code}
-shrink (T w i0 i1 i2 i3) =
-  [ T a b c d e | a <- shrink w,  b <- shrink i0
-                , c <- shrink i1, d <- shrink i2
-                , e <- shrink i3 ]
+shrink (T i0 i1 i2 i3 i4) =
+  [ T a b c d e | a <- shrink i0, b <- shrink i1
+                , c <- shrink i2, d <- shrink i3
+                , e <- shrink i4 ]
 \end{code}
@@
 \begin{code}
-newtype J = J { getInt :: Int16 }
+newtype J = J \{ getInt :: Int16 \}
 \end{code}
@@
 \item \ttp{opaque} is false when the type of the value is an ``interesting
-  type''; informally, this is a type other than a primitive type like \ttp{Int}
-  or \ttp{Char}.  See Section~\ref{sec:base} for a full discussion.
+  type''; informally, this is a type other than a primitive type like \ttp{Int},
+  \ttp{Char}, or \ttp{Bool}.  See Section~\ref{sec:base} for a full discussion.
 \item \ttp{subVals} returns a tree of all non opaque-type sub-values.  A tree is a
@@
 \section*{Acknowledgments}
-I thank Andy Gill, Joe Hurd, John Launchbury, Simon Winwood, the anonymous
-reviewers of ICFP'13 (from which an early draft of this paper was rejected), and
-especially John Hughes for reading an earlier draft of this paper.  Their deep
-insights and bug catches dramatically improved both the form and content.
+I thank Andy Gill, Joe Hurd, John Launchbury, Simon Winwood, Tritan Ravitch, the
+anonymous reviewers of ICFP'13 (from which an earlier draft of this paper was
+rejected), and especially John Hughes.  Their deep insights and bug catches
+dramatically improved the presentation and content.
 

### 2014-05-09 regression/Div0/Div0.hs

@@ -40,10 +40,14 @@ instance Arbitrary Exp where
     mkM n = oneof [ liftM2 Add mkM' mkM'
                   , liftM2 Div mkM' mkM' ]
       where mkM' = mkM =<< choose (0,n-1)
-
+#ifdef qc
   shrink (C i)       = map C (shrink i)
   shrink (Add e0 e1) = [e0, e1]
   shrink (Div e0 e1) = [e0, e1]
+#endif
+#ifdef qcGen
+  shrink = genericShrink
+#endif
 
 -- property: so long as 0 isn't in the divisor, we won't try to divide by 0.
 -- It's false: something might evaluate to 0 still.
@@ -90,7 +94,7 @@ main = do
   [file', rnds'] <- getArgs
   let rnds = read rnds' :: Int
   let file  = read file' :: String
-#ifdef qc
+#if defined(qc) || defined(qcGen)
   test file rnds $ runQC' proxy stdArgs prop_div size
 #else
   test file rnds $ runSC scStdArgs prop_div size

### 2014-05-09 regression/Heap/Heap_Program.hs

@@ -2,6 +2,7 @@
 
 {-# LANGUAGE ScopedTypeVariables, TemplateHaskell, DeriveDataTypeable, StandaloneDeriving #-}
 {-# LANGUAGE DeriveGeneric #-}
+{-# LANGUAGE DeriveDataTypeable #-}
 
 {-# OPTIONS_GHC -fno-warn-missing-signatures #-}
 {-# OPTIONS_GHC -fno-warn-orphans #-}
@@ -54,7 +55,7 @@ instance (SC.SubTypes a, Arbitrary a, Generic a)
 instance (SC.SubTypes a, Ord a, Arbitrary a, Generic a)
          => SC.SubTypes (HeapPP a)
 
-instance (Ord a, Arbitrary a) => Arbitrary (Heap a) where
+instance (Ord a, Arbitrary a, Typeable a) => Arbitrary (Heap a) where
   arbitrary = do p <- arbitrary :: Gen (HeapP a)
                  return $ heap p
 
@@ -139,7 +140,7 @@ heap (SafeRemoveMin p) = case removeMin (heap p) of
 heap (Merge p q)       = heap p `merge` heap q
 heap (FromList xs)     = fromList xs
 
-instance Arbitrary a => Arbitrary (HeapP a) where
+instance (Typeable a, Arbitrary a) => Arbitrary (HeapP a) where
   arbitrary = sized arbHeapP
    where
     arbHeapP s =
@@ -162,7 +163,7 @@ instance Arbitrary a => Arbitrary (HeapP a) where
       s1 = s-1
       s2 = s`div`2
 
-
+#ifdef qc
   shrink (Unit x)          = [ Unit x' | x' <- shrink x ]
   shrink (FromList xs)     = [ Unit x | x <- xs ]
                           ++ [ FromList xs' | xs' <- shrink xs ]
@@ -175,18 +176,25 @@ instance Arbitrary a => Arbitrary (HeapP a) where
                           ++ [ Merge p' q | p' <- shrink p ]
                           ++ [ Merge p q' | q' <- shrink q ]
   shrink _                 = []
+#endif
+#ifdef qcGen
+  shrink = genericShrink
+#endif
 
 data HeapPP a = HeapPP (HeapP a) (Heap a)
  deriving ( Show, Read, Typeable, Generic )
 
-instance (Ord a, Arbitrary a) => Arbitrary (HeapPP a) where
+instance (Ord a, Arbitrary a, Typeable a) => Arbitrary (HeapPP a) where
[…]

### 2014-05-09 regression/List/List.hs

@@ -31,7 +31,7 @@ main = do
   [file', rnds'] <- getArgs
   let rnds = read rnds' :: Int
   let file  = read file' :: String
-#ifdef qc
+#if defined(qc) || defined(qcGen)
   test file rnds $ runQC' (proxy :: Proxy [Int]) stdArgs prop_rev length
 #else
   test file rnds $ runSC scStdArgs prop_rev length

### 2014-05-09 regression/PaperExample1/PaperExample1.hs

@@ -8,7 +8,7 @@
 
 module Main where
 
---import Test
+import Test
 import Test.SmartCheck
 import Test.QuickCheck
 #ifdef small
@@ -103,6 +103,9 @@ instance Arbitrary T where
                   , b <- shrink i1, c <- shrink i2
                   , d <- shrink i3, e <- shrink i4 ]
 #endif
+#if defined(qcGen)
+  shrink = genericShrink
+#endif
 
 -- Feat --------------------------------
 #ifdef feat
@@ -115,12 +118,12 @@ toList (T i0 i1 i2 i3 i4) =
 #if defined(qcjh) || defined(qcNone) || defined(qc10) || defined(qc20)
   (map . map) (fromIntegral . getInt) [i0, i1, i2, i3, i4]
 #else
-  (map . map) fromIntegral [i0, i1, i2, i3, i4]
+  [i0, i1, i2, i3, i4]
 #endif
 
 
 pre :: T -> Bool
-pre t = all ((>) 256 . sum) (toList t)
+pre t = all ((< 256) . sum) (toList t)
 
 post :: T -> Bool
 post t = (sum . concat) (toList t) < 5 * 256
@@ -142,20 +145,20 @@ prop_small t = pre t S.==> post t
 size :: T -> Int
 size t = sum $ map length (toList t)
 
--- main :: IO ()
--- main = do
---   [file', rnds'] <- getArgs
---   let rnds = read rnds' :: Int
---   let file  = read file' :: String
--- #ifdef feat
---   test file rnds $ runQC' proxy stdArgs {maxSuccess = 10000} prop size
--- #endif
--- #ifdef smart
---   test file rnds $ runSC scStdArgs prop size
--- #endif
--- #if defined(qcNone) || defined(qc10) || defined(qc20) || defined(qcjh) || defined(qcNaive)
---   test file rnds $ runQC' proxy stdArgs prop size
--- #endif
+main :: IO ()
+main = do
+  [file', rnds'] <- getArgs
+  let rnds = read rnds' :: Int
+  let file  = read file' :: String
+#ifdef feat
+  test file rnds $ runQC' proxy stdArgs {maxSuccess = 10000} prop size
+#endif
+#ifdef smart
+  test file rnds $ runSC scStdArgs prop size
+#endif
+#if defined(qcNone) || defined(qc10) || defined(qc20) || defined(qcjh) || defined(qcNaive) || defined(qcGen)
[…]

### 2014-05-09 regression/Parser/Parser.hs

@@ -105,23 +105,32 @@ instance Arbitrary Var where
 
 instance Arbitrary Lang where
   arbitrary = Lang <$> nonEmpty arbitrary <*> nonEmpty arbitrary
-#ifndef feat
+#ifdef qc
   shrink (Lang m f) = map go (shrink (m, f))
     where go (a,b) = Lang a b
+#endif
+#ifdef qcGen
+  shrink = genericShrink
 #endif
 
 instance Arbitrary Mod where
   arbitrary = Mod <$> nonEmpty arbitrary <*> nonEmpty arbitrary
-#ifndef feat
+#ifdef qc
   shrink (Mod a b) = map go (shrink (a, b))
     where go (x,y) = Mod x y
+#endif
+#ifdef qcGen
+  shrink = genericShrink
 #endif
 
 instance Arbitrary Func where
   arbitrary = Func <$> arbitrary <*> nonEmpty arbitrary <*> nonEmpty arbitrary
-#ifndef feat
+#ifdef qc
   shrink (Func f a st) = map go (shrink (a, st))
     where go (x, s) = Func f x s
+#endif
+#ifdef qcGen
+  shrink = genericShrink
 #endif
 
 instance Arbitrary Stmt where
@@ -132,11 +141,14 @@ instance Arbitrary Stmt where
     let a1 = Alloc v e
     let a2 = Return e
     elements [a0, a1, a2]
-#ifndef feat
+#ifdef qc
   shrink stmt = case stmt of
     Assign v e -> map (Assign v) (shrink e)
     Alloc v e  -> map (Alloc v) (shrink e)
     Return e   -> map Return (shrink e)
+#endif
+#ifdef qcGen
+  shrink = genericShrink
 #endif
 
 instance Arbitrary Exp where
@@ -166,7 +178,7 @@ instance Arbitrary Exp where
                     , Not  <$> goa
                     , And  <$> goa <*> gob
                     ]
-#ifndef feat
+#ifdef qc
   shrink e = case e of
     Int  i    -> map Int (shrink i)
     Bool b    -> map Bool (shrink b)
@@ -177,6 +189,9 @@ instance Arbitrary Exp where
     Not e0    -> map Not (shrink e0)
     And e0 e1 -> map (uncurry And) (zip (shrink e0) (shrink e1))
     Or e0 e1  -> map (uncurry Or) (zip (shrink e0) (shrink e1))
+#endif
+#ifdef qcGen
+  shrink = genericShrink
 #endif
 
[…]

### 2014-05-10 examples/Comp.hs
build: deprecate examples.

@@ -1,93 +1,0 @@
-{-# LANGUAGE DeriveDataTypeable #-}
-{-# LANGUAGE DeriveGeneric #-}
-
-module Comp where
-
-import Test.QuickCheck
-import Test.SmartCheck
-import Control.Monad
-
-import GHC.Generics
-import Data.Typeable
-
------------------------------------------------------------------
-
-data I = I Int
-  deriving (Eq, Show, Typeable, Generic)
-
-instance Arbitrary I where
-  arbitrary = liftM I arbitrary
-
-instance SubTypes I
-
-instance Ord I where
-  compare (I i) (I j) = compare i j
-
-data X = X0 I
-       | X1 I
-  deriving (Show, Typeable, Generic)
-
-instance Arbitrary X where
-  arbitrary = oneof [ liftM X0 arbitrary
-                    , liftM X1 arbitrary ]
-
-instance SubTypes X
-
-instance Ord X where
-  compare (X0 i) (X0 j) = compare i j
-  compare (X0 i) (X1 j) = compare i j
-  compare (X1 i) (X0 j) = compare i j
-  compare (X1 i) (X1 j) = compare i j
-
-instance Eq X where
-  (X0 i) == (X0 j) = i == j
-  (X0 i) == (X1 j) = i == j
-  (X1 i) == (X0 j) = i == j
-  (X1 i) == (X1 j) = i == j
-
-data A = A0 Int
-       | A1 Int
-  deriving (Eq, Show, Typeable, Generic)
-
-instance Ord A where
-  compare (A0 i) (A0 j) = compare i j
-  compare (A0 i) (A1 j) = compare i j
-  compare (A1 i) (A0 j) = compare i j
-  compare (A1 i) (A1 j) = compare i j
-
-instance Arbitrary A where
-  arbitrary = oneof [ liftM A0 arbitrary
-                    , liftM A1 arbitrary ]
-
-instance SubTypes A
-
-data L = L [Int]
-  deriving (Show, Typeable, Generic)
-
-instance Arbitrary L where
-  arbitrary = liftM L arbitrary
-
-instance SubTypes L
-
-data B = B L X
-  deriving (Show, Typeable, Generic)
-
-instance Arbitrary B where
-  arbitrary = liftM2 B arbitrary arbitrary
-
-instance SubTypes B
-
-comp :: X -> X -> A -> Bool
-comp a0 a1 _ = a0 /= a1
-
-comp2 :: X -> Bool
-comp2 a0 = a0 < a0
-
[…]

### 2014-05-10 examples/Lambda.hs

@@ -1,443 +1,0 @@
--- From QuickCheck test suite.
-
-{-# LANGUAGE ScopedTypeVariables, TemplateHaskell #-}
-{-# LANGUAGE DeriveGeneric #-}
-{-# LANGUAGE DeriveDataTypeable #-}
-module Main where
-
---------------------------------------------------------------------------
--- imports
-
-import Data.Functor ((<$>))
-import Data.Typeable
-import GHC.Generics
-import Test.SmartCheck
-
-import Test.QuickCheck
-
-import Data.Maybe
-
-import Control.Monad
-  ( liftM
-  , liftM2
-  )
-
-import Data.Char
-  ( toUpper
-  )
-
-import Data.Set (Set)
-import qualified Data.Set as Set
-
-import Debug.Trace
-
---------------------------------------------------------------------------
--- types for lambda expressions
-
--- variables
-
-newtype Var = MkVar String
-  deriving ( Eq, Ord, Show, Read, Typeable, Generic )
-
-instance SubTypes Var
-
--- instance Show Var where
---   show (MkVar s) = s
-
-varList :: [Var]
-varList = [ MkVar s
-          | let vs = [ c:v | v <- "" : vs, c <- ['a'..'z'] ]
-          , s <- vs
-          ]
-
-instance Arbitrary Var where
-  arbitrary = growingElements [ MkVar [c] | c <- ['a'..'z'] ]
-
--- constants
-
-newtype Con = MkCon String
-  deriving ( Eq, Ord, Show, Read, Typeable, Generic )
-
-instance SubTypes Con
-
--- instance Show Con where
---   show (MkCon s) = s
-
-instance Arbitrary Con where
-  arbitrary = growingElements [ MkCon [c] | c <- ['A'..'Z'] ]
-
--- expressions
-
-data Exp
-  = Lam Var Exp
-  | App Exp Exp
-  | Var Var
-  | Con Con
- deriving ( Eq, Ord, Show, Read, Typeable, Generic )
-
-instance SubTypes Exp
-
--- instance Show Exp where
---   showsPrec n (Lam x t) = showParen (n>0) (showString "\\" . shows x . showString "." . shows t)
---   showsPrec n (App s t) = showParen (n>1) (showsPrec 1 s . showString " " . showsPrec 2 t)
---   showsPrec _ (Var x)   = shows x
---   showsPrec _ (Con c)   = shows c
-
-instance Arbitrary Exp where
-  arbitrary = sized arbExp
[…]

### 2014-05-10 examples/LambdaCalc.hs

@@ -1,141 +1,0 @@
-{-# LANGUAGE DeriveDataTypeable #-}
-{-# LANGUAGE DeriveGeneric #-}
-
--- Copied from <http://augustss.blogspot.com/2007/10/simpler-easier-in-recent-paper-simply.html>
-
-module LambdaCalc where
-
-import Data.List
-import Data.Typeable
-
-import Control.Monad
-import GHC.Generics
-
-import Test.QuickCheck
-
-import Test.SmartCheck
-
-type Sym = String
-
-data Expr
-        = Var Sym
-        | App Expr Expr
-        | Lam Sym Expr
-        deriving (Eq, Read, Show, Typeable, Generic)
-
-freeVars :: Expr -> [Sym]
-freeVars (Var v) = [v]
-freeVars (App f a) = freeVars f `union` freeVars a
-freeVars (Lam i e) = freeVars e \\ [i]
-
-subst :: Sym -> Expr -> Expr -> Expr
-subst v x b = sub b
-  where sub e@(Var i) = if i == v then x else e
-        sub (App f a) = App (sub f) (sub a)
-        sub (Lam i e) =
-            if v == i then
-                Lam i e
-            else if i `elem` fvx then
-                let i' = cloneSym e i
-                    e' = substVar i i' e
-                in  Lam i' (sub e')
-            else
-                Lam i (sub e)
-        fvx = freeVars x
-        cloneSym e i = loop i
-           where loop i' = if i' `elem` vs then loop (i ++ "'") else i'
-                 vs = fvx ++ freeVars e
-
-substVar :: Sym -> Sym -> Expr -> Expr
-substVar v v' e = subst v (Var v') e
-
-alphaEq :: Expr -> Expr -> Bool
-alphaEq (Var v)   (Var v')    = v == v'
-alphaEq (App f a) (App f' a') = alphaEq f f' && alphaEq a a'
-alphaEq (Lam v e) (Lam v' e') = alphaEq e (substVar v' v e')
-alphaEq _ _ = False
-
-nf :: Expr -> Expr
-nf ee = spine ee []
-  where spine (App f a) as = spine f (a:as)
-        spine (Lam v e) [] = Lam v (nf e)
-        spine (Lam v e) (a:as) = spine (subst v a e) as
-        spine f as = app f as
-        app f as = foldl App f (map nf as)
-
-betaEq :: Expr -> Expr -> Bool
-betaEq e1 e2 = alphaEq (nf e1) (nf e2)
-
-z,s,m,n :: Expr
[…]

### 2014-05-10 examples/Large.hs

@@ -1,113 +1,0 @@
-{-# LANGUAGE DeriveDataTypeable #-}
-{-# LANGUAGE DeriveGeneric #-}
-
--- | Try to generate a very large counterexample.
-
-module Large where
-
-import Test.SmartCheck
-import Test.QuickCheck
-import Test.LazySmallCheck hiding (Property, test, (==>))
-
-import GHC.Generics hiding (P, C)
-import Data.Typeable
-import Control.Monad
-
-import Data.Int
-
------------------------------------------------------------------
-
--- Let's try to generate a product type of long lists when all we need is a
--- single element to have a long list.
-
--- Container so that we don't have base types.
-data A = A Int16
-  deriving (Read, Show, Typeable, Generic)
-
-instance Serial Int16 where
-  series d = drawnFrom [(-d')..d']
-    where d' = fromIntegral d
-
-instance SubTypes A
-
-instance Arbitrary A where
-  arbitrary = liftM A arbitrary
-  shrink (A i) = map A (shrink i)
-
-instance Serial A where
-  series = cons1 A
-
-data B = B [A] [A] [A] [A]
-  deriving (Read, Show, Typeable, Generic)
-
-instance SubTypes B
-
--- qc/shrink takes over 1m seconds
-instance Arbitrary B where
-  arbitrary = liftM4 B arbitrary arbitrary arbitrary arbitrary
---  shrink (B a b c d) = [ B w x y z | w <- shrink a, x <- shrink b, y <- shrink c, z <- shrink d ]
-
-instance Serial B where
-  series = cons4 B 
-
-add :: [A] -> Int16
-add = sum . map (\(A i) -> i)
-
-pre :: B -> Bool
-pre (B a b c d) = and $ map pre' [a, b, c, d]
-  where
-  pre' x = add x < 16
-
-test :: B -> Bool
-test (B a b c d) = add a + add b + add c + add d < 64
-
-prop_p :: B -> Property
-prop_p p = pre p ==> test p
-
-main :: IO ()
-main = smartCheck scStdArgs { runForall = False, runExists = False } prop_p
-  -- smartCheck scStdArgs { format = PrintString 
-  --                      , scMaxDepth = Just 20
-  --                      } 
-  --   prop_p
-
-
--- sc :: Int -> IO ()
--- sc n = smallCheck n test
-
-t0 :: Int -> Bool
-t0 a = a < 99999999999
-
----------------------
-
-
[…]

### 2014-05-10 examples/MutualRecData.hs

@@ -1,60 +1,0 @@
-{-# LANGUAGE DeriveDataTypeable #-}
-{-# LANGUAGE DeriveDataTypeable #-}
-{-# LANGUAGE DeriveGeneric #-}
-
-module MutualRecData where
-
-import Test.SmartCheck
-import Test.QuickCheck hiding (Result)
-
-import Data.Tree
-import Control.Monad.State
-import Data.Typeable
-
-import GHC.Generics
-
---------------------------------------------------------------------------------
-
-data M = M N N Int | P
-  deriving (Typeable, Show, Eq, Read, Generic)
-
-instance SubTypes M
-
-data N = N M Int String
-  deriving (Typeable, Show, Eq, Read, Generic)
-
-instance SubTypes N
-
---------------------------------------------------------------------------------
-
-instance Arbitrary M where
-  arbitrary =
-    sized $ \n -> if n == 0 then return P
-                    else oneof [ return P
-                               , liftM3 M (resize (n-1) arbitrary)
-                                          (resize (n-1) arbitrary)
-                                          arbitrary
-                               ]
-
-instance Arbitrary N where
-  arbitrary = liftM3 N arbitrary arbitrary arbitrary
-
---------------------------------------------------------------------------------
-
-prop0 :: M -> Bool
-prop0 (M _ _ a) = a < 100
-prop0 _         = True
-
-mutRecTest :: IO ()
-mutRecTest = smartCheck args prop0
-  where
-  args = scStdArgs { qcArgs = stdArgs {maxSuccess = 1000} }
-
---------------------------------------------------------------------------------
-
-xx :: M
-xx = M (N (M (N P 1 "goo") (N P 7 "foo") 8) 3 "hi") (N P 4 "bye") 6
-yy :: Forest Int
-yy = [Node 0 [Node 1 [], Node 2 []], Node 3 [Node 4 [], Node 5 [Node 6 []]]]
-
---------------------------------------------------------------------------------

### 2014-05-10 examples/NewType.hs

@@ -1,34 +1,0 @@
-{-# LANGUAGE DeriveDataTypeable #-}
-{-# LANGUAGE DeriveGeneric #-}
-{-# LANGUAGE FlexibleInstances #-}
-
-module NewType where
-
-import Test.QuickCheck
-import Test.SmartCheck
-
-import Control.Monad
-import Data.Word
-
-import GHC.Generics
-import Data.Typeable
-
------------------------------------------------------------------
-
-prop :: Maybe I -> Bool
-prop Nothing      = True
-prop (Just (I i)) = (100 > i)
-
-data I = I Word8
-  deriving (Eq, Read, Show, Typeable, Generic)
-
-instance Arbitrary I where
-  arbitrary    = liftM I arbitrary
-  shrink (I i) = map I (shrink i)
-
-instance SubTypes I
-
-sc :: IO ()
-sc = smartCheck scStdArgs prop
-
------------------------------------------------------------------

### 2014-05-10 examples/Protocol.hs

@@ -1,134 +1,0 @@
-{-# LANGUAGE DeriveDataTypeable #-}
-{-# LANGUAGE DeriveGeneric #-}
-
--- A made up protocol translation.
-
-module Protocol where
-
-import Prelude hiding (last, id)
-import Data.List hiding (last)
-import Control.Monad
-import Test.QuickCheck
-import Test.SmartCheck
-
-import GHC.Generics
-import Data.Typeable
-
--- Input format
-
-data Input = Input
-  { header :: Header }
-  deriving (Generic, Typeable, Show, Eq, Read)
-
-instance SubTypes Input
-
-data Header = Header
-  { name   :: String
-  , id     :: String
-  , size   :: Int -- total number of Strings and Ints in the fields.
-  , fields :: Fields
-  , misc   :: [String] }
-  deriving (Generic, Typeable, Show, Eq, Read)
-
-instance SubTypes Header
-
-data Fields = Fields
-  { f0   :: Ins
-  , f1   :: Ins
-  , f2   :: Ins
-  , f3   :: Ins
-  , f4   :: Ins
-  , f5   :: Ins
-  , f6   :: Ins
-  , f7   :: Ins
-  , f8   :: Ins
-  , f9   :: Ins
-  , fn   :: Ins' }
-  deriving (Generic, Typeable, Show, Eq, Read)
-
-instance SubTypes Fields
-
-data Ins = Strs [String]
-         | Ints [Int]
-         | Else [String]
-  deriving (Generic, Typeable, Show, Eq, Read)
-
-instance SubTypes Ins
-
-data Ins' = Strs' [String]
-         | Ints' [Int]
-         | Else' [String]
-  deriving (Generic, Typeable, Show, Eq, Read)
-
-instance SubTypes Ins'
-
--- Output format
-data Output = Output
-  { outSize      :: Int
-  , collapseStrs :: [String]
-  , collapseInts :: [Int]
-  }
-  deriving Show
-
-instance Arbitrary Ins where
-  arbitrary = oneof [ liftM Strs arbitrary
-                    , liftM Ints arbitrary
-                    , liftM Else arbitrary ]
-
-instance Arbitrary Ins' where
-  arbitrary = oneof [ liftM Strs' arbitrary
-                    , liftM Ints' arbitrary
-                    , liftM Else' arbitrary ]
-
-instance Arbitrary Fields where
[…]

### 2014-05-10 examples/Tests.hs

@@ -1,11 +1,0 @@
-module Main where
-
-import Div0
-import MutualRecData
-import Heap_Program
-
-main :: IO ()
-main = do
-  divTest
-  mutRecTest
-  heapProgramTest

### 2014-05-10 paper/paper.tex
paper: updates.

@@
 \usepackage{pgf}
+\usepackage{nth}
 
@@
 
-\newcommand{\sci}[2]{\ensuremath{#1 \times 10^{#2}}}
+%% \newcommand{\sci}[2]{\ensuremath{#1 \times 10^{#2}}}
+\newcommand{\sci}[2]{\ensuremath{#1e^{#2}}}
 \begin{document}
@@
 large counterexample.  We are left facing the \emph{shrinking problem}: assume
-there exist a set of inputs $i_0, \ i_1, \ldots, \i_n$, such that for program
+there exist a set of inputs $i_0, \ i_1, \ldots, \ i_n$, such that for program
 $P$ (modeled as a function mapping inputs to output) and property $\phi$,
-$\phi(P(i_0, \ i_1, \ldots, \i_n))$ is false.  The shrinking problem is to find
-the smallest set of inputs $i'_0, \ i'_1, \ldots, \i'_n$, according to some
-measure, such that $\phi(P(i\_0, \ i\_1, \ldots, \i\_n))$ is also false.
+$\phi(P(i_0, \ i_1, \ldots, \ i_n))$ is false.  The shrinking problem is to find
+the smallest set of inputs $i'_0, \ i'_1, \ldots, \ i'_n$, according to some
+measure, such that $\phi(P(i\_0, \ i\_1, \ldots, \ i\_n))$ is also false.
 
@@
 example, a property stating that pretty-printing an AST and then parsing it
-results in the original AST may fail due to multiple bugs.  During testing, it
-is useful to discover counterexamples arising from each bug in one go.  In
-practice, the problem is solved by discovering a counterexample \ttp{cex},
-abstracting it, and then adding a new precondition to the property that
-informally says ``omit counterexamples of form \ttp{cex}.''  Adding
-preconditions manually is laborious and may cause the programmer to make
-premature fixes to the program, if she believes she has isolated the error
-before she actually does.
+results in the original AST may fail due to multiple bugs, such that each bug in
+isolation is sufficient to cause failure.  During testing, it is useful to
+discover counterexamples arising from all the bugs in one go.  In practice, the
+problem is solved by discovering a counterexample \ttp{cex}, abstracting it, and
+then adding a new precondition to the property that informally says ``omit
+counterexamples of form \ttp{cex}.''  Adding preconditions manually is laborious
+and may cause the programmer to make premature fixes to the program, if she
+believes she has isolated the error before she actually does.
 
@@
 
+The most recent released version of QuickCheck provides a \ttp{genericShrink}
+function for algebraic data types, so the user merely has to declare
+%
+\begin{code}
+shrink = genericShrink
+\end{code}
+%
+\noindent
+to obtain shrink instances.  (The implementation relies on the same generics
+package that SmartCheck does; see Section~\ref{sec:preliminaries}.)
+
 %% is unable to verify the property from Figure~\ref{fig:initial}, for much the
@@
   \begin{center}
-    \begin{tabular}{|r||c|c|c|c|}
+    \begin{tabular}{|r||c|c|c|c|c|c|}
 \hline
-           & QC none     & QC trunc     & QC tuple    & SmartCheck\\
-\hline \hline
-Mean       & 83, 0.013s  & 40, 0.064s   & 10, 0.031s  & 6, 0.021s\\
-\hline
-Std. dev.  & 22, 0.014   & 20, 0.334    & 4,  0.039   & 4, 0.009\\
-% Median     & 0.009s, 67  & 0.17s, 33  & 1.81s, 13  & 0.10s, 6\\
-%% \hline
-%% MAD        & 0.02s, 13  & 0.09s, 8   & 0.62s, 3   & 0.02s, 2\\
-\hline
-95\%       & 119, 0.023s  & 84, 0.153s  & 16, 0.052s  & 13, 0.036s\\
-\hline
+
+              & \multicolumn{2}{c}{Mean} & \multicolumn{2}{c}{Std. dev.} &
+\multicolumn{2}{c}{95\%}\\\hline
+              & size & time  & size & time & size & time \\\hline\hline
+QC none       & 83 & 0.013   & 22 & 0.014   & 119 & 0.023\\\hline
+QC trunc      & 40 & 0.064   & 20 & 0.334   & 84 & 0.153\\\hline
+QC tuple      & 10 & 0.031   & 4 &  0.039   & 16 & 0.052\\\hline
+QC generic    & &        &        &           & &\\\hline
+SmartCheck    & 6 & 0.021s    &  4 & 0.009   & 13 & 0.036\\\hline
+
     \end{tabular}
@@
 The results of our discussion above are summarized in Table~\ref{table:results}
-and the corresponding graph in Figure~\ref{fig:graphs}. They show three
+and the corresponding graph in Figure~\ref{fig:graphs}. They show four
 approaches to shrinking with QuickCheck: no shrinking (QC none), to provide a
 baseline, truncated shrinking (QC trunc) with \ttp{Int16} shrinking disabled, in
-which each list is truncated to 10 elements, and tuple shrinking (QC tuple),
-also with \ttp{Int16} shrinking disabled.  We omit SmallCheck, since as noted,
-it is not even feasible for the example with default instances.\footnote{All
-  results reported in the paper are with a GHC-7.6.3-compiled program, using
-  \ttp{-O2}, on a 2.8 GHz 4~core i7 with 16GB memory under OSX v. 10.9.2.}  We
-graph the final counterexample size, measured in the total number of \ttp{Int16}
-values, and execution time for counterexample discovery and shrinking.
+which each list is truncated to 10 elements, tuple shrinking (QC tuple), also
+with \ttp{Int16} shrinking disabled, and finally shrinking using QuickCheck's
+\ttp{genericShrink} (qc generic).  We omit SmallCheck, since as noted, it is not
+even feasible for the example with default instances.\footnote{All results
+  reported in the paper are with a GHC-7.6.3-compiled program, using \ttp{-O2},
+  on a 2.8 GHz 4~core i7 with 16GB memory under OSX v. 10.9.2.}  We graph the
+final counterexample size, measured in the total number of \ttp{Int16} values,
+and execution time for counterexample discovery and shrinking.
 Figure~\ref{fig:graphs} shows how many of the 1000 runs resulted in a final
@@
   type''; informally, this is a type other than a primitive type like \ttp{Int},
-  \ttp{Char}, or \ttp{Bool}.  See Section~\ref{sec:base} for a full discussion.
-\item \ttp{subVals} returns a tree of all non opaque-type sub-values.  A tree is a
-  value with the type
+  \ttp{Char}, or \ttp{Bool}, and may be user-defined.  See
+  Section~\ref{sec:base} for a full discussion.
+\item \ttp{subVals} returns a tree of all non opaque-type sub-values.  A tree
+  has the type
 %
@@
 subVals (B (B L\sub{0} L\sub{1}) L\sub{2}) =
-  Node (B L\sub{0} L\sub{1}) [Node L\sub{0} [], Node L\sub{1} []]
-       (Node L\sub{1} [])
+  Node (B (B L\sub{0} L\sub{1}) L\sub{2})
+       [ Node (B L\sub{0} L\sub{1}) [Node L\sub{0} [], Node L\sub{1} []]
+       , Node L\sub{2} []
+       ]
 \end{code}
@@
 \ttp{cex}, and the property \ttp{prop}.  The reduction begins at the first
-proper sub-value of \ttp{cex}; call it \ttp{v}.  When the index \ttp{idx}
-becomes out-of-bounds and returns \ttp{Nothing}, the algorithm terminates.
-Otherwise, a list of new random values are generated.
+proper sub-value of \ttp{cex}; call it \ttp{v} (this is an optimization
+described in Section~\ref{sec:variant}).  When the index \ttp{idx} becomes
+out-of-bounds and returns \ttp{Nothing}, the algorithm terminates.  Otherwise, a
+list of new random values are generated.
 %
@@
 
-\subsubsection{Variant Counterexample Hypothesis}
+\subsubsection{Variant Counterexample Hypothesis}\label{sec:variant}
 A motivation for the design of the reduction algorithm is something we call the
@@
 Recall that the algorithm begins at the \emph{first} sub-value of the
-counterexample rather than the 0th sub-value.  The 0th sub-value of \ttp{cex} is
-\ttp{cex} itself.  No invariant of the algorithm would be violated by beginning
-with the 0th sub-value, and in particular, the algorithm would still terminate.
-However, we begin with a strict sub-value because of the hypothesis.
+counterexample rather than the zeroth sub-value.  The zeroth sub-value of
+\ttp{cex} is \ttp{cex} itself.  No invariant of the algorithm would be violated
+by beginning with the zeroth sub-value, and in particular, the algorithm would
+still terminate.  However, we begin with a strict sub-value because of the
+hypothesis.
 
@@
 may potentially fail to discover a smaller counterexample.  However, in
-Sections~\ref{sec:existential} and~\ref{sec:precondition}, we describe an
+Sections~\ref{sec:existential} and~\ref{sec:precondition}, we describe 
 approaches to generalize counterexamples based on discovering new counterexample
@@
 We generalize the idea of ignoring primitive types by introducing the concept of
-\emph{opaque types}.  If the reduction algorithm encounters a opaque type, it is
+\emph{opaque types}.  If the reduction algorithm encounters an opaque type, it is
 ignored.  Opaque types include the primitive types mentioned above, but in the
-user can declare any substructure in a data type to be a opaque type by
+user can declare any substructure in a data type to be an opaque type by
 providing custom instances.  Doing so effectively treats values from that type
@@
 Opaque types can be conditional.  For example, the user may want lists to be
-shrunk in general, unless the element of the list are opaque themselves.  Such a
+shrunk in general, unless the elements of the list are opaque themselves.  Such a
 definition is possible.
@@
 divSubTerms (C _)         = True
-odivSubTerms (Div _ (C 0)) = False
+divSubTerms (Div _ (C 0)) = False
 divSubTerms (Add e0 e1)   =  divSubTerms e0
@@
 the debugging process.  A characterizing formula reduces the noise in favor of
-the signal by abstracting away portions of large counterexample that are
+the signal by abstracting away portions of a large counterexample that are
 irrelevant to why it violates the property.
@@
 new counterexamples.
+
+In SmartCheck, the universal and existential generalization phases run after a
+counterexample has been minimized.  Precondition strengthening is used when
+iteratively generating multiple counterexamples, so it is interspersed with
+counterexample reduction.
 
@@
 
-The extrapolation algorithm is shown in Figure~\ref{fig:universal}; let sketch
+The extrapolation algorithm is shown in Figure~\ref{fig:universal}; let us sketch
 its specification.  The algorithm is similar to the reduction algorithm in
@@
 \noindent
-Then the following would all counterexamples:
+Then the following would all be counterexamples:
 %
@@
 the fourth equation.
+
+SmartCheck can run in a real-eval-print loop (REPL), discovering new
+counterexamples, shrinking them, generalizing them, and then repeating.  In each
+iteration of the REPL, the property precondition is strengthened, requiring that
+\ttp{matchesShape} does not hold on the current counterexample candidate and
+\emph{any} previous counterexample generalization already discovered.
 
@@
 \noindent
-Recall that through the entire process, we never had to learn how GenI works,
-what the property meant, or how to write a custom shrink method!
-
-``Retro-fitting'' SmartCheck to GenI, which is a large software system, required
-some boilerplate work.  Fortunately, because the program already uses
-QuickCheck, \ttp{arbitrary} instances were already defined, but we had to hunt
-down the data types used to write \ttp{instance Subtypes D}, for each data type
-\ttp{D} (indeed, the job could have largely been done with Sed and Awk).
+Through the entire process, we never had to learn how GenI works, what the
+property meant, or how to write a custom shrink method!
+
+%% ``Retro-fitting'' SmartCheck to GenI, which is a large software system, required
+%% some boilerplate work.  Fortunately, because the program already uses
+%% QuickCheck, \ttp{arbitrary} instances were already defined, but we had to hunt
+%% down the data types used to write \ttp{instance Subtypes D}, for each data type
+%% \ttp{D} (indeed, the job could have largely been done with Sed and Awk).
 
@@
 constructors) and the time required for counterexample generation and shrinking
-in seconds.  The results are presented in Table~\ref{table:results}.  Again, we
+in seconds.  The results are presented in Table~\ref{table:benchmarks}.  Again, we
 summarize the mean, median, standard deviation, and the results at the 95th
@@
 
+\todo{generic or hand-written?}
+
 \begin{table}[ht]
-\footnotesize
+\scriptsize
   \begin{center}
-    \begin{tabular}{|r|r||c|c|}
+    \begin{tabular}{|r|r||c|c|c|c|c|c|}
+
 \hline
-\multicolumn{2}{|c||}{} & QuickCheck & SmartCheck \\
-\hline \hline
-
-\multirow{4}{*}{Reverse}
-  & Mean       & 2, 0.002s  & 2, \sci{4.8}{4}s  \\
-  \cline{2-4}
-  & Std. dev.  & 0, 0.002   & 0, \sci{8.1}{4}  \\
-  \cline{2-4}
-  & 95\%       & 2, 0.007s  & 2, \sci{5.8}{4}s  \\
-  \hline
-
-\multirow{4}{*}{Div0}
-  & Mean       & 10, 0.001s  & 5, 0.001s  \\
-  \cline{2-4}
-  & Std. dev.  & 6, 0.002    & 0, 0.001   \\
-  \cline{2-4}
-  & 95\%       & 23, 0.002s  & 5, 0.002s \\
-  \hline
-
-\multirow{4}{*}{Heap}
-  & Mean       & 15, 0.016s  & 7, 0.007s  \\
-  \cline{2-4}
-  & Std. dev.  & 1, 0.025    & 2, 0.003   \\
-  \cline{2-4}
-  & 95\%       & 19, 0.045s  & 12, 0.015s  \\
-  \hline
-
-\multirow{4}{*}{Parser}
-  & Mean       & 191, 0.013s  & 16, 0.241s  \\
-  \cline{2-4}
-  & Std. dev.  & 195, 0.018   & 75, 0.800   \\
-  \cline{2-4}
-  & 95\%       & 577, 0.044s  & 14, 0.521s  \\
-  \hline
+  \multicolumn{2}{|c||}{} & \multicolumn{2}{c|}{Mean} & \multicolumn{2}{c|}{Std. dev.} &
+\multicolumn{2}{c|}{95\%}\\\cline{3-8}
+  \multicolumn{2}{|c||}{} & size & time & size & time & size & time \\\hline\hline
+
+\multirow{2}{*}{Reverse}
+  & QC & 2 & 0.002         & 0 & 0.002         & 2 & 0.007 \\\cline{2-8}
+  & SC & 2 & \sci{4.8}{-4} & 0 & \sci{8.1}{-4} & 2 & \sci{5.8}{-4}\\\hline\hline
+
+\multirow{2}{*}{Div0}
+  & QC & 10 & 0.001 & 6 & 0.002 & 23 & 0.002 \\\cline{2-8}
+  & SC & 5 & 0.001  & 0 & 0.001 & 5 & 0.002  \\\hline\hline
+
+\multirow{2}{*}{Heap}
+  & QC & 15 & 0.016  & 1 & 0.025 & 19 & 0.045 \\\cline{2-8}
+  & SC & 7  & 0.007  & 2 & 0.003 & 12 & 0.015 \\\hline\hline
+
+\multirow{2}{*}{Parser}
+  & QC &  191 & 0.013 & 195 & 0.018 & 577 & 0.044 \\\cline{2-8}
+  & SC &  16  & 0.241 & 75  & 0.800 & 14  & 0.521 \\\hline
 
@@
     constructors).}
-  \label{table:results}
+  \label{table:benchmarks}
 \end{table}

### 2014-05-10 regression/Div0/Div0.hs
regression: build updates, add QC generic.

@@ -6,11 +6,13 @@
 
 module Main where
 
+#if defined(qc) || defined(qcGen) || defined(smart)
 import Test
+import System.Environment
+#endif
 import Test.QuickCheck
 import Test.SmartCheck
 import Control.Monad
-import System.Environment
 
 import GHC.Generics
 import Data.Typeable
@@ -89,6 +91,7 @@ divTest = smartCheck args prop_div
                    , runForall  = True
                    }
 
+#if defined(qc) || defined(qcGen) || defined(smart)
 main :: IO ()
 main = do
   [file', rnds'] <- getArgs
@@ -98,6 +101,7 @@ main = do
   test file rnds $ runQC' proxy stdArgs prop_div size
 #else
   test file rnds $ runSC scStdArgs prop_div size
+#endif
 #endif
 
 --------------------------------------------------------------------------------

### 2014-05-10 regression/Heap/Heap_Program.hs

@@ -14,9 +14,11 @@ module Main where
 --------------------------------------------------------------------------
 -- imports
 
+#if defined(qc) || defined(qcGen) || defined(smart)
 import Test
-import Test.SmartCheck
 import System.Environment
+#endif
+import Test.SmartCheck
 
 import Test.QuickCheck
 import Test.QuickCheck.Poly
@@ -227,14 +229,16 @@ sizeH hp = case hp of
 l :: HeapP OrdA
 l = FromList [OrdA 2, OrdA 1]
 
+#if defined(qc) || defined(qcGen) || defined(smart)
 main :: IO ()
 main = do
   [file', rnds'] <- getArgs
   let rnds = read rnds' :: Int
   let file  = read file' :: String
-#ifdef defined(qc) || defined(qcGen)
+#if defined(qc) || defined(qcGen)
   test file rnds $ runQC' proxy stdArgs prop_ToSortedList (sizePP :: HeapPP OrdA -> Int)
 #else
   test file rnds $ runSC scStdArgs prop_ToSortedList sizePP
+#endif
 #endif
 

### 2014-05-10 regression/List/List.hs

@@ -6,10 +6,14 @@
 
 module Main where
 
+
+#if defined(qc) || defined(qcGen) || defined(smart)
 import Test
+import System.Environment
+#endif
+
 import Test.QuickCheck
 import Test.SmartCheck
-import System.Environment
 
 -----------------------------------------------------------------
 
@@ -26,6 +30,7 @@ revTest = smartCheck args prop_rev
                    , runForall  = True
                    }
 
+#if defined(qc) || defined(qcGen) || defined(smart)
 main :: IO ()
 main = do
   [file', rnds'] <- getArgs
@@ -35,6 +40,7 @@ main = do
   test file rnds $ runQC' (proxy :: Proxy [Int]) stdArgs prop_rev length
 #else
   test file rnds $ runSC scStdArgs prop_rev length
+#endif
 #endif
 
 --------------------------------------------------------------------------------

### 2014-05-10 regression/PaperExample1/PaperExample1.hs

@@ -8,7 +8,11 @@
 
 module Main where
 
+#if defined(qcjh) || defined(qcNone) || defined(qc10) || defined(qc20) || defined(feat) || defined(qcGen) || defined(smart) || defined(small)
 import Test
+import System.Environment
+#endif
+
 import Test.SmartCheck
 import Test.QuickCheck
 #ifdef small
@@ -21,7 +25,6 @@ import Data.Typeable
 
 import Data.Int
 import Control.Monad
-import System.Environment
 
 #ifdef feat
 import Test.Feat
@@ -145,6 +148,7 @@ prop_small t = pre t S.==> post t
 size :: T -> Int
 size t = sum $ map length (toList t)
 
+#if defined(qcjh) || defined(qcNone) || defined(qc10) || defined(qc20) || defined(feat) || defined(qcGen) || defined(smart) || defined(small)
 main :: IO ()
 main = do
   [file', rnds'] <- getArgs
@@ -159,6 +163,7 @@ main = do
 #if defined(qcNone) || defined(qc10) || defined(qc20) || defined(qcjh) || defined(qcNaive) || defined(qcGen)
   test file rnds $ runQC' proxy stdArgs prop size
 #endif
+#endif
 
 #ifdef smart
 -- Tester (not part of the benchmark).
@@ -169,3 +174,4 @@ smtChk = smartCheck scStdArgs { scMaxForall = 20
                               , format = PrintString
                               } prop
 #endif
+

### 2014-05-10 regression/Parser/Parser.hs

@@ -9,10 +9,13 @@ module Main where
 
 import Prelude hiding (showList, mod)
 
+#if defined(qc) || defined(qcGen) || defined(smart)
 import Test
+import System.Environment
+#endif
+
 import Test.QuickCheck
 import Test.SmartCheck
-import System.Environment
 import Data.List
 
 import GHC.Generics
@@ -438,6 +441,8 @@ scargs = scStdArgs { qcArgs  = stdArgs
                    , scMaxSize =  5
                    , scMaxReduce = 10
                    }
+
+#if defined(qc) || defined(qcGen) || defined(smart)
 main :: IO ()
 main = do
   [file', rnds'] <- getArgs
@@ -452,7 +457,7 @@ main = do
 #ifdef smart
   test file rnds $ runSC scargs prop_parse size
 #endif
-
+#endif
 --------------------------------------------------------------------------------
 
 {-

### 2014-05-11 paper/paper.tex
paper: submitted version.

@@
 \title{SmartCheck: Automatic and Efficient Counterexample Reduction and Generalization}
-\subtitle{UNPUBLISHED DRAFT}
+%\subtitle{UNPUBLISHED DRAFT}
 
@@
 
-As shown in the motivating example in Section~\ref{sec:example}, SmartCheck
-can shrink more than an order-of-magnitude faster than even the most efficient hand-written
-shrink implementations for some properties and programs.  As well, the time
-required to shrink scales linearly with the size of the counterexample---other
-common shrink implementations often scale in polynomial time or worse.
-(Shrinking time is important in real-world uses; see
-Section~\ref{sec:experiments} for examples.)
+%% As shown in the motivating example in Section~\ref{sec:example}, SmartCheck
+%% generates smaller counterexamples faster than hand-written or generated
+%% QuickCheck shrinking, for some properties and programs.  (Shrinking time is
+%% important in real-world uses; see Section~\ref{sec:experiments} for examples.)
 
@@
 which is required when using QuickCheck; we will show a poor
-design before arriving at a ``canonical'' manual solution.  However, even the
-canonical solution produces significantly larger counterexamples and takes
-significantly longer to produce them than SmartCheck, even though SmartCheck's
-shrinking strategy is completely automatic and generic.
+design before arriving at a ``canonical'' manual solution.%%   However, even the
+%% canonical solution produces significantly larger counterexamples and takes
+%% significantly longer to produce them than SmartCheck, even though SmartCheck's
+%% shrinking strategy is completely automatic and generic.
 
@@
 
-The most recent released version of QuickCheck provides a \ttp{genericShrink}
-function for algebraic data types, so the user merely has to declare
+In early 2014, QuickCheck added the ability to derive \ttp{shrink} instances for
+algebraic data type using the same generic package that SmartCheck does; see
+Section~\ref{sec:preliminaries}.  (The first release of the SmartCheck software
+was mid-2012.)  With this new feature, the user declares
 %
@@
 \noindent
-to obtain shrink instances.  (The implementation relies on the same generics
-package that SmartCheck does; see Section~\ref{sec:preliminaries}.)
+to obtain efficient shrink instances.
 
@@
               & size & time  & size & time & size & time \\\hline\hline
-QC none       & 83 & 0.013   & 22 & 0.014   & 119 & 0.023\\\hline
+none          & 83 & 0.013   & 22 & 0.014   & 119 & 0.023\\\hline
 QC trunc      & 40 & 0.064   & 20 & 0.334   & 84 & 0.153\\\hline
 QC tuple      & 10 & 0.031   & 4  & 0.039   & 16 & 0.052\\\hline
-QC generic    & &        &        &           & &\\\hline
+QC generic    & 13 & 0.204   & 7  & 0.946   & 28 & 0.543\\\hline
 SmartCheck    & 6 & 0.021s   & 4  & 0.009   & 13 & 0.036\\\hline
@@
 The results of our discussion above are summarized in Table~\ref{table:results}
-and the corresponding graph in Figure~\ref{fig:graphs}. They show four
-approaches to shrinking with QuickCheck: no shrinking (QC none), to provide a
-baseline, truncated shrinking (QC trunc) with \ttp{Int16} shrinking disabled, in
-which each list is truncated to 10 elements, tuple shrinking (QC tuple), also
-with \ttp{Int16} shrinking disabled, and finally shrinking using QuickCheck's
-\ttp{genericShrink} (qc generic).  We omit SmallCheck, since as noted, it is not
-even feasible for the example with default instances.\footnote{All results
+and the corresponding graph in Figure~\ref{fig:graphs}.  First, we show the
+results of no shrinking (labeled `none') as a baseline in terms of worst-case
+value size.  Then we show four approaches to shrinking, as described above:
+truncated shrinking (`QC trunc') with \ttp{Int16} shrinking disabled, in which
+each list is truncated to 10 elements, tuple shrinking (`QC tuple'), also with
+\ttp{Int16} shrinking disabled, and finally shrinking using QuickCheck's
+\ttp{genericShrink} (`QC generic').\footnote{All results
   reported in the paper are with a GHC-7.6.3-compiled program, using \ttp{-O2},
@@
 In Table~\ref{table:results}, we summarize the graphed values and provide
-execution time in seconds as well.  For both execution time and final value
-size, we summarize the mean, median, standard deviation, and the results at the
-95th percentile.  (While we provide the standard deviations, note that the plots
-are not necessarily Gaussian.)
+execution time (including both counterexample discovery and shrinking) in
+seconds as well.  For both execution time and final value size, we summarize the
+mean, median, standard deviation, and the results at the 95th percentile.
+(While we provide the standard deviations, note that the plots are not
+necessarily Gaussian.)
 
@@
 quickly---even smaller and faster than the best hand-tuned solution---all
-without writing any problem-specific code.  Note that the ``no shrinking''
-example provides a lower-bound on execution time (only including counterexample
+without writing any problem-specific code.  Note that the `none' example
+provides a lower-bound on execution time (only including counterexample
 discovery time) and an upper-bound on counterexample size.  In this example, the
 hand-written ``tuple shrink'' algorithm is only slightly slower and produces
-slightly larger results.  In Section~\ref{sec:benchmarks}, we provide additional
-benchmarks, but now we turn to the design of SmartCheck.
+slightly larger results than SmartCheck.  (QuickCheck's generic shrinking under
+performs the ``tuple shrink'' likely because the generic shrink implementation
+shrinks the \ttp{Int16} values.)
+
+In Section~\ref{sec:benchmarks}, we provide additional benchmarks, but now we
+turn to the design of SmartCheck.
 
@@
 %% always clear which implementation to compare against.
-We now describe in detail SmartCheck's algorithm for generic shrinking.
 
@@
 In this section, we describe how to efficiently and generically shrink algebraic
-data values.  Recall the basic idea behind the \ttp{shrink}
-method of the \ttp{Arbitrary} class: generate a list of values, each of which is
-smaller than the current counterexample.  Each of the new values generated
-may not bear any relationship to the original counterexample other than being
-smaller.  SmartCheck pursues an approach that searches for smaller but
-structurally similar counterexamples, as we make precise below.
+data values.  Recall the basic idea behind the \ttp{shrink} method of the
+\ttp{Arbitrary} class: generate a list of values, each of which is smaller than
+the current counterexample.  Each of the new values generated may not bear any
+relationship to the original counterexample other than being smaller.
+SmartCheck pursues an approach that searches for smaller but structurally
+similar counterexamples, as we make precise below.  Perhaps the most significant
+difference between SmartCheck's reduction algorithm and QuickCheck shrink
+implementations is that the latter is deterministic.  SmartCheck combines
+counterexample search with shrinking.
 
@@
 \ttp{SubTypes} type class.
-
-\todo{reference into with QC, no int shrinking}
 
@@
 
-\todo{how long unanswered?}
-
 \noindent
@@
 collected a small number of benchmarks, in addition to the more involved
-case-studies described earlier in this section.
+case-studies described earlier in this section.  However, these are contrived
+insofar as initial counterexamples for them are discovered quickly.
 
 The benchmarks presented, in addition to the motivating example presented in
-Section~\ref{sec:example}, compare standard SmartCheck against hand-written
-shrink implementations.  The benchmarks are as follows:
+Section~\ref{sec:example}, compare standard SmartCheck against QuickCheck's
+generic shrink implementation, which is, in general, as good or better than
+hand-written shrink implementations.  The benchmarks are as follows:
 
@@
   \item \emph{Parser}, a parser/pretty-printer for a toy imperative language
-    containing a parser bug that switches the arguments of an binary operator.
+    containing a parser bug that switches the arguments of an binary expression
+    operator.
 \end{itemize}
@@
 
-\todo{generic or hand-written?}
-
 \begin{table}[ht]
@@
 \multirow{2}{*}{Reverse}
-  & QC & 2 & 0.002         & 0 & 0.002         & 2 & 0.007 \\\cline{2-8}
-  & SC & 2 & \sci{4.8}{-4} & 0 & \sci{8.1}{-4} & 2 & \sci{5.8}{-4}\\\hline\hline
+  & QC & 2 & 0.002       & 0 & 0.002       & 2 & 0.003 \\\cline{2-8}
+  & SC & 2 & \sci{4}{-4} & 0 & \sci{5}{-4} & 2 & \sci{7}{-4}\\\hline\hline
 
 \multirow{2}{*}{Div0}
-  & QC & 10 & 0.001 & 6 & 0.002 & 23 & 0.002 \\\cline{2-8}
-  & SC & 5 & 0.001  & 0 & 0.001 & 5 & 0.002  \\\hline\hline
+  & QC & 5  & 0.004 & 1 & 0.006 & 7  & 0.015 \\\cline{2-8}
+  & SC & 5  & 0.001 & 0 & 0.001 & 5  & 0.001  \\\hline\hline
 
 \multirow{2}{*}{Heap}
-  & QC & 15 & 0.016  & 1 & 0.025 & 19 & 0.045 \\\cline{2-8}
-  & SC & 7  & 0.007  & 2 & 0.003 & 12 & 0.015 \\\hline\hline
+  & QC & 19 & \sci{9}{-4}  & 9 & 0.001 & 36 & 0.001 \\\cline{2-8}
+  & SC & 7  & 0.006        & 2 & 0.002 & 10 & 0.010 \\\hline\hline
 
 \multirow{2}{*}{Parser}
-  & QC &  191 & 0.013 & 195 & 0.018 & 577 & 0.044 \\\cline{2-8}
-  & SC &  16  & 0.241 & 75  & 0.800 & 14  & 0.521 \\\hline
+  & QC &  4   & 0.024 & 0  & 0.020 & 4 & 0.061 \\\cline{2-8}
+  & SC &  6   & 0.116 & 2  & 0.083 & 9 & 0.230 \\\hline
 
@@
 The \emph{Reverse} benchmark essentially provides a lower-bound on the benefit
-of using SmartCheck to discover small counterexamples, since a list of length
-two falsifies the property (SmartCheck is still useful in generalizing the
-counterexample, however!).  Perhaps surprisingly, SmartCheck still slightly
-outperforms QuickCheck.  The other two benchmarks have much larger counterexamples,
-where SmartCheck's benefits are more pronounced.  For example, consider the
-parser example in which the initial counterexample to shrink may be in a large
-parse tree.  SmartCheck performs significantly better than optimal hand-written
-shrink instances, but even SmartCheck can sometimes return large
-counterexamples---this is why the mean counterexample size is larger than the
-size at the 95th percentile: there are a small number of large values not
-successfully shrunk by SmartCheck.
+shrinking in general, since the original counterexamples are generally close to
+being minimal.  Surprisingly, SmartCheck slightly outperforms QuickCheck in
+efficiency.  The other three benchmarks have larger counterexamples, so the
+benefit of shrinking is more pronounced.
+
+SmartCheck finds smaller counterexamples in the \emph{Div0} and \emph{Heap}
+benchmarks, while QuickCheck shrinking finds smaller counterexamples faster in
+the \emph{Parser} example.  The \emph{Parser} example is one in which
+SmartCheck's strategy of generating random values during counterexample
+reduction is not a benefit, and so this is why it likely under performs a
+generic \ttp{shrink} method.  Moreover, because SmartCheck uses QuickCheck as a
+front-end, the two approaches are complementary: QuickCheck shrinking can be
+used instead of or in combination with SmartCheck reduction.
 
@@
 
-Other researchers have also investigated counterexample shrinking.  Zeller
-and Hildebrandt describe an application of greedy search to shrink
+Zeller and Hildebrandt describe an application of greedy search to shrink
 counterexamples they call ``delta-debugging'' (DD)~\cite{dd}.  The authors apply

### 2014-06-17 paper/paper.tex
paper updates based on reviewer comments.

@@
 
+\newcommand{\xmonad}{\textsc{xmonad}}
+
 %% \conferenceinfo{ICFP'12,} {September 9--15, 2012, Copenhagen, Denmark.}
@@
 large counterexample.  We are left facing the \emph{shrinking problem}: assume
-there exist a set of inputs $i_0, \ i_1, \ldots, \ i_n$, such that for program
+there exist inputs $i_0, \ i_1, \ldots, \ i_n$, such that for program
 $P$ (modeled as a function mapping inputs to output) and property $\phi$,
 $\phi(P(i_0, \ i_1, \ldots, \ i_n))$ is false.  The shrinking problem is to find
-the smallest set of inputs $i'_0, \ i'_1, \ldots, \ i'_n$, according to some
-measure, such that $\phi(P(i\_0, \ i\_1, \ldots, \ i\_n))$ is also false.
-
-In general, the problem too expensive to solve, since it requires showing that
-for a set of failing inputs, there exist no strictly smaller inputs that also
-cause the property to fail.  Such a demonstration requires exhaustive testing or
-proof.  Solutions to the shrinking problem are therefore approximated to
-efficiently discover sufficiently small inputs causing the input to fail.
+the smallest inputs $i'_0, \ i'_1, \ldots, \ i'_n$, according to some
+measure, such that $\phi(P(i'_0, \ i'_1, \ldots, \ i'_n))$ is also false.
+
+In general, the problem is too expensive to solve optimally, since it requires
+showing that for a set of failing inputs, there exist no strictly smaller inputs
+that also cause the property to fail.  Such a demonstration requires exhaustive
+testing or proof.  Solutions to the shrinking problem are therefore approximated
+to efficiently discover sufficiently small inputs causing the input to fail.
 Usually, the shrunk inputs are derived algorithmically (but perhaps
@@
 
-To address the shrinking problem, QuickCheck defines a type class
+Originally, QuickCheck defines a type class
 \ttp{Arbitrary} that presents a method \ttp{arbitrary} for generating random
@@
 relatively small counterexample returned when QuickChecking a property in (a
-bug-injected version of) Xmonad, a popular X11 window manager written in
+bug-injected version of) \xmonad, a popular X11 window manager written in
 Haskell~\cite{xmonad}:
@@
 Section~\ref{sec:experiments}, we discuss some of our experiences with using
-SmartCheck, including checking properties from XMonad and a natural language
+SmartCheck, including checking properties from \xmonad and a natural language
 processing library.
@@
 The input to the program satisfies the invariant \ttp{pre}, that the sum of
-values in each list of \ttp{Int16}s is less than or equal to 256.  Assuming this,
-we want to show \ttp{post} holds, that the sum of all the values from \ttp{T} is
-less than $5 * 256$, where five is the number of fields in \ttp{T}.  At first
-glance, the property seems reasonable.  But we have forgotten about underflow;
-for example, the value
+values in each list of \ttp{Int16}s is less than or equal to 256.  Assuming
+this, we want to show \ttp{post} holds, that the sum of all the values from
+\ttp{T} is less than $5 * 256$, where five is the number of fields in \ttp{T}.
+At first glance, the property seems reasonable.  But we have forgotten about
+underflow; for example, since $(-20000 + -20000) \ mod \ (2^{15}) = 25536$, and
+$25536 \ge 5*256$, the value
 %
@@
 
-The first way we will control the complexity is focusing on structurally
-shrinking the data type rather than shrinking the base types.  We prevent
-QuickCheck from using its default shrink instance for \ttp{Int16} by defining a
-newtype:
+The first way we will try to contain the search space is by preventing the
+\ttp{shrink} method from operating on numeric types.  The standard approach for
+doing so is by defining a newtype wrapper:
 %
@@
 Next, a programmer might try to control the complexity by truncating lists using
-truncation where
 %
@@
 \begin{code}
-shrink (T w i0 i1 i2 i3) =
-  [ T a b c d e | a <- tk w
-                , b <- tk i0, c <- tk i1
-                , d <- tk i2, e <- tk i3 ]
+shrink (T i0 i1 i2 i3 i4) =
+  [ T a b c d e | a <- tk i0
+                , b <- tk i1, c <- tk i2
+                , d <- tk i3, e <- tk i4 ]
   where tk x = take 10 (shrink x)
@@
 \begin{code}
-shrink (T w i0 i1 i2 i3) = map go xs
-  where xs = shrink (w, i0, i1, i2, i3)
-        go (w', i0', i1', i2', i3')
-           = T w' i0' i1' i2' i3'
+shrink (T i0 i1 i2 i3 i4) = map go xs
+  where xs = shrink (i0, i1, i2, i3, i4)
+        go (i0', i1', i2', i3', i4')
+           = T i0' i1' i2' i3' i4'
 \end{code}
@@
 seconds as well.  For both execution time and final value size, we summarize the
-mean, median, standard deviation, and the results at the 95th percentile.
+mean, standard deviation, and the results at the 95th percentile.
 (While we provide the standard deviations, note that the plots are not
@@
 \begin{itemize}
-\item \ttp{size} returns the size of a value---intuitively, the number of
-  constructors contained within it,
+\item \ttp{size} returns the size of a value---the number of
+  constructors contained within it.
 \item \ttp{index} returns a \emph{sub-value} at a breadth-first index in a
-  value,
-\item \ttp{replace} replaces a sub-value at a particular focus (returning the
-  original value if the index is out-of-bounds),
+  value.
+\item \ttp{replace} replaces a sub-value at a particular focus, returning the
+  original value if the index is out-of-bounds.
 \item \ttp{constr} returns a string representation of the constructor tagging
-  the value,
-\item \ttp{constrs} returns the list of all possible constructors from the
-  value's type,
+  the value.
+\item \ttp{constrs} returns the list of all possible constructor names from the
+  value's type.
 \item \ttp{opaque} is false when the type of the value is an ``interesting
@@
 
-index tree 4  = (Just . SubVal) (B\sub{4} L\sub{6} L\sub{8})
+index tree 3  = (Just . SubVal) (B\sub{4} L\sub{6} L\sub{8})
 index tree 12 = Nothing
@@
 
-subVals (B (B L\sub{0} L\sub{1}) L\sub{2}) =
-  Node (B (B L\sub{0} L\sub{1}) L\sub{2})
-       [ Node (B L\sub{0} L\sub{1}) [Node L\sub{0} [], Node L\sub{1} []]
+subVals (B\sub{0} (B\sub{1} L\sub{0} L\sub{1}) L\sub{2}) = SubVal
+  (Node (B\sub{0} (B\sub{1} L\sub{0} L\sub{1}) L\sub{2})
+        [ Node (B\sub{1} L\sub{0} L\sub{1}) [Node L\sub{0} [], Node L\sub{1} []]
         , Node L\sub{2} []
-       ]
+        ])
 \end{code}
@@
 
-In our implementation, \ttp{constr} and \ttp{constrs} depends on GHC
+In our implementation, the \ttp{SubTypes} class and its methods depend on GHC
 Generics~\cite{generics}, which we describe in Section~\ref{sec:implementation}.
 For simplicity, we omit here Generics-specific super-class constraints on the
-\ttp{SubTypes} class.  Moreover, our presentation simplifies the implementation
-(Section~\ref{sec:implementation}) somewhat to improve the presentation.
+\ttp{SubTypes} class here.  Moreover, our presentation simplifies the
+implementation (Section~\ref{sec:implementation}) somewhat to improve the
+presentation.
 
@@
 structurally smaller sub-values and build a new value from that, leaving
-the remainder of the tree unchanged.  Intuitively, by a \emph{structurally
+the remainder of the tree unchanged.  By a \emph{structurally
   smaller value}, we mean one with fewer constructors.  We continue until we
@@
                  (scMaxReduce args) v
-         maybe (reduce' (idx+1)) (reduce args prop)
-               (test cex idx vs prop)
+         case test cex idx vs prop of
+           Nothing -> reduce' (idx+1)
+           Just a  -> reduce args prop a
     | otherwise = return cex
@@
 Recall that the algorithm begins at the \emph{first} sub-value of the
-counterexample rather than the zeroth sub-value.  The zeroth sub-value of
-\ttp{cex} is \ttp{cex} itself.  No invariant of the algorithm would be violated
-by beginning with the zeroth sub-value, and in particular, the algorithm would
-still terminate.  However, we begin with a strict sub-value because of the
-hypothesis.
+counterexample rather than the zeroth sub-value so that the variant of the
+counterexample remains the same.  No invariant of the algorithm would be
+violated by beginning with the zeroth sub-value, and in particular, the
+algorithm would still terminate.
 
@@
 \emph{opaque types}.  If the reduction algorithm encounters an opaque type, it is
-ignored.  Opaque types include the primitive types mentioned above, but in the
+ignored.  Opaque types include the primitive types mentioned above, but the
 user can declare any substructure in a data type to be an opaque type by
@@
 Opaque types can be conditional.  For example, the user may want lists to be
-shrunk in general, unless the elements of the list are opaque themselves.  Such a
-definition is possible.
+shrunk in general, unless the elements of the list are opaque themselves.  Such
+a definition is possible.  Sometimes shrinking primitive types is imperative;
+for example, to determine if there is a relationship between two integers in a
+counterexample.
 
@@
     | Just v <- index cex idx
-    = maybe (test' v idx) (reduceOpt args prop)
-            (testHole v)
+    = case testHole v of
+        Nothing -> test' v idx
+        Just a  -> reduceOpt args prop a
     | otherwise = return cex
@@
     vs <- newVals (getSize v) (scMaxReduce args) v
-    maybe (reduce' (idx+1)) (reduceOpt args prop)
-          (test cex idx vs prop)
+    case test cex idx vs prop of
+      Nothing -> reduce' (idx+1)
+      Just a  -> reduceOpt args prop a
 
@@
 that implements this optimization.  The function \ttp{testHole} tests the
-current sub-value and if it fails the property, the function tests the sub-value
-directly.
+current sub-value and if it fails the property, then we run the execution
+algorithm on the sub-value directly.
 
@@
 
-In some cases, a formula may be returned that is overly general.  For example,
-consider the counterexample in which both arguments of the outermost \ttp{Add}
-constructor contain values causing the failure:
+Second, in some cases, a formula may be returned that is overly general.  For
+example, consider the counterexample in which both arguments of the outermost
+\ttp{Add} constructor contain values causing the failure:
 %
@@
 \noindent
-Since no matter what random value the first field of \ttp{Add} is replaced with,
-the property fails by the second field and vice versa for the second.  The
-universal generalization algorithm might return the formula
+Since no matter what random value the first field of the outermost \ttp{Add}
+constructor is replaced with, the property fails by \ttp{Add}'s second field,
+and vice versa for replacing the second field.  Consequently, the universal
+generalization algorithm might return the formula
 \noindent
@@
 \noindent
-An over-generalized formula is only a problem when generalizing without
-minimizing the counterexample first, as pointed out to the author by John
-Hughes, so in practice with SmartCheck, this form of unsoundness does not arise.
-However, if one were to generalize without shrinking first, the reader should
-read a universally quantified formula as shorthand for quantifying each variable
-independently and taking the conjunction of formulas.  For example, instead of
+%% An over-generalized formula is only a problem when generalizing without
+%% minimizing the counterexample first, as pointed out to the author by John
+%% Hughes, so in practice with SmartCheck, this form of unsoundness does not arise.
+%% However, if one were to generalize without shrinking first, 
+The reader should read a universally quantified formula as shorthand for
+quantifying each variable independently and taking the conjunction of formulas.
+For example, instead of
 %
@@
   \end{code}
-  \caption{Universal sub-value generation algorithm.}
+  \caption{Existential sub-value generation algorithm.}
   \label{fig:constrs}
@@
 generalized.  The basic idea of the algorithm is to determine whether two values
-have the same ``shape''.  Intuitively, we consider two values to have the same
+have the same ``shape''.  We consider two values to have the same
 shape if in their respective tree representations, their constructors at each
@@
 \ttp{e1} contains the constructor \ttp{C} and \ttp{e2} contains the constructor
-\ttp{Add}.  However, index 0 has been generalized, then we ignore the sub-value
+\ttp{Add}.  If, however, index 0 has been generalized, then we ignore the sub-value
 at index 0, and they match (the third equation).  The same reasoning holds for
@@
 %
-Default instances are provided for common Prelude types and some additional
+Predefined instances are provided for common Prelude types and some additional
 ones, including all types for which QuickCheck provides instances.
 
-SmartCheck does not implement a counterexample discovery algorithm or an
-\ttp{arbitrary} method.  Any initial counterexample can be passed in or
-generated, and the algorithm depends on QuickCheck's \ttp{arbitrary} to generate new
-well-typed values.
+SmartCheck does not implement a counterexample discovery algorithm itself.  An
+initial counterexample can be passed in explicitly, or SmartCheck will use
+QuickCheck as a library to generate an initial counterexample to analyze.
 
@@
 \noindent
-We find that for very large and nested data structures, a tree representation
+We find that for very large data structures, a tree representation
 aids in visually parsing the value.
@@
 \section{Experiments}\label{sec:experiments}
-We describe two experiments using SmartCheck, including an XMonad property and a
+We describe two experiments using SmartCheck, including an \xmonad property and a
 property about a natural language processing library.
@@
 
-\subsection{XMonad}
-Recall from the introduction the XMonad example.  The XMonad window manager is a
+\subsection{\xmonad}
+Recall from the introduction the \xmonad example.  The \xmonad window manager is a
 large software project with many contributors, so naturally, a QuickCheck
 test harness is included to help ensure new commits do not introduce bugs.  At
-the heart of XMonad is a \ttp{StackSet} data type that encodes the relationship
-between windows, work spaces, and which window has the focus.  XMonad contains
+the heart of \xmonad is a \ttp{StackSet} data type that encodes the relationship
+between windows, work spaces, and which window has the focus.  \xmonad contains
 properties to ensure the correct manipulation of \ttp{StackSet}s.  Due to having
-one large data-structure that is essential to the entire program, XMonad is a
+one large data-structure that is essential to the entire program, \xmonad is a
 perfect candidate for SmartCheck.
 
-XMonad passes all of its QuickCheck tests, but let us see what might happen to a
-new contributor if things go awry.  Suppose a developer defines a deletion
-function to delete a window, if it exists.  There is a deletion function that
-exists in XMonad, which is quite complex, given the amount of state that is managed
-by \ttp{StackSet}.  However, one function used in deletion is to filter the
-stack of windows associated with each workspace defined:
+\xmonad passes all of its QuickCheck tests, but let us see what might happen to
+a new contributor if things go awry.  Suppose a developer defines a deletion
+function to delete a window, if it exists.  An existing deletion function in
+\xmonad exists, which is quite complex, given the amount of state that is
+managed by \ttp{StackSet}.  However, one function used in deletion is to filter
+the stack of windows associated with each workspace defined:
 %
@@
 Let us examine what was generalized.  In our test run, we chose to treat data
-maps as opaque, so the forth element of \ttp{StackSet} is not generalized, but
-is simply the empty map, so looks uninteresting.  The second and third fields of
-\ttp{StackSet} are generalized, but the first one is not.  So there is something
+maps as opaque, so the fourth element of \ttp{StackSet} is not generalized, but
+is simply the empty map, which looks uninteresting.  The second and third fields of
+\ttp{StackSet} are generalized, but the first one is not.  There is something
 particular about it.  So the culprit is one of the small constants (1 and -1) or
@@
 In 2012, a question was posted on the programming message board
-Stack Overflow asking about how to shrink large nested data
+Stack Overflow asking about how to shrink large data
 types.\footnote{\url{http://stackoverflow.com/questions/8788542/how-do-i-get-good-small-shrinks-out-of-quickcheck}}
@@
 80-character lines of output.  Most of the savings in the counterexample size
-were due to universal generalization, like in the XMonad case: entire record
+were due to universal generalization, like in the \xmonad case: entire record
 fields are abstracted away.  From that, we (syntactically) shrunk the
-counterexample by-hand further by naming common sub-expressions.
+counterexample by hand further by naming common sub-expressions.
 
 We were able to send a substantially reduced and generalized counterexample to
-the message poster, making the cause of the bug more obvious.  The author responded:
+the message poster, making the cause of the bug more obvious.  The author
+responded (in private communication):
 \begin{quote}
@@
 in seconds.  The results are presented in Table~\ref{table:benchmarks}.  Again, we
-summarize the mean, median, standard deviation, and the results at the 95th
+summarize the mean, standard deviation, and the results at the 95th
 percentile.  (While we provide the standard deviations, note that the plots are

### 2014-06-17 paper/Paper.hs

@@ -102,8 +102,9 @@ reduce args prop cex = reduce' 1
     | Just v <- index cex idx
     = do vs <- newVals (getSize v)
                  (scMaxReduce args) v
-         maybe (reduce' (idx+1)) (reduce args prop)
-               (test cex idx vs prop)
+         case test cex idx vs prop of
+           Nothing -> reduce' (idx+1)
+           Just a  -> reduce args prop a
     | otherwise = return cex
 
 test :: SubTypes a => a -> Idx -> [SubVal]
@@ -124,14 +125,16 @@ reduceOpt args prop cex = reduce' 1
   where
   reduce' idx
     | Just v <- index cex idx
-    = maybe (test' v idx) (reduceOpt args prop)
-            (testHole v)
+    = case testHole v of
+        Nothing -> test' v idx
+        Just a  -> reduceOpt args prop a
     | otherwise = return cex
 
   test' v idx = do
     vs <- newVals (getSize v) (scMaxReduce args) v
-    maybe (reduce' (idx+1)) (reduceOpt args prop)
-          (test cex idx vs prop)
+    case test cex idx vs prop of
+      Nothing -> reduce' (idx+1)
+      Just a  -> reduceOpt args prop a
 
   testHole (SubVal a) = do
     a' <- cast a :: Maybe a

### 2014-06-17 paper/paper.tex
paper: Added Regehr citation.

@@
 power of QuickCheck to \emph{randomly} discover structurally smaller
-counterexamples; in short, this is the first treatment we know of that
-integrates counterexample generation with counterexample reduction.
-Furthermore, neither paper explores the idea of counterexample generalization.
+counterexamples.  The idea of randomization in test-case reduction was developed
+at approximately the same time as SmartCheck and first reported in the domain of
+reducing C~programs that demonstrate compiler bugs~\cite{regehr}.  We believe our
+work is the first to explore the technique in a more general setting.
+Furthermore, we believe ours is the first to explore the idea of counterexample
+generalization.
 

### 2014-06-23 paper/paper.tex
paper: final version edits.

@@
 
-%% \conferenceinfo{ICFP'12,} {September 9--15, 2012, Copenhagen, Denmark.}
-%% \CopyrightYear{2012}
-%% \copyrightdata{978-1-4503-1054-3/12/09}
+\exclusivelicense
+\conferenceinfo{Haskell~'14}{September 6, 2014, Gothenburg, Sweden.}
+\CopyrightYear{2014}
+\copyrightdata{978-1-4503-3041-1/14/09}
+\doi{2633357.2633365}
 
@@
 
-%% \category{D.2.4}{Software/Program Verification}{Reliability}
+\category{D.2.5}{Software Engineering}{Testing and Debugging}
 
@@
 
-%% \keywords
-%% embedded domain-specific language, compiler, verification
+\keywords
+property-based testing; test-case generalization; delta-debugging
 
@@
 
+\begin{figure}[ht]
+\begin{code}
+\scriptsize
+StackSet \{current = Screen \{workspace = Workspace
+\{tag = NonNegative \{getNonNegative = 0\}, layout = -1, stack =
+Just (Stack \{focus = `S', up ="", down = ""\})\}, screen = 1,
+screenDetail = 1\}, visible = [Screen \{workspace = Workspace
+\{tag = NonNegative \{getNonNegative = 2\}, layout = -1, stack =
+Nothing\}, screen = 2, screenDetail = -1\},Screen \{workspace =
+Workspace \{tag = NonNegative \{getNonNegative = 3\}, layout = -1,
+stack = Nothing\}, screen = 0, screenDetail = -1\}], hidden =
+[Workspace \{tag = NonNegative \{getNonNegative = 1\}, layout = -1,
+stack = Just (Stack \{focus = `NUL', up = "", down = ""\})\},
+Workspace \{tag = NonNegative \{getNonNegative = 4\}, layout = -1,
+stack = Just (Stack \{focus = `I', up = "", down = ""\})\}],
+floating = fromList []\}
+\end{code}
+  \caption{Counterexample returned by QuickCheck for a bug-injected version of \xmonad.}
+  \label{fig:xmonadex}
+\end{figure}
+
 In addition, shrinking is one side of the coin when it comes to making
@@
 abstraction is often implicitly made by the programmer in determining the reason
-why counterexamples fail the property.  For example, the following is a
-relatively small counterexample returned when QuickChecking a property in (a
-bug-injected version of) \xmonad, a popular X11 window manager written in
-Haskell~\cite{xmonad}:
-%
-\medskip%
-\begin{sloppypar}
-%\footnotesize
-\noindent%
-\ttp{StackSet \{current = Screen \{workspace = Workspace \{tag = NonNegative
-      \{getNonNegative = 0\}, layout = -1, stack = Just (Stack \{focus = 'S', up =
-        "", down = ""\})\}, screen = 1, screenDetail = 1\}, visible = [Screen
-    \{workspace = Workspace \{tag = NonNegative \{getNonNegative = 2\}, layout = -1,
-        stack = Nothing\}, screen = 2, screenDetail = -1\},Screen \{workspace =
-      Workspace \{tag = NonNegative \{getNonNegative = 3\}, layout = -1, stack =
-        Nothing\}, screen = 0, screenDetail = -1\}], hidden = [Workspace \{tag =
-      NonNegative \{getNonNegative = 1\}, layout = -1, stack = Just (Stack \{focus
-        = `\char`\\NUL', up = "", down = ""\})\},Workspace \{tag = NonNegative
-      \{getNonNegative = 4\}, layout = -1, stack = Just (Stack \{focus = 'I', up =
-        "", down = ""\})\}], floating = fromList []\} }
-\end{sloppypar}
-\medskip%
-\noindent
-(This counterexample uses Haskell's default \ttp{Show} instances, which uses
-record syntax.)  Programmers may be familiar with having to debug
-a ``wall of text'' as shown above.  What if instead a
-formula like the following were returned, stating that for any well-typed
-values $x_0$, $x_1$, $x_2$, and $x_3$, tested, a counterexample is found?
+why counterexamples fail the property.  For example, Figure~\ref{fig:xmonadex}
+contains a relatively small counterexample returned when QuickChecking a
+property in (a bug-injected version of) \xmonad, a popular X11 window manager
+written in Haskell~\cite{xmonad}.  (This counterexample uses Haskell's default
+\ttp{Show} instances, which uses record syntax.)  Programmers may be familiar
+with having to debug a ``wall of text'' as shown above.  What if instead a
+formula like the following were returned, stating that for any well-typed values
+$x_0$, $x_1$, $x_2$, and $x_3$, tested, a counterexample is found?
+
+%% %
+
+%% \medskip%
+%% \begin{sloppypar}
+%% %\footnotesize
+%% \noindent%
+%% \ttp{StackSet \{current = Screen \{workspace = Workspace \{tag = NonNegative
+%%       \{getNonNegative = 0\}, layout = -1, stack = Just (Stack \{focus = 'S', up =
+%%         "", down = ""\})\}, screen = 1, screenDetail = 1\}, visible = [Screen
+%%     \{workspace = Workspace \{tag = NonNegative \{getNonNegative = 2\}, layout = -1,
+%%         stack = Nothing\}, screen = 2, screenDetail = -1\},Screen \{workspace =
+%%       Workspace \{tag = NonNegative \{getNonNegative = 3\}, layout = -1, stack =
+%%         Nothing\}, screen = 0, screenDetail = -1\}], hidden = [Workspace \{tag =
+%%       NonNegative \{getNonNegative = 1\}, layout = -1, stack = Just (Stack \{focus
+%%         = `\char`\\NUL', up = "", down = ""\})\},Workspace \{tag = NonNegative
+%%       \{getNonNegative = 4\}, layout = -1, stack = Just (Stack \{focus = 'I', up =
+%%         "", down = ""\})\}], floating = fromList []\} }
+%% \end{sloppypar}
+%% \medskip%
+%% \noindent
 
@@
 0 and returning \ttp{Nothing} in that case:
+
+\pagebreak
+
 %
@@
          | Div Exp Exp
-
+\end{code}
+%
+\begin{code}
 eval :: Exp -> Maybe Int

### 2014-06-25 paper/paper.tex
final paper edits.

@@
 \usepackage{nth}
+\usepackage{xspace}
 
@@
 
-\newcommand{\xmonad}{\textsc{xmonad}}
+\newcommand{\xmonad}{\textsc{xmonad}\xspace}
 
@@
 \begin{abstract}
-QuickCheck is a powerful library for automatic test-case generation.  Because
-QuickCheck performs random testing, some of the counterexamples discovered are
-very large.  QuickCheck provides an interface for the user to write shrink
-functions to attempt to reduce the size of counterexamples.  Hand-written
-implementations of shrink can be complex, inefficient, and consist of
-significant boilerplate code.  Furthermore, shrinking is only one aspect in
-debugging: counterexample generalization is the process of extrapolating from
-individual counterexamples to a class of counterexamples, often requiring a
-flash of insight from the programmer.  To improve counterexample reduction and
-generalization, we introduce \emph{SmartCheck}.  SmartCheck is a debugging tool
-that reduces algebraic data using generic search heuristics to efficiently find
-smaller counterexamples.  In addition to shrinking, SmartCheck also
-automatically generalizes counterexamples to formulas representing classes of
-counterexamples.  SmartCheck has been implemented for Haskell and is freely
-available.
+QuickCheck is a powerful library for automatic test-case generation.
+Because QuickCheck performs random testing, some of the counterexamples discovered are very large.
+QuickCheck provides an interface for the user to write shrink functions to attempt to reduce the size of counterexamples.
+Hand-written implementations of shrink can be complex, inefficient, and consist of significant boilerplate code.
+Furthermore, shrinking is only one aspect in debugging: counterexample generalization is the process of extrapolating from individual counterexamples to a class of counterexamples, often requiring a flash of insight from the programmer.
+To improve counterexample reduction and generalization, we introduce \emph{SmartCheck}.
+SmartCheck is a debugging tool that reduces algebraic data using generic search heuristics to efficiently find smaller counterexamples.
+In addition to shrinking, SmartCheck also automatically generalizes counterexamples to formulas representing classes of counterexamples.
+SmartCheck has been implemented for Haskell and is freely available.
 \end{abstract}
@@
 The QuickCheck testing framework was a revolutionary step-forward in
-type-directed random testing~\cite{qc,monadic}.  Originally designed for Haskell,
+property-based testing~\cite{qc,monadic}.  Originally designed for Haskell,
 QuickCheck has been ported to other languages and is a now a widely-used testing
@@
 That is, automated testing helps find a counterexample, but it may be a very
-large counterexample.  We are left facing the \emph{shrinking problem}: assume
-there exist inputs $i_0, \ i_1, \ldots, \ i_n$, such that for program
-$P$ (modeled as a function mapping inputs to output) and property $\phi$,
-$\phi(P(i_0, \ i_1, \ldots, \ i_n))$ is false.  The shrinking problem is to find
-the smallest inputs $i'_0, \ i'_1, \ldots, \ i'_n$, according to some
-measure, such that $\phi(P(i'_0, \ i'_1, \ldots, \ i'_n))$ is also false.
+large counterexample.  We are left facing the \emph{shrinking problem}: consider
+program $P$ modeled as a function mapping an input vector $\vec{i}$ to an output
+value.  Consider property $\phi$ modeled as a function mapping the inputs of $P$
+to a Boolean value that has the form $pre(\vec{i}) \Longrightarrow
+post(P(\vec{i}))$, where $pre$ is a precondition on inputs and $post$ is a
+postcondition on the program output.  Let $<$ be a partial order over the set of
+possible inputs.  Given inputs $\vec{i}$ such that $\phi(\vec{i})$ is false, the
+shrinking problem is to find inputs $\vec{j}$ such that $\phi(\vec{j})$ is false
+and there exist no inputs $\vec{k}$ such that $\vec{k} < \vec{j}$ and
+$\phi(\vec{k})$ is false.
 
@@
 Usually, the shrunk inputs are derived algorithmically (but perhaps
-nondeterministically) from a known larger set of inputs resulting in failure,
-and the inputs may not be minimal.
+nondeterministically) from known larger inputs resulting in failure.
 
@@
 counterexamples more understandable: the other side is extrapolation from
-individual counterexamples to a class of counterexamples.  This leap of
-abstraction is often implicitly made by the programmer in determining the reason
-why counterexamples fail the property.  For example, Figure~\ref{fig:xmonadex}
-contains a relatively small counterexample returned when QuickChecking a
-property in (a bug-injected version of) \xmonad, a popular X11 window manager
-written in Haskell~\cite{xmonad}.  (This counterexample uses Haskell's default
-\ttp{Show} instances, which uses record syntax.)  Programmers may be familiar
-with having to debug a ``wall of text'' as shown above.  What if instead a
-formula like the following were returned, stating that for any well-typed values
-$x_0$, $x_1$, $x_2$, and $x_3$, tested, a counterexample is found?
+individual counterexamples to a class of counterexamples characterizing the bug.
+This leap of abstraction is often implicitly made by the programmer in
+determining the reason why counterexamples fail the property.  For example,
+Figure~\ref{fig:xmonadex} contains a relatively small counterexample returned
+when using QuickCheck to test a property in (a bug-injected version of) \xmonad,
+a popular X11 window manager written in Haskell~\cite{xmonad}.  (This
+counterexample uses Haskell's default \ttp{Show} instances, which uses record
+syntax.)  Programmers may be familiar with having to debug a ``wall of text'' as
+shown above.  What if instead a formula like the following were returned,
+stating that for any well-typed values $x_0$, $x_1$, $x_2$, and $x_3$, tested, a
+counterexample is found?
 
@@
 Motivated by the problems of reducing and generalizing large counterexamples, we
-developed SmartCheck.  SmartCheck takes a counterexample produced by some
-oracle and generically minimizes and generalizes the counterexample.
+developed SmartCheck.  SmartCheck takes a counterexample produced by some oracle
+and generically minimizes and generalizes the counterexample.  After presenting
+some preliminary definitions in Section~\ref{sec:preliminaries}, in
+Section~\ref{sec:shrinking}, we describe SmartCheck's generic counterexample
+reduction algorithm.
 
@@
 %% important in real-world uses; see Section~\ref{sec:experiments} for examples.)
-
-After presenting some preliminary definitions in
-Section~\ref{sec:preliminaries}, in Section~\ref{sec:shrinking}, we describe
-SmartCheck's generic counterexample reduction algorithm.  The algorithm is
-particularly novel in that it generates new random sub-values during the
-reduction phase (see related work in Section~\ref{sec:related}).
 
@@
 simple property can result in large counterexamples without any shrinking.  Then
-we show the difficulty in designing an efficient shrink implementation manually,
-which is required when using QuickCheck; we will show a poor
-design before arriving at a ``canonical'' manual solution.%%   However, even the
+we show the difficulty in designing an efficient shrink implementation.  We will
+show a poor design before arriving at a ``canonical'' manual solution.
+
+%% %%
+%% However, even the
 %% canonical solution produces significantly larger counterexamples and takes
@@
 troublesome is that with QuickCheck's default \ttp{Arbitrary} instances, the
-length of potential counterexamples returned by \ttp{shrink} can easily exceed
+length of potential counterexamples returned by \ttp{shrink} can exceed
 $10^{10}$, which is an intractable number of tests for QuickCheck to analyze.
@@
 In early 2014, QuickCheck added the ability to derive \ttp{shrink} instances for
-algebraic data type using the same generic package that SmartCheck does; see
-Section~\ref{sec:preliminaries}.  (The first release of the SmartCheck software
-was mid-2012.)  With this new feature, the user declares
+algebraic data types.  (The first release of the SmartCheck software was
+mid-2012.)  With this new feature, the user declares
 %
@@
 
+index tree 0  = tree
 index tree 3  = (Just . SubVal) (B\sub{4} L\sub{6} L\sub{8})
@@
 value returned by \ttp{index a i} as the \emph{$i$th sub-value of $a$}, so for
-example, (\ttp{B\sub{4} L\sub{6} L\sub{8}}) is the 4th sub-value of \ttp{tree}.
+example, (\ttp{B\sub{4} L\sub{6} L\sub{8}}) is the 3rd sub-value of \ttp{tree}.
 An invariant of \ttp{index} is that for any value \ttp{a}, and for the smallest
@@
 
-The hypothesis may be incorrect for some properties, in which case SmartCheck
-may potentially fail to discover a smaller counterexample.  However, in
-Sections~\ref{sec:existential} and~\ref{sec:precondition}, we describe
-approaches to generalize counterexamples based on discovering new counterexample
-variants.  These generalization techniques are executed in an (optional)
-generalization phase, run after the reduction phase, in which this hypothesis is
-implemented.
+Incidentally, QuickCheck's generic shrink implementation is fundamentally build
+around the variant counterexample hypothesis.  For a given counterexample
+\ttp{cex}, smaller candidates produced by \ttp{shrink} contain only the
+constructors found in \ttp{cex} as opposed to just the outermost constructor, as
+in SmartCheck.  Sometimes shrinking gets ``stuck'' at a local minimum due to a
+lack of entropy in generating smaller candidates.
+
+The hypothesis may be unsuitable for some properties, in which case SmartCheck
+(and especially QuickCheck) may potentially fail to discover a smaller
+counterexample.  However, in Sections~\ref{sec:existential}
+and~\ref{sec:precondition}, we describe approaches to generalize counterexamples
+based on discovering new counterexample variants.  These generalization
+techniques are executed in an (optional) generalization phase, run after the
+reduction phase, in which this hypothesis is implemented.
 
@@
 0 and returning \ttp{Nothing} in that case:
-
-\pagebreak
-
 %
@@
   prop cex}) returns (\ttp{Just True}) if \ttp{cex} passes \ttp{prop}'s
-precondition but fails the property; (\ttp{Just False}) if \ttp{cex}
+precondition but fails the property, (\ttp{Just False}) if \ttp{cex}
 non-trivially satisfies \ttp{prop}, and \ttp{Nothing} if \ttp{cex} fails
@@
 
-Like in the reduction algorithm user-specified flags bound the behavior of the
+Like in the reduction algorithm, user-specified flags bound the behavior of the
 algorithm.  We bound the size of values to generate by the flag \ttp{scMaxSize},
@@
 The algorithm's complexity is $\mathcal{O}(n)$, where $n$ is the number of
-constructors in the counterexample.  Again, we assume that the cost for each
+constructors in the counterexample.  Again, we assume that the cost for
 generating random values and testing them at each index is constant.
@@
 is sound.  The existential claim is only that for each variant, there exists at
-least one counterexample, so inductive reasoning is not required.
+least one counterexample.
 
@@
 matchesShape e1 (e2, [])  == False
-matchesShape e1 (e2, [0]) == True
-matchesShape e3 (e2, [0]) == True
+matchesShape e1 (e2, [1]) == True
+matchesShape e3 (e2, [1]) == True
 \end{code}
@@
 The first equation holds because we ignore opaque types, so the arguments to
-\ttp{C} are considered equal.  The second equation fails because at index 0,
+\ttp{C} are considered equal.  The second equation fails because at index 1,
 \ttp{e1} contains the constructor \ttp{C} and \ttp{e2} contains the constructor
-\ttp{Add}.  If, however, index 0 has been generalized, then we ignore the sub-value
-at index 0, and they match (the third equation).  The same reasoning holds for
+\ttp{Add}.  If, however, index 1 has been generalized, then we ignore the sub-value
+at index 1, and they match (the third equation).  The same reasoning holds for
 the fourth equation.
@@
 generics library to encode generic traversals.  Specifically, SmartCheck uses
-``GHC Generics'', a recently-developed generics library for
-Haskell~\cite{generics}.  The generics library allows algebraic data types to
-automatically derive the type class \ttp{SubTypes} presented in
-Section~\ref{sec:preliminaries}.  One limitation with GHC Generics is that it
-does not (currently) support generalized algebraic data types~\cite{gadts}.
+``GHC Generics'', a generics library for Haskell~\cite{generics}.  The library
+allows algebraic data types to automatically derive the type class
+\ttp{SubTypes} presented in Section~\ref{sec:preliminaries}.  One limitation
+with GHC Generics is that it does not (currently) support generalized algebraic
+data types~\cite{gadts}.
 
@@
 We describe two experiments using SmartCheck, including an \xmonad property and a
-property about a natural language processing library.
+property about a natural language processing library.  Then we present a small
+set of benchmarks comparing SmartCheck and QuickCheck.
 
@@
 \noindent
-QuickCheck returns the large value shown in the introduction.  That value is a
-relatively small counterexample, but even the smallest \ttp{StackSet} values are
-somewhat visually overwhelming due to the number of fields within it.  Recall
-the value returned by SmartCheck after generalization:
+QuickCheck returns the large value shown in Figure~\ref{fig:xmonadex}.  That
+value is a relatively small counterexample, but even the smallest \ttp{StackSet}
+values are somewhat visually overwhelming due to the number of fields within it.
+Recall the value returned by SmartCheck after generalization:
 %
@@
   \item \emph{Parser}, a parser/pretty-printer for a toy imperative language
-    containing a parser bug that switches the arguments of an binary expression
-    operator.
+    containing a parser bug that switches the arguments of disjunction
+    expressions.
 \end{itemize}
@@
 \multirow{2}{*}{Parser}
-  & QC &  4   & 0.024 & 0  & 0.020 & 4 & 0.061 \\\cline{2-8}
-  & SC &  6   & 0.116 & 2  & 0.083 & 9 & 0.230 \\\hline
+  & QC &  4   & 0.010 & 0  & 0.006 & 4  & 0.023 \\\cline{2-8}
+  & SC &  7   & 0.182 & 3  & 0.124 & 12 & 0.418 \\\hline
 
@@
 
-The \emph{Reverse} benchmark essentially provides a lower-bound on the benefit
+The \emph{Reverse} benchmark essentially provides a lower-bound on the benefit of
 shrinking in general, since the original counterexamples are generally close to
@@
 benchmarks, while QuickCheck shrinking finds smaller counterexamples faster in
-the \emph{Parser} example.  The \emph{Parser} example is one in which
-SmartCheck's strategy of generating random values during counterexample
-reduction is not a benefit, and so this is why it likely under performs a
-generic \ttp{shrink} method.  Moreover, because SmartCheck uses QuickCheck as a
-front-end, the two approaches are complementary: QuickCheck shrinking can be
-used instead of or in combination with SmartCheck reduction.
+the \emph{Parser} example.  The example is one in which SmartCheck's
+counterexample reduction strategy is less optimal than QuickCheck's.  Recall
+from Section~\ref{sec:variant} that QuickCheck's generic shrink implementation
+generates candidates that contain a subset of the constructors from original
+counterexample.  In the parser example, the bug is localized in the
+counterexample, arising from a single expression in the program.  SmartCheck
+wastes effort generating new programs using new constructors. SmartCheck is
+better suited, however, at avoiding local minima for other properties and
+programs.
 
@@
 power of QuickCheck to \emph{randomly} discover structurally smaller
-counterexamples.  The idea of randomization in test-case reduction was developed
-at approximately the same time as SmartCheck and first reported in the domain of
-reducing C~programs that demonstrate compiler bugs~\cite{regehr}.  We believe our
-work is the first to explore the technique in a more general setting.
-Furthermore, we believe ours is the first to explore the idea of counterexample
-generalization.
+counterexamples.  The idea of randomization in test-case reduction was
+independently developed at approximately the same time as SmartCheck and first
+published in the domain of reducing C~programs that demonstrate compiler
+bugs~\cite{regehr}. We believe our work is the first to explore
+the idea of counterexample generalization.
 
@@
 relationships between variables in executions of a program~\cite{daikon}.
-SmartCheck does not attempt to infer properties like these tools do, so it is in
-one sense less general.  However, SmartCheck's three kinds of counterexample
-generalization are novel.
+SmartCheck does not attempt to infer properties like these tools dox.
 
